### 2020-04-30 paper/latex/paper.tex
Add paper.tex

% -----------------------------------------------
% Template for ISMIR Papers
% 2020 version, based on previous ISMIR templates

% Requirements :
% * 6+n page length maximum
% * 4MB maximum file size
% * Copyright note must appear in the bottom left corner of first page
% * Clearer statement about citing own work in anonymized submission
% (see conference website for additional details)
% -----------------------------------------------

\documentclass{article}
\usepackage[T1]{fontenc} % add special characters (e.g., umlaute)
\usepackage[utf8]{inputenc} % set utf-8 as default input encoding
\usepackage{ismir,amsmath,cite,url}
\usepackage{graphicx}
\usepackage{color}

% Optional: To use hyperref, uncomment the following.
% \usepackage[bookmarks=false,hidelinks]{hyperref}
% Mind the bookmarks=false option; bookmarks are incompatible with ismir.sty.

\usepackage{lineno}
\linenumbers

% Title.
% ------
\title{Deep learning approaches on musical assisted-orchestration: an evaluation study}

% Note: Please do NOT use \thanks or a \footnote in any of the author markup

% Single address
% To use with only one author or several with the same address
% ---------------
%\oneauthor
% {Names should be omitted for double-blind reviewing}
% {Affiliations should be omitted for double-blind reviewing}

% Two addresses
% --------------
%\twoauthors
%  {First author} {School \\ Department}
%  {Second author} {Company \\ Address}

%% To make customize author list in Creative Common license, uncomment and customize the next line
%  \def\authorname{First Author, Second Author}

% Three addresses
% --------------
\threeauthors
  {First Author} {Affiliation1 \\ {\tt author1@ismir.edu}}
  {Second Author} {\bf Retain these fake authors in\\\bf submission to preserve the formatting}
  {Third Author} {Affiliation3 \\ {\tt author3@ismir.edu}}

%% To make customize author list in Creative Common license, uncomment and customize the next line
%  \def\authorname{First Author, Second Author, Third Author}

% Four or more addresses
% OR alternative format for large number of co-authors
% ------------
%\multauthor
%{First author$^1$ \hspace{1cm} Second author$^1$ \hspace{1cm} Third author$^2$} { \bfseries{Fourth author$^3$ \hspace{1cm} Fifth author$^2$ \hspace{1cm} Sixth author$^1$}\\
%  $^1$ Department of Computer Science, University , Country\\
%$^2$ International Laboratories, City, Country\\
%$^3$  Company, Address\\
%{\tt\small CorrespondenceAuthor@ismir.edu, PossibleOtherAuthor@ismir.edu}
%}
%\def\authorname{First author, Second author, Third author, Fourth author, Fifth author, Sixth author}

\sloppy % please retain sloppy command for improved formatting

\begin{document}

%
\maketitle
%
\begin{abstract}
The abstract should be placed at the top left column and should contain about 150-200 words.
\end{abstract}
%
\section{Introduction}\label{sec:introduction}

\section{Our model}

\subsection{Baseline}
In order to have a baseline to compare our results against, we attempted to solve the classification problem using various linear classifiers. The classifiers we tested were Support Vector Machine (SVM), Random Forest (RF), and K-Nearest Neighbors (KNN). We used the implementations provided in the scikit-learn library for each classifier. For SVM, we used SVC with an RBF kernel. For RF, we set the maximum depth of each tree to be 15. Each classifier was wrapped in a MultiOutputClassifier to achieve the multi-label nature of this problem. We found SVM to have the highest accuracy of the three classifiers across all experiments.

We started by simplifying the problem to classifying only the instruments and not the pitch. This had the benefit of both reducing the number of classes and increasing the number of samples per class.

\subsection{CNN with LSTM}

\subsection{ResNet}

\section{Orchestration experiments}

\section{Evaluation and Conclusions}

\subsection{Interpreting the Latent Space}

\section{Future steps}

\section{Citations}

All bibliographical references should be listed at the end,
inside a section named ``REFERENCES,'' numbered and in alphabetical order.
All references listed should be cited in the text.
When referring to a document, type the number in square brackets
\cite{Author:00}, or for a range \cite{Author:00,Someone:10,Someone:04}.

When the following words appear in the conference publication titles, please abbreviate them: Proceedings $\rightarrow$ Proc.; Record $\rightarrow$ Rec.; Symposium $\rightarrow$ Symp.; Technical Digest $\rightarrow$ Tech. Dig.; Technical Paper $\rightarrow$ Tech. Paper; First $\rightarrow$ 1st; Second $\rightarrow$ 2nd; Third $\rightarrow$ 3rd; Fourth/nth $\rightarrow$ 4th/nth.

\textcolor{red}{As submission is double blind, refer to your own published work in the third person. That is, use ``In the previous work of \cite{Someone:10},'' not ``In our previous work \cite{Someone:10}.'' If you cite your other papers that are not widely available (e.g., a journal paper under review), use anonymous author names in the citation, e.g., an author of the form ``A. Anonymous.''}

% For bibtex users:
\bibliography{ISMIRtemplate}

% For non bibtex users:
%\begin{thebibliography}{citations}
%
%\bibitem {Author:00}
%E. Author.
%``The Title of the Conference Paper,''
%{\it Proceedings of the International Symposium
%on Music Information Retrieval}, pp.~000--111, 2000.
%
%\bibitem{Someone:10}
%A. Someone, B. Someone, and C. Someone.
%``The Title of the Journal Paper,''
%{\it Journal of New Music Research},
%Vol.~A, No.~B, pp.~111--222, 2010.
%
%\bibitem{Someone:04} X. Someone and Y. Someone. {\it Title of the Book},
%    Editorial Acme, Porto, 2012.
%
%\end{thebibliography}

\end{document}

### 2020-04-30 paper/latex/paper.tex
Finshing up baseline section

@@
 In order to have a baseline to compare our results against, we attempted to solve the classification problem using various linear classifiers.
-The classifiers we tested were Support Vector Machine (SVM), Random Forest (RF), and K-Nearest Neighbors (KNN).
+The classifiers we tested were Support Vector Machine (SVM), Random Forest, and K-Nearest Neighbors.
 We used the implementations provided in the scikit-learn library for each classifier.
 For SVM, we used SVC with an RBF kernel.
-For RF, we set the maximum depth of each tree to be 15.
+For Random Forest, we set the maximum depth of each tree to be 15.
 Each classifier was wrapped in a MultiOutputClassifier to achieve the multi-label nature of this problem.
 We found SVM to have the highest accuracy of the three classifiers across all experiments.
+All of the following baselin experiments used 50,000 generated samples with a train-test split of 60/40.
+Each sample is a combination of one or more instruments and is four seconds in length.
+The features used are the mel frequency cepstral coefficients (MFCCs) of the resulting combination, with a total of 100 coefficients per sample.
 
-We started by simplifying the problem to classifying only the instruments and not the pitch. This had the benefit of both reducing the number of classes and increasing the number of samples per class.
+We started by simplifying the problem to classifying only the instruments and not the pitch.
+This had the benefit of both reducing the number of classes and increasing the number of samples per class.
+We found that SVM was able to very accurately identify the instrument given an input that had only one instrument present; for this case the accuracy was 99.8\%.
+However as soon as multiple instruments were present in the input, the accuracy dropped significantly.
+With two instruments, accuracy was 55.4\%, with three it was 19.6\% and with 10 instruments the accuracy was 0.5\%.
+
+To better approximate the problem of identifying instrument and pitch, we then attempted to classify the instrument and pitch class.
+That is, which octave the pitch was in did not matter, only the pitch class.
+The input was a combination of two instruments drawn from a possible twelve instruments, and the classifier attempted to identify which instruments were present and for two of those instruments, say Flute and Violin, which pitch classes were present.
+If another instrument was present that was not Flute or Violin, the classifier would attempt to identify that instrument, but not its pitch class.
+The best results from this setting of the problem was SVM with 30.2\% accuracy, which was a result of classifying the pitch class of Flute and Violin.
+Depending on which two instruments had their pitch class identified, the accuracy varied greatly.
+For a Violin and Cello accuracy was 20.9\%, and for Trombone and Bass Tuba accuracy was 2.2\%.
+
+For a slight modification on this setting, we no longer attempted to identify which instrument was present if it was not one of the two instruments whose pitch class was being identified.
+Instead, the classifier would simply identify that an instrument that was not one of the two was present.
+Since this is a simpler problem, it lead to increased accuracies.
+Flute and Violin went from 30.2\% to 38.8\%, Oboe and French Horn from 27.3\% to 39.9\%, and Trombone and Bass Tuba from 2.2\% to 3.5\%.
+\textit{(I think we should maybe put a table of this data instead of listing it all out)} Random Forest performed significantly worse for this setting.
+Flute and Violin had an accuracy of 9.8\% and Oboe and French Horn was 17.5\%.
+For this reason, we stopped testing with Random Forest and continued with SVM only.
+
+We then performed this same experiment with three instruments having their pitch class identified.
+Each input was a combination of three instruments drawn from a possible twelve instruments.
+For three instruments specified in that experiment, pitch class was identified.
+Flute, Oboe, and Violin reached an accuracy of 11.1\%, and Bass Tuba, Trumpet, and Trombone was 0.5\%.
+As we increased the number of instruments whose pitch classes was being identified, the accuracy continued to drop.
+For classifying the pitch class of four instruments: Oboe, French Horn, Violin, and Flute, the accuracy was 2.7\%.
+
+This was still a simplified version of the problem, as we were identifying only the pitch class of a few instruments.
+However, the linear classifiers were unable to achieve accurate results as the number of instruments increased.
+Therefore, we did not attempt the full setting of the problem, in which individual pitches are classified for all instruments, with linear classifiers.
 
@@
 \section{Orchestration experiments}
+List of experiments:
+
+- CNN with varying n
+
+- ResNet with n=10
 

### 2020-05-01 paper/latex/paper.tex
Put in Carmine's notes for each section

@@
 \section{Our model}
+- deep learning trained for classification and then used for orchestraion
+- changing number of instruments
 
@@
 \subsection{ResNet}
+- SVM was good but CNN was better and ResNet was best
+- review all experiments we did and choose key plots
 
 \section{Orchestration experiments}
-List of experiments:
+- set of 10 targets
+- orchestra fix to 10
+- qualitative evaluation: acoustic inspection of the solutions
+- quantitative evaluation: distance in feature space
+- comparison table between Orchidea and our model
 
-- CNN with varying n
+We performed experiments varying the number instruments used to form combinations, and varying the number of instruments used in each combination.
+(insert table of data showing the results of CNN with 2,3,4 etc instruments).
+An arbitrary number of samples can be used in the solution, since the $n$ highest probabilities can be taken from the output, leading to $n$ samples in the solution.
 
-- ResNet with n=10
+We tested our models using 15 targets for orchestration.
+These targets are not made from combinations of the input data, and some are not even recordings of instruments.
+Among the targets are recordings of bells, a car horn, a gong, and an environmental recording of a boat docking.
+By passing these samples into a model, an orchestration of the target is created from the TinySOL samples that were used to train the network.
 
 \section{Evaluation and Conclusions}
+- the approach seems promising for orchestration
+- many things used in Orchidea are not implemented here (symbolic constraints, sparsity,...)
+- CNN seems better for timbre
+- ResNet seems better for pitch (what are timbre and pitch??)
 
 \subsection{Interpreting the Latent Space}
+- the system finds filters like...
 
 \section{Future steps}
+- using conditioning to impose symbolic constraints
+- variable size solutions
+- joint networks for orchestral size detection and orchestral families (see paper)
 

### 2020-05-01 paper/latex/paper.tex
Add to experiments section

@@
 
+We did not train our model to classify the dynamics of a sample despite TinySOL having pp, mf, and ff recordings for each sample.
+If we had, then each class would have only one data point.
+Instead, when the model is used for orchestration, the dynamic is determined by the probability of that sample as output by the model.
+If the model output a probability higher than $0.66$ for a sample, the ff version of the sample was used.
+If it was between $0.33$ and $0.66$, then the mf version, and if less than $0.33$ the pp dynamic.
+
 \subsection{Baseline}
@@
 We tested our models using 15 targets for orchestration.
-These targets are not made from combinations of the input data, and some are not even recordings of instruments.
-Among the targets are recordings of bells, a car horn, a gong, and an environmental recording of a boat docking.
+Two of the targets were made from TinySOL samples, but the rest are not combinations of input data, and some are not even recordings of instruments.
+Among the targets are recordings of bells, a car horn, a gong, and a recording of a boat docking.
 By passing these samples into a model, an orchestration of the target is created from the TinySOL samples that were used to train the network.
+For orchestrating these targets, we used both the CNN and ResNet models, each trained on combinations of 10 instruments from a possible 10 instruments.
+The CNN was trained for 49 epochs and the ResNet for 24.
+
+We evaluated our orchestrations both qualitatively and quantitatively.
+Qualitative inspection was done through an acoustic inspection of the solution, paying close attention to timbre and pitch.
+For targets that had harmonic content, it was noted if the partials present in the target were also represented in the orchestrated solution.
+For example, one of the samples of a bell had partials that represented a C sharp minor chord.
+The ResNet .
+For a target that contained specific notes, the solution from ResNet contained both the note in the target and its partials: a sample that was two octaves higher, the fifth an octave up, and a minor third two octaves up (mix\_ObA4\_BnC3).
 

### 2020-05-04 paper/latex/paper.tex
Add quantitative evaluation

@@
 
-We performed experiments varying the number instruments used to form combinations, and varying the number of instruments used in each combination.
-(insert table of data showing the results of CNN with 2,3,4 etc instruments).
-An arbitrary number of samples can be used in the solution, since the $n$ highest probabilities can be taken from the output, leading to $n$ samples in the solution.
+Our final model was trained on data where each input was a combination of ten instruments, but we performed experiments with varying numbers of instruments used in combination.
+(insert table of data showing the results of CNN with 2,3,4 etc instruments) An arbitrary number of samples can be used in the solution, since the $n$ highest probabilities can be taken from the output, leading to $n$ samples in the solution.
 
@@
 
+For quantitative evaluation, we orchestrated the same targets with OrchIdea, then compared the distance between our solutions and the targets and OrchIdea's solutions and the targets.
+The distance metric we used was
+$$d(x^t, \tilde{x}^t) =\lambda_1 \sum_k \delta_k^t(x_k^t - \tilde{x	}_k^t) + \sum_k \delta_k^t|x_k^t - \tilde{x	}_k^t|$$
+$$\delta_k^t = 
+\begin{cases}
+1, \text{if} (x_k - \tilde{x}_k) < 0 \\
+0, \text{otherwise}
+\end{cases} $$
 \section{Evaluation and Conclusions}
@@
 - joint networks for orchestral size detection and orchestral families (see paper)
+
+Future steps in this project include implementing various methods that are present in OrchIdea.
+
+Our current model orchestrates all targets using the same number of samples, and this does not take into account the density of different targets.
+The solution to this is to allow sparse solutions in which the model decides how many samples should be used to best represent the target.
+This allows a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds.
+
+Partial filtering is a method that would aid our model in orchestrating the harmonics of a target.
+The dominant harmonic partials of the target are identified, and the search space is limited to only include samples of those pitches.
+For example, if the target is a recording of an instrument playing a C4, then the partials identified may be C4, C5, G5, and E6.
+The model would then only consider samples of these pitches to be used in the solution.
+This leads to a solution whose harmonics are much closer to the target, which is an important part of aural similarity.
 

### 2020-05-05 paper/latex/paper.tex
add 'cnn with lstm' and 'resnet' part

@@
 \subsection{CNN with LSTM}
+The first deep model we tested was Convolutional Neural Network(CNN).
+CNN shows good performance on classification assignment for its strong ability to extract spacial features, so we implemented a basic CNN with three convolutional layers and two fully connected layers.
+Each convolutional layer is followed by a BatchNorm layer, a ReLU activation layer and a 2 * 2  MaxPool layer with a stride of 2.
+The kernel size is 3 * 3 with a stride of 1 and a padding of 1.
+The number of filters are 8, 16, 32.
+This setting is simple but we can make a good comparison between it and the baseline.
+
+- figure
+
+Given that audio is one kind of sequential data, it is easy to consider Long Short-Term Memory network(LSTM) as test model.
+In our setting, we added one LSTM layer with 32 outputs units following three convolutional layers.
+The convolutional layers did not change the sequence relationship.
+Instead, it scaled down the size of feature maps.
+After LSTM layer, we reshaped the output into the same size before inputing.
+Then we implemented another convolutional layer with the same structure and kernel size.
+Finally, we flattened out the outputs and fed them into fully connected layers with Dropout layer.
+The probability of an element to be zeroed is 0.5 and we took sigmoid as activation function.
+We used binary classification for each class and treated them independently.
 
@@
 - review all experiments we did and choose key plots
+
+To make our setting more reasonable, we took the well-known deep residual network(ResNet) as backbone in our experiments.
+Specifically, we used 18-layer ResNet.
+Besides, to make the model more suitable to our problem, we reset the output channel numbers of each block with 32, 64, 32, 32 respectively.
 

### 2020-05-05 paper/latex/paper.tex
add data generation section

@@
 
+\section{Data generation}
+We used the TinySOL database to create our input data.
+\textbf{(Should we talk more about TinySOL here?)} When training a model, the number of samples $n$ to be used in each combination is specified.
+To create the training data, $n$ instruments are randomly selected from the list of specified instruments, and then a TinySOL sample is randomly selected for each chosen instrument.
+These samples are then combined to play simultaneously.
+This melspectrogram is then taken of this combinations, creating the features to be fed to the model.
+We created the melspectrogram by \textbf{(Alejandro can you write this of how you generated the melspectrograms using the STFT?)}
+
 \section{Our model}
@@
 
-We did not train our model to classify the dynamics of a sample despite TinySOL having pp, mf, and ff recordings for each sample.
+Our idea is perform orchestration using deep neural networks trained for classification.
+By generating input data that is combinations of different instruments and pitches, the network learns how to take a sound that is a complex combination of notes and timbres and deconstruct it into its original parts.
+When a target sound is then fed to the network, it attempts to classify which samples are present.
+By taking the samples that have the highest probability of being in the target, an orchestrated solution can be created.
+
+We did not train our model to classify the dynamics of a sample despite TinySOL having pianissimo, mezzo forte, and fortissimo recordings for each sample.
 If we had, then each class would have only one data point.
 Instead, when the model is used for orchestration, the dynamic is determined by the probability of that sample as output by the model.
-If the model output a probability higher than $0.66$ for a sample, the ff version of the sample was used.
-If it was between $0.33$ and $0.66$, then the mf version, and if less than $0.33$ the pp dynamic.
+If the model output a probability higher than $0.66$ for a sample, the fortissimo version of the sample was used.
+If it was between $0.33$ and $0.66$, then the mezzo forte version, and if less than $0.33$ the pianissimo dynamic.
 
@@
 \subsection{CNN with LSTM}
+
 The first deep model we tested was Convolutional Neural Network(CNN).
@@
 For example, one of the samples of a bell had partials that represented a C sharp minor chord.
-The ResNet .
 For a target that contained specific notes, the solution from ResNet contained both the note in the target and its partials: a sample that was two octaves higher, the fifth an octave up, and a minor third two octaves up (mix\_ObA4\_BnC3).
@@
 The distance metric we used was
-$$d(x^t, \tilde{x}^t) =\lambda_1 \sum_k \delta_k^t(x_k^t - \tilde{x	}_k^t) + \sum_k \delta_k^t|x_k^t - \tilde{x	}_k^t|$$
+
+$$d(x^t, \tilde{x}^t) =\lambda_1 \sum_k \delta_k^t(x_k^t - \tilde{x}_k^t) + \sum_k \delta_k^t|x_k^t - \tilde{x	}_k^t|$$
 $$\delta_k^t = 
@@
 \end{cases} $$
+
 \section{Evaluation and Conclusions}
@@
 - ResNet seems better for pitch (what are timbre and pitch??)
+
+Based on our accuracies from training the model and the results of orchestrating 15 targets, our approach seems promising for orchestration.
+When comparing our method with OrchIdea, it is important to note that many of processes OrchIdea uses to optimize its solution were not used in our model.
+This includes important aspects such as sparsity, partial filtering, \textbf{(Carmine any more to add?)}.
+Our model also lacks the implementation of symbolic constraints, which is an important part of assisted orchestration.
+
+(Alejandro, add how there were certain instruments that performed poorly. Add how ResNet was awful for Flute and always output Flute C4)
+
+We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets.
+Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the pitch(es) of the target.
 

### 2020-05-07 paper/latex/paper.tex
Edit data gen section

@@
 \section{Data generation}
-We used the TinySOL database to create our input data.
-\textbf{(Should we talk more about TinySOL here?)} When training a model, the number of samples $n$ to be used in each combination is specified.
+We used the TinySOL database to create our input data. \textbf{(Should we talk more about TinySOL here?)}
+
+When training a model, the number of samples $n$ to be used in each combination is specified.
 To create the training data, $n$ instruments are randomly selected from the list of specified instruments, and then a TinySOL sample is randomly selected for each chosen instrument.
-These samples are then combined to play simultaneously.
-This melspectrogram is then taken of this combinations, creating the features to be fed to the model.
+This gives $n$ different samples, each of a different intrument.
+These samples are combined to form a soundfile that has each sample playing simultaneously.
+The melspectrogram is taken of this combination, creating the features to be fed to the model.
 We created the melspectrogram by \textbf{(Alejandro can you write this of how you generated the melspectrograms using the STFT?)}
@@
 
-Our idea is perform orchestration using deep neural networks trained for classification.
+Our idea is to perform orchestration using deep neural networks trained for classification.
 By generating input data that is combinations of different instruments and pitches, the network learns how to take a sound that is a complex combination of notes and timbres and deconstruct it into its original parts.
@@
 
-The first deep model we tested was Convolutional Neural Network(CNN).
+The first deep model we tested was a Convolutional Neural Network (CNN).
 CNN shows good performance on classification assignment for its strong ability to extract spacial features, so we implemented a basic CNN with three convolutional layers and two fully connected layers.
@@
 
+\textbf{(Add results of quantitatively comparing our results to OrchIdea)}
+
 \section{Evaluation and Conclusions}
@@
 
-(Alejandro, add how there were certain instruments that performed poorly. Add how ResNet was awful for Flute and always output Flute C4)
+\textbf{(Alejandro, add how there were certain instruments that performed poorly. Add how ResNet was awful for Flute and always output Flute C4)}
 

### 2020-05-07 paper/latex/paper.tex
minor edits

@@
 The melspectrogram is taken of this combination, creating the features to be fed to the model.
-We created the melspectrogram by \textbf{(Alejandro can you write this of how you generated the melspectrograms using the STFT?)}
+We created the melspectrogram by \textbf{(Alejandro can you write this of how you generated the melspectrograms using the STFT?)}.
+We chose to use the melspectrogram because of its ability to represent timbre, a key part of assisted orchestration.
 
@@
 Our idea is to perform orchestration using deep neural networks trained for classification.
-By generating input data that is combinations of different instruments and pitches, the network learns how to take a sound that is a complex combination of notes and timbres and deconstruct it into its original parts.
-When a target sound is then fed to the network, it attempts to classify which samples are present.
+By generating input data that is a mixture of different instruments and pitches, the network learns how to take a sound that is a complex combination of notes and timbres and deconstruct it into its original parts.
+When a target sound is then fed to the network, it attempts to classify which samples would best represent the target.
 By taking the samples that have the highest probability of being in the target, an orchestrated solution can be created.
@@
 The first deep model we tested was a Convolutional Neural Network (CNN).
-CNN shows good performance on classification assignment for its strong ability to extract spacial features, so we implemented a basic CNN with three convolutional layers and two fully connected layers.
+CNN shows good performance on classification assignment for its strong ability to extract spatial features, so we implemented a basic CNN with three convolutional layers and two fully connected layers.
 Each convolutional layer is followed by a BatchNorm layer, a ReLU activation layer and a 2 * 2  MaxPool layer with a stride of 2.
@@
 
-Given that audio is one kind of sequential data, it is easy to consider Long Short-Term Memory network(LSTM) as test model.
+Given that audio is one kind of sequential data, it is easy to consider Long Short-Term Memory network (LSTM) as test model.
 In our setting, we added one LSTM layer with 32 outputs units following three convolutional layers.
@@
 For targets that had harmonic content, it was noted if the partials present in the target were also represented in the orchestrated solution.
-For example, one of the samples of a bell had partials that represented a C sharp minor chord.
-For a target that contained specific notes, the solution from ResNet contained both the note in the target and its partials: a sample that was two octaves higher, the fifth an octave up, and a minor third two octaves up (mix\_ObA4\_BnC3).
 
@@
 We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets.
-Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the pitch(es) of the target.
+Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the pitch of the target.
+For a target that contained specific notes, the solution from ResNet contained both the note in the target and its partials: a sample that was two octaves higher, the fifth an octave up, and a minor third two octaves up (mix\_ObA4\_BnC3).
+Clearly, ResNet was able to identify the partials of the target and recreate them in the solution.
 

### 2020-05-08 paper/latex/paper.tex
Changes from call
added abstract and changed title

@@
 % ------
-\title{Deep learning approaches on musical assisted-orchestration: an evaluation study}
+\title{Neural models for computer-assisted musical orchestration: a preliminary study}
 
@@
 \begin{abstract}
-The abstract should be placed at the top left column and should contain about 150-200 words.
+In this paper we will explore preliminary neural models for the task of computer-assisted musical orchestration.
+After an introduction on the problem, we will show how we decided to model musical orchestration as a classification task.
+In this context, we will propose two deep learning models and we will show how they perform as classifiers for musical instruments recognition by comparing them with specific baselines.
+We will the show how they perform, both qualitative and quantitative, in the task of computer-assisted orchestration by comparing them with state-of-the-art systems.
+We will highlight, finally, benefits and problems of neural approaches for orchestration and we will propose possible future steps.
+
 \end{abstract}

### 2020-05-08 paper/latex/paper.tex
Add to data gen section that samples are 4 secs

@@
 To create the training data, $n$ instruments are randomly selected from the list of specified instruments, and then a TinySOL sample is randomly selected for each chosen instrument.
-This gives $n$ different samples, each of a different intrument.
+This gives $n$ different samples, each of a different instrument.
 These samples are combined to form a soundfile that has each sample playing simultaneously.
+The audio was trimmed to four seconds in length so our data would all be of the same size.
 The melspectrogram is taken of this combination, creating the features to be fed to the model.
@@
 \subsection{Baseline}
-In order to have a baseline to compare our results against, we attempted to solve the classification problem using various linear classifiers.
+In order to have a baseline to compare our results against, we attempted to solve the classification problem using various parametric classifiers.
 The classifiers we tested were Support Vector Machine (SVM), Random Forest, and K-Nearest Neighbors.
@@
 This was still a simplified version of the problem, as we were identifying only the pitch class of a few instruments.
-However, the linear classifiers were unable to achieve accurate results as the number of instruments increased.
-Therefore, we did not attempt the full setting of the problem, in which individual pitches are classified for all instruments, with linear classifiers.
+However, the parametric classifiers were unable to achieve accurate results as the number of instruments increased.
+Therefore, we did not attempt the full setting of the problem, in which individual pitches are classified for all instruments, with parametric classifiers.
 
@@
 
+Although the data we trained the model on was fixed to four seconds of audio, the target samples can be of arbitrary length.
+By changing the mel hop length, which is the distance between the frames of the melspectrogram, an audio signal of any length can be represented as a matrix of the same size as the training data.
+
+\subsection{Evaluation}
+-evaluation of orchestrations (not evaluation of model)
+
 We evaluated our orchestrations both qualitatively and quantitatively.
@@
 
-$$d(x^t, \tilde{x}^t) =\lambda_1 \sum_k \delta_k^t(x_k^t - \tilde{x}_k^t) + \sum_k \delta_k^t|x_k^t - \tilde{x	}_k^t|$$
-$$\delta_k^t = 
+$$d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k|$$
+
+$$\delta_{k1} = 
 \begin{cases}
-1, \text{if} (x_k - \tilde{x}_k) < 0 \\
+1, \text{if   } x_k > \tilde{x}_k \\
 0, \text{otherwise}
 \end{cases} $$
+
+$$ \delta_{k2} = 1 - \delta_{k1}$$
+
+This metric computes the difference between the FFT of the target and the FFT of the solution.
 

### 2020-05-08 paper/latex/paper.tex
fix distance eqn
Add baseline table

@@
 We found SVM to have the highest accuracy of the three classifiers across all experiments.
-All of the following baselin experiments used 50,000 generated samples with a train-test split of 60/40.
+All of the following baseline experiments used 50,000 generated samples with a train-test split of 60/40.
 Each sample is a combination of one or more instruments and is four seconds in length.
 The features used are the mel frequency cepstral coefficients (MFCCs) of the resulting combination, with a total of 100 coefficients per sample.
+We chose MFCCs because of their ability to describe the timbre of a sample using a relatively small number of coefficients.
 
@@
 Since this is a simpler problem, it lead to increased accuracies.
-Flute and Violin went from 30.2\% to 38.8\%, Oboe and French Horn from 27.3\% to 39.9\%, and Trombone and Bass Tuba from 2.2\% to 3.5\%.
-\textit{(I think we should maybe put a table of this data instead of listing it all out)} Random Forest performed significantly worse for this setting.
-Flute and Violin had an accuracy of 9.8\% and Oboe and French Horn was 17.5\%.
-For this reason, we stopped testing with Random Forest and continued with SVM only.
+The accuracies from this experiment are outlined in Table 1.
 
@@
 Therefore, we did not attempt the full setting of the problem, in which individual pitches are classified for all instruments, with parametric classifiers.
+
+\begin{table}[h!]
+  \begin{center}
+    \label{tab:table1}
+    \begin{tabular}{c|c|c|c} 
+    	  \hline
+      \textbf{Instr. 1} & \textbf{Instr. 2} & \textbf{SVM Accuracy} & \textbf{RF Accuracy}\\
+      \hline
+      Violin & Flute & 38.8\% & 9.8\% \\
+      \hline
+      Violin & Trumpet & 33.8\% & 9.1\% \\
+      \hline
+      Violin & Cello & 34.8\% & 6.3\% \\
+      \hline
+      Cello & Viola & 32.1\% & 5.8\% \\
+      \hline
+      Oboe & French Horn & \textbf{39.9\%} & \textbf{17.5\%} \\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{Comparison of results between SVM and Random Forest.
+Each data point is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment.
+For the samples drawn from one of the two instruments, the pitch class of that sample is identified.
+For a sample not from one of the two instruments, the classifier simply attempted to identify that a sample from one of the non-specified instruments was present.
+In this setting, there are 25 classes: 24 from the 12 pitch classes from 2 instruments, and 1 class that specified whether an instrument that is not one of the two specified is present.}
+\end{table}
 
@@
 For quantitative evaluation, we orchestrated the same targets with OrchIdea, then compared the distance between our solutions and the targets and OrchIdea's solutions and the targets.
-The distance metric we used was
+The distance metric we use is defined in \eqref{distance}.
 
-$$d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k|$$
+\begin{equation}\label{distance}
+d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k| \\
+\end{equation}
 
-$$\delta_{k1} = 
+$$
+\delta_{k1} = 
 \begin{cases}
@@
 0, \text{otherwise}
-\end{cases} $$
+\end{cases} 
+$$
 
-$$ \delta_{k2} = 1 - \delta_{k1}$$
+$$
+\delta_{k2} = 
+\begin{cases}
+1, \text{if   } x_k < \tilde{x}_k \\
+0, \text{otherwise}
+\end{cases} 
+$$
 

### 2020-05-08 paper/latex/paper.tex
Add orch eval table

@@
 
-\begin{table}[h!]
+\begin{table}
   \begin{center}
     \label{tab:table1}
-    \begin{tabular}{c|c|c|c} 
+    \begin{tabular}{|c|c|c|c|} 
     	  \hline
@@
 
+\begin{table}
+  \begin{center}
+    \label{tab:table2}
+    \begin{tabular}{|c|c|} 
+    	  \hline
+      \textbf{Model} & \textbf{Average Distance} \\
+      \hline
+      CNN with LSTM &   \\
+      \hline
+      ResNet &  \\
+      \hline
+      OrchIdea &  \\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{Comparison of the orchestration solutions of our two models, CNN with LSTM and ResNet, against OrchIdea's solutions.
+The distance between the target and solution are calculated for each target using \eqref{distance} and then averaged.}
+\end{table}
+
 \section{Evaluation and Conclusions}
@@
 Based on our accuracies from training the model and the results of orchestrating 15 targets, our approach seems promising for orchestration.
-When comparing our method with OrchIdea, it is important to note that many of processes OrchIdea uses to optimize its solution were not used in our model.
+When comparing our method with OrchIdea, it is important to note that many of the processes OrchIdea uses to optimize its solution were not used in our model.
 This includes important aspects such as sparsity, partial filtering, \textbf{(Carmine any more to add?)}.
@@
 
-\textbf{(Alejandro, add how there were certain instruments that performed poorly. Add how ResNet was awful for Flute and always output Flute C4)}
+In order to have comparable solutions, we did not allow OrchIdea to create sparse solutions, meaning that OrchIdea was forced to use all 10 instruments in each solution.
+This creates a more fair comparison since our model is optimize solutions by using fewer than 10 samples.
+
+\textbf{(Alejandro, add how there were certain instruments that performed really poorly or really well. Add how ResNet was awful for Flute and always output Flute C7)}
 

### 2020-05-09 paper/latex/paper.tex
Add info about TinySOL

@@
 \section{Data generation}
-We used the TinySOL database to create our input data. \textbf{(Should we talk more about TinySOL here?)}
+We used the TinySOL database to create our input data.
+TinySOL is a subset of the Studio On Line (SOL) database created by IRCAM.
+TinySOL contains 2,478 samples from 12 instruments.
+The instruments come from different orchestral families: strings, woodwinds, and brass are all represented in TinySOL.
+Each TinySOL sample is one instrument playing one note in the "ordinary" playing style.
 

### 2020-05-10 paper/latex/paper.tex
Update evaluation section

@@
 By passing these samples into a model, an orchestration of the target is created from the TinySOL samples that were used to train the network.
-For orchestrating these targets, we used both the CNN and ResNet models, each trained on combinations of 10 instruments from a possible 10 instruments.
-The CNN was trained for 49 epochs and the ResNet for 24.
+
+For orchestrating these targets, we used both the CNN and ResNet models.
+The CNN was trained on 200,000 samples for 49 epochs and the ResNet on 400,000 samples for 24 epochs.
+The instruments used to train the models are: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet.
+The resulting orchestrations are made from 10 samples drawn from these instruments.
 
@@
 We evaluated our orchestrations both qualitatively and quantitatively.
-Qualitative inspection was done through an acoustic inspection of the solution, paying close attention to timbre and pitch.
+Qualitative evaluation was done through an acoustic inspection of the solution, paying close attention to both timbre and pitch.
 For targets that had harmonic content, it was noted if the partials present in the target were also represented in the orchestrated solution.
 
-For quantitative evaluation, we orchestrated the same targets with OrchIdea, then compared the distance between our solutions and the targets and OrchIdea's solutions and the targets.
-The distance metric we use is defined in \eqref{distance}.
+For quantitative evaluation, we used the distance metric defined in \eqref{distance} to calculate the distance between the target samples and our orchestrated solutions.
+We then orchestrated the same targets with OrchIdea and calculated the distance for OrchIdea's solutions.
+A comparison of these results is in Table 2.
 

### 2020-05-11 paper/latex/paper.tex
add model figure

@@
 
-The first deep model we tested was a Convolutional Neural Network (CNN).
-CNN shows good performance on classification assignment for its strong ability to extract spatial features, so we implemented a basic CNN with three convolutional layers and two fully connected layers.
-Each convolutional layer is followed by a BatchNorm layer, a ReLU activation layer and a 2 * 2  MaxPool layer with a stride of 2.
-The kernel size is 3 * 3 with a stride of 1 and a padding of 1.
+The first deep model we examined was a Convolutional Neural Network (CNN).
+CNN shows good performance on classification assignment for its strong ability to extract local features, so we implemented a basic CNN with three convolutional layers and two fully connected layers.
+Each convolutional layer is followed by a BatchNorm layer, a ReLU activation layer and a $2 \times 2$  MaxPool layer with a stride of 2.
+The kernel size is $3 \times 3$ with a stride of 1 and a padding of 1.
 The number of filters are 8, 16, 32.
@@
 
-- figure
+\begin{figure*}[ht]
+\centering
+\includegraphics[scale=0.35]{figs/CNN_with_LSTM.png}
+\caption{Model Architecture}
+\label{cnn_lstm}
+\end{figure*}
 
-Given that audio is one kind of sequential data, it is easy to consider Long Short-Term Memory network (LSTM) as test model.
+Given that audio is one kind of sequential data, it is easy to consider Long Short-Term Memory network (LSTM) as a choice of modeling sequential data.
 In our setting, we added one LSTM layer with 32 outputs units following three convolutional layers.
@@
 The probability of an element to be zeroed is 0.5 and we took sigmoid as activation function.
-We used binary classification for each class and treated them independently.
+Since there is no competition relationship between different classes for each class can be classification result equally, it is reasonable for us to take sigmoid activations and used binary classification for each class to treat them independently.
+We minimize the binary cross entropy(BCE) loss between the predictions and targets.
+We implemented our neural networks with PyTorch and trained them with Adam optimizer.
+The model architecture is shown in Figure\ref{cnn_lstm}.
 
@@
 To make our setting more reasonable, we took the well-known deep residual network(ResNet) as backbone in our experiments.
-Specifically, we used 18-layer ResNet.
-Besides, to make the model more suitable to our problem, we reset the output channel numbers of each block with 32, 64, 32, 32 respectively.
+Specifically, we used 18-layer ResNet, which allows information to pass directly from the input to the final layer of each block.
+Besides, to make the model more suitable to our problem, we reduce the number of parameters and reset the output channel numbers of each block with 32, 64, 32, 32 respectively.
+We took the same loss function-BCE and optimizer-Adam as those in training the CNN with LSTM model.
 

### 2020-05-11 OrchDataset_2.py
Added to the paper

@@ -1,0 +1,296 @@
+import itertools
+from functools import reduce
+from generateDB_2 import generateDBLabels
+import librosa.display
+import random
+import math
+from torch.utils import data
+import torch
+import numpy as np
+import os
+import pickle
+import random
+import hashlib
+
+from parameters import GLOBAL_PARAMS
+
+def showSample(s):
+    print(s[1])
+    librosa.display.specshow(s[0], x_axis='time',y_axis='mel',sr=44100)
+
+class IndexIncrementor:
+    '''
+        This class provides a way to iterate over the set of features contained in 
+        a RawDatabase object to ensure that all pitch bins are uniformly chosen.
+        This helps generating a homogeneous database that doesn't miss some dynamics.
+        
+        A certain degree of randomness is added in the sense that the step and starting point
+        are randomized.
+    '''
+    def __init__(self, N, base, offset=None):
+        self.N = N
+        if type(base) == int:
+            self.base = [base]*N
+        elif type(base) == list and len(base) == N:
+            self.base = base
+        else:
+            raise Exception("Wrong argument for IndexIncrementor (must be int or list of size N)")
+        self.max_count = reduce((lambda x,y:x*y), self.base)
+        self.restart(offset)
+    
+    def findStep(self):
+        self.step = int(random.randint(int(0.05*self.max_count),int(0.95*self.max_count)))
+        while math.gcd(self.step, self.max_count) > 1:
+            self.step += 1
+    
+    def add(self, e):
+        r = e
+        i = 0
+        while i < self.N and r != 0:
+            r,self.index[i] = divmod(self.index[i]+r,self.base[i])
+            i += 1
+            
+    def incr(self):
+        self.add(self.step)
+        self.count += 1
+        
+    def restart(self, offset = None):
+        self.count = 0
+        self.index = [0]*self.N
+        self.findStep()
+        if offset is None:
+            self.add(random.randint(0,self.max_count-1))
+        elif offset >= 0:
+            self.add(offset)
+        else:
+            raise Exception("Wrong offset")
+    
+    def isOver(self):
+        return self.count >= self.max_count
+            
+class HashTable:
+    def __init__(self, max_depth, depth=0):
+        self.depth = depth
+        self.max_depth = max_depth
+        if max_depth == depth+1:
+            self.data = []
+        else:
+            self.data = {}
+            for i in '0123456789abcdef':
+                self.data[i] = HashTable(max_depth,depth+1)
+            
[…]

### 2020-05-11 apmoihfpouazefb.py

@@ -1,0 +1,141 @@
+from parameters import GLOBAL_PARAMS, SimParams
+from OrchDataset import OrchDataSet, RawDatabase
+import numpy as np
+import matplotlib.pyplot as plt
+from model import OrchMatchNet
+import os
+import librosa
+import random
+
+import torch
+import torch.nn as nn
+import torchvision.transforms as transforms
+import torch.nn.functional as F
+from torch.autograd import Variable
+from torch.utils import data
+
+PATH = './model/ResNet/run1'
+train_list = list(range(1))
+
+
+
+GLOBAL_PARAMS.load_parameters(PATH)
+
+tot_size = sum(len(GLOBAL_PARAMS.lab_class[k]) for k in GLOBAL_PARAMS.lab_class)
+    
+def class_encoder(list_samp):
+    label = [0 for i in range(tot_size)]
+    for s in list_samp:
+        label[GLOBAL_PARAMS.lab_class[s['instrument']][s['pitch_name']]] = 1
+    return np.array(label).astype(np.float32)
+
+def evaluate(preds, labels):
+    if preds.shape != labels.shape:
+        print("[Error]: size difference")
+    # compute the label-based accuracy
+    result = {}
+
+    result['acc'] = np.sum(preds*labels)/max(1.0, np.sum(labels))
+    pitch_acc = {}
+    for i in GLOBAL_PARAMS.lab_class:
+        l = [GLOBAL_PARAMS.lab_class[i][x] for x in GLOBAL_PARAMS.lab_class[i]]
+        f = np.zeros(preds.shape, dtype = np.float32)
+        f[:,min(l):max(l)+1] = 1.0
+        f = labels*f
+        pitch_acc[i] = np.sum(preds*f)/max(1.0, np.sum(f))
+    result['pitch_acc'] = pitch_acc
+
+    return result
+
+
+def testMix(mix):
+    samps = []
+    for i in mix:
+        instr,pitch,nuance=i
+        p = './TinySOL/{}/ordinario'.format(instr)
+        l = [x for x in os.listdir(p) if pitch in x and nuance in x]
+        if len(l) != 1:
+            raise Exception('Wrong sample {}'.format(i))
+        y,sr = librosa.load(p + '/' + l[0], sr=None)
+        nb_samples = int(GLOBAL_PARAMS.TIME_LENGTH*sr)
+        if len(y) < nb_samples:
[…]

### 2020-05-11 parameters.py

@@ -1,9 +1,9 @@
 import pickle
 
-resume_model = False
+resume_model = True
 model_path = './model'
-model_run_resume = 5
-model_epoch_resume = 3
+model_run_resume = 3
+model_epoch_resume = 9
 
 class SimParams:
     def __init__(self):
@@ -15,17 +15,17 @@ class SimParams:
         self.TIME_LENGTH = 4
         self.FEATURE_TYPE = 'mel'
         
-        self.N = 10
+        self.N = 2
         self.nb_samples = 400000
         self.rdm_granularity = 10
         self.nb_pitch_range = 8
-        self.instr_filter = ['Hn','Ob','Vn','Va', 'Vc', 'Fl', 'Tbn', 'Bn', 'TpC', 'ClBb'][:self.N]
+        self.instr_filter = ['Hn','Ob','Vn','Va', 'Vc', 'Fl', 'Tbn', 'Bn', 'TpC', 'ClBb']
         self.batch_size = 16
         self.model_type = 'resnet'
-        self.nb_epoch = 100
-        self.rain_proportion = 0.8
+        self.nb_epoch = 50
+        self.train_proportion = 0.8
         
-        self.coeff_freq_shift_data_augment = 0.005 # For data augmentation, proportional change to sampling rate
+        self.coeff_freq_shift_data_augment = 0.008 # For data augmentation, proportional change to sampling rate
         self.delay_offset_avg = 0.005 # Average of offset applied to delay filters (ie the position of the first impulse)
         self.delay_period_avg = 0.002 # Average of the period of the impulses for delaying
         self.delay_feedback_avg = 0.5 # Average of the feedback factor of delay impulses

### 2020-05-11 showOutputs.py

@@ -6,7 +6,7 @@ from matplotlib.colors import ListedColormap
 
 from train import getPosNMax, prediction
 
-path = './outputs.npy'
+path = './model/run2/outputs.npy'
 
 viridis = cm.get_cmap('viridis')
 prec = 0.2

### 2020-05-11 test.py

@@ -200,15 +200,22 @@ def combine(soundlist, target):
     return mixed_file, sr
 
 
+def next_power_of_2(x):
+    return 1 if x == 0 else 2**(x - 1).bit_length()
+
+
 def compute_distance(target, solution):
     target, _ = librosa.load(target['path'], sr=None)
 
     # if the target is longer than the solution, must trim the target
     if (target.size > solution.size):
         target = target[:solution.size]
+        N = next_power_of_2(target.size)
+    else:
+        N = next_power_of_2(solution.size)
 
-    target_fft = np.fft.fft(target)
-    solution_fft = np.fft.fft(solution)
+    target_fft = np.abs(np.fft.rfft(target, N))
+    solution_fft = np.abs(np.fft.rfft(solution, N))
 
     lambda_1 = 1
     lambda_2 = 1
@@ -216,11 +223,12 @@ def compute_distance(target, solution):
     sum_1 = 0
     sum_2 = 0
     for i in range(target_fft.size):
-        a = target_fft[i]
-        b = solution_fft[i]
-        if a - b < 0:
-            sum_1 += a - b
-            sum_2 += abs(a - b)
+        k_target = target_fft[i]
+        k_solution = solution_fft[i]
+        if k_target >= k_solution:
+            sum_1 += k_target - k_solution
+        else:
+            sum_2 += abs(k_target - k_solution)
     distance = lambda_1 * sum_1 + lambda_2 * sum_2
     return float(distance)
 

### 2020-05-11 train.py

@@ -17,7 +17,7 @@ from matplotlib.animation import FuncAnimation
 from model import OrchMatchNet
 from parameters import GLOBAL_PARAMS
 from parameters import resume_model, model_epoch_resume, model_path, model_run_resume
-from OrchDataset import OrchDataSet,RawDatabase
+from OrchDataset_2 import OrchDataSet,RawDatabase
 
 class Timer:
     def __init__(self, size_buffer):
@@ -68,8 +68,6 @@ class DataSaver:
         r['loss_min'] = loss_min
         np.save(self.path, r)
         self.count += 1
-        
-ds = DataSaver('./outputs.npy')
 
 timer = Timer(1000)
 
@@ -86,9 +84,16 @@ def main(rdb = None):
     else:
         raw_db = rdb
     train_dataset = OrchDataSet(raw_db,class_encoder, GLOBAL_PARAMS.FEATURE_TYPE)
-    train_dataset.generate(GLOBAL_PARAMS.N,int(GLOBAL_PARAMS.train_proportion*GLOBAL_PARAMS.nb_samples))
+    if resume_model:
+        train_dataset.load(model_path+'/run{}/trainset.pkl'.format(model_run_resume))
+    else:
+        train_dataset.generate(GLOBAL_PARAMS.N,int(GLOBAL_PARAMS.train_proportion*GLOBAL_PARAMS.nb_samples))
     test_dataset = OrchDataSet(raw_db,class_encoder, GLOBAL_PARAMS.FEATURE_TYPE)
-    test_dataset.generate(GLOBAL_PARAMS.N,GLOBAL_PARAMS.nb_samples-int(GLOBAL_PARAMS.train_proportion*GLOBAL_PARAMS.nb_samples))
+    
+    if resume_model:
+        test_dataset.load(model_path+'/run{}/testset.pkl'.format(model_run_resume))
+    else:
+        test_dataset.generate(GLOBAL_PARAMS.N,GLOBAL_PARAMS.nb_samples-int(GLOBAL_PARAMS.train_proportion*GLOBAL_PARAMS.nb_samples))
 
     # load data
     train_load = torch.utils.data.DataLoader(dataset=train_dataset,
@@ -156,10 +161,12 @@ def train(model, save_path, optimizer, train_load, test_load, start_epoch, out_n
     total_loss = 0
 
     weight_decay = 0.01
[…]

### 2020-05-11 paper/latex/paper.tex
Added to the paper

--- 2020-05-09 paper/latex/paper.tex
@@
 % ------
-\title{Neural models for computer-assisted musical orchestration: a preliminary study}
+\title{Deep learning approaches on musical assisted-orchestration: an evaluation study}
 
@@
 \begin{abstract}
-In this paper we will explore preliminary neural models for the task of computer-assisted musical orchestration.
-After an introduction on the problem, we will show how we decided to model musical orchestration as a classification task.
-In this context, we will propose two deep learning models and we will show how they perform as classifiers for musical instruments recognition by comparing them with specific baselines.
-We will the show how they perform, both qualitative and quantitative, in the task of computer-assisted orchestration by comparing them with state-of-the-art systems.
-We will highlight, finally, benefits and problems of neural approaches for orchestration and we will propose possible future steps.
-
+The abstract should be placed at the top left column and should contain about 150-200 words.
 \end{abstract}
@@
 \section{Data generation}
+
+\subsection{Sources}
 We used the TinySOL database to create our input data.
-TinySOL is a subset of the Studio On Line (SOL) database created by IRCAM.
-TinySOL contains 2,478 samples from 12 instruments.
-The instruments come from different orchestral families: strings, woodwinds, and brass are all represented in TinySOL.
-Each TinySOL sample is one instrument playing one note in the "ordinary" playing style.
+This database contains samples of notes played using the ordinario style and with three different dynamics for each pair of instrument and pitch : \textit{pp}, \textit{mf} and \textit{ff}.
+The exact list of samples is shown in Tab.\ref{tab_pitch}.
+\textbf{(Should we talk more about TinySOL here?)}
 
-When training a model, the number of samples $n$ to be used in each combination is specified.
-To create the training data, $n$ instruments are randomly selected from the list of specified instruments, and then a TinySOL sample is randomly selected for each chosen instrument.
-This gives $n$ different samples, each of a different instrument.
-These samples are combined to form a soundfile that has each sample playing simultaneously.
-The audio was trimmed to four seconds in length so our data would all be of the same size.
-The melspectrogram is taken of this combination, creating the features to be fed to the model.
-We created the melspectrogram by \textbf{(Alejandro can you write this of how you generated the melspectrograms using the STFT?)}.
-We chose to use the melspectrogram because of its ability to represent timbre, a key part of assisted orchestration.
+Since the goal was to train the model to detect multiple instruments playing simultaneously, the samples used for training were generated by mixing a random combination of instruments and pitches.
+More precisely, given the number $N$ of instruments playing at the same time, the data was generated as follows:
+
+\begin{itemize}
+\item An orchestra of $N$ instruments is chosen in advance (it is the same for all samples).
+\item For each instrument in the orchestra, a random note is picked among the corresponding available samples in TinySOL.
+\item For each chosen note, a random dynamic is picked.
+\item The samples are added and normalized by the number of instruments.
+\item The Mel spectrogram of the mixture is then computed and used as input of the model.
+\end{itemize}
+
+The choice of Mel spectrograms as input features of the model is very commmon in music information \textbf{\{Insert reference\}} and can be considered to be an accurate representation of the way humans perceive sound and music.\\
+
+Samples had a sample rate of $44100Hz$ and were all padded or chopped to be exactly 4 seconds long.
+The Mel spectrograms were then generated using an FFT hop length of 2048 samples (e.g.
+the window of each FFT was $46ms$ wide), and a number of Mel bins equal to 128.
+Therefore, all the features fed to the model were matrices of size $128\times 345$.
+We used Librosa package in Python to compute the features, and more details on the exact computations can be found in \textbf{\{Reference to Librosa\}} The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10ms$ could be perceived by the human ear.
+
+\begin{table}
+\begin{tabular}{|c|c|c|}
+\hline
+Instrument & Abbreviation & Range \\
+\hline
+Violin & Vn & G3-E7\\
+Cello & Vc & C2-C6\\
+Viola & Va & C3-C7\\
+Trumpet C & TpC & F\#3-C6\\
+Trombone & Tbn & A\#0;A\#1-C5\\
+Oboe & Ob & A\#3-A6\\
+Horn & Hn & G1-F5\\
+Flute & Fl & B3-D7\\
+Clarinette Bb & ClBb & D3-G6\\
+Contrabass & Cb & E1-C5\\
+Bass Tuba & BTb & F\#1-F4\\
+Bassoon & Bn & A\#1-D\#5\\
+\hline
+\end{tabular}
+\begin{caption}Table containing the pitch ranges in the TinySOL database, used for training. Hyphen means all the notes between the two explicited notes. \label{tab_pitch}
+\end{caption}
+\end{table}
+
+\subsection{Data augmentation}
+
+In order to increase the variability of the samples, we used two methods of data augmentation on the samples.\\
+
+The first one was done on the mixture of the samples, by slightly shifting the frequency of the signal.
+In practice, the sample rate used to compute the Mel spectrogram was modified to be within $5\%$ of the acual $44100Hz$.\\
+
+The second one was done on the Mel spectrogram, by setting to 0 random columns and rows of the feature matrices.
+For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$ reset columns and $3.45$ reset rows.
+This method of data augmentation aimed to be more resilient to the possible errors of recording, due either to a setup that induces interferences and cancels specific frequences, or to a default in the microphone that fails to record a small temporal portion.
 
@@
 Our idea is to perform orchestration using deep neural networks trained for classification.
-By generating input data that is a mixture of different instruments and pitches, the network learns how to take a sound that is a complex combination of notes and timbres and deconstruct it into its original parts.
-When a target sound is then fed to the network, it attempts to classify which samples would best represent the target.
-By taking the samples that have the highest probability of being in the target, an orchestrated solution can be created.
+By generating input data that is combinations of different instruments and pitches, the network learns how to take a sound that is a complex combination of notes and timbres and deconstruct it into its original parts.
+Then, when a network is given a target sound as input, it will attempt to apply the same classification rules, outputing a vector of the probabilities of each sample being in the target sound.
+By taking the samples that have the highest probability, an orchestrated solution can be created.
 
@@
 If it was between $0.33$ and $0.66$, then the mezzo forte version, and if less than $0.33$ the pianissimo dynamic.
+\textbf{\{I am not sure this hypothesis can be used accurately\}}
 
 \subsection{Baseline}
-In order to have a baseline to compare our results against, we attempted to solve the classification problem using various parametric classifiers.
+In order to have a baseline to compare our results against, we attempted to solve the classification problem using various linear classifiers.
 The classifiers we tested were Support Vector Machine (SVM), Random Forest, and K-Nearest Neighbors.
@@
 We found SVM to have the highest accuracy of the three classifiers across all experiments.
-All of the following baseline experiments used 50,000 generated samples with a train-test split of 60/40.
+All of the following baselin experiments used 50,000 generated samples with a train-test split of 60/40.
 Each sample is a combination of one or more instruments and is four seconds in length.
 The features used are the mel frequency cepstral coefficients (MFCCs) of the resulting combination, with a total of 100 coefficients per sample.
-We chose MFCCs because of their ability to describe the timbre of a sample using a relatively small number of coefficients.
 
@@
 Since this is a simpler problem, it lead to increased accuracies.
-The accuracies from this experiment are outlined in Table 1.
+Flute and Violin went from 30.2\% to 38.8\%, Oboe and French Horn from 27.3\% to 39.9\%, and Trombone and Bass Tuba from 2.2\% to 3.5\%.
+\textit{(I think we should maybe put a table of this data instead of listing it all out)} Random Forest performed significantly worse for this setting.
+Flute and Violin had an accuracy of 9.8\% and Oboe and French Horn was 17.5\%.
+For this reason, we stopped testing with Random Forest and continued with SVM only.
 
@@
 This was still a simplified version of the problem, as we were identifying only the pitch class of a few instruments.
-However, the parametric classifiers were unable to achieve accurate results as the number of instruments increased.
-Therefore, we did not attempt the full setting of the problem, in which individual pitches are classified for all instruments, with parametric classifiers.
-
-\begin{table}
-  \begin{center}
-    \label{tab:table1}
-    \begin{tabular}{|c|c|c|c|} 
-    	  \hline
-      \textbf{Instr. 1} & \textbf{Instr. 2} & \textbf{SVM Accuracy} & \textbf{RF Accuracy}\\
-      \hline
-      Violin & Flute & 38.8\% & 9.8\% \\
-      \hline
-      Violin & Trumpet & 33.8\% & 9.1\% \\
-      \hline
-      Violin & Cello & 34.8\% & 6.3\% \\
-      \hline
-      Cello & Viola & 32.1\% & 5.8\% \\
-      \hline
-      Oboe & French Horn & \textbf{39.9\%} & \textbf{17.5\%} \\
-      \hline
-    \end{tabular}
-  \end{center}
-  \caption{Comparison of results between SVM and Random Forest.
-Each data point is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment.
-For the samples drawn from one of the two instruments, the pitch class of that sample is identified.
-For a sample not from one of the two instruments, the classifier simply attempted to identify that a sample from one of the non-specified instruments was present.
-In this setting, there are 25 classes: 24 from the 12 pitch classes from 2 instruments, and 1 class that specified whether an instrument that is not one of the two specified is present.}
-\end{table}
+However, the linear classifiers were unable to achieve accurate results as the number of instruments increased.
+Therefore, we did not attempt the full setting of the problem, in which individual pitches are classified for all instruments, with linear classifiers.
 
@@
 The first deep model we tested was a Convolutional Neural Network (CNN).
-CNN shows good performance on classification assignment for its strong ability to extract spatial features, so we implemented a basic CNN with three convolutional layers and two fully connected layers.
+CNN shows good performance on classification assignment for its strong ability to extract spacial features, so we implemented a basic CNN with three convolutional layers and two fully connected layers.
 Each convolutional layer is followed by a BatchNorm layer, a ReLU activation layer and a 2 * 2  MaxPool layer with a stride of 2.
@@
 This setting is simple but we can make a good comparison between it and the baseline.
+\\
 
-- figure
+\begin{figure}
+\includegraphics[scale=0.45]{CNN_LSTM_diagram.png}
+\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
+\end{figure}
 
-Given that audio is one kind of sequential data, it is easy to consider Long Short-Term Memory network (LSTM) as test model.
-In our setting, we added one LSTM layer with 32 outputs units following three convolutional layers.
-The convolutional layers did not change the sequence relationship.
-Instead, it scaled down the size of feature maps.
-After LSTM layer, we reshaped the output into the same size before inputing.
-Then we implemented another convolutional layer with the same structure and kernel size.
-Finally, we flattened out the outputs and fed them into fully connected layers with Dropout layer.
-The probability of an element to be zeroed is 0.5 and we took sigmoid as activation function.
-We used binary classification for each class and treated them independently.
+Given that audio is one kind of sequential data, it is easy to consider Long Short-Term Memory network(LSTM) as test model.
+In our setting, we added one LSTM layer with 32 outputs units following three convolutional layers following the same pattern as the CNN described earlier.
+After LSTM layer, we reshaped the output into the same size as the output of the last convulational layer, which was fed to another convolutional layer with the same structure and kernel size, yielding a tensor of dimensions $32\times 8 \times 21$.
+Finally, we flattened out the outputs and fed them into fully a connected layer with Dropout layer, followed by another fully connected layer.
+Finally, the sigmoid function was applied to th final layer in order to get a probability between 0 and 1 for each individual label.\\
 
 \subsection{ResNet}
-- SVM was good but CNN was better and ResNet was best
-- review all experiments we did and choose key plots
+
+The second deep model that we used was the well known deep residual network ResNet.
 
@@
 Besides, to make the model more suitable to our problem, we reset the output channel numbers of each block with 32, 64, 32, 32 respectively.
+
+\subsection{Classification results}
+
+During training, the loss function used to optimize the inner parameters of the model was binary cross entropy, as it is the common choice for multiclass multilabel classification frameworks such as this one.
+However, the value of the loss function alone is difficult to interpret.
+This is why, we used a complementary function $f$ only for evaluation, to match the notion of accuracy of the model with the intuitive approach.
+This function compares a vector of expected outputs $\overline{X}$ with the estimated output from the model $\hat{X}$ by using the following function
+
+\begin{equation}
+f(\overline{X}, \hat{X}) = \frac{1}{N}<\overline{X}, M_N(\hat{X})>
+\label{eval}
+\end{equation}
+
+where
+
+\begin{equation}
+M_N(\hat{X})_i = \left\{\begin{array}{ll}
+1 \text{ if } i \in I_N(\hat{X})\\
+0 \text{ otherwise}
+\end{array}\right.
+\label{NMax}
+\end{equation}
+and $I_N(\hat{X})$ is the set of the indices containing the $N$ first maximums of the vector $\hat{X}$.\\
+
+More specifically, the function $M_N(\hat{X})$ takes as an input a vector of probabilities (values between 0 and 1) and outputs a vector where only the positions of the $N$ first maxima are set to 1.
+This new vector would be the orchestration of $N$ instruments given by the model.
+Thus, the function $f$ simply outputs the proportion of the estimated orchestration that matches the expected one.\\
+
+Different experiments were made by varying the number of instruments in the orchestra $N$ but always matching this number with the number of instruments in each mixture sample.
+Tab.
+\ref{exp_tab} shows the orchestra for each different value of $N$ used in the experiments.\\
+
+\begin{table}
+\begin{tabular}{|c|l|}
+\hline
+$N$ & Orchestra\\
+\hline
+4 & Hn, Ob, Vn, Va\\
+5 & Hn, Ob, Vn, Va, Vc\\
+6 & Hn, Ob, Vn, Va, Vc, Fl\\
+7 & Hn, Ob, Vn, Va, Vc, Fl, Tbn\\
+8 & Hn, Ob, Vn, Va, Vc, Fl, Tbn, Bn\\
+9 & Hn, Ob, Vn, Va, Vc, Fl, Tbn, Bn, TpC\\
+10 & Hn, Ob, Vn, Va, Vc, Fl, Tbn, Bn, TpC, ClBb\\
+\hline
+\end{tabular}
+
+\begin{caption}
+Orchestra for each $N$ used in the experiments.
+\label{exp_tab}
+\end{caption}
+\end{table}
+
+Then, for both CNN with LSTM and ResNet, we computed the maximum accuracy over the epochs.
+Tab.
+\ref{epoch} compiles the number of samples per epoch and the total number of epochs used for each model and each value of $N$.
+Tab.
+\ref{class_res} shows the maximum accuracy computed using the function $f$ explicited in (\ref{eval}).
+
+\begin{table}
+\begin{tabular}{|c|c|c|c|}
+\hline
+Model & $N$ & Nb epochs & Nb samples\\
+\hline
+CNN-LSTM & 4-7 & 50 & 100000 \\
+CNN-LSTM & 8-10 & 50 & 200000 \\
+ResNet & 4-8 & 20 & 200000 \\
+ResNet & 9-10 & 20 & 400000 \\
+\hline
+\end{tabular}
+
+\begin{caption}
+Number of epochs and number of samples per epoch used for training each model with a given value of the size of the orchestra $N$.
+\label{epoch}
+\end{caption}
+\end{table}
+
+\begin{table}
+\begin{tabular}{|c|l|}
+\hline
+$N$ & Orchestra\\
+\hline
+4 & Hn, Ob, Vn, Va\\
+5 & Hn, Ob, Vn, Va, Vc\\
+6 & Hn, Ob, Vn, Va, Vc, Fl\\
+7 & Hn, Ob, Vn, Va, Vc, Fl, Tbn\\
+8 & Hn, Ob, Vn, Va, Vc, Fl, Tbn, Bn\\
+9 & Hn, Ob, Vn, Va, Vc, Fl, Tbn, Bn, TpC\\
+10 & Hn, Ob, Vn, Va, Vc, Fl, Tbn, Bn, TpC, ClBb\\
+\hline
+\end{tabular}
+
+\begin{caption}
+Orchestra for each $N$ used in the experiments.
+\label{class_res}
+\end{caption}
+\end{table}
 
@@
 
-Although the data we trained the model on was fixed to four seconds of audio, the target samples can be of arbitrary length.
-By changing the mel hop length, which is the distance between the frames of the melspectrogram, an audio signal of any length can be represented as a matrix of the same size as the training data.
-
-\subsection{Evaluation}
--evaluation of orchestrations (not evaluation of model)
-
 We evaluated our orchestrations both qualitatively and quantitatively.
@@
 For targets that had harmonic content, it was noted if the partials present in the target were also represented in the orchestrated solution.
+For example, one of the samples of a bell had partials that represented a C sharp minor chord.
+For a target that contained specific notes, the solution from ResNet contained both the note in the target and its partials: a sample that was two octaves higher, the fifth an octave up, and a minor third two octaves up (mix\_ObA4\_BnC3).
 
 For quantitative evaluation, we orchestrated the same targets with OrchIdea, then compared the distance between our solutions and the targets and OrchIdea's solutions and the targets.
-The distance metric we use is defined in \eqref{distance}.
+The distance metric we used was
 
-\begin{equation}\label{distance}
-d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k| \\
-\end{equation}
-
-$$
-\delta_{k1} = 
+$$d(x^t, \tilde{x}^t) =\lambda_1 \sum_k \delta_k^t(x_k^t - \tilde{x}_k^t) + \sum_k \delta_k^t|x_k^t - \tilde{x	}_k^t|$$
+$$\delta_k^t = 
 \begin{cases}
-1, \text{if   } x_k > \tilde{x}_k \\
+1, \text{if} (x_k - \tilde{x}_k) < 0 \\
 0, \text{otherwise}
-\end{cases} 
-$$
-
-$$
-\delta_{k2} = 
-\begin{cases}
-1, \text{if   } x_k < \tilde{x}_k \\
-0, \text{otherwise}
-\end{cases} 
-$$
-
-This metric computes the difference between the FFT of the target and the FFT of the solution.
+\end{cases} $$
 
 \textbf{(Add results of quantitatively comparing our results to OrchIdea)}
-
-\begin{table}
-  \begin{center}
-    \label{tab:table2}
-    \begin{tabular}{|c|c|} 
-    	  \hline
-      \textbf{Model} & \textbf{Average Distance} \\
-      \hline
-      CNN with LSTM &   \\
-      \hline
-      ResNet &  \\
-      \hline
-      OrchIdea &  \\
-      \hline
-    \end{tabular}
-  \end{center}
-  \caption{Comparison of the orchestration solutions of our two models, CNN with LSTM and ResNet, against OrchIdea's solutions.
-The distance between the target and solution are calculated for each target using \eqref{distance} and then averaged.}
-\end{table}
 
@@
 Based on our accuracies from training the model and the results of orchestrating 15 targets, our approach seems promising for orchestration.
-When comparing our method with OrchIdea, it is important to note that many of the processes OrchIdea uses to optimize its solution were not used in our model.
+When comparing our method with OrchIdea, it is important to note that many of processes OrchIdea uses to optimize its solution were not used in our model.
 This includes important aspects such as sparsity, partial filtering, \textbf{(Carmine any more to add?)}.
@@
 
-In order to have comparable solutions, we did not allow OrchIdea to create sparse solutions, meaning that OrchIdea was forced to use all 10 instruments in each solution.
-This creates a more fair comparison since our model is optimize solutions by using fewer than 10 samples.
-
-\textbf{(Alejandro, add how there were certain instruments that performed really poorly or really well. Add how ResNet was awful for Flute and always output Flute C7)}
+\textbf{(Alejandro, add how there were certain instruments that performed poorly. Add how ResNet was awful for Flute and always output Flute C4)}
 
 We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets.
-Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the pitch of the target.
-For a target that contained specific notes, the solution from ResNet contained both the note in the target and its partials: a sample that was two octaves higher, the fifth an octave up, and a minor third two octaves up (mix\_ObA4\_BnC3).
-Clearly, ResNet was able to identify the partials of the target and recreate them in the solution.
+Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the pitch(es) of the target.
 

### 2020-05-11 paper/latex/paper.tex
Added plots

@@
 \begin{table}
-\begin{tabular}{|c|l|}
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
 \hline
-$N$ & Orchestra\\
+Instruments & \multicolumn{2}{c|}{$N=4$} & \multicolumn{2}{c|}{$N=5$}& \multicolumn{2}{c|}{$N=6$}& \multicolumn{2}{c|}{$N=7$}& \multicolumn{2}{c|}{$N=8$}& \multicolumn{2}{c|}{$N=9$}& \multicolumn{2}{c|}{$N=10$}\\
+\cline{2-3} & C & R & C & R & C & R & C & R & C & R & C & R & C & R\\
 \hline
-4 & Hn, Ob, Vn, Va\\
-5 & Hn, Ob, Vn, Va, Vc\\
-6 & Hn, Ob, Vn, Va, Vc, Fl\\
-7 & Hn, Ob, Vn, Va, Vc, Fl, Tbn\\
-8 & Hn, Ob, Vn, Va, Vc, Fl, Tbn, Bn\\
-9 & Hn, Ob, Vn, Va, Vc, Fl, Tbn, Bn, TpC\\
-10 & Hn, Ob, Vn, Va, Vc, Fl, Tbn, Bn, TpC, ClBb\\
+All & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+Hn & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+Ob & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+Vn & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+Va & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+Vc & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+Fl & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+Tbn & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+Bn & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+TpC & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+ClBb & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
 \hline
@@
 \begin{caption}
-Orchestra for each $N$ used in the experiments.
+Best test accuracy obtained for each model and value of $N$ over the total number of epochs.
+The accuracy is computed using $f$ define in (\ref{eval}).
+The first row correspond to the overall accuracy of the model, regrouping all the instruments.
+C is CNN with LSTM, R is ResNet.
 \label{class_res}
@@
 
+\section{Figures}
+
+\begin{figure}
+\includegraphics[scale=0.6]{figs/CNN4.png}
+\end{figure}
+\begin{figure}
+\includegraphics[scale=0.6]{figs/CNN5.png}
+\end{figure}
+\begin{figure}
+\includegraphics[scale=0.6]{figs/CNN6.png}
+\end{figure}
+\begin{figure}
+\includegraphics[scale=0.6]{figs/CNN7.png}
+\end{figure}
+\begin{figure}
+\includegraphics[scale=0.6]{figs/CNN8.png}
+\end{figure}
+\begin{figure}
+\includegraphics[scale=0.6]{figs/CNN9.png}
+\end{figure}
+\begin{figure}
+\includegraphics[scale=0.6]{figs/CNN10.png}
+\end{figure}
+
 % For bibtex users:

### 2020-05-11 paper/latex/paper.tex
added back intro and changed title

@@
 \begin{abstract}
-The abstract should be placed at the top left column and should contain about 150-200 words.
+In this paper we will explore preliminary neural models for the task of computer-assisted musical orchestration.
+After an introduction on the problem, we will show how we decided to model musical orchestration as a classification task.
+In this context, we will propose two deep learning models and we will show how they perform as classifiers for musical instruments recognition by comparing them with specific baselines.
+We will the show how they perform, both qualitative and quantitative, in the task of computer-assisted orchestration by comparing them with state-of-the-art systems.
+We will highlight, finally, benefits and problems of neural approaches for orchestration and we will propose possible future steps.
 \end{abstract}
@@
 
-\begin{figure}
+\begin{figure*}
 \includegraphics[scale=0.45]{CNN_LSTM_diagram.png}
 \caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
-\end{figure}
+\end{figure*}
 

### 2020-05-12 paper/latex/paper.tex
Redo commit 'changes from call'
Redo commit 'added abstract and changed title'
Redo commit 'Add to data gen section that samples are 4 secs'

@@
 
-\begin{figure*}
+\begin{figure}
 \includegraphics[scale=0.45]{CNN_LSTM_diagram.png}
 \caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
-\end{figure*}
+\end{figure}
 
@@
 
+Although the data we trained the model on was fixed to four seconds of audio, the target samples can be of arbitrary length.
+By changing the mel hop length, which is the distance between the frames of the melspectrogram, an audio signal of any length can be represented as a matrix of the same size as the training data.
+
+\subsection{Evaluation}
+
 We evaluated our orchestrations both qualitatively and quantitatively.
@@
 
-$$d(x^t, \tilde{x}^t) =\lambda_1 \sum_k \delta_k^t(x_k^t - \tilde{x}_k^t) + \sum_k \delta_k^t|x_k^t - \tilde{x	}_k^t|$$
-$$\delta_k^t = 
+For quantitative evaluation, we orchestrated the same targets with OrchIdea, then compared the distance between our solutions and the targets and OrchIdea's solutions and the targets.
+The distance metric we used was
+
+$$d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k|$$
+
+$$\delta_{k1} = 
 \begin{cases}
-1, \text{if} (x_k - \tilde{x}_k) < 0 \\
+1, \text{if   } x_k > \tilde{x}_k \\
 0, \text{otherwise}
 \end{cases} $$
+
+$$ \delta_{k2} = 1 - \delta_{k1}$$
+
+This metric computes the difference between the FFT of the target and the FFT of the solution.
 

### 2020-05-12 paper/latex/paper.tex
Redo commit 'fix distance eqn'
Redo commit 'Add baseline table'

@@
 
+\begin{table}
+  \begin{center}
+    \label{tab:table1}
+    \begin{tabular}{c|c|c|c} 
+    	  \hline
+      \textbf{Instr. 1} & \textbf{Instr. 2} & \textbf{SVM Accuracy} & \textbf{RF Accuracy}\\
+      \hline
+      Violin & Flute & 38.8\% & 9.8\% \\
+      \hline
+      Violin & Trumpet & 33.8\% & 9.1\% \\
+      \hline
+      Violin & Cello & 34.8\% & 6.3\% \\
+      \hline
+      Cello & Viola & 32.1\% & 5.8\% \\
+      \hline
+      Oboe & French Horn & \textbf{39.9\%} & \textbf{17.5\%} \\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{Comparison of results between SVM and Random Forest.
+Each data point is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment.
+For the samples drawn from one of the two instruments, the pitch class of that sample is identified.
+For a sample not from one of the two instruments, the classifier simply attempted to identify that a sample from one of the non-specified instruments was present.
+In this setting, there are 25 classes: 24 from the 12 pitch classes from 2 instruments, and 1 class that specified whether an instrument that is not one of the two specified is present.}
+\end{table}
+
 \subsection{CNN with LSTM}
@@
 For quantitative evaluation, we orchestrated the same targets with OrchIdea, then compared the distance between our solutions and the targets and OrchIdea's solutions and the targets.
-The distance metric we used was
+The distance metric we used is defined in Eqn.
+\eqref{distance}
 
-$$d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k|$$
+\begin{equation}\label{distance}
+d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k| \\
+\end{equation}
 
-$$\delta_{k1} = 
+$$
+\delta_{k1} = 
 \begin{cases}

### 2020-05-12 paper/latex/paper.tex
more edits from last commit

@@
 Since this is a simpler problem, it lead to increased accuracies.
-Flute and Violin went from 30.2\% to 38.8\%, Oboe and French Horn from 27.3\% to 39.9\%, and Trombone and Bass Tuba from 2.2\% to 3.5\%.
-\textit{(I think we should maybe put a table of this data instead of listing it all out)} Random Forest performed significantly worse for this setting.
-Flute and Violin had an accuracy of 9.8\% and Oboe and French Horn was 17.5\%.
-For this reason, we stopped testing with Random Forest and continued with SVM only.
+The accuracies from this experiment are outlined in Table 1.
+Since Random Forest performed worse than SVM is every experiment, we stopped testing with Random Forest and used SVM from that point on.
 

### 2020-05-12 paper/latex/paper.tex
Redo commit 'Add orch eval table'

@@
     \label{tab:table1}
-    \begin{tabular}{c|c|c|c} 
+    \begin{tabular}{|c|c|c|c|} 
     	  \hline
@@
 
+In order to have comparable solutions, we did not allow OrchIdea to create sparse solutions, meaning that OrchIdea was forced to use all 10 instruments in each solution.
+This creates a more fair comparison since our model is optimize solutions by using fewer than 10 samples.
+
 \begin{equation}\label{distance}
@@
 
-\textbf{(Add results of quantitatively comparing our results to OrchIdea)}
+\begin{table}
+  \begin{center}
+    \label{tab:table2}
+    \begin{tabular}{|c|c|} 
+    	  \hline
+      \textbf{Model} & \textbf{Average Distance} \\
+      \hline
+      CNN with LSTM &   \\
+      \hline
+      ResNet &  \\
+      \hline
+      OrchIdea &  \\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{Comparison of the orchestration solutions of our two models, CNN with LSTM and ResNet, against OrchIdea's solutions.
+The distance between the target and solution are calculated for each target using \eqref{distance} and then averaged.}
+\end{table}
 

### 2020-05-12 paper/latex/paper.tex
Redo commit 'Add info about TinySOL'

@@
 We used the TinySOL database to create our input data.
-This database contains samples of notes played using the ordinario style and with three different dynamics for each pair of instrument and pitch : \textit{pp}, \textit{mf} and \textit{ff}.
+TinySOL is a subset of the Studio On Line (SOL) database created by IRCAM \textbf{Reference needed}.
+TinySOL contains 1,529 samples from 12 instruments.
+The instruments come from different orchestral families: strings, woodwinds, and brass.
+Notes are played using the ordinario style and with three different dynamics for each pair of instrument and pitch : \textit{pp}, \textit{mf} and \textit{ff}.
 The exact list of samples is shown in Tab.\ref{tab_pitch}.
-\textbf{(Should we talk more about TinySOL here?)}
 

### 2020-05-12 paper/latex/paper.tex
Redo commit 'Update evaluation section'

@@
 
-For quantitative evaluation, we orchestrated the same targets with OrchIdea, then compared the distance between our solutions and the targets and OrchIdea's solutions and the targets.
-The distance metric we used was
-
-For quantitative evaluation, we orchestrated the same targets with OrchIdea, then compared the distance between our solutions and the targets and OrchIdea's solutions and the targets.
-The distance metric we used is defined in Eqn.
-\eqref{distance}
+For quantitative evaluation, we used the distance metric defined in \eqref{distance} to calculate the distance between the target samples and our orchestrated solutions.
+We then orchestrated the same targets with OrchIdea and calculated the distance for OrchIdea's solutions.
+A comparison of these results is in Tab.
+\ref{orch_eval}.
 
@@
   \begin{center}
-    \label{tab:table2}
+    \label{orch_eval}
     \begin{tabular}{|c|c|} 

### 2020-05-12 paper/latex/paper.tex
Make diagram full page
re commit model part

@@
 Finally, we flattened out the outputs and fed them into fully a connected layer with Dropout layer, followed by another fully connected layer.
-Finally, the sigmoid function was applied to th final layer in order to get a probability between 0 and 1 for each individual label.\\
+Finally, the sigmoid function was applied to the final layer in order to get a probability between 0 and 1 for each individual label.
+Since there is no competition relationship between different classes for each class can be classification result equally, it is reasonable for us to take sigmoid activations and used binary classification for each class to treat them independently.
+We minimize the binary cross entropy(BCE) loss between the predictions and targets.
+We implemented our neural networks with PyTorch and trained them with Adam optimizer.
+The model architecture is shown in Figure\ref{cnn_lstm}.\\
 
@@
 
-$$
+\begin{equation}
 \delta_{k1} = 
@@
 \end{cases} 
-$$
+\end{equation}
 
-$$
+\begin{equation}
 \delta_{k2} = 
@@
 \end{cases}
-$$
+\end{equation}
 
-This metric computes the difference between the FFT of the target and the FFT of the solution.
+This metric takes in the FFT of the target, $x$ and the FFT of the solution $\tilde{x}$ and computes the difference between them at each bin.
 

### 2020-05-12 paper/latex/paper.tex
Revise data section

@@
 
-\section{Data generation}
+\section{Data}
 
-\subsection{Sources}
+\subsection{Data Source and Generation}
 We used the TinySOL database to create our input data.
@@
 The instruments come from different orchestral families: strings, woodwinds, and brass.
-Notes are played using the ordinario style and with three different dynamics for each pair of instrument and pitch : \textit{pp}, \textit{mf} and \textit{ff}.
-The exact list of samples is shown in Tab.\ref{tab_pitch}.
+Each sample is one instrument playing a single note in the ordinario style, with one of three dynamics: \textit{pp}, \textit{mf}, or \textit{ff}.
+The instruments and ranges over which they were recorded is summarized in Tab.\ref{tab_pitch}.
 
-Since the goal was to train the model to detect multiple instruments playing simultaneously, the samples used for training were generated by mixing a random combination of instruments and pitches.
-More precisely, given the number $N$ of instruments playing at the same time, the data was generated as follows:
+We generated all the data we used to train our models.
+For a chosen $N$, each input to our model is a combination of $N$ TinySOL samples.
+An orchestra of $N$ instruments is chosen and the data is generated in the following way:
 
 \begin{itemize}
-\item An orchestra of $N$ instruments is chosen in advance (it is the same for all samples).
-\item For each instrument in the orchestra, a random note is picked among the corresponding available samples in TinySOL.
-\item For each chosen note, a random dynamic is picked.
-\item The samples are added and normalized by the number of instruments.
-\item The Mel spectrogram of the mixture is then computed and used as input of the model.
+\item For each instrument in the orchestra, a random TinySOL sample for that instrument is chosen. This leads to a random combination of both notes and dynamics.
+\item The chosen samples are combined to be played simultaneously and normalized by the number of instruments.
+\item The mel spectrogram of the mixture is computed to be used as the features input to the model.
 \end{itemize}
 
-The choice of Mel spectrograms as input features of the model is very commmon in music information \textbf{\{Insert reference\}} and can be considered to be an accurate representation of the way humans perceive sound and music.\\
+The choice of using the mel spectrogram as input features of the model is very commmon in music information \textbf{\{Insert reference\}} and can be considered to be an accurate representation of the way humans perceive sound and music.
 
-Samples had a sample rate of $44100Hz$ and were all padded or chopped to be exactly 4 seconds long.
+Samples had a sample rate of $44100$Hz and were all padded or chopped to be exactly 4 seconds long.
 The Mel spectrograms were then generated using an FFT hop length of 2048 samples (e.g.
-the window of each FFT was $46ms$ wide), and a number of Mel bins equal to 128.
+the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128.
 Therefore, all the features fed to the model were matrices of size $128\times 345$.
-We used Librosa package in Python to compute the features, and more details on the exact computations can be found in \textbf{\{Reference to Librosa\}} The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10ms$ could be perceived by the human ear.
+We used Librosa package in Python to compute the features, and more details on the exact computations can be found in \textbf{\{Reference to Librosa\}} The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms could be perceived by the human ear.
 
@@
 
-\subsection{Data augmentation}
+\subsection{Data Augmentation}
 
@@
 - changing number of instruments
+
+- explain why we generated our data in the way we did
+Since our goal is to train a model to detect multiple instruments and pitches playing simultaneously, the data used for training is generated by mixing a random combination of instruments and pitches.
 
@@
 \end{equation}
-and $I_N(\hat{X})$ is the set of the indices containing the $N$ first maximums of the vector $\hat{X}$.\\
+and $I_N(\hat{X})$ is the set of the indices containing the $N$ first maximums of the vector $\hat{X}$.
 
@@
 This new vector would be the orchestration of $N$ instruments given by the model.
-Thus, the function $f$ simply outputs the proportion of the estimated orchestration that matches the expected one.\\
+Thus, the function $f$ simply outputs the proportion of the estimated orchestration that matches the expected one.
 
@@
 Tab.
-\ref{exp_tab} shows the orchestra for each different value of $N$ used in the experiments.\\
+\ref{exp_tab} shows the orchestra for each different value of $N$ used in the experiments.
+
+\textbf{(Alejandro, add how there were certain instruments that performed poorly and some that were great. Add how ResNet was awful for Flute and always output Flute C7)}
 
@@
 
-\section{Evaluation and Conclusions}
+\section{Conclusion}
 - the approach seems promising for orchestration
@@
 Our model also lacks the implementation of symbolic constraints, which is an important part of assisted orchestration.
-
-\textbf{(Alejandro, add how there were certain instruments that performed poorly. Add how ResNet was awful for Flute and always output Flute C4)}
 

### 2020-05-12 paper/latex/paper.tex
Revise data augmentation

@@
 
-In order to increase the variability of the samples, we used two methods of data augmentation on the samples.\\
+In order to increase variability in our generated data, we used two methods of data augmentation.
 
-The first one was done on the mixture of the samples, by slightly shifting the frequency of the signal.
-In practice, the sample rate used to compute the Mel spectrogram was modified to be within $5\%$ of the acual $44100Hz$.\\
+The first one was done on the soundfile resulting from the mixture of the samples, before the mel spectrogram was taken.
+We performed a slight pitch shift by shifting the frequency of the signal.
+We did this by ...
+\textbf{Alejandro, can you add how you did the pitch shift (and maybe explain why changing the SR leads to a pitch shift)}In practice, the sample rate used to compute the Mel spectrogram was modified to be within $5\%$ of the actual $44100$Hz.
 
-The second one was done on the Mel spectrogram, by setting to 0 random columns and rows of the feature matrices.
-For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$ reset columns and $3.45$ reset rows.
-This method of data augmentation aimed to be more resilient to the possible errors of recording, due either to a setup that induces interferences and cancels specific frequences, or to a default in the microphone that fails to record a small temporal portion.
+The second augmentation was performed on the feature matrix itself, the mel spectrogram.
+We chose random columns and rows of the matrix to zero out.
+For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$  columns and $3.45$ rows being zero-ed out.
+This method of data augmentation aimed to be more resilient to the possible errors of recording, due either to a setup that induces interferences and cancels specific frequencies, or to a default in the microphone that fails to record a small temporal portion.
 
 \section{Our model}
-- deep learning trained for classification and then used for orchestraion
-- changing number of instruments
-
-- explain why we generated our data in the way we did
-Since our goal is to train a model to detect multiple instruments and pitches playing simultaneously, the data used for training is generated by mixing a random combination of instruments and pitches.
 
 Our idea is to perform orchestration using deep neural networks trained for classification.
-By generating input data that is combinations of different instruments and pitches, the network learns how to take a sound that is a complex combination of notes and timbres and deconstruct it into its original parts.
-Then, when a network is given a target sound as input, it will attempt to apply the same classification rules, outputing a vector of the probabilities of each sample being in the target sound.
+In our setting, each class is an instrument-note pair.
+Our model takes in the mel spectrogram of a soundfile that is a combination of different instruments and pitches, and attempts to identify which instrument-pitch pairs are present.
+The network learns how to take a sound that is a complex combination of notes and timbres and deconstruct it into its original parts.
+Then, when a network is given a target sound as input, it will attempt to apply the same classification rules, outputting a vector of the probabilities of each sample being in the target sound.
 By taking the samples that have the highest probability, an orchestrated solution can be created.
@@
 We did not train our model to classify the dynamics of a sample despite TinySOL having pianissimo, mezzo forte, and fortissimo recordings for each sample.
-If we had, then each class would have only one data point.
+If we had, then each class would be an instrument-pitch-dynamic tuple and would have only one data point.
 Instead, when the model is used for orchestration, the dynamic is determined by the probability of that sample as output by the model.
@@
 If it was between $0.33$ and $0.66$, then the mezzo forte version, and if less than $0.33$ the pianissimo dynamic.
+The idea behind this is that samples that are the most represented in the target should be the loudest in the orchestrated solution.
 \textbf{\{I am not sure this hypothesis can be used accurately\}}

### 2020-05-13 paper/latex/paper.tex
Add paper.bib

@@
 The classifiers we tested were Support Vector Machine (SVM), Random Forest, and K-Nearest Neighbors.
-We used the implementations provided in the scikit-learn library for each classifier.
+We used the implementations provided in the scikit-learn library for Python.
 For SVM, we used SVC with an RBF kernel.
@@
 We found SVM to have the highest accuracy of the three classifiers across all experiments.
-All of the following baselin experiments used 50,000 generated samples with a train-test split of 60/40.
+All of the following baseline experiments used 50,000 generated samples with a train-test split of 60/40.
 Each sample is a combination of one or more instruments and is four seconds in length.
 The features used are the mel frequency cepstral coefficients (MFCCs) of the resulting combination, with a total of 100 coefficients per sample.
+We chose to use MFCCs for this setting because using the mel spectrogram leads to a number of features that is too high for the parametric classifiers.
 
@@
 The first deep model we tested was a Convolutional Neural Network (CNN).
-CNN shows good performance on classification assignment for its strong ability to extract spacial features, so we implemented a basic CNN with three convolutional layers and two fully connected layers.
-Each convolutional layer is followed by a BatchNorm layer, a ReLU activation layer and a 2 * 2  MaxPool layer with a stride of 2.
-The kernel size is 3 * 3 with a stride of 1 and a padding of 1.
+CNNs show good performance on classification problems for their ability to extract spatial features.
+\textbf{Reference needed}.
+Our architecture contained three convolutional layers and two fully connected layers.
+Each convolutional layer is followed by a BatchNorm layer, a ReLU activation layer and a $2 \times 2$  MaxPool layer with a stride of 2.
+The kernel size is $3 \times 3$ with a stride of 1 and a padding of 1.
 The number of filters are 8, 16, 32.
-This setting is simple but we can make a good comparison between it and the baseline.
-\\
 
@@
 
-Given that audio is one kind of sequential data, it is easy to consider Long Short-Term Memory network(LSTM) as test model.
+Given the sequential nature of audio, a Long Short-Term Memory (LSTM) network provides a way to learn long term dependencies in the data \cite{LSTM}.
+
  In our setting, we added one LSTM layer with 32 outputs units following three convolutional layers following the same pattern as the CNN described earlier.
@@
 % For bibtex users:
-\bibliography{ISMIRtemplate}
+\bibliography{paper}
 

### 2020-05-13 paper/latex/paper.tex
Revise CNN section

@@
 By taking the samples that have the highest probability, an orchestrated solution can be created.
+Both of our deep models use the Adam optimizer, Binary Cross-Entropy loss, and were written using PyTorch.
 
@@
 
-The first deep model we tested was a Convolutional Neural Network (CNN).
+The first deep model we tested was a Convolutional Neural Network (CNN) with a Long Short-Term Memory (LSTM) unit.
 CNNs show good performance on classification problems for their ability to extract spatial features.
 \textbf{Reference needed}.
-Our architecture contained three convolutional layers and two fully connected layers.
+Long Short-Term Memory (LSTM) units provide a way to learn long term dependencies in the data \cite{LSTM}, which is relevent given the sequential nature of audio.
+
+Our architecture contains four convolutional layers and two fully connected layers.
 Each convolutional layer is followed by a BatchNorm layer, a ReLU activation layer and a $2 \times 2$  MaxPool layer with a stride of 2.
 The kernel size is $3 \times 3$ with a stride of 1 and a padding of 1.
-The number of filters are 8, 16, 32.
+The number of filters are 8, 16, 32, and 32.
+
+Following the first three convolutional layers, there is an LSTM layer with 32 outputs.
+After the LSTM layer, there is a final convolutional layer yielding a tensor of dimensions $32\times 8 \times 21$.
+We flatten the outputs and feed them into fully a connected layer with Dropout, then another another fully connected layer.
+Finally, the sigmoid function is applied to the final layer.
+
+Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class. The model architecture is shown in Figure \ref{cnn_lstm}.
 
@@
 
-Given the sequential nature of audio, a Long Short-Term Memory (LSTM) network provides a way to learn long term dependencies in the data \cite{LSTM}.
-
- In our setting, we added one LSTM layer with 32 outputs units following three convolutional layers following the same pattern as the CNN described earlier.
-After LSTM layer, we reshaped the output into the same size as the output of the last convulational layer, which was fed to another convolutional layer with the same structure and kernel size, yielding a tensor of dimensions $32\times 8 \times 21$.
-Finally, we flattened out the outputs and fed them into fully a connected layer with Dropout layer, followed by another fully connected layer.
-Finally, the sigmoid function was applied to the final layer in order to get a probability between 0 and 1 for each individual label.
-Since there is no competition relationship between different classes for each class can be classification result equally, it is reasonable for us to take sigmoid activations and used binary classification for each class to treat them independently.
-We minimize the binary cross entropy(BCE) loss between the predictions and targets.
-We implemented our neural networks with PyTorch and trained them with Adam optimizer.
-The model architecture is shown in Figure\ref{cnn_lstm}.\\
-
 \subsection{ResNet}
@@
 
-To make our setting more reasonable, we took the well-known deep residual network(ResNet) as backbone in our experiments.
+To make our setting more reasonable, we took the well-known deep residual network (ResNet) as backbone in our experiments.
 Specifically, we used 18-layer ResNet, which allows information to pass directly from the input to the final layer of each block.
 Besides, to make the model more suitable to our problem, we reduce the number of parameters and reset the output channel numbers of each block with 32, 64, 32, 32 respectively.
-We took the same loss function-BCE and optimizer-Adam as those in training the CNN with LSTM model.
 
@@
 However, the value of the loss function alone is difficult to interpret.
-This is why, we used a complementary function $f$ only for evaluation, to match the notion of accuracy of the model with the intuitive approach.
+For this reason we created a complementary function $f$ to be used only for evaluation.
 This function compares a vector of expected outputs $\overline{X}$ with the estimated output from the model $\hat{X}$ by using the following function
@@
 \end{equation}
-and $I_N(\hat{X})$ is the set of the indices containing the $N$ first maximums of the vector $\hat{X}$.
+and $I_N(\hat{X})$ is the set of indices containing the $N$ first maximums of the vector $\hat{X}$.
 
-More specifically, the function $M_N(\hat{X})$ takes as an input a vector of probabilities (values between 0 and 1) and outputs a vector where only the positions of the $N$ first maxima are set to 1.
+More specifically, the function $M_N(\hat{X})$ takes as an input a vector of probabilities and outputs a vector where only the positions of the $N$ first maxima are set to 1.
 This new vector would be the orchestration of $N$ instruments given by the model.
@@
 \section{Orchestration experiments}
-- set of 10 targets
-- orchestra fix to 10
-- qualitative evaluation: acoustic inspection of the solutions
-- quantitative evaluation: distance in feature space
-- comparison table between Orchidea and our model
 
@@
   \caption{Comparison of the orchestration solutions of our two models, CNN with LSTM and ResNet, against OrchIdea's solutions.
-The distance between the target and solution are calculated for each target using \eqref{distance} and then averaged.}
+The distance between the target and solution are calculated for each target using Eqn.
+\eqref{distance} and then averaged.}
 \end{table}
@@
 \begin{figure}
-\includegraphics[scale=0.6]{figs/CNN4.png}
-\end{figure}
-\begin{figure}
-\includegraphics[scale=0.6]{figs/CNN5.png}
-\end{figure}
-\begin{figure}
-\includegraphics[scale=0.6]{figs/CNN6.png}
-\end{figure}
-\begin{figure}
-\includegraphics[scale=0.6]{figs/CNN7.png}
-\end{figure}
-\begin{figure}
-\includegraphics[scale=0.6]{figs/CNN8.png}
-\end{figure}
-\begin{figure}
-\includegraphics[scale=0.6]{figs/CNN9.png}
-\end{figure}
-\begin{figure}
 \includegraphics[scale=0.6]{figs/CNN10.png}

### 2020-05-13 paper/latex/paper.tex
Add introduction

@@
 \section{Introduction}\label{sec:introduction}
+
+Computer-assisted orchestration is an important problem ... \textbf{Carmine can you add more?}.
+
+We attempt to use neural networks to solve the assisted orchestration problem.
+Our idea is to frame this as a classification problem, and then use these classifiers to perform orchestrations.
+Our classifiers will be trained on sounds that are combinations of multiple instruments and notes, leading to a model that can take a complex mixture of timbres and deconstruct it into individual instrument-note pairs.
+In our setting of the problem, each class is an instrument-note pair.
+When a new sound, one that is not simply a mixture of notes, is fed to the model, the predictions are used to create an orchestrated solution.
+
+We found that even on a simplified version of this classification problem, parametric classifiers were unable to meaningfully identify instrument-note pairs as the number of instruments used in the combinations increased.
+Therefore a need for neural networks arose.
+The two architectures we experimented with are a Convolutional Neural Network with a Long Short-Term Memory unit, and ResNet, a well known residual network.
+After training these models on combinations of 10 instruments, we orchestrated 15 target sounds by feeding them to the networks.
+We then compared our orchestrations to OrchIdea's solutions for the same targets.
 

### 2020-05-13 paper/latex/paper.tex
Added figures

@@
 \ref{epoch} compiles the number of samples per epoch and the total number of epochs used for each model and each value of $N$.
-Tab.
-\ref{class_res} shows the maximum accuracy computed using the function $f$ explicited in (\ref{eval}).
+Fig.
+\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of instruments $N$ to detect simultaneously.\\
+
+From this plot, we see that ResNet achieves better performances regardless of the number of instruments used in the combination, with
+
+For more details on the accuracy Fig. \ref{best_acc_cnn} and Fig. \ref{best_acc_resnet} show the maximum accuracy computed using the function $f$ explicited in (\ref{eval}).
 
@@
 
-\begin{table}
+\begin{figure}
+\includegraphics[scale=0.57]{figs/CNN_vs_ResNet.png}
+\caption{Best overall accuracy and for CNN and ResNet depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
+\end{figure}
+
+\iffalse
+\begin{table*}
+\center{
 \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
@@
 \label{class_res}
-\end{caption}
-\end{table}
+\end{caption}}
+\end{table*}
+\fi
 
@@
 
+\begin{figure*}
+\includegraphics[scale=0.6]{figs/Acc_vs_N_CNN.png}
+\caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM depending on the number of instruments in the combinations. \label{best_acc_cnn}}
+\end{figure*}
+
+\begin{figure*}
+\includegraphics[scale=0.6]{figs/Acc_vs_N_ResNet.png}
+\caption{Best overall accuracy and for each instrument obtained by the ResNet depending on the number of instruments in the combinations. \label{best_acc_resnet}}
+\end{figure*}
+
 % For bibtex users:

### 2020-05-13 plot.py

@@ -1,5 +1,60 @@
 import matplotlib.pyplot as plt
 import pickle
+import os
+import parameters
+
+def f(l):
+    Y = [0,0,0,0,0,0]
+    for i in l:
+        Y[X.index(i[0])] = i[1]['acc']
+    return Y
+
+def plot_best_acc(path):
+    
+    list_accs = []
+    for i in os.listdir(path):
+        accs = {}
+        MM = len([x for x in os.listdir('{}/{}'.format(path, i)) if 'result' in x])
+        epochs = list(range(4,MM*5,5))
+        max_epoch = -1
+        
+        parameters.GLOBAL_PARAMS.load_parameters('{}/{}'.format(path, i))
+    
+        for j in epochs:
+            f = open('{}/{}/result{}.pkl'.format(path, i, j), 'rb')
+            r = pickle.load(f)
+            if len(accs.keys()) == 0:
+                accs['acc'] = 0
+                for k in r['pitch_acc']:
+                    accs[k] = 0
+            
+            if r['acc'] > accs['acc']:
+                max_epoch = j
+                accs['acc'] = r['acc']
+                for k in r['pitch_acc']:
+                    accs[k] = r['pitch_acc'][k]
+            f.close()
+            
+        list_accs.append((parameters.GLOBAL_PARAMS.N, accs, max_epoch, epochs[-1]))
+    plt.figure(figsize=(12,7))
+    for i in list_accs[0][1].keys():
+        accs = []
+        for x in list_accs:
+            accs.append((x[0], x[1][i]))
+            
+        accs.sort(key = lambda x : x[0])
+        
+        if i == 'acc':
+            plt.plot([x[0] for x in accs], [x[1] for x in accs], '-o', label = "Overall accuracy")
+        else:
+            plt.plot([x[0] for x in accs], [x[1] for x in accs], '--x', label = "{} accuracy".format(i))
+        plt.grid()
+    plt.xlabel("Number of instruments in the combinations")
+    plt.ylabel("Accuracy")
+    plt.title("Maximum test accuracy achieved vs number of instruments N".format(i))
+    plt.legend()
+        
+    return list_accs
 
 def plot_run(path, run_nb):
     res = {'acc' : []}

### 2020-05-13 test.py
Compare distances of orchidea solutions

@@ -283,84 +283,84 @@ def make_fake_targets(num_classes):
 
 
 if __name__ == "__main__":
-    # print('--------------------------')
-    # print('MODEL TYPE:', model_type)
-    # print('NUMBER OF INSTRUMENTS:', len(instr_filter))
-    # print('N:', n)
-    # print('--------------------------')
+    print('--------------------------')
+    print('MODEL TYPE:', model_type)
+    print('NUMBER OF INSTRUMENTS:', len(instr_filter))
+    print('N:', n)
+    print('--------------------------')
 
 
-    # rdb = RawDatabase(tinysol_path, GLOBAL_PARAMS.rdm_granularity, instr_filter)
+    rdb = RawDatabase(tinysol_path, GLOBAL_PARAMS.rdm_granularity, instr_filter)
 
-    # lab_class = GLOBAL_PARAMS.lab_class
+    lab_class = GLOBAL_PARAMS.lab_class
 
-    # # calculate num classes
-    # num_classes = 0
-    # for k in rdb.db:
-    #     a = set()
-    #     for l in rdb.db[k]:
-    #         for e in l:
-    #             a.add(e['pitch_name'])
-    #     for p in rdb.pr:
-    #         if p in a:
-    #             num_classes += 1
+    # calculate num classes
+    num_classes = 0
+    for k in rdb.db:
+        a = set()
+        for l in rdb.db[k]:
+            for e in l:
+                a.add(e['pitch_name'])
+        for p in rdb.pr:
+            if p in a:
+                num_classes += 1
 
     
-    # def class_encoder(list_samp):
-    #     label = [0 for i in range(num_classes)]
-    #     for s in list_samp:
-    #         label[lab_class[s['instrument']][s['pitch_name']]] = 1
-    #     return np.array(label).astype(np.float32)
+    def class_encoder(list_samp):
+        label = [0 for i in range(num_classes)]
+        for s in list_samp:
+            label[lab_class[s['instrument']][s['pitch_name']]] = 1
+        return np.array(label).astype(np.float32)
 
-    # def evaluate(preds, labels):
-    #     if preds.shape != labels.shape:
-    #         print("[Error]: size difference")
-    #     # compute the label-based accuracy
-    #     result = {}
+    def evaluate(preds, labels):
+        if preds.shape != labels.shape:
+            print("[Error]: size difference")
+        # compute the label-based accuracy
+        result = {}
 
[…]

### 2020-05-13 paper/latex/paper.tex

@@
 
-Different experiments were made by varying the number of instruments in the orchestra $N$ but always matching this number with the number of instruments in each mixture sample.
-Tab.
-\ref{exp_tab} shows the orchestra for each different value of $N$ used in the experiments.
+Different experiments were made by varying the number $N$ of instruments in each mixture sample.
+We used an orchestra of 10 instruments containing Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet C and Clarinet Bb.
+
+Then, for both CNN with LSTM and ResNet, we computed the maximum accuracy over the epochs.
+For the CNN model, all the trainings with different values of $N$ were made using a dataset containing 200000 samples over 50 epochs, whereas for the ResNet, the dataset contained 400000 samples and was trained over 20 epochs.
+Fig.
+\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of instruments $N$.\\
 
@@
 
-\begin{table}
-\begin{tabular}{|c|l|}
-\hline
-$N$ & Orchestra\\
-\hline
-4 & Hn, Ob, Vn, Va\\
-5 & Hn, Ob, Vn, Va, Vc\\
-6 & Hn, Ob, Vn, Va, Vc, Fl\\
-7 & Hn, Ob, Vn, Va, Vc, Fl, Tbn\\
-8 & Hn, Ob, Vn, Va, Vc, Fl, Tbn, Bn\\
-9 & Hn, Ob, Vn, Va, Vc, Fl, Tbn, Bn, TpC\\
-10 & Hn, Ob, Vn, Va, Vc, Fl, Tbn, Bn, TpC, ClBb\\
-\hline
-\end{tabular}
+From this plot, we see that ResNet achieves better performances than the CNN regardless of the number of instruments used in the combination.
+This result isn't surprising \textbf{\color{red} Interpretation of why ResNet is better than CNN}
 
-\begin{caption}
-Orchestra for each $N$ used in the experiments.
-\label{exp_tab}
-\end{caption}
-\end{table}
-
-Then, for both CNN with LSTM and ResNet, we computed the maximum accuracy over the epochs.
-Tab.
-\ref{epoch} compiles the number of samples per epoch and the total number of epochs used for each model and each value of $N$.
 Fig.
-\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of instruments $N$ to detect simultaneously.\\
-
-From this plot, we see that ResNet achieves better performances regardless of the number of instruments used in the combination, with
-
-For more details on the accuracy Fig. \ref{best_acc_cnn} and Fig. \ref{best_acc_resnet} show the maximum accuracy computed using the function $f$ explicited in (\ref{eval}).
-
-\begin{table}
-\begin{tabular}{|c|c|c|c|}
-\hline
-Model & $N$ & Nb epochs & Nb samples\\
-\hline
-CNN-LSTM & 4-7 & 50 & 100000 \\
-CNN-LSTM & 8-10 & 50 & 200000 \\
-ResNet & 4-8 & 20 & 200000 \\
-ResNet & 9-10 & 20 & 400000 \\
-\hline
-\end{tabular}
-
-\begin{caption}
-Number of epochs and number of samples per epoch used for training each model with a given value of the size of the orchestra $N$.
-\label{epoch}
-\end{caption}
-\end{table}
+\ref{best_acc_cnn} and Fig.
+\ref{best_acc_resnet} show the maximum accuracy computed using the function $f$ explicited in (\ref{eval}).
+Note that for ResNet, the variance of accuracy is much smaller until $N$ reaches 5 before becoming similar to the CNN.
+Another key observation is that both models struggle to classify the same instruments.
+Flute, Trombone and Trumpet yield the worst results for both models, being notably worst \textbf{\color{red} Use same colors for instruments}
 

### 2020-05-14 paper/latex/paper.tex
Major edits

@@
 % Optional: To use hyperref, uncomment the following.
-% \usepackage[bookmarks=false,hidelinks]{hyperref}
+\usepackage[bookmarks=false,hidelinks]{hyperref}
 % Mind the bookmarks=false option; bookmarks are incompatible with ismir.sty.
@@
 % --------------
-\threeauthors
-  {First Author} {Affiliation1 \\ {\tt author1@ismir.edu}}
-  {Second Author} {\bf Retain these fake authors in\\\bf submission to preserve the formatting}
-  {Third Author} {Affiliation3 \\ {\tt author3@ismir.edu}}
+%\threeauthors
+%  {First Author} {Affiliation1 \\ {\tt author1@ismir.edu}}
+%  {Second Author} {\bf Retain these fake authors in\\\bf submission to preserve the formatting}
+%  {Third Author} {Affiliation3 \\ {\tt author3@ismir.edu}}
 
@@
 % ------------
-%\multauthor
-%{First author$^1$ \hspace{1cm} Second author$^1$ \hspace{1cm} Third author$^2$} { \bfseries{Fourth author$^3$ \hspace{1cm} Fifth author$^2$ \hspace{1cm} Sixth author$^1$}\\
-%  $^1$ Department of Computer Science, University , Country\\
-%$^2$ International Laboratories, City, Country\\
-%$^3$  Company, Address\\
-%{\tt\small CorrespondenceAuthor@ismir.edu, PossibleOtherAuthor@ismir.edu}
-%}
-%\def\authorname{First author, Second author, Third author, Fourth author, Fifth author, Sixth author}
+\multauthor
+{First author$^1$ \hspace{1cm} Second author$^1$ \hspace{1cm} Third author$^2$} { \bfseries{Fourth author$^3$ \hspace{1cm} Fifth author$^2$}\\
+  $^1$ Department of Computer Science, University , Country\\
+$^2$ International Laboratories, City, Country\\
+$^3$  Company, Address\\
+{\tt\small CorrespondenceAuthor@ismir.edu, PossibleOtherAuthor@ismir.edu}
+}
+\def\authorname{First author, Second author, Third author, Fourth author, Fifth author}
 
@@
 We attempt to use neural networks to solve the assisted orchestration problem.
-Our idea is to frame this as a classification problem, and then use these classifiers to perform orchestrations.
+Our model must be able to take in a sound and identify what instruments playing which notes best represents the sound.
+Our idea is to frame this as a classification problem, where each class is an instrument-note pair.
 Our classifiers will be trained on sounds that are combinations of multiple instruments and notes, leading to a model that can take a complex mixture of timbres and deconstruct it into individual instrument-note pairs.
-In our setting of the problem, each class is an instrument-note pair.
 When a new sound, one that is not simply a mixture of notes, is fed to the model, the predictions are used to create an orchestrated solution.
+The features we use are the mel spectrogram of the input sound, because of its ability to represent the way we hear sounds and its common use in MIR\textbf{Reference}.
 
 We found that even on a simplified version of this classification problem, parametric classifiers were unable to meaningfully identify instrument-note pairs as the number of instruments used in the combinations increased.
-Therefore a need for neural networks arose.
-The two architectures we experimented with are a Convolutional Neural Network with a Long Short-Term Memory unit, and ResNet, a well known residual network.
-After training these models on combinations of 10 instruments, we orchestrated 15 target sounds by feeding them to the networks.
-We then compared our orchestrations to OrchIdea's solutions for the same targets.
+Therefore the need for neural networks arose.
+The two architectures we used are a Convolutional Neural Network (CNN) with a Long Short-Term Memory (LSTM) unit, and ResNet, a well known residual architecture.
+We chose to use a CNN because of its success in audio classification \cite{Hershey2016}.
+Our inclusion of an LSTM is because of its ability to learn long term dependencies in data \cite{Hochreiter1997}, which is important given the temporal nature of audio.
+After training these models on varying combinations of 10 instruments, we orchestrated 15 target sounds by feeding them to the networks and synthesizing a solution from the most likely samples output from the system.
+We then compared our orchestrations to OrchIdea's solutions for the same targets through both qualitative and quantitative means.
+For quantitative comparison, we computed a distance metric on the FFTs of the target and solution, comparing the distance between OrchIdea's solution and the target and our model's solution and the target.
 
+Our codebase can be found at: \url{https://github.com/dzluke/DeepOrchestration}.
 \section{Data}
@@
 Flute & Fl & B3-D7\\
-Clarinette Bb & ClBb & D3-G6\\
+Clarinet Bb & ClBb & D3-G6\\
 Contrabass & Cb & E1-C5\\
@@
 \end{tabular}
-\begin{caption}Table containing the pitch ranges in the TinySOL database, used for training. Hyphen means all the notes between the two explicited notes. \label{tab_pitch}
-\end{caption}
+\caption{Table showing the pitch ranges present in the TinySOL database.}
 \end{table}
@@
 
-We did not train our model to classify the dynamics of a sample despite TinySOL having pianissimo, mezzo forte, and fortissimo recordings for each sample.
-If we had, then each class would be an instrument-pitch-dynamic tuple and would have only one data point.
-Instead, when the model is used for orchestration, the dynamic is determined by the probability of that sample as output by the model.
-If the model output a probability higher than $0.66$ for a sample, the fortissimo version of the sample was used.
-If it was between $0.33$ and $0.66$, then the mezzo forte version, and if less than $0.33$ the pianissimo dynamic.
-The idea behind this is that samples that are the most represented in the target should be the loudest in the orchestrated solution.
-\textbf{\{I am not sure this hypothesis can be used accurately\}}
-
 \subsection{Baseline}
@@
 The classifiers we tested were Support Vector Machine (SVM), Random Forest, and K-Nearest Neighbors.
-We used the implementations provided in the scikit-learn library for Python.
-For SVM, we used SVC with an RBF kernel.
+We used the implementations provided in the scikit-learn library for Python \cite{scikit-learn}.
+For SVM, we used SVC with a non-linear RBF kernel.
 For Random Forest, we set the maximum depth of each tree to be 15.
-Each classifier was wrapped in a MultiOutputClassifier to achieve the multi-label nature of this problem.
-We found SVM to have the highest accuracy of the three classifiers across all experiments.
 All of the following baseline experiments used 50,000 generated samples with a train-test split of 60/40.
@@
 We chose to use MFCCs for this setting because using the mel spectrogram leads to a number of features that is too high for the parametric classifiers.
+We found SVM to have the highest accuracy of the three classifiers across all experiments.
 
-We started by simplifying the problem to classifying only the instruments and not the pitch.
+We found that the parametric classifiers were not able to solve the full setting of our problem which had 424 classes, so we started by simplifying the problem to classifying only the instruments and not the pitch.
 This had the benefit of both reducing the number of classes and increasing the number of samples per class.
@@
 With two instruments, accuracy was 55.4\%, with three it was 19.6\% and with 10 instruments the accuracy was 0.5\%.
+K-Nearest Neighbors performed significantly worse than SVM, so we did not attempt any further testing with KNN.
 
-To better approximate the problem of identifying instrument and pitch, we then attempted to classify the instrument and pitch class.
+To better approximate the problem of identifying instrument and pitch, we performed experiments in which two instruments were selected and for input data that contained samples from one of those two instruments, both the instrument and pitch class of the sample would be classified.
 That is, which octave the pitch was in did not matter, only the pitch class.
-The input was a combination of two instruments drawn from a possible twelve instruments, and the classifier attempted to identify which instruments were present and for two of those instruments, say Flute and Violin, which pitch classes were present.
-If another instrument was present that was not Flute or Violin, the classifier would attempt to identify that instrument, but not its pitch class.
-The best results from this setting of the problem was SVM with 30.2\% accuracy, which was a result of classifying the pitch class of Flute and Violin.
-Depending on which two instruments had their pitch class identified, the accuracy varied greatly.
-For a Violin and Cello accuracy was 20.9\%, and for Trombone and Bass Tuba accuracy was 2.2\%.
-
-For a slight modification on this setting, we no longer attempted to identify which instrument was present if it was not one of the two instruments whose pitch class was being identified.
-Instead, the classifier would simply identify that an instrument that was not one of the two was present.
-Since this is a simpler problem, it lead to increased accuracies.
+The input was a combination of two instruments drawn from a possible twelve instruments, and the classifier attempted to identify which instruments were present and for the specified instruments, say Violin and Viola, which pitch classes were present.
+If another instrument was present in the input combination that was not Violin or Viola, the classifier would simply identify that an instrument that was not one of the two was present.
 The accuracies from this experiment are outlined in Table 1.
@@
 We then performed this same experiment with three instruments having their pitch class identified.
-Each input was a combination of three instruments drawn from a possible twelve instruments.
-For three instruments specified in that experiment, pitch class was identified.
 Flute, Oboe, and Violin reached an accuracy of 11.1\%, and Bass Tuba, Trumpet, and Trombone was 0.5\%.
@@
     	  \hline
-      \textbf{Instr. 1} & \textbf{Instr. 2} & \textbf{SVM Accuracy} & \textbf{RF Accuracy}\\
+      \textbf{Instr. 1} & \textbf{Instr. 2} & \textbf{SVM Acc.} & \textbf{RF Acc.}\\
       \hline
@@
   \end{center}
-  \caption{Comparison of results between SVM and Random Forest.
+  \caption{Comparison of accuracies between SVM and Random Forest.
 Each data point is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment.
@@
 The first deep model we tested was a Convolutional Neural Network (CNN) with a Long Short-Term Memory (LSTM) unit.
-CNNs show good performance on classification problems for their ability to extract spatial features.
-\textbf{Reference needed}.
-Long Short-Term Memory (LSTM) units provide a way to learn long term dependencies in the data \cite{LSTM}, which is relevent given the sequential nature of audio.
+CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey2016}.
+Long Short-Term Memory (LSTM) units provide a way to learn long term dependencies in the data \cite{Hochreiter1997}, which is relevant given the sequential nature of audio.
 
@@
 \subsection{ResNet}
-
+\textbf{Carmine, can you add more to this section about why ResNet is good, and maybe some details about the implementation?}
 The second deep model that we used was the well known deep residual network ResNet.
@@
 
-Our final model was trained on data where each input was a combination of ten instruments, but we performed experiments with varying numbers of instruments used in combination.
-(insert table of data showing the results of CNN with 2,3,4 etc instruments) An arbitrary number of samples can be used in the solution, since the $n$ highest probabilities can be taken from the output, leading to $n$ samples in the solution.
+For orchestrating, we trained our models on combinations of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet.
+A target sound is input to the model, and the 10 classes with the highest probability are extracted.
+These 10 classes represent the instrument-note pairs that are most represented in the target, and can be from any combination of the 10 instruments.
+The dynamic of each sample is inferred from its probability.
+
+We did not train our model to classify the dynamics of a sample despite TinySOL having pianissimo, mezzo forte, and fortissimo recordings for each sample.
+If we had, then each class would have only one data point: the TinySOL sample ir represents.
+By not initially ignoring dynamics, this led to each class having three data points: the instrument-note at \textit{pp}, \textit{mf}, and \textit{ff} When the model is used for orchestration, the dynamic is determined by the probability of that sample as output by the model.
+If the model outputs a probability higher than $0.66$ for a sample, the fortissimo version of the sample was used.
+If it was between $0.33$ and $0.66$, then the mezzo forte version, and if less than $0.33$ the pianissimo dynamic.
+The idea behind this is that samples that are the most represented in the target should be the loudest in the orchestrated solution.
 
@@
 Among the targets are recordings of bells, a car horn, a gong, and a recording of a boat docking.
-By passing these samples into a model, an orchestration of the target is created from the TinySOL samples that were used to train the network.
-For orchestrating these targets, we used both the CNN and ResNet models, each trained on combinations of 10 instruments from a possible 10 instruments.
-The CNN was trained for 49 epochs and the ResNet for 24.
 
@@
 We evaluated our orchestrations both qualitatively and quantitatively.
-Qualitative inspection was done through an acoustic inspection of the solution, paying close attention to timbre and pitch.
+Qualitative evaluation was done through an acoustic inspection of the solution, paying close attention to timbre and pitch.
 For targets that had harmonic content, it was noted if the partials present in the target were also represented in the orchestrated solution.
-For example, one of the samples of a bell had partials that represented a C sharp minor chord.
-For a target that contained specific notes, the solution from ResNet contained both the note in the target and its partials: a sample that was two octaves higher, the fifth an octave up, and a minor third two octaves up (mix\_ObA4\_BnC3).
+For example, one of the samples of a bell had partials that represented a C\# minor chord, so we checked if the orchestration contained the notes of the chord.
+If a target had specific notes, we identified if the note or its partials were present.
+For example there is target of an Oboe playing an A4 and a Bassoon playing a C\#3, the solution from ResNet contained the partials of the Bassoon's note: C\#3, G\#4, C\#5, and E6 were all included on various instruments in the solution.
 
-For quantitative evaluation, we used the distance metric defined in \eqref{distance} to calculate the distance between the target samples and our orchestrated solutions.
+For quantitative evaluation, we used the distance metric defined in \eqref{distance} to calculate differences in timbre between targets and solutions.
+The equation takes in the full FFT of the target and full FFT of the solution.
+Then for each bin of the FFT, it calculates the absolute difference between the values.
+The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.
+
+In the summations, all the terms with a $\lambda_1$ will be distances calculated when there was more energy in the target than the solution, since $\delta_{k1} = 1$ only when $x_k \ge \tilde{x}_k$.
+By the same reasoning, all the terms with a $\lambda_2$ are distances when the solution provided more energy to a frequency than the target; it overshot the frequency.
+Therefore, the relation between $\lambda_1$ and $\lambda_2$ determines whether a solution is penalized more for undershooting or overshooting the target.
+
+We calculated the distance between the target samples and our orchestrated solutions.
 We then orchestrated the same targets with OrchIdea and calculated the distance for OrchIdea's solutions.
@@
 In order to have comparable solutions, we did not allow OrchIdea to create sparse solutions, meaning that OrchIdea was forced to use all 10 instruments in each solution.
-This creates a more fair comparison since our model is optimize solutions by using fewer than 10 samples.
+This creates a more fair comparison since our model is unable to create sparse solutions.
 
@@
 \begin{cases}
-1, \text{if   } x_k > \tilde{x}_k \\
+1, \text{if   } x_k \ge \tilde{x}_k \\
 0, \text{otherwise}
@@
 
-This metric takes in the FFT of the target, $x$ and the FFT of the solution $\tilde{x}$ and computes the difference between them at each bin.
-
-\begin{table}
+\begin{table*}
   \begin{center}
     \label{orch_eval}
-    \begin{tabular}{|c|c|} 
+    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|} 
     	  \hline
-      \textbf{Model} & \textbf{Average Distance} \\
+      \textbf{Model} & \textbf{Ob + Bn} & \textbf{Bn} & \textbf{Bass Clarinet} & \textbf{Bell1} & \textbf{Bell2} & \textbf{Yan} & \textbf{Car Horn} & \textbf{Boat} & \textbf{Wind Harp} \\
       \hline
-      CNN with LSTM &   \\
+      CNN with LSTM & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
       \hline
-      ResNet &  \\
+      ResNet & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
       \hline
-      OrchIdea &  \\
+      OrchIdea & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0  \\
       \hline
@@
   \end{center}
-  \caption{Comparison of the orchestration solutions of our two models, CNN with LSTM and ResNet, against OrchIdea's solutions.
-The distance between the target and solution are calculated for each target using Eqn.
-\eqref{distance} and then averaged.}
-\end{table}
+  \caption{Quantitative comparison of orchestrations.}
+\end{table*}
+\begin{table*}
+  \begin{center}
+    \label{orch_eval2}
+    \begin{tabular}{|c|c|c|c|c|c|c|c|} 
+    	  \hline
+      \textbf{Model} & \textbf{Chord1} & \textbf{Reed Bone} & \textbf{Chord2} & \textbf{Gong} & \textbf{Scream} & \textbf{Trompe Attack} & \textbf{Average}  \\
+      \hline
+      CNN with LSTM & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+      \hline
+      ResNet & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+      \hline
+      OrchIdea & 0 & 0 & 0 & 0 & 0 & 0 & 0   \\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{Continuation of Table 3}
+\end{table*}
 
 \section{Conclusion}
-- the approach seems promising for orchestration
-- many things used in Orchidea are not implemented here (symbolic constraints, sparsity,...)
-- CNN seems better for timbre
-- ResNet seems better for pitch (what are timbre and pitch??)
 
-Based on our accuracies from training the model and the results of orchestrating 15 targets, our approach seems promising for orchestration.
-When comparing our method with OrchIdea, it is important to note that many of processes OrchIdea uses to optimize its solution were not used in our model.
-This includes important aspects such as sparsity, partial filtering, \textbf{(Carmine any more to add?)}.
+Classification through deep learning seems promising for orchestration because of our model's ability to classify individual instruments and pitches out of dense combinations of samples.
+While our model is not able to outperform OrchIdea, it shows consistent results.
+When comparing our methods to OrchIdea, it is important to note that many of processes OrchIdea uses to optimize its solution are not present in our model.
 Our model also lacks the implementation of symbolic constraints, which is an important part of assisted orchestration.
@@
 We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets.
-Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the pitch(es) of the target.
+Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.
 
 \subsection{Interpreting the Latent Space}
-- the system finds filters like...
 
 \section{Future steps}
-- using conditioning to impose symbolic constraints
-- variable size solutions
-- joint networks for orchestral size detection and orchestral families (see paper)
 
@@
 This leads to a solution whose harmonics are much closer to the target, which is an important part of aural similarity.
-
-\section{Citations}
-
-All bibliographical references should be listed at the end,
-inside a section named ``REFERENCES,'' numbered and in alphabetical order.
-All references listed should be cited in the text.
-When referring to a document, type the number in square brackets
-\cite{Author:00}, or for a range \cite{Author:00,Someone:10,Someone:04}.
-
-When the following words appear in the conference publication titles, please abbreviate them: Proceedings $\rightarrow$ Proc.; Record $\rightarrow$ Rec.; Symposium $\rightarrow$ Symp.; Technical Digest $\rightarrow$ Tech.
-Dig.; Technical Paper $\rightarrow$ Tech.
-Paper; First $\rightarrow$ 1st; Second $\rightarrow$ 2nd; Third $\rightarrow$ 3rd; Fourth/nth $\rightarrow$ 4th/nth.
-
-\textcolor{red}{As submission is double blind, refer to your own published work in the third person.
-That is, use ``In the previous work of \cite{Someone:10},'' not ``In our previous work \cite{Someone:10}.'' If you cite your other papers that are not widely available (e.g., a journal paper under review), use anonymous author names in the citation, e.g., an author of the form ``A.
-Anonymous.''}
-
-\section{Figures}
 

### 2020-05-14 model.py
Added code for showing latent space and added to the paper

@@ -1,3 +1,5 @@
+import torch
+import numpy as np
 import torch.nn as nn
 import torch.nn.functional as F
 
@@ -60,36 +62,47 @@ class CNN(nn.Module):
         self.activation = F.sigmoid
 
     def forward(self, x):
-        #print(x.shape)
         out = self.layer1(x)
-        #print(out.shape)
         out = self.layer2(out)
-        #print(out.shape)
         out = self.layer3(out)
-        #print(out.shape)
 
         out = out.view(-1, self.lstm_shape[0], self.lstm_shape[1])
-        #print(out.shape)
         out, _ = self.lstm(out)
 
-        #print(out.shape)
         out = out.view(-1, 32, self.lstm_shape[0], self.lstm_shape[1])
-#        print(out.shape)
         out = self.layer4(out)
-        #print(out.shape)
-
-        #print(out.shape)
+        
         out = out.contiguous().view(out.size()[0], -1)
-        #print(out.shape)
         out = self.fc1(out)
-        #print(out.shape)
         out = self.fc3(out)
-        #print(out.shape)
         out = self.activation(out)
-        #print(out.shape)
 
         return out
+    
+    def getLatentSpace(self, x):
+        res = []
+        
+        out = self.layer1(x)
+        res.append(out.detach().cpu().numpy())
+        out = self.layer2(out)
+        res.append(out.detach().cpu().numpy())
+        out = self.layer3(out)
+        res.append(out.detach().cpu().numpy())
 
+        out = out.view(-1, self.lstm_shape[0], self.lstm_shape[1])
+        out, _ = self.lstm(out)
+
+        out = out.view(-1, 32, self.lstm_shape[0], self.lstm_shape[1])
+        res.append(out.detach().cpu().numpy())
+        out = self.layer4(out)
+        res.append(out.detach().cpu().numpy())
+        
+        out = out.contiguous().view(out.size()[0], -1)
+        out = self.fc1(out)
+        out = self.fc3(out)
+        out = self.activation(out)
+        res.append(out.detach().cpu().numpy())
+        return res
 
 class OrchMatchNet(nn.Module):
     def __init__(self, out_num, model_select, features_dim):
@@ -103,6 +116,9 @@ class OrchMatchNet(nn.Module):
         out = self.net(x)
 
         return out
+    
+    def getLatentSpace(self, feature):
+        return self.net.getLatentSpace(feature)
 
 if __name__ == "__main__":
     cnn = CNN(out_num=505)

### 2020-05-14 paper/latex/paper.tex

@@
 \usepackage{lineno}
-\linenumbers
+%\linenumbers
+\newcommand{\carmine}[1]{\textbf{\color{red} Carmine => #1}}
 
@@
 
-Computer-assisted orchestration is an important problem ... \textbf{Carmine can you add more?}.
+Computer-assisted orchestration is an important problem ... \carmine{Can you add more?}.
 
+The approach studied in \cite{Cella19} consists in finding a good orchestration for any given sound by successive iterations, exploring a feature space defined by a fixed feature function $f$.
+This method is currently considered state of the art and was implemented in Orch-Idea.\\
+
+In this paper, we try a different approach, using Deep Neural networks.
+The main idea is to train a model to classify combinations of real instruments, and use it then for orchestration.
+By training it with real combinations of instruments, such as \{Flute C6, Bassoon C4, Bassoon G4\}, the model will hopefully acquire the ability to identify the presence of each instrument and its associaed pitch by building the appropriate filter.
+Thus, when a sound is given as input, those filters will identify what are the best instruments that match the target sound.\\
+
+This method is motivated by the good results obtained in previous research on musical instruments identification \cite{Benetos07}\cite{Kitahara05}\cite{Chetry05} and the more recent use of deep neural networks for musical classification \cite{lostanlen16}\cite{Bian}\cite{Brunner}\cite{Defossez19}.
+\carmine{Can you check if those references are good to use in this context?}\\
+
+\iffalse
 We attempt to use neural networks to solve the assisted orchestration problem.
@@
 When a new sound, one that is not simply a mixture of notes, is fed to the model, the predictions are used to create an orchestrated solution.
-The features we use are the mel spectrogram of the input sound, because of its ability to represent the way we hear sounds and its common use in MIR\textbf{Reference}.
+
+\fi
+
+The features we use are the mel spectrogram of the input sound, because of its ability to represent the way we hear sounds and its common use in MIR\textbf{Reference}.\\
 
@@
 Therefore the need for neural networks arose.
-The two architectures we used are a Convolutional Neural Network (CNN) with a Long Short-Term Memory (LSTM) unit, and ResNet, a well known residual architecture.
-We chose to use a CNN because of its success in audio classification \cite{Hershey2016}.
-Our inclusion of an LSTM is because of its ability to learn long term dependencies in data \cite{Hochreiter1997}, which is important given the temporal nature of audio.
-After training these models on varying combinations of 10 instruments, we orchestrated 15 target sounds by feeding them to the networks and synthesizing a solution from the most likely samples output from the system.
+The two architectures we used are a Convolutional Neural Network (CNN) with a Long Short-Term Memory (LSTM) unit, and ResNet, a well known residual architecture that already yielded good results for image classification \cite{He15}.
+We chose to use a CNN because of its success in audio classification \cite{Hershey16}.
+Our inclusion of an LSTM is because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.\\
+
+After training these models on varying combinations of 10 instruments, we used them for orchestration.
+More specifically, we orchestrated 15 target sounds by feeding them to the networks and synthesizing a solution from the most likely samples output from the system.
 We then compared our orchestrations to OrchIdea's solutions for the same targets through both qualitative and quantitative means.
 For quantitative comparison, we computed a distance metric on the FFTs of the target and solution, comparing the distance between OrchIdea's solution and the target and our model's solution and the target.
+\carmine{Do we need to reference the use of that distance?}
 
@@
 \subsection{Data Source and Generation}
+
 We used the TinySOL database to create our input data.
@@
 Each sample is one instrument playing a single note in the ordinario style, with one of three dynamics: \textit{pp}, \textit{mf}, or \textit{ff}.
-The instruments and ranges over which they were recorded is summarized in Tab.\ref{tab_pitch}.
+The instruments and ranges over which they were recorded is summarized in Tab.\ref{tinysol}.
+\carmine{Can you review this part about TinySOL and add more details if needed?}\\
 
-We generated all the data we used to train our models.
-For a chosen $N$, each input to our model is a combination of $N$ TinySOL samples.
-An orchestra of $N$ instruments is chosen and the data is generated in the following way:
+We generated all the data we used to train our models based on the samples of TinySOL.
+For a given $N$, each input to our model is a combination of $N$ TinySOL samples chosen among an orchestra of 10 instruments.
+The data is generated in the following way:
 
 \begin{itemize}
-\item For each instrument in the orchestra, a random TinySOL sample for that instrument is chosen. This leads to a random combination of both notes and dynamics.
-\item The chosen samples are combined to be played simultaneously and normalized by the number of instruments.
+\item We iteratively choose a random instrument, pitch and dynamic to pick the sample.
+We repeat that step $N$ times.
+Note that we don't allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures.
+\item The chosen samples are additively combined to be played simultaneously and normalized by the number of instruments.
 \item The mel spectrogram of the mixture is computed to be used as the features input to the model.
@@
 
-The choice of using the mel spectrogram as input features of the model is very commmon in music information \textbf{\{Insert reference\}} and can be considered to be an accurate representation of the way humans perceive sound and music.
+The choice of using the mel spectrogram as input features of the model is very commmon in music information \cite{Stevens37} and can be considered to be an accurate representation of the way humans perceive sound and music.\carmine{Do we need more references for mel spectrogram?}\\
 
@@
 Therefore, all the features fed to the model were matrices of size $128\times 345$.
-We used Librosa package in Python to compute the features, and more details on the exact computations can be found in \textbf{\{Reference to Librosa\}} The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms could be perceived by the human ear.
+We used Librosa package in Python to compute the features, and more details on the exact computations can be found in \cite{mcfee15} The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms could be perceived by the human ear \carmine{Do you have a reference for that?
+I don't find anything on the topic}.
 
@@
 \end{tabular}
-\caption{Table showing the pitch ranges present in the TinySOL database.}
+\caption{Table showing the pitch ranges present in the TinySOL database.\label{tinysol}}
 \end{table}
@@
 
-In order to increase variability in our generated data, we used two methods of data augmentation.
+In order to increase variability in our generated data, we used two methods of data augmentation \cite{Salamon17}\cite{Bhardwaj17} : pitch shifting and partial feature dropout.\\
 
-The first one was done on the soundfile resulting from the mixture of the samples, before the mel spectrogram was taken.
+The first one was done on the TinySOL samples.
 We performed a slight pitch shift by shifting the frequency of the signal.
-We did this by ...
-\textbf{Alejandro, can you add how you did the pitch shift (and maybe explain why changing the SR leads to a pitch shift)}In practice, the sample rate used to compute the Mel spectrogram was modified to be within $5\%$ of the actual $44100$Hz.
+We did this by reading the samples with different sample rates.
+Python package librosa automatically converts the data coming from the samples to match the specified sample rate \cite{mcfee15}.
+Thus, a small difference in sample rate will slightly modify the perceived pitch if played at the normal sample rate.
+In practice, the sample rates used for this data augmentation were within $5\%$ of the actual $44100$Hz.\\
 
@@
 For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$  columns and $3.45$ rows being zero-ed out.
-This method of data augmentation aimed to be more resilient to the possible errors of recording, due either to a setup that induces interferences and cancels specific frequencies, or to a default in the microphone that fails to record a small temporal portion.
+This method of data augmentation aimed to be more resilient to the possible errors of recording \carmine{Is there any other reference to justify this sentence?}, due either to a setup that induces interferences and cancels specific frequencies, or to a default in the microphone that fails to record a small temporal portion.
 
@@
 The first deep model we tested was a Convolutional Neural Network (CNN) with a Long Short-Term Memory (LSTM) unit.
-CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey2016}.
-Long Short-Term Memory (LSTM) units provide a way to learn long term dependencies in the data \cite{Hochreiter1997}, which is relevant given the sequential nature of audio.
+CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey16}.
+Long Short-Term Memory (LSTM) units provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.
+
+\carmine{Do we need any more justification on why we used that model?}
 
@@
 
+\carmine{Does this architecture need to be explained in more details?}
+
 Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class. The model architecture is shown in Figure \ref{cnn_lstm}.
@@
 \subsection{ResNet}
-\textbf{Carmine, can you add more to this section about why ResNet is good, and maybe some details about the implementation?}
+\carmine{Can you add more to this section about why ResNet is good, and maybe some details about the implementation?}
 The second deep model that we used was the well known deep residual network ResNet.
@@
 
-During training, the loss function used to optimize the inner parameters of the model was binary cross entropy, as it is the common choice for multiclass multilabel classification frameworks such as this one.
+During training, the loss function used to optimize the inner parameters of the model was binary cross entropy, as it is the common choice for multiclass multilabel classification frameworks such as this one \carmine{Is there a specific paper justifying that claim?
+Or is it obvious enough?}.
 However, the value of the loss function alone is difficult to interpret.
@@
 
-Different experiments were made by varying the number $N$ of instruments in each mixture sample.
+Different experiments were made by varying the number $N$ of samples in each mixture sample.
 We used an orchestra of 10 instruments containing Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet C and Clarinet Bb.
@@
 Fig.
-\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of instruments $N$.\\
+\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$.\\
 
-\textbf{(Alejandro, add how there were certain instruments that performed poorly and some that were great. Add how ResNet was awful for Flute and always output Flute C7)}
+\begin{figure}
+\center{
+\includegraphics[scale=0.4]{figs/CNN_vs_ResNet.png}}
+\caption{Best overall accuracy for CNN with LSTM and ResNet depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
+\end{figure}
 
-From this plot, we see that ResNet achieves better performances than the CNN regardless of the number of instruments used in the combination.
-This result isn't surprising \textbf{\color{red} Interpretation of why ResNet is better than CNN}
+From this plot, we see that ResNet achieves better performances than the CNN regardless of the number of samples used in the combination.
+This result isn't surprising as residual networks usually perform better for this kind of classification problems \carmine{It would need a citation on why ResNet was expected to be better}
 
@@
 Note that for ResNet, the variance of accuracy is much smaller until $N$ reaches 5 before becoming similar to the CNN.
-Another key observation is that both models struggle to classify the same instruments.
-Flute, Trombone and Trumpet yield the worst results for both models, being notably worst \textbf{\color{red} Use same colors for instruments}
+The results on both figures show consistency on the relative accuracy of instruments, which is a first step towards the validation of this method \carmine{Can I say that?}.
+Flute, Trombone and Trumpet yield the worst results for both models, being notably worst, which could be explained by the nature of their spectra.
+For the Flute, since it has very few harmonics, it is difficult to identify it among other instruments, especially if they are richer.
+On the other hand, brasses have a very rich spectra and a strong noisy component that excites almost all the frequencies.\\
+Strings give consistent results, having for both models similar accuracies, an interesting point to notice is the very high accuracy of Oboe on both models.
+This could indicate that there is an optimal balance of harmonies that maximizes the probability of being detected in such classification framework.\carmine{Please review those interpretations}
 
 \begin{figure}
-\includegraphics[scale=0.57]{figs/CNN_vs_ResNet.png}
-\caption{Best overall accuracy and for CNN and ResNet depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
+\includegraphics[scale=0.5]{figs/Acc_vs_N_CNN.png}
+\caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM depending on the number of instruments in the combinations. \label{best_acc_cnn}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.5]{figs/Acc_vs_N_ResNet.png}
+\caption{Best overall accuracy and for each instrument obtained by the ResNet depending on the number of instruments in the combinations. \label{best_acc_resnet}}
 \end{figure}
@@
 When comparing our methods to OrchIdea, it is important to note that many of processes OrchIdea uses to optimize its solution are not present in our model.
-Our model also lacks the implementation of symbolic constraints, which is an important part of assisted orchestration.
+Our model also lacks the implementation of symbolic constraints, which is an important part of assisted orchestration.\\
 
@@
 
+After training the CNN with LSTM for $N=10$ samples per combination, it is possible to visualize the filters applied to the input layer after layer.
+Figs.
+\ref{latent0}-\ref{latent5} show the successive outputs of each layer of the convolutional network.
+\carmine{I put here all the intermediary features of the CNN with LSTM.
+I don't find any particular conclusion other than saying that LSTM seems to focus more on vertical components rather than horizontal, ie time over frequency}
+
+\begin{figure}
+\includegraphics[scale=0.5]{figs/latent_space_layer0.png}
+\caption{Initial feature matrix used as input of the model \label{latent0}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.5]{figs/latent_space_layer1.png}
+\caption{Intermediary features after the first convolutional layer of the CNN with LSTM. \label{latent1}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.5]{figs/latent_space_layer2.png}
+\caption{Intermediary features after the second convolutional layer of the CNN with LSTM. \label{latent2}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.5]{figs/latent_space_layer3.png}
+\caption{Intermediary features after the third convolutional layer of the CNN with LSTM. \label{latent3}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.5]{figs/latent_space_layer4.png}
+\caption{Intermediary features after the LSTM layer. \label{latent4}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=0.5]{figs/latent_space_layer5.png}
+\caption{Intermediary features after final convolutional layer of the CNN with LSTM. \label{latent5}}
+\end{figure}
+
 \section{Future steps}
 
-Future steps in this project include implementing various methods that are present in OrchIdea.
+Future steps in this project include implementing various methods that are present in OrchIdea.\\
 
@@
 The solution to this is to allow sparse solutions in which the model decides how many samples should be used to best represent the target.
-This allows a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds.
+This allows a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds.\\
 
@@
 
+\iffalse
 \begin{figure}
@@
 \end{figure}
-
-\begin{figure*}
-\includegraphics[scale=0.6]{figs/Acc_vs_N_CNN.png}
-\caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM depending on the number of instruments in the combinations. \label{best_acc_cnn}}
-\end{figure*}
-
-\begin{figure*}
-\includegraphics[scale=0.6]{figs/Acc_vs_N_ResNet.png}
-\caption{Best overall accuracy and for each instrument obtained by the ResNet depending on the number of instruments in the combinations. \label{best_acc_resnet}}
-\end{figure*}
+\fi
 

### 2020-05-14 test.py
Implement OrchIdea ratio calculations

@@ -27,8 +27,8 @@ to be the path to this folder
 # path to TinySOL data
 tinysol_path = './TinySOL'
 
-# cnn or resnet
-model_type = 'resnet'
+# path to target samples
+target_path = './target_samples'
 
 # number of samples to be used in solution
 n = 10
@@ -36,22 +36,6 @@ n = 10
 # instruments to be used (all instruments will be used)
 instr_filter = ['Hn', 'Ob', 'Vn', 'Va',
                 'Vc', 'Fl', 'Tbn', 'Bn', 'TpC', 'ClBb']
-
-# path to target samples
-target_path = './target_samples'
-
-# path to store solutions as .wav
-solutions_path = './orchestrated_targets/{}_n={}'.format(model_type, n)
-
-# The argument must be the folder where the params.pkl file is
-GLOBAL_PARAMS.load_parameters('./orchestrated_targets/params_{}'.format(model_type))
-
-# path to a trained version of the model
-if model_type == 'cnn':
-    state_path = './orchestrated_targets/params_cnn/epoch_49.pth'
-else:
-    state_path = './orchestrated_targets/params_resnet/epoch_24.pth'
-
 
 # if true, doesn't orchestrate targets but instead creates combinations of TinySOL samples to input to the model
 sanity_check = False
@@ -104,14 +88,14 @@ def test(model, state_path, data, targets):
             target['distance'] = compute_distance(target, mixed_file)
         
         # write to text file
-        f.write('Target: {}; Distance: {:,.2f}\nSamples used: {}\n\n'.format(target['name'], target['distance'], target['classes']))
+        f.write('Target: {}; Distance: {:,.8f}\nSamples used: {}\n\n'.format(target['name'], target['distance'], target['classes']))
 
     # compute avg distance
     sum = 0
     for target in targets:
         sum += target['distance']
     sum /= len(targets)
-    f.write('Average distance: {:,.2f}'.format(sum))
+    f.write('Average distance: {:,.8f}'.format(sum))
     f.close()
 
         
@@ -217,8 +201,11 @@ def compute_distance(target, solution):
     target_fft = np.abs(np.fft.rfft(target, N))
     solution_fft = np.abs(np.fft.rfft(solution, N))
 
-    lambda_1 = 0.1
-    lambda_2 = 2
+    target_fft /= (np.max(target_fft) * N)
+    solution_fft /= (np.max(solution_fft) * N)
+
[…]

### 2020-05-14 paper/latex/paper.tex
Revisions throughout paper, format target table

@@
 \begin{itemize}
-\item We iteratively choose a random instrument, pitch and dynamic to pick the sample.
-We repeat that step $N$ times.
+\item $N$ random TinySOL samples are selected, leading to a variety of instruments, pitches, and dynamics.
 Note that we don't allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures.
-\item The chosen samples are additively combined to be played simultaneously and normalized by the number of instruments.
-\item The mel spectrogram of the mixture is computed to be used as the features input to the model.
+\item The chosen samples are combined to be played simultaneously and normalized by the number of instruments.
+\item The mel spectrogram of the mixture is computed to be used as the features that are input to the model.
 \end{itemize}
 
-The choice of using the mel spectrogram as input features of the model is very commmon in music information \cite{Stevens37} and can be considered to be an accurate representation of the way humans perceive sound and music.\carmine{Do we need more references for mel spectrogram?}\\
+The choice of using the mel spectrogram as input features of the model is very common in music information \cite{Stevens37} and can be considered to be an accurate representation of the way humans perceive sound and music.\carmine{Do we need more references for mel spectrogram?}\\
 
-Samples had a sample rate of $44100$Hz and were all padded or chopped to be exactly 4 seconds long.
-The Mel spectrograms were then generated using an FFT hop length of 2048 samples (e.g.
+The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly 4 seconds long.
+The Mel spectrogram is generated using an FFT hop length of 2048 samples (e.g.
 the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128.
 Therefore, all the features fed to the model were matrices of size $128\times 345$.
-We used Librosa package in Python to compute the features, and more details on the exact computations can be found in \cite{mcfee15} The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms could be perceived by the human ear \carmine{Do you have a reference for that?
+We used the Librosa package in Python to compute the features, and more details on the exact computations can be found in \cite{mcfee15} The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms could be perceived by the human ear \carmine{Do you have a reference for that?
 I don't find anything on the topic}.
@@
 We did this by reading the samples with different sample rates.
-Python package librosa automatically converts the data coming from the samples to match the specified sample rate \cite{mcfee15}.
+Librosa automatically converts the data coming from the samples to match the specified sample rate \cite{mcfee15}.
 Thus, a small difference in sample rate will slightly modify the perceived pitch if played at the normal sample rate.
@@
 Then, for both CNN with LSTM and ResNet, we computed the maximum accuracy over the epochs. 
-For the CNN model, all the trainings with different values of $N$ were made using a dataset containing 200000 samples over 50 epochs, whereas for the ResNet, the dataset contained 400000 samples and was trained over 20 epochs.
+The CNN was trained on 200,000 generated samples over 50 epochs, and ResNet was trained on 400,000 generated samples over 20 epochs.
 Fig.
@@
 \ref{best_acc_cnn} and Fig.
-\ref{best_acc_resnet} show the maximum accuracy computed using the function $f$ explicited in (\ref{eval}).
-Note that for ResNet, the variance of accuracy is much smaller until $N$ reaches 5 before becoming similar to the CNN.
+\ref{best_acc_resnet} show the maximum accuracy computed using the function $f$ from Eqn.
+\eqref{eval}.
+For ResNet, the variance in accuracy is much smaller until $N$ reaches 5, at which point it becomes similar to the CNN.
 The results on both figures show consistency on the relative accuracy of instruments, which is a first step towards the validation of this method \carmine{Can I say that?}.
-Flute, Trombone and Trumpet yield the worst results for both models, being notably worst, which could be explained by the nature of their spectra.
+Flute, Trombone and Trumpet yield the worst results for both models.
+Their low accuracy could be explained by the nature of their spectra.
 For the Flute, since it has very few harmonics, it is difficult to identify it among other instruments, especially if they are richer.
-On the other hand, brasses have a very rich spectra and a strong noisy component that excites almost all the frequencies.\\
-Strings give consistent results, having for both models similar accuracies, an interesting point to notice is the very high accuracy of Oboe on both models.
+On the other hand, brasses have a very rich spectra and a strong noisy component that excites almost all the frequencies.
+
+Strings give similar results across both models.
+An interesting point to notice is the very high accuracy of Oboe on both models.
 This could indicate that there is an optimal balance of harmonies that maximizes the probability of being detected in such classification framework.\carmine{Please review those interpretations}
@@
 Best test accuracy obtained for each model and value of $N$ over the total number of epochs.
-The accuracy is computed using $f$ define in (\ref{eval}).
+The accuracy is computed using $f$ defined in \eqref{eval}.
 The first row correspond to the overall accuracy of the model, regrouping all the instruments.
@@
 These 10 classes represent the instrument-note pairs that are most represented in the target, and can be from any combination of the 10 instruments.
-The dynamic of each sample is inferred from its probability.
 
 We did not train our model to classify the dynamics of a sample despite TinySOL having pianissimo, mezzo forte, and fortissimo recordings for each sample.
-If we had, then each class would have only one data point: the TinySOL sample ir represents.
-By not initially ignoring dynamics, this led to each class having three data points: the instrument-note at \textit{pp}, \textit{mf}, and \textit{ff} When the model is used for orchestration, the dynamic is determined by the probability of that sample as output by the model.
+If we had, then each class would have only one data point: the TinySOL sample it represents.
+When the model is used for orchestration, the dynamic is determined by the probability of that sample as output by the model.
 If the model outputs a probability higher than $0.66$ for a sample, the fortissimo version of the sample was used.
@@
     \label{orch_eval}
-    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|} 
+    \begin{tabular}{|c|c|c|c|c|c|c|c|c|} 
     	  \hline
-      \textbf{Model} & \textbf{Ob + Bn} & \textbf{Bn} & \textbf{Bass Clarinet} & \textbf{Bell1} & \textbf{Bell2} & \textbf{Yan} & \textbf{Car Horn} & \textbf{Boat} & \textbf{Wind Harp} \\
+      \textbf{Model} & \textbf{Ob + Bn} & \textbf{Bn} & \textbf{Bass Clarinet} & \textbf{Bell1} & \textbf{Bell2} & \textbf{Yan} & \textbf{Car Horn} & \textbf{Boat}  \\
       \hline
-      CNN with LSTM & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+      CNN with LSTM & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
       \hline
-      ResNet & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+      ResNet & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
       \hline
-      OrchIdea & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0  \\
+      OrchIdea & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+      \hline
+      \hline
+      \textbf{Model} & \textbf{Wind Harp} & \textbf{Chord1} & \textbf{Reed Bone} & \textbf{Chord2} & \textbf{Gong} & \textbf{Scream} & \textbf{Trompe Attack} & \textbf{Average}  \\
+      \hline
+      CNN with LSTM & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+      \hline
+      ResNet & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+      \hline
+      OrchIdea & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
       \hline
@@
   \caption{Quantitative comparison of orchestrations.}
-\end{table*}
-\begin{table*}
-  \begin{center}
-    \label{orch_eval2}
-    \begin{tabular}{|c|c|c|c|c|c|c|c|} 
-    	  \hline
-      \textbf{Model} & \textbf{Chord1} & \textbf{Reed Bone} & \textbf{Chord2} & \textbf{Gong} & \textbf{Scream} & \textbf{Trompe Attack} & \textbf{Average}  \\
-      \hline
-      CNN with LSTM & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-      \hline
-      ResNet & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-      \hline
-      OrchIdea & 0 & 0 & 0 & 0 & 0 & 0 & 0   \\
-      \hline
-    \end{tabular}
-  \end{center}
-  \caption{Continuation of Table 3}
 \end{table*}

### 2020-05-14 paper/latex/paper.tex
Final revisions

@@
 The approach studied in \cite{Cella19} consists in finding a good orchestration for any given sound by successive iterations, exploring a feature space defined by a fixed feature function $f$.
-This method is currently considered state of the art and was implemented in Orch-Idea.\\
+This method is currently considered state of the art and was implemented in Orch-Idea.
 
@@
 By training it with real combinations of instruments, such as \{Flute C6, Bassoon C4, Bassoon G4\}, the model will hopefully acquire the ability to identify the presence of each instrument and its associaed pitch by building the appropriate filter.
-Thus, when a sound is given as input, those filters will identify what are the best instruments that match the target sound.\\
+Thus, when a sound is given as input, those filters will identify what are the best instruments that match the target sound.
 
 This method is motivated by the good results obtained in previous research on musical instruments identification \cite{Benetos07}\cite{Kitahara05}\cite{Chetry05} and the more recent use of deep neural networks for musical classification \cite{lostanlen16}\cite{Bian}\cite{Brunner}\cite{Defossez19}.
-\carmine{Can you check if those references are good to use in this context?}\\
+\carmine{Can you check if those references are good to use in this context?}
 
@@
 
-The features we use are the mel spectrogram of the input sound, because of its ability to represent the way we hear sounds and its common use in MIR\textbf{Reference}.\\
+The features we use are the mel spectrogram of the input sound, because of its ability to represent the way we hear sounds and its common use in MIR\textbf{Reference}.
 
@@
 We chose to use a CNN because of its success in audio classification \cite{Hershey16}.
-Our inclusion of an LSTM is because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.\\
+Our inclusion of an LSTM is because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.
 
@@
 The instruments and ranges over which they were recorded is summarized in Tab.\ref{tinysol}.
-\carmine{Can you review this part about TinySOL and add more details if needed?}\\
+\carmine{Can you review this part about TinySOL and add more details if needed?}
 
@@
 
-The choice of using the mel spectrogram as input features of the model is very common in music information \cite{Stevens37} and can be considered to be an accurate representation of the way humans perceive sound and music.\carmine{Do we need more references for mel spectrogram?}\\
+The choice of using the mel spectrogram as input features of the model is very common in music information \cite{Stevens37} and can be considered to be an accurate representation of the way humans perceive sound and music.\carmine{Do we need more references for mel spectrogram?}
 
@@
 
-In order to increase variability in our generated data, we used two methods of data augmentation \cite{Salamon17}\cite{Bhardwaj17} : pitch shifting and partial feature dropout.\\
+In order to increase variability in our generated data, we used two methods of data augmentation \cite{Salamon17}\cite{Bhardwaj17} : pitch shifting and partial feature dropout.
 
@@
 Thus, a small difference in sample rate will slightly modify the perceived pitch if played at the normal sample rate.
-In practice, the sample rates used for this data augmentation were within $5\%$ of the actual $44100$Hz.\\
+In practice, the sample rates used for this data augmentation were within $5\%$ of the actual $44100$Hz.
 
@@
 For orchestrating, we trained our models on combinations of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet.
+This leads to 424 classes, each one representing an instrument-note pair.
 A target sound is input to the model, and the 10 classes with the highest probability are extracted.
-These 10 classes represent the instrument-note pairs that are most represented in the target, and can be from any combination of the 10 instruments.
+These 10 classes are the instrument-note pairs that are most represented in the target, and can be from any combination of the 10 instruments.
 
@@
 We evaluated our orchestrations both qualitatively and quantitatively.
+We compared our solutions to OrchIdea, the state-of-the-art program for computer-assisted orchestration.
+In order to have comparable solutions, we did not allow OrchIdea to create sparse solutions, meaning that OrchIdea was forced to use all 10 instruments in each solution.
+This creates a more fair comparison since our model is unable to create sparse solutions.
+
 Qualitative evaluation was done through an acoustic inspection of the solution, paying close attention to timbre and pitch.
@@
 
-For quantitative evaluation, we used the distance metric defined in \eqref{distance} to calculate differences in timbre between targets and solutions.
+For quantitative evaluation, we used the distance metric defined in Eqn.
+\eqref{distance} to calculate differences in timbre between targets and solutions.
 The equation takes in the full FFT of the target and full FFT of the solution.
@@
 
-In the summations, all the terms with a $\lambda_1$ will be distances calculated when there was more energy in the target than the solution, since $\delta_{k1} = 1$ only when $x_k \ge \tilde{x}_k$.
-By the same reasoning, all the terms with a $\lambda_2$ are distances when the solution provided more energy to a frequency than the target; it overshot the frequency.
+In the first summation, $\lambda_1$ is multiplied by all the distances calculated when there was more energy in the target than the solution, since $\delta_{k1} = 1$ only when $x_k \ge \tilde{x}_k$.
+Similarly in the second summation, $\lambda_2$ is multiplied by all the distances when the solution provided more energy to a frequency than the target; it overshot the frequency.
 Therefore, the relation between $\lambda_1$ and $\lambda_2$ determines whether a solution is penalized more for undershooting or overshooting the target.
@@
 \ref{orch_eval}.
-
-In order to have comparable solutions, we did not allow OrchIdea to create sparse solutions, meaning that OrchIdea was forced to use all 10 instruments in each solution.
-This creates a more fair comparison since our model is unable to create sparse solutions.
 
@@
 When comparing our methods to OrchIdea, it is important to note that many of processes OrchIdea uses to optimize its solution are not present in our model.
-Our model also lacks the implementation of symbolic constraints, which is an important part of assisted orchestration.\\
+Our model also lacks the implementation of symbolic constraints, which is an important part of assisted orchestration.
 
@@
 
-Future steps in this project include implementing various methods that are present in OrchIdea.\\
+Future steps in this project include implementing various methods that are present in OrchIdea.
 
@@
 The solution to this is to allow sparse solutions in which the model decides how many samples should be used to best represent the target.
-This allows a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds.\\
+This allows a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds.
 

### 2020-05-14 test.py
Use new CNN and ResNet models

@@ -289,7 +289,7 @@ if __name__ == "__main__":
 
         # The argument must be the folder where the params.pkl file is
         GLOBAL_PARAMS.load_parameters(
-            './orchestrated_targets/params_{}'.format(model_type))
+            './orchestrated_targets/new_models/{}_params'.format(model_type))
 
         # path to a trained version of the model
         if model_type == 'cnn':

### 2020-05-14 paper/latex/paper.tex
Add orchestration results data

@@
       \hline
-      CNN with LSTM & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+      CNN with LSTM & 0.17 & 0.28 & 0.70 & 0.55 & 0.26 & \textbf{1.10} & 0.68 & \textbf{1.12} \\
       \hline
-      ResNet & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+      ResNet & 0.34 & 0.50 & 0.48 & 0.59 & 0.45 & 0.90 & 0.49 & \textbf{1.16} \\
       \hline
-      OrchIdea & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+      OrchIdea & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 \\
       \hline
@@
       \hline
-      CNN with LSTM & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+      CNN with LSTM & 0.55 & - & 0.70 & - & 0.73 & \textbf{1.14} & 0.79 & 0.59\\
       \hline
-      ResNet & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+      ResNet & 0.61 & - & 0.51 & - & 0.71 & \textbf{1.03} & \textbf{1.05} & 0.55 \\
       \hline
-      OrchIdea & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+      OrchIdea & 1 & - & 1 & - & 1 & 1 & 1 & 1 \\
       \hline
@@
   \end{center}
-  \caption{Quantitative comparison of orchestrations.}
+  \caption{Quantitative comparison of orchestrations as ratios to OrchIdea.
+Eqn.
+\eqref{distance} was used with $\lambda_1 = 0.5$ and $\lambda_2 = 10$ to compute distances between orchestrations and targets.
+What is shown is the the ratio between the distance calculated and OrchIdea's distance for the same target.
+A value less than 1 means that the model performed worse (i.e.
+had a larger distance), and a value greater than 1 means the model performed better than OrchIdea.
+"Average" is the ratio of the average distances for the model across all targets.}
 \end{table*}

### 2020-05-15 paper/latex/paper.tex
Moved figures to figs folder and reduced size of diagram
several modification for the final revision - not finished

--- 2020-05-14 paper/latex/paper.tex
@@
 % ------
-\title{Neural models for computer-assisted musical orchestration: a preliminary study}
+\title{Neural models for target-based computer-assisted musical orchestration: a preliminary study}
 
@@
 \begin{abstract}
-In this paper we will explore preliminary neural models for the task of computer-assisted musical orchestration.
-After an introduction on the problem, we will show how we decided to model musical orchestration as a classification task.
-In this context, we will propose two deep learning models and we will show how they perform as classifiers for musical instruments recognition by comparing them with specific baselines.
-We will the show how they perform, both qualitative and quantitative, in the task of computer-assisted orchestration by comparing them with state-of-the-art systems.
-We will highlight, finally, benefits and problems of neural approaches for orchestration and we will propose possible future steps.
+In this paper we will explore preliminary neural models for the task of target-based computer-assisted musical orchestration.
+After an introduction on computer-assisted orchestration, we will show how we model this  musical problem as a classification task.
+In this context, we will propose two deep learning models and we will show, first, how they perform as classifiers for musical instruments recognition by comparing them with specific baselines.
+We will then show how they perform, both qualitative and quantitative, in the task of computer-assisted orchestration by comparing them with state-of-the-art systems.
+We will highlight, finally, benefits and problems of neural approaches for assisted orchestration and we will propose possible future steps.
 \end{abstract}
@@
 
-The approach studied in \cite{Cella19} consists in finding a good orchestration for any given sound by successive iterations, exploring a feature space defined by a fixed feature function $f$.
-This method is currently considered state of the art and was implemented in Orch-Idea.
+The approach studied in \cite{Cella18, Cella2020} consists in finding a good orchestration for any given sound by searching combinations of sounds from a database with a multi-objective optimization heuristics and a constraint solver that are jointly optimized.
+Both the target sound and the sounds in the database are embedded in a feature space defined by a fixed feature function $f$ and each generated combination of sounds is evaluated by using a specific metric.
+This method is currently considered state of the art and is implemented in the \emph{Orchidea} toolbox for assisted orchestration (\url{www.orch-idea.org}); we will describe this approach in more detail in the next sections.
 
-In this paper, we try a different approach, using Deep Neural networks.
-The main idea is to train a model to classify combinations of real instruments, and use it then for orchestration.
-By training it with real combinations of instruments, such as \{Flute C6, Bassoon C4, Bassoon G4\}, the model will hopefully acquire the ability to identify the presence of each instrument and its associaed pitch by building the appropriate filter.
-Thus, when a sound is given as input, those filters will identify what are the best instruments that match the target sound.
+In this paper, we try a different approach by experimenting with deep neural architectures.
+The main idea is to train a model to classify combinations of real instruments and use it then for orchestration.
+A typical solution for assisted orchestration is a set of triples \emph{instrument-pitch-dynamics} such as \{\texttt{Flute C6 pp, Bassoon C4 mf, Bassoon G4 ff}\}.
+By training a neural network with real combinations of instrumental notes, it will acquire the ability to identify the presence of each instrument and its associated pitch by building the appropriate latent representation.
+Thus, when an unknown target sound is given as input, the network will identify which are the best instruments-pitches pairs to match the target sound, and it will be able to  deconstruct a complex mixture of timbres into individual instrument-note pairs\footnote{The dynamics is retrived using the probability distributions generated as output by the network; more details on this will be given in the next sections.}.
 
-This method is motivated by the good results obtained in previous research on musical instruments identification \cite{Benetos07}\cite{Kitahara05}\cite{Chetry05} and the more recent use of deep neural networks for musical classification \cite{lostanlen16}\cite{Bian}\cite{Brunner}\cite{Defossez19}.
-\carmine{Can you check if those references are good to use in this context?}
+This method is motivated by the good results obtained in previous research on musical instruments identification \cite{Benetos07, Kitahara05} and the more recent use of deep neural networks for musical classification \cite{lostanlen16, Bian19}.
 
-\iffalse
-We attempt to use neural networks to solve the assisted orchestration problem.
-Our model must be able to take in a sound and identify what instruments playing which notes best represents the sound.
-Our idea is to frame this as a classification problem, where each class is an instrument-note pair.
-Our classifiers will be trained on sounds that are combinations of multiple instruments and notes, leading to a model that can take a complex mixture of timbres and deconstruct it into individual instrument-note pairs.
-When a new sound, one that is not simply a mixture of notes, is fed to the model, the predictions are used to create an orchestrated solution.
-
-\fi
-
-The features we use are the mel spectrogram of the input sound, because of its ability to represent the way we hear sounds and its common use in MIR\textbf{Reference}.
-
-We found that even on a simplified version of this classification problem, parametric classifiers were unable to meaningfully identify instrument-note pairs as the number of instruments used in the combinations increased.
-Therefore the need for neural networks arose.
-The two architectures we used are a Convolutional Neural Network (CNN) with a Long Short-Term Memory (LSTM) unit, and ResNet, a well known residual architecture that already yielded good results for image classification \cite{He15}.
-We chose to use a CNN because of its success in audio classification \cite{Hershey16}.
-Our inclusion of an LSTM is because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.
+In this paper we do preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit in the middle and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}.
+We chose to use a CNN because of its success in audio classification \cite{Hershey16} and we decided to include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.
 
@@
 More specifically, we orchestrated 15 target sounds by feeding them to the networks and synthesizing a solution from the most likely samples output from the system.
-We then compared our orchestrations to OrchIdea's solutions for the same targets through both qualitative and quantitative means.
-For quantitative comparison, we computed a distance metric on the FFTs of the target and solution, comparing the distance between OrchIdea's solution and the target and our model's solution and the target.
-\carmine{Do we need to reference the use of that distance?}
+We then compared our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means.
+For quantitative comparison, we computed the distance metric suggested in \cite{Cella2020} on the FFTs of the target and solution, comparing the distance between Orchidea's solution and the target and our model's solution and the target.
 
-Our codebase can be found at: \url{https://github.com/dzluke/DeepOrchestration}.
-\section{Data}
+Next sections will be as follows: section \ref{sec:models} will describe the dataset we used for experiments and the models we trained as classifiers, showing the classification results.
+Sections \ref{sec:orchestration} and \ref{sec:conclusions} will show, respectively, the results of orchestration experiments using the trained models and will present the conclusions of our preliminary study on neural approaches for target-based computer-assisted orchestration.
 
-\subsection{Data Source and Generation}
+The codebase for this paper and some audio examples can be found at: \url{https://github.com/dzluke/DeepOrchestration}.
 
-We used the TinySOL database to create our input data.
-TinySOL is a subset of the Studio On Line (SOL) database created by IRCAM \textbf{Reference needed}.
+\section{Neural models}
+\label{sec:models}
+
+\subsection{Form orchestration to classification}
+
+In this paper, we model assisted orchestration as a classification problem. The general methodology is as follows:
+\begin{enumerate}
+\item we train specific models to classify instrument-pitch pairs on sound combinations of a database of instrument notes, up to ten simultaneous instruments;
+\item we then pick the best classifier and we feed into it an unknown sound to be classified;
+\item since the output of the classifier will be made of instrument-pitch pairs, we use this information to synthesize an orchestration for the target sound;
+\item finally, we evaluate the generated orchestration against state-of-the-art systems for computer-assisted orchestrations.
+\end{enumerate}
+
+In other words, the classifiers learn how to take a complex combination of pitches and timbres and deconstruct it into its original parts.
+
+A complete orchestration solution, however, would normally be made by triples of instrument-pitch-dynamics.
+This is difficult, however, since to frame this problem as classification, we would need to have a very high number of classes.
+Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in \ref{sec:dataset}), each class would be represented by a single sample.
+For these reasons, our models have been trained on simplified versions of the original problem.
+
+On one hand, the baseline classifiers have been trained to recognize either the instrument only for a total of 12 classes or the pair instrument-pitch \emph{class} (a pitch class is the pitch of a note without the corresponding octave, e.g.
+C, D\#, G, etc.) for a total of 25 classes.
+
+On the other hand, for the neural models, we reframed the classification problem by using full pitch and trained them to recognize the pair instruments-pitch for a total of 424 classes (i.e.
+the neural classifiers would classify instruments and pitches with respective octaves).
+We want to point out, however, that this is still a simplified problem compared to assisted orchestration.
+Even in the case of instrument-pitch pair classification, indeed, we would miss the dynamics of the note.
+To cope with this problem, in the orchestration phase, we quantize the probability functions generated by the models.
+When a network is given a target sound as input, it will attempt to apply the same classification rules, outputting a vector of the probabilities of each sample being in the target sound.
+Since the classes only encode instrument-pitch pairs and since we also need the dynamics, we quantize the probability function generated by the network into three echelons coresponding, respectively, to the dynamic levels \emph{pianissimo}, \emph{mezzoforte} and \emph{fortissimo} (see section \ref{sec:orchestration}).
+By taking the instruments and pitches that have the highest probability and by using quantized probabilities to retrieve dynamics, a full orchestrated solution can be created.
+Figure \ref{fig:method} illustrates the described ideas.
+
+\begin{figure}
+\includegraphics[scale=0.5]{figs/Acc_vs_N_CNN.png}
+\caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM depending on the number of instruments in the combinations. \label{best_acc_cnn}}
+\end{figure}
+
+\begin{figure}
+\center{
+\includegraphics[scale=0.5]{figs/CNN_vs_ResNet.png}}
+\caption{Best overall accuracy for CNN with LSTM and ResNet depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
+\end{figure}
+
+\subsection{Dataset}
+\label{sec:dataset}
+
+To create the input data for training the classifiers, we used the \emph{TinySOL} database.
+TinySOL is a subset of the Studio On Line (SOL) database created by IRCAM \cite{Cella2020b}.
 TinySOL contains 1,529 samples from 12 instruments.
 The instruments come from different orchestral families: strings, woodwinds, and brass.
-Each sample is one instrument playing a single note in the ordinario style, with one of three dynamics: \textit{pp}, \textit{mf}, or \textit{ff}.
-The instruments and ranges over which they were recorded is summarized in Tab.\ref{tinysol}.
-\carmine{Can you review this part about TinySOL and add more details if needed?}
+Each sample is one instrument playing a single note in the \emph{ordinario} playing style, with one of three dynamics: \textit{pp}, \textit{mf}, or \textit{ff}, for example \texttt{Flute-C4-pp} or \texttt{Clarinet-D5-mf}.
+The instruments and ranges over which they were recorded is summarized in table \ref{tab:ranges}.
 
-We generated all the data we used to train our models based on the samples of TinySOL.
-For a given $N$, each input to our model is a combination of $N$ TinySOL samples chosen among an orchestra of 10 instruments.
-The data is generated in the following way:
+For a given number of instruments $N$, each input to our models is a combination of $N$ TinySOL samples chosen among an orchestra of 10 instruments. The data is generated in the following way:
 
 \begin{itemize}
-\item $N$ random TinySOL samples are selected, leading to a variety of instruments, pitches, and dynamics.
-Note that we don't allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures.
-\item The chosen samples are combined to be played simultaneously and normalized by the number of instruments.
-\item The mel spectrogram of the mixture is computed to be used as the features that are input to the model.
+\item $N$ random TinySOL samples are selected, leading to a variety of instruments, pitches, and dynamics; we did not allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures;
+\item the chosen samples are combined to be played simultaneously and normalized by the number of instruments;
+\item the Mel frequency cepstral coefficients (MFCCs) and the Mel spectrogram of the mixture are computed to be used as the input features for the baseline classifiers and the neural classifiers respectively.
 \end{itemize}
 
-The choice of using the mel spectrogram as input features of the model is very common in music information \cite{Stevens37} and can be considered to be an accurate representation of the way humans perceive sound and music.\carmine{Do we need more references for mel spectrogram?}
+The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly 4 seconds long.
 
-The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly 4 seconds long.
+The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{Stevens37} and can be considered to be an accurate representation of sound and musical signals.
 The Mel spectrogram is generated using an FFT hop length of 2048 samples (e.g.
 the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128.
-Therefore, all the features fed to the model were matrices of size $128\times 345$.
-We used the Librosa package in Python to compute the features, and more details on the exact computations can be found in \cite{mcfee15} The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms could be perceived by the human ear \carmine{Do you have a reference for that?
-I don't find anything on the topic}.
+Therefore, the Mel features fed to the model were matrices of size $128\times 345$.
+We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}.
+The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed.
 
 \begin{table}
+\begin{center}
 \begin{tabular}{|c|c|c|}
@@
 \end{tabular}
-\caption{Table showing the pitch ranges present in the TinySOL database.\label{tinysol}}
+\caption{Table showing the pitch ranges present in the TinySOL database.\label{tab:ranges}}
+\end{center}
 \end{table}
 
-\subsection{Data Augmentation}
+\subsection{Data augmentation}
 
-In order to increase variability in our generated data, we used two methods of data augmentation \cite{Salamon17}\cite{Bhardwaj17} : pitch shifting and partial feature dropout.
+In order to increase variability in our generated data, we also used two methods of data augmentation as described in \cite{Salamon17, hardwaj17}; more specifically, we used pitch shifting and partial feature dropout.
 
-The first one was done on the TinySOL samples.
-We performed a slight pitch shift by shifting the frequency of the signal.
-We did this by reading the samples with different sample rates.
-Librosa automatically converts the data coming from the samples to match the specified sample rate \cite{mcfee15}.
-Thus, a small difference in sample rate will slightly modify the perceived pitch if played at the normal sample rate.
+The former has been directly applied on the TinySOL samples each time the are selected to generate a new combination.
+We performed a small pitch shift by reading the samples with different sample rates: a small difference in sample rate will slightly modify the duration and the perceived pitch if played at the normal sample rate.
 In practice, the sample rates used for this data augmentation were within $5\%$ of the actual $44100$Hz.
 
-The second augmentation was performed on the feature matrix itself, the mel spectrogram.
+The latter was performed on the feature matrix itself of input samples, the Mel spectrogram.
 We chose random columns and rows of the matrix to zero out.
 For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$  columns and $3.45$ rows being zero-ed out.
-This method of data augmentation aimed to be more resilient to the possible errors of recording \carmine{Is there any other reference to justify this sentence?}, due either to a setup that induces interferences and cancels specific frequencies, or to a default in the microphone that fails to record a small temporal portion.
-
-\section{Our model}
-
-Our idea is to perform orchestration using deep neural networks trained for classification.
-In our setting, each class is an instrument-note pair.
-Our model takes in the mel spectrogram of a soundfile that is a combination of different instruments and pitches, and attempts to identify which instrument-pitch pairs are present.
-The network learns how to take a sound that is a complex combination of notes and timbres and deconstruct it into its original parts.
-Then, when a network is given a target sound as input, it will attempt to apply the same classification rules, outputting a vector of the probabilities of each sample being in the target sound.
-By taking the samples that have the highest probability, an orchestrated solution can be created.
-Both of our deep models use the Adam optimizer, Binary Cross-Entropy loss, and were written using PyTorch.
+This method of data augmentation aimed to be more resilient to the possible variations in the recording of the instruments.
 
 \subsection{Baseline}
-In order to have a baseline to compare our results against, we attempted to solve the classification problem using various parametric classifiers.
-The classifiers we tested were Support Vector Machine (SVM), Random Forest, and K-Nearest Neighbors.
-We used the implementations provided in the scikit-learn library for Python \cite{scikit-learn}.
-For SVM, we used SVC with a non-linear RBF kernel.
-For Random Forest, we set the maximum depth of each tree to be 15.
-All of the following baseline experiments used 50,000 generated samples with a train-test split of 60/40.
-Each sample is a combination of one or more instruments and is four seconds in length.
-The features used are the mel frequency cepstral coefficients (MFCCs) of the resulting combination, with a total of 100 coefficients per sample.
-We chose to use MFCCs for this setting because using the mel spectrogram leads to a number of features that is too high for the parametric classifiers.
+
+In order to have a baseline to compare our classification results and to get a sense on the complexity of the problem, we initially used various parametric classifiers.
+The classifiers we tested were support vector machine (SVM), random forest (RF), and K-nearest neighbours (KNN).
+We used the implementations provided in the \emph{scikit-learn }library for Python \cite{scikit-learn}.
+More specifically, for SVM we used a non-linear RBF kernel and for RF we set the maximum depth of each tree to be 15.
+All of the baseline experiments used 50,000 generated samples with a train-test split of 60/40.
+Each sample was made by a combination of one to ten instruments and was four seconds in length.
+In this case, differently from the neural models, the features used are the MFCCs of the resulting combination; we chose to use MFCCs for this setting, instead of the Mel spectrogram, to have a number of features that is more appropriate for the parametric classifiers.
 We found SVM to have the highest accuracy of the three classifiers across all experiments.
 
-We found that the parametric classifiers were not able to solve the full setting of our problem which had 424 classes, so we started by simplifying the problem to classifying only the instruments and not the pitch.
+The complete problem of classification of the pairs instrument-pitch has 424 total classes.
+This problem is very hard for parametric classifiers, so we initially switched to the simpler problem of classifying only the instruments and not the pitch.
 This had the benefit of both reducing the number of classes and increasing the number of samples per class.
 We found that SVM was able to very accurately identify the instrument given an input that had only one instrument present; for this case the accuracy was 99.8\%.
-However as soon as multiple instruments were present in the input, the accuracy dropped significantly.
-With two instruments, accuracy was 55.4\%, with three it was 19.6\% and with 10 instruments the accuracy was 0.5\%.
-K-Nearest Neighbors performed significantly worse than SVM, so we did not attempt any further testing with KNN.
+However, as soon as multiple instruments were present in the input, the accuracy dropped significantly.
+With two instruments, accuracy was 55.4\%, with three it was 19.6\%.
+KNN performed significantly worse than SVM, so we did not attempt any further testing with it.
 
-To better approximate the problem of identifying instrument and pitch, we performed experiments in which two instruments were selected and for input data that contained samples from one of those two instruments, both the instrument and pitch class of the sample would be classified.
-That is, which octave the pitch was in did not matter, only the pitch class.
+To better approximate our original problem of identifying instrument and pitch, however, we performed experiments in which two instruments were selected and for input data that contained samples from one of those two instruments, both the instrument and pitch \emph{class} of the sample would be classified.
+The pitch class is the note name without the octave, e.g.
+C, D\#, G, etc.
 The input was a combination of two instruments drawn from a possible twelve instruments, and the classifier attempted to identify which instruments were present and for the specified instruments, say Violin and Viola, which pitch classes were present.
 If another instrument was present in the input combination that was not Violin or Viola, the classifier would simply identify that an instrument that was not one of the two was present.
-The accuracies from this experiment are outlined in Table 1.
-Since Random Forest performed worse than SVM is every experiment, we stopped testing with Random Forest and used SVM from that point on.
+The accuracies from this experiment are outlined in table \ref{tab:baselines}.
+Since RF performed worse than SVM is every experiment, we stopped testing with it and used SVM from that point on.
 
@@
   \begin{center}
-    \label{tab:table1}
     \begin{tabular}{|c|c|c|c|} 
@@
   \end{center}
-  \caption{Comparison of accuracies between SVM and Random Forest.
+  \caption{Comparison of accuracies between SVM and RF.
 Each data point is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment.
@@
 For a sample not from one of the two instruments, the classifier simply attempted to identify that a sample from one of the non-specified instruments was present.
-In this setting, there are 25 classes: 24 from the 12 pitch classes from 2 instruments, and 1 class that specified whether an instrument that is not one of the two specified is present.}
+In this setting, there are 25 classes: 24 from the 12 pitch classes from 2 instruments, and 1 class that specified whether an instrument that is not one of the two specified is present.
+\label{tab:baselines}}
 \end{table}
+
+In the rest of this section, we will show the classification results of two neural models.
 
@@
 
-The first deep model we tested was a Convolutional Neural Network (CNN) with a Long Short-Term Memory (LSTM) unit.
+The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit.
 CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey16}.
-Long Short-Term Memory (LSTM) units provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.
+LSTM units provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.
 
-\carmine{Do we need any more justification on why we used that model?}
-
-Our architecture contains four convolutional layers and two fully connected layers.
+Our architecture is made of four convolutional layers and two fully connected layers.
 Each convolutional layer is followed by a BatchNorm layer, a ReLU activation layer and a $2 \times 2$  MaxPool layer with a stride of 2.
@@
 
-\carmine{Does this architecture need to be explained in more details?}
-
 Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class. The model architecture is shown in Figure \ref{cnn_lstm}.
@@
 \begin{figure*}
-\includegraphics[scale=0.45]{CNN_LSTM_diagram.png}
+\begin{center}
+\includegraphics[scale=0.4]{CNN_LSTM_diagram.png}
 \caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
+\end{center}
 \end{figure*}
@@
 \subsection{ResNet}
-\carmine{Can you add more to this section about why ResNet is good, and maybe some details about the implementation?}
-The second deep model that we used was the well known deep residual network ResNet.
 
-To make our setting more reasonable, we took the well-known deep residual network (ResNet) as backbone in our experiments.
+The second and more deep model that we trained as classifier was the well known deep residual network \emph{ResNet}.
 Specifically, we used 18-layer ResNet, which allows information to pass directly from the input to the final layer of each block.
-Besides, to make the model more suitable to our problem, we reduce the number of parameters and reset the output channel numbers of each block with 32, 64, 32, 32 respectively.
+Besides, to make the model more suitable to our problem, we decided to reduce the number of parameters and to reset the output channel numbers of each block with 32, 64, 32, 32 respectively.
 
@@
 
-During training, the loss function used to optimize the inner parameters of the model was binary cross entropy, as it is the common choice for multiclass multilabel classification frameworks such as this one \carmine{Is there a specific paper justifying that claim?
-Or is it obvious enough?}.
+During training, the loss function used to optimize the inner parameters of the model was binary cross entropy, as it is the common choice for multiclass multilabel classification frameworks.
 However, the value of the loss function alone is difficult to interpret.
@@
 Different experiments were made by varying the number $N$ of samples in each mixture sample.
-We used an orchestra of 10 instruments containing Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet C and Clarinet Bb.
+We used an orchestra of 10 instruments containing Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet in C and Clarinet in Bb.
 
@@
 
-\begin{figure}
-\center{
-\includegraphics[scale=0.4]{figs/CNN_vs_ResNet.png}}
-\caption{Best overall accuracy for CNN with LSTM and ResNet depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
-\end{figure}
-
-From this plot, we see that ResNet achieves better performances than the CNN regardless of the number of samples used in the combination.
-This result isn't surprising as residual networks usually perform better for this kind of classification problems \carmine{It would need a citation on why ResNet was expected to be better}
+ResNet outperforms the CNN regardless of the number of samples used in the combination.
+This result is consistent with previous research \cite{He15}, as residual networks usually perform well in classification problems.
 
@@
 For ResNet, the variance in accuracy is much smaller until $N$ reaches 5, at which point it becomes similar to the CNN.
-The results on both figures show consistency on the relative accuracy of instruments, which is a first step towards the validation of this method \carmine{Can I say that?}.
+The results on both figures show consistency on the relative accuracy of instruments, which was for us the first step towards the validation of this method.
 Flute, Trombone and Trumpet yield the worst results for both models.
-Their low accuracy could be explained by the nature of their spectra.
-For the Flute, since it has very few harmonics, it is difficult to identify it among other instruments, especially if they are richer.
-On the other hand, brasses have a very rich spectra and a strong noisy component that excites almost all the frequencies.
+
+While it is not easy to explain these differences in accuracy, we hypothesize this being related to the nature of peculiar spectral and temporal morphology of each instrument.
+For example, flute notes tend to exhibit a steep spectral rolloff, with most of the energy captured by the first few partials.
+Moreover, the noisy nature of the transient portions of these note is not well represented by frequency-based descriptions such as Mel spectra.
+These two factors combined, could make the disentanglement of the flute from the analyzed combination more difficult.
 
@@
 An interesting point to notice is the very high accuracy of Oboe on both models.
-This could indicate that there is an optimal balance of harmonies that maximizes the probability of being detected in such classification framework.\carmine{Please review those interpretations}
-
-\begin{figure}
-\includegraphics[scale=0.5]{figs/Acc_vs_N_CNN.png}
-\caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM depending on the number of instruments in the combinations. \label{best_acc_cnn}}
-\end{figure}
+This could indicate that there is an optimal spectral shape that maximizes the probability of being detected in such classification framework.
 
@@
 
-\iffalse
-\begin{table*}
-\center{
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
-\hline
-Instruments & \multicolumn{2}{c|}{$N=4$} & \multicolumn{2}{c|}{$N=5$}& \multicolumn{2}{c|}{$N=6$}& \multicolumn{2}{c|}{$N=7$}& \multicolumn{2}{c|}{$N=8$}& \multicolumn{2}{c|}{$N=9$}& \multicolumn{2}{c|}{$N=10$}\\
-\cline{2-3} & C & R & C & R & C & R & C & R & C & R & C & R & C & R\\
-\hline
-All & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
-Hn & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
-Ob & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
-Vn & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
-Va & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
-Vc & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
-Fl & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
-Tbn & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
-Bn & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
-TpC & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
-ClBb & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
-\hline
-\end{tabular}
+\section{Orchestration experiments}
+\label{sec:orchestration}
 
-\begin{caption}
-Best test accuracy obtained for each model and value of $N$ over the total number of epochs.
-The accuracy is computed using $f$ defined in \eqref{eval}.
-The first row correspond to the overall accuracy of the model, regrouping all the instruments.
-C is CNN with LSTM, R is ResNet.
-\label{class_res}
-\end{caption}}
-\end{table*}
-\fi
+After training the neural models for classification, we finally tested them for the task of target-based computer-assisted orchestration.
+For this purpose, we decided to use the models trained on combinations of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet.
+The total number of classes in this setting is 424, each one representing an instrument-note pair.
 
-\section{Orchestration experiments}
-
-For orchestrating, we trained our models on combinations of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet.
-This leads to 424 classes, each one representing an instrument-note pair.
-A target sound is input to the model, and the 10 classes with the highest probability are extracted.
+To orchestrate, a target sound is input to the model, and the 10 classes with the highest probability are extracted.
 These 10 classes are the instrument-note pairs that are most represented in the target, and can be from any combination of the 10 instruments.
 
-We did not train our model to classify the dynamics of a sample despite TinySOL having pianissimo, mezzo forte, and fortissimo recordings for each sample.
-If we had, then each class would have only one data point: the TinySOL sample it represents.
-When the model is used for orchestration, the dynamic is determined by the probability of that sample as output by the model.
-If the model outputs a probability higher than $0.66$ for a sample, the fortissimo version of the sample was used.
-If it was between $0.33$ and $0.66$, then the mezzo forte version, and if less than $0.33$ the pianissimo dynamic.
-The idea behind this is that samples that are the most represented in the target should be the loudest in the orchestrated solution.
+Since we decided not to train our models to classify the dynamics of a sample (despite TinySOL having \emph{pianissimo}, \emph{mezzoforte}, and \emph{fortissimo} recordings), the dynamics is determined by the probability of each sample as output by the model.
+If the model outputs a probability higher than $0.66$ for a sample, the fortissimo version of the sample is used.
+If the probability is between $0.33$ and $0.66$, then the mezzoforte version is selected and if it is less than $0.33$ the pianissimo version is used.
+The idea behind this quantization, is that samples that are the most represented in the target should appear as the loudest in the orchestrated solution.
 
-We tested our models using 15 targets for orchestration.
-Two of the targets were made from TinySOL samples, but the rest are not combinations of input data, and some are not even recordings of instruments.
-Among the targets are recordings of bells, a car horn, a gong, and a recording of a boat docking.
+In order to test our models for orchestration, we used 15 targets from the Orchidea distribution.
+These targets represent a variety of signal types but are mostly static, in the sense that they do not change sensibly over time.
 
-Although the data we trained the model on was fixed to four seconds of audio, the target samples can be of arbitrary length.
-By changing the mel hop length, which is the distance between the frames of the melspectrogram, an audio signal of any length can be represented as a matrix of the same size as the training data.
+Some of the targets were made of instrumental samples: a sum of oboe and bassoon notes, single bassoon and bass clarinet notes, two multiphonics of basson and two symphonic chords.
+The other targets there are: the recordings of two bells, of a car horn, of a gong, of a screaming voice, of a wind harp, of the attack of a brass instrument and the recording of a boat docking.
+
+Although the data we trained the model on was fixed to four seconds of audio, the target samples can be of arbitrary length since we adapt the representation to the size of the training data by changing the hop length during the computation of the Mel spectrogram.
 
@@
 
-We evaluated our orchestrations both qualitatively and quantitatively.
-We compared our solutions to OrchIdea, the state-of-the-art program for computer-assisted orchestration.
-In order to have comparable solutions, we did not allow OrchIdea to create sparse solutions, meaning that OrchIdea was forced to use all 10 instruments in each solution.
+We evaluated our orchestrations both qualitatively and quantitatively by comparing our solutions to Orchidea's solutions, the state-of-the-art program for computer-assisted orchestration.
+In order to have a more fair comparison, we did not allow Orchidea to create sparse solutions, meaning that OrchIdea was forced to use all 10 instruments in each solution.
 This creates a more fair comparison since our model is unable to create sparse solutions.
@@
     \label{orch_eval}
-    \begin{tabular}{|c|c|c|c|c|c|c|c|c|} 
+    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c}
      \hline
-      \textbf{Model} & \textbf{Ob + Bn} & \textbf{Bn} & \textbf{Bass Clarinet} & \textbf{Bell1} & \textbf{Bell2} & \textbf{Yan} & \textbf{Car Horn} & \textbf{Boat}  \\
+      \textbf{Model} & \textbf{Ob + Bn} & \textbf{Bn} & \textbf{Bass cl.} & \textbf{Bell 1} & \textbf{Bell 2} & \textbf{Multiph. 1} & \textbf{Car horn} & \textbf{Boat} $\ldots$ \\
       \hline
-      CNN with LSTM & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+      CNN with LSTM & 0.17 & 0.28 & 0.70 & 0.55 & 0.26 & \textbf{1.10} & 0.68 & \textbf{1.12} $\ldots$  \\
       \hline
-      ResNet & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-      \hline
-      OrchIdea & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
+      ResNet & 0.34 & 0.50 & 0.48 & 0.59 & 0.45 & 0.90 & 0.49 & \textbf{1.16} $\ldots$ \\
       \hline
       \hline
-      \textbf{Model} & \textbf{Wind Harp} & \textbf{Chord1} & \textbf{Reed Bone} & \textbf{Chord2} & \textbf{Gong} & \textbf{Scream} & \textbf{Trompe Attack} & \textbf{Average}  \\
+      \textbf{$\ldots$} & \textbf{Wind harp} & \textbf{Chord 1} & \textbf{Multiph. 2} & \textbf{Chord 2} & \textbf{Gong} & \textbf{Scream} & \textbf{Brass} & \textbf{Average}  \\
       \hline
-      CNN with LSTM & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+      $\ldots$ & 0.55 & 0.79 & 0.70 & 0.57 & 0.73 & \textbf{1.14} & 0.79 & 0.71\\
       \hline
-      ResNet & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
-      \hline
-      OrchIdea & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+      $\ldots$  & 0.61 & 0.86 & 0.51 & 0.37 & 0.71 & \textbf{1.03} & \textbf{1.05} & 0.66 \\
       \hline
@@
   \end{center}
-  \caption{Quantitative comparison of orchestrations.}
+  \caption{Quantitative comparison of orchestrations as ratios to Orchidea.
+Eqn.
+\eqref{distance} was used with $\lambda_1 = 0.5$ and $\lambda_2 = 10$ to compute distances between orchestrations and targets.
+What is shown is the ratio between the distance calculated and Orchidea's distance for the same target.
+A value less than 1 means that the model performed worse (i.e.
+had a larger distance), and a value greater than 1 means the model performed better than Orchidea.
+The last column shows the ratio of the average distances for the model across all targets.}
 \end{table*}
 
-\section{Conclusion}
+\section{Conclusions}
+\label{sec:conclusions}
 
@@
 Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.
-
-\subsection{Interpreting the Latent Space}
-
-After training the CNN with LSTM for $N=10$ samples per combination, it is possible to visualize the filters applied to the input layer after layer.
-Figs.
-\ref{latent0}-\ref{latent5} show the successive outputs of each layer of the convolutional network.
-\carmine{I put here all the intermediary features of the CNN with LSTM.
-I don't find any particular conclusion other than saying that LSTM seems to focus more on vertical components rather than horizontal, ie time over frequency}
-
-\begin{figure}
-\includegraphics[scale=0.5]{figs/latent_space_layer0.png}
-\caption{Initial feature matrix used as input of the model \label{latent0}}
-\end{figure}
-
-\begin{figure}
-\includegraphics[scale=0.5]{figs/latent_space_layer1.png}
-\caption{Intermediary features after the first convolutional layer of the CNN with LSTM. \label{latent1}}
-\end{figure}
-
-\begin{figure}
-\includegraphics[scale=0.5]{figs/latent_space_layer2.png}
-\caption{Intermediary features after the second convolutional layer of the CNN with LSTM. \label{latent2}}
-\end{figure}
-
-\begin{figure}
-\includegraphics[scale=0.5]{figs/latent_space_layer3.png}
-\caption{Intermediary features after the third convolutional layer of the CNN with LSTM. \label{latent3}}
-\end{figure}
-
-\begin{figure}
-\includegraphics[scale=0.5]{figs/latent_space_layer4.png}
-\caption{Intermediary features after the LSTM layer. \label{latent4}}
-\end{figure}
-
-\begin{figure}
-\includegraphics[scale=0.5]{figs/latent_space_layer5.png}
-\caption{Intermediary features after final convolutional layer of the CNN with LSTM. \label{latent5}}
-\end{figure}
-
-\section{Future steps}
 
@@
 
+%\subsection{Interpreting the Latent Space}
+%
+%After training the CNN with LSTM for $N=10$ samples per combination, it is possible to visualize the filters applied to the input layer after layer.
+Figs.
+\ref{latent0}-\ref{latent5} show the successive outputs of each layer of the convolutional network.
+\carmine{I put here all the intermediary features of the CNN with LSTM.
+I don't find any particular conclusion other than saying that LSTM seems to focus more on vertical components rather than horizontal, ie time over frequency}
+%
+%\begin{figure}
+%\includegraphics[scale=0.5]{figs/latent_space_layer0.png}
+%\caption{Initial feature matrix used as input of the model \label{latent0}}
+%\end{figure}
+%
+%\begin{figure}
+%\includegraphics[scale=0.5]{figs/latent_space_layer1.png}
+%\caption{Intermediary features after the first convolutional layer of the CNN with LSTM. \label{latent1}}
+%\end{figure}
+%
+%\begin{figure}
+%\includegraphics[scale=0.5]{figs/latent_space_layer2.png}
+%\caption{Intermediary features after the second convolutional layer of the CNN with LSTM. \label{latent2}}
+%\end{figure}
+%
+%\begin{figure}
+%\includegraphics[scale=0.5]{figs/latent_space_layer3.png}
+%\caption{Intermediary features after the third convolutional layer of the CNN with LSTM. \label{latent3}}
+%\end{figure}
+%
+%\begin{figure}
+%\includegraphics[scale=0.5]{figs/latent_space_layer4.png}
+%\caption{Intermediary features after the LSTM layer. \label{latent4}}
+%\end{figure}
+%
+%\begin{figure}
+%\includegraphics[scale=0.5]{figs/latent_space_layer5.png}
+%\caption{Intermediary features after final convolutional layer of the CNN with LSTM. \label{latent5}}
+%\end{figure}
+
 \iffalse
@@
 % For bibtex users:
-\bibliography{paper}
+%\bibliography{paper}
 
-% For non bibtex users:
-%\begin{thebibliography}{citations}
-%
-%\bibitem {Author:00}
-%E. Author.
-%``The Title of the Conference Paper,''
-%{\it Proceedings of the International Symposium
-%on Music Information Retrieval}, pp.~000--111, 2000.
-%
-%\bibitem{Someone:10}
-%A. Someone, B. Someone, and C. Someone.
-%``The Title of the Journal Paper,''
-%{\it Journal of New Music Research},
-%Vol.~A, No.~B, pp.~111--222, 2010.
-%
-%\bibitem{Someone:04} X. Someone and Y. Someone. {\it Title of the Book},
-%    Editorial Acme, Porto, 2012.
-%
-%\end{thebibliography}
+\begin{thebibliography}{99}
+
+\bibitem{Benetos07}
+Emmanouil Benetos, Margarita Kotti, and Constantine Kotropoulos.
+\newblock Large scale musical instrument identification.
+\newblock 2007.
+
+\bibitem{Bhardwaj17}
+Siddharth Bhardwaj.
+\newblock Audio data augmentation with respect to musical instrument
+  recognition.
+\newblock Master's thesis, 2017.
+
+\bibitem{Bian19}
+Wenhao Bian, Jie Wang, Bojin Zhuang, Shaojun~Wang Jiankui~Yang, and Jing Xiao.
+\newblock Audio-based music classification with densenet and data augmentation.
+\newblock ArXiv (https://arxiv.org/abs/1906.11620v1), 2019.
+
+\bibitem{Cella18}
+Carmine~Emanuele Cella and Philippe Esling.
+\newblock Open-Source Modular Toolbox for Computer-Aided Orchestration.
+\newblock Timbre conference, 2018.
+
+\bibitem{Cella19}
+Carmine~Emanuele Cella, Jon Gillick, and David Bamman.
+\newblock Estimating unobserved audio features for targed-based orchestration.
+\newblock ISMIR, 2019.
+
+\bibitem{Cella2020}
+Carmine~Emanuele Cella.
+\newblock Orchidea: a comprehensive framework for target-based assisted orchestration.
+\newblock in preparation; available on request from the author, 2020.
+
+\bibitem{Cella2020b}
+Carmine~Emanuele Cella, Daniele Ghisi, Vincent Lostanlen, Fabi\'en Levy, Yan Maresz and Joshua Fineberg.
+\newblock OrchideaSOL: a dataset of extended instrumental techniques for computer-aided orchestration.
+\newblock ICMC (under review), 2020.
+
+\bibitem{Chetry05}
+Nicolas Chetry, Mike Davies, and Mark Sandler.
+\newblock Musical instrument identification using lsf and k-means.
+\newblock 2005.
+
+\bibitem{He15}
+Kaiming He, Xiangyu Zhang, Shaoqing Ren, and Jian Sun.
+\newblock Deep residual learning for image recognition.
+\newblock 2015.
+
+\bibitem{Hershey16}
+Shawn Hershey, Sourish Chaudhuri, Daniel P.~W. Ellis, Jort~F. Gemmeke, Aren
+  Jansen, R.~Channing Moore, Manoj Plakal, Devin Platt, Rif~A. Saurous, Bryan
+  Seybold, Malcolm Slaney, Ron~J. Weiss, and Kevin Wilson.
+\newblock Cnn architectures for large-scale audio classification.
+\newblock 2016.
+
+\bibitem{Hochreiter97}
+Sepp Hochreiter and J\"{u}rgen Schmidhuber.
+\newblock Long short-term memory.
+\newblock {\em Neural Comput.}, 9(8):1735–1780, November 1997.
+
+\bibitem{Kitahara05}
+Tetsuro Kitahara, Masataka Goto, and Hiroshi~G. Okuno.
+\newblock Pitch-dependent identification of musical instrument sounds.
+\newblock 2005.
+
+\bibitem{lostanlen16}
+Vincent Lostanlen and Carmine-Emanuele Cella.
+\newblock Deep convolutional networks on the pitch spiral for musical
+  instrument recognition, 2016.
+
+\bibitem{mcfee15}
+Brian McFee, Colin Raffel, Dawen Liang, Daniel~P.W. Ellis, Matt McVicar, Eric
+  Battenberg, and Oriol Nieto.
+\newblock librosa: Audio and music signal analysis in python.
+\newblock 2015.
+
+\bibitem{scikit-learn}
+F.~Pedregosa, G.~Varoquaux, A.~Gramfort, V.~Michel, B.~Thirion, O.~Grisel,
+  M.~Blondel, P.~Prettenhofer, R.~Weiss, V.~Dubourg, J.~Vanderplas, A.~Passos,
+  D.~Cournapeau, M.~Brucher, M.~Perrot, and E.~Duchesnay.
+\newblock Scikit-learn: Machine learning in {P}ython.
+\newblock {\em Journal of Machine Learning Research}, 12:2825--2830, 2011.
+
+\bibitem{Salamon17}
+Justin Salamon and Juan~Pablo Bello.
+\newblock Deep convolutional neural networks and data augmentation for
+  environmental sound classification.
+\newblock {\em IEEE Signal Processing Letters}, 24(3):279–283, Mar 2017.
+
+\bibitem{Stevens37}
+S.~S. Stevens, J.~Volkmann, , and E.~B. Newman.
+\newblock A scale for the measurement of the psychological magnitude pitch.
+\newblock 1937.
+
+\end{thebibliography}
 

### 2020-05-15 paper/latex/paper.tex

@@
 \begin{center}
-\includegraphics[scale=0.4]{CNN_LSTM_diagram.png}
+\includegraphics[scale=0.4]{figs/CNN_LSTM_diagram.png}
 \caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
@@
 
-Some of the targets were made of instrumental samples: a sum of oboe and bassoon notes, single bassoon and bass clarinet notes, two multiphonics of basson and two symphonic chords.
+Some of the targets were made of instrumental samples: a sum of oboe and bassoon notes, single bassoon and bass clarinet notes, two multiphonics of bassoon and two symphonic chords.
 The other targets there are: the recordings of two bells, of a car horn, of a gong, of a screaming voice, of a wind harp, of the attack of a brass instrument and the recording of a boat docking.

### 2020-05-16 paper/latex/paper.tex
final version before proof read

@@
 \newcommand{\carmine}[1]{\textbf{\color{red} Carmine => #1}}
+\renewcommand{\baselinestretch}{.98}
 
@@
 \begin{abstract}
-In this paper we will explore preliminary neural models for the task of target-based computer-assisted musical orchestration.
-After an introduction on computer-assisted orchestration, we will show how we model this  musical problem as a classification task.
-In this context, we will propose two deep learning models and we will show, first, how they perform as classifiers for musical instruments recognition by comparing them with specific baselines.
+In this paper we will do a preliminary exploration on how neural networks can be used for the task of target-based computer-assisted musical orchestration.
+We will show how it possible to model this  musical problem as a classification task and we will propose two deep learning models.
+We will show, first, how they perform as classifiers for musical instruments recognition by comparing them with specific baselines.
 We will then show how they perform, both qualitative and quantitative, in the task of computer-assisted orchestration by comparing them with state-of-the-art systems.
-We will highlight, finally, benefits and problems of neural approaches for assisted orchestration and we will propose possible future steps.
+Finally, we will highlight benefits and problems of neural approaches for assisted orchestration and we will propose possible future steps.
 \end{abstract}
@@
 
-Computer-assisted orchestration is an important problem ... \carmine{Can you add more?}.
+The development of computational tools to assist and inspire the musical composition process constitutes an import research area known as \emph{Computer-Assisted Composition (CAC)} \cite{FerVic2013, Ari2005}.
+Within CAC, target-based computer-assisted orchestration is an important example how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.
 
-The approach studied in \cite{Cella18, Cella2020} consists in finding a good orchestration for any given sound by searching combinations of sounds from a database with a multi-objective optimization heuristics and a constraint solver that are jointly optimized.
-Both the target sound and the sounds in the database are embedded in a feature space defined by a fixed feature function $f$ and each generated combination of sounds is evaluated by using a specific metric.
-This method is currently considered state of the art and is implemented in the \emph{Orchidea} toolbox for assisted orchestration (\url{www.orch-idea.org}); we will describe this approach in more detail in the next sections.
+Target-based computer-assisted orchestration can be thought of as the process of searching for the best combinations of sounds in a database of instrumental samples, to match a target sound under a specific similarity metric and specific sets of constraints.
+A solution to this problem is a set of orchestral scores that represent the mixtures of audio samples in the database, ranked by similarity with the target sound.
+Valid orchestral scores may contain several instruments sounding simultaneously, selected out of a large number of possible combinations of sounds from the database.
 
-In this paper, we try a different approach by experimenting with deep neural architectures.
+\begin{figure}
+\includegraphics[scale=0.17]{figs/method.png}
+\caption{An overview of the proposed method for assisted orchestration with neural models.
+Instruments and pitches are determined as peaks of the output probability distribution, while the dynamics are computed by quantizing the probabilities.
+\label{fig:method}}
+\end{figure}
+
+The approach studied in \cite{Carpentier2010} consists in finding a good orchestration for any given sound by searching combinations of sounds from a database with a multi-objective optimization heuristics and a constraint solver that are jointly optimized.
+Both the target sound and the sounds in the database are embedded in a feature space defined by a fixed feature function and each generated combination of sounds is evaluated by using a specific metric.
+This method has been substantially improved in \cite{Cella18, Cella2020} and is implemented in the \emph{Orchidea} toolbox for assisted orchestration (\url{www.orch-idea.org}), currently considered the state-of-the-art system for assisted orchestration.
+
+In this paper, we try a different approach to this problem by experimenting with deep neural architectures.
 The main idea is to train a model to classify combinations of real instruments and use it then for orchestration.
@@
 By training a neural network with real combinations of instrumental notes, it will acquire the ability to identify the presence of each instrument and its associated pitch by building the appropriate latent representation.
-Thus, when an unknown target sound is given as input, the network will identify which are the best instruments-pitches pairs to match the target sound, and it will be able to  deconstruct a complex mixture of timbres into individual instrument-note pairs\footnote{The dynamics is retrived using the probability distributions generated as output by the network; more details on this will be given in the next sections.}.
+Thus, when an unknown target sound is given as input, the network will identify which are the best instruments to match the target sound, and it will be able to  deconstruct a complex mixture of timbres into individual instrument notes.
+
+%\footnote{The dynamics is retrived using the probability distributions generated as output by the network; more details on this will be given in the next sections.}.
 
@@
 In this paper we do preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit in the middle and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}.
-We chose to use a CNN because of its success in audio classification \cite{Hershey16} and we decided to include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.
+We chose to use a CNN because of its success in audio classification \cite{Hershey17} and we decided to include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.
 
@@
 We then compared our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means.
-For quantitative comparison, we computed the distance metric suggested in \cite{Cella2020} on the FFTs of the target and solution, comparing the distance between Orchidea's solution and the target and our model's solution and the target.
+For quantitative comparison, we computed a specific distance metric on the FFTs of the target and solution, comparing the distance between Orchidea's solution and the target and our model's solution and the target.
 
 Next sections will be as follows: section \ref{sec:models} will describe the dataset we used for experiments and the models we trained as classifiers, showing the classification results.
-Sections \ref{sec:orchestration} and \ref{sec:conclusions} will show, respectively, the results of orchestration experiments using the trained models and will present the conclusions of our preliminary study on neural approaches for target-based computer-assisted orchestration.
+Sections \ref{sec:orchestration} and \ref{sec:conclusions} will present, respectively, the results of orchestration experiments using the trained models and the conclusions of our preliminary study on neural approaches for target-based computer-assisted orchestration.
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.22]{figs/CNN_LSTM_diagram.png}
+\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
+\end{center}
+\end{figure}
+
 \subsection{Form orchestration to classification}
@@
 \begin{enumerate}
-\item we train specific models to classify instrument-pitch pairs on sound combinations of a database of instrument notes, up to ten simultaneous instruments;
+\item we train specific models to classify the instruments present in combinations of sounds from a database of instrument notes, up to ten simultaneous instruments;
 \item we then pick the best classifier and we feed into it an unknown sound to be classified;
-\item since the output of the classifier will be made of instrument-pitch pairs, we use this information to synthesize an orchestration for the target sound;
+\item since the output of the classifier will be in the form of the probability that specific instruments are present in the sound, we use this information to synthesize an orchestration for the target sound;
 \item finally, we evaluate the generated orchestration against state-of-the-art systems for computer-assisted orchestrations.
 \end{enumerate}
+
+\begin{figure}
+\center{
+\includegraphics[scale=0.4]{figs/CNN_vs_ResNet.png}}
+\caption{Best overall accuracy for CNN with LSTM and ResNet depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
+\end{figure}
 
@@
 A complete orchestration solution, however, would normally be made by triples of instrument-pitch-dynamics.
-This is difficult, however, since to frame this problem as classification, we would need to have a very high number of classes.
+It is difficult to frame this problem as classification, since we would need to have a very high number of classes.
 Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in \ref{sec:dataset}), each class would be represented by a single sample.
-For these reasons, our models have been trained on simplified versions of the original problem.
+For these reasons, our models have been trained on simplified versions of the original problem and we used different strategies to provide the remaining information needed to generate orchestrations.
 
-On one hand, the baseline classifiers have been trained to recognize either the instrument only for a total of 12 classes or the pair instrument-pitch \emph{class} (a pitch class is the pitch of a note without the corresponding octave, e.g.
+In order to have a baseline to compare our classification results and to get a sense on the complexity of the problem, we used various parametric classifiers along with the neural models.
+
+Three baseline classifiers have been trained to recognize either the instrument only for a total of 12 classes or the pair instrument-pitch \emph{class} (a pitch class is the pitch of a note without the corresponding octave, e.g.
 C, D\#, G, etc.) for a total of 25 classes.
 
-On the other hand, for the neural models, we reframed the classification problem by using full pitch and trained them to recognize the pair instruments-pitch for a total of 424 classes (i.e.
+Two neural classifiers, instead, have been trained on the more complex problem of recognizing instrument-pitch pairs, for a total of 424 classes (i.e.
 the neural classifiers would classify instruments and pitches with respective octaves).
@@
 
-\begin{figure}
-\includegraphics[scale=0.5]{figs/Acc_vs_N_CNN.png}
-\caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM depending on the number of instruments in the combinations. \label{best_acc_cnn}}
-\end{figure}
-
-\begin{figure}
-\center{
-\includegraphics[scale=0.5]{figs/CNN_vs_ResNet.png}}
-\caption{Best overall accuracy for CNN with LSTM and ResNet depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
-\end{figure}
-
 \subsection{Dataset}
@@
 The instruments come from different orchestral families: strings, woodwinds, and brass.
-Each sample is one instrument playing a single note in the \emph{ordinario} playing style, with one of three dynamics: \textit{pp}, \textit{mf}, or \textit{ff}, for example \texttt{Flute-C4-pp} or \texttt{Clarinet-D5-mf}.
-The instruments and ranges over which they were recorded is summarized in table \ref{tab:ranges}.
+Each sample is one instrument playing a single note in the \emph{ordinario} playing style, with one of three dynamics: \textit{pp}, \textit{mf}, or \textit{ff} (for example \texttt{Flute-C4-pp} or \texttt{Clarinet-D5-mf}).
+
+%The instruments and ranges over which they were recorded are summarized in table \ref{tab:ranges}.
 
@@
 
-The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{Stevens37} and can be considered to be an accurate representation of sound and musical signals.
+The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an accurate representation of sound and musical signals.
 The Mel spectrogram is generated using an FFT hop length of 2048 samples (e.g.
@@
 
-\begin{table}
-\begin{center}
-\begin{tabular}{|c|c|c|}
-\hline
-Instrument & Abbreviation & Range \\
-\hline
-Violin & Vn & G3-E7\\
-Cello & Vc & C2-C6\\
-Viola & Va & C3-C7\\
-Trumpet C & TpC & F\#3-C6\\
-Trombone & Tbn & A\#0;A\#1-C5\\
-Oboe & Ob & A\#3-A6\\
-Horn & Hn & G1-F5\\
-Flute & Fl & B3-D7\\
-Clarinet Bb & ClBb & D3-G6\\
-Contrabass & Cb & E1-C5\\
-Bass Tuba & BTb & F\#1-F4\\
-Bassoon & Bn & A\#1-D\#5\\
-\hline
-\end{tabular}
-\caption{Table showing the pitch ranges present in the TinySOL database.\label{tab:ranges}}
-\end{center}
-\end{table}
+%\begin{table}
+%\begin{center}
+%\begin{tabular}{|c|c|c|}
+%\hline
+%\textbf{Instrument} & \textbf{Abbreviation} & \textbf{Range} \\
+%\hline
+%Violin & Vn & G3-E7\\
+%Cello & Vc & C2-C6\\
+%Viola & Va & C3-C7\\
+%Trumpet C & TpC & F\#3-C6\\
+%Trombone & Tbn & A\#0;A\#1-C5\\
+%Oboe & Ob & A\#3-A6\\
+%Horn & Hn & G1-F5\\
+%Flute & Fl & B3-D7\\
+%Clarinet Bb & ClBb & D3-G6\\
+%Contrabass & Cb & E1-C5\\
+%Bass Tuba & BTb & F\#1-F4\\
+%Bassoon & Bn & A\#1-D\#5\\
+%\hline
+%\end{tabular}
+%\caption{Table showing the pitch ranges present in the TinySOL database.\label{tab:ranges}}
+%\end{center}
+%\end{table}
 
@@
 
-In order to increase variability in our generated data, we also used two methods of data augmentation as described in \cite{Salamon17, hardwaj17}; more specifically, we used pitch shifting and partial feature dropout.
+In order to increase variability in the generated data for the neural models, we also used two methods of data augmentation as described in \cite{Salamon17, Bhardwaj17}; more specifically, we used pitch shifting and partial feature dropout.
 
-The former has been directly applied on the TinySOL samples each time the are selected to generate a new combination.
+The former has been directly applied on the TinySOL samples each time they are selected to generate a new combination.
 We performed a small pitch shift by reading the samples with different sample rates: a small difference in sample rate will slightly modify the duration and the perceived pitch if played at the normal sample rate.
@@
 \subsection{Baseline}
-
-In order to have a baseline to compare our classification results and to get a sense on the complexity of the problem, we initially used various parametric classifiers.
-The classifiers we tested were support vector machine (SVM), random forest (RF), and K-nearest neighbours (KNN).
+The baseline classifiers that we tested are support vector machine (SVM), random forest (RF), and K-nearest neighbours (KNN).
 We used the implementations provided in the \emph{scikit-learn }library for Python \cite{scikit-learn}.
@@
 However, the parametric classifiers were unable to achieve accurate results as the number of instruments increased.
-Therefore, we did not attempt the full setting of the problem, in which individual pitches are classified for all instruments, with parametric classifiers.
+Therefore, we did not attempt, with parametric classifiers, the full setting of the problem in which individual pitches are classified for all instruments.
 
@@
 
-In the rest of this section, we will show the classification results of two neural models.
+In the rest of this section, we will show the classification results of the two neural models.
 
@@
 
-\begin{figure*}
-\begin{center}
-\includegraphics[scale=0.4]{figs/CNN_LSTM_diagram.png}
-\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
-\end{center}
-\end{figure*}
-
 \subsection{ResNet}
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.4]{figs/Acc_vs_N_CNN.png}
+\caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM depending on the number of instruments in the combinations. \label{best_acc_cnn}}
+\end{center}
+\end{figure}
+
 During training, the loss function used to optimize the inner parameters of the model was binary cross entropy, as it is the common choice for multiclass multilabel classification frameworks.
 However, the value of the loss function alone is difficult to interpret.
-For this reason we created a complementary function $f$ to be used only for evaluation.
+For this reason we created a complementary function $f$ to be used for evaluation only.
 This function compares a vector of expected outputs $\overline{X}$ with the estimated output from the model $\hat{X}$ by using the following function
@@
 \begin{figure}
-\includegraphics[scale=0.5]{figs/Acc_vs_N_ResNet.png}
+\begin{center}
+\includegraphics[scale=0.40]{figs/Acc_vs_N_ResNet.png}
 \caption{Best overall accuracy and for each instrument obtained by the ResNet depending on the number of instruments in the combinations. \label{best_acc_resnet}}
+\end{center}
 \end{figure}
@@
 For this purpose, we decided to use the models trained on combinations of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet.
-The total number of classes in this setting is 424, each one representing an instrument-note pair.
+The total number of classes in this setting is 424, each one representing an instrument-pitch pair.
 
 To orchestrate, a target sound is input to the model, and the 10 classes with the highest probability are extracted.
-These 10 classes are the instrument-note pairs that are most represented in the target, and can be from any combination of the 10 instruments.
+These 10 classes are the instrument-pitch pairs that are most represented in the target, and can be from any combination of the 10 instruments.
 
@@
 Although the data we trained the model on was fixed to four seconds of audio, the target samples can be of arbitrary length since we adapt the representation to the size of the training data by changing the hop length during the computation of the Mel spectrogram.
-
-\subsection{Evaluation}
-
-We evaluated our orchestrations both qualitatively and quantitatively by comparing our solutions to Orchidea's solutions, the state-of-the-art program for computer-assisted orchestration.
-In order to have a more fair comparison, we did not allow Orchidea to create sparse solutions, meaning that OrchIdea was forced to use all 10 instruments in each solution.
-This creates a more fair comparison since our model is unable to create sparse solutions.
-
-Qualitative evaluation was done through an acoustic inspection of the solution, paying close attention to timbre and pitch.
-For targets that had harmonic content, it was noted if the partials present in the target were also represented in the orchestrated solution.
-For example, one of the samples of a bell had partials that represented a C\# minor chord, so we checked if the orchestration contained the notes of the chord.
-If a target had specific notes, we identified if the note or its partials were present.
-For example there is target of an Oboe playing an A4 and a Bassoon playing a C\#3, the solution from ResNet contained the partials of the Bassoon's note: C\#3, G\#4, C\#5, and E6 were all included on various instruments in the solution.
-
-For quantitative evaluation, we used the distance metric defined in Eqn.
-\eqref{distance} to calculate differences in timbre between targets and solutions.
-The equation takes in the full FFT of the target and full FFT of the solution.
-Then for each bin of the FFT, it calculates the absolute difference between the values.
-The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.
-
-In the first summation, $\lambda_1$ is multiplied by all the distances calculated when there was more energy in the target than the solution, since $\delta_{k1} = 1$ only when $x_k \ge \tilde{x}_k$.
-Similarly in the second summation, $\lambda_2$ is multiplied by all the distances when the solution provided more energy to a frequency than the target; it overshot the frequency.
-Therefore, the relation between $\lambda_1$ and $\lambda_2$ determines whether a solution is penalized more for undershooting or overshooting the target.
-
-We calculated the distance between the target samples and our orchestrated solutions.
-We then orchestrated the same targets with OrchIdea and calculated the distance for OrchIdea's solutions.
-A comparison of these results is in Tab.
-\ref{orch_eval}.
-
-\begin{equation}\label{distance}
-d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k| \\
-\end{equation}
-
-\begin{equation}
-\delta_{k1} = 
-\begin{cases}
-1, \text{if   } x_k \ge \tilde{x}_k \\
-0, \text{otherwise}
-\end{cases} 
-\end{equation}
-
-\begin{equation}
-\delta_{k2} = 
-\begin{cases}
-1, \text{if   } x_k < \tilde{x}_k \\
-0, \text{otherwise}
-\end{cases}
-\end{equation}
 
@@
 
-\section{Conclusions}
-\label{sec:conclusions}
+\subsection{Evaluation}
 
-Classification through deep learning seems promising for orchestration because of our model's ability to classify individual instruments and pitches out of dense combinations of samples.
-While our model is not able to outperform OrchIdea, it shows consistent results.
-When comparing our methods to OrchIdea, it is important to note that many of processes OrchIdea uses to optimize its solution are not present in our model.
-Our model also lacks the implementation of symbolic constraints, which is an important part of assisted orchestration.
+We evaluated our orchestrations both qualitatively and quantitatively by comparing our solutions to the solutions generated by Orchidea, the state-of-the-art system for computer-assisted orchestration.
 
-We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets.
-Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.
-
-Future steps in this project include implementing various methods that are present in OrchIdea.
-
-Our current model orchestrates all targets using the same number of samples, and this does not take into account the density of different targets.
-The solution to this is to allow sparse solutions in which the model decides how many samples should be used to best represent the target.
-This allows a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds.
-
-Partial filtering is a method that would aid our model in orchestrating the harmonics of a target.
+Orchidea implements many advanced features that are not supported by our models.
+For example, it is able to apply symbolic constraints to the search, hence allowing only specific instrumental combinations or playing styles.
+It is also able to reduce the search space by applying harmonic analysis on the target sound.
 The dominant harmonic partials of the target are identified, and the search space is limited to only include samples of those pitches.
@@
 This leads to a solution whose harmonics are much closer to the target, which is an important part of aural similarity.
+Orchidea's solutions, moreover, can use any number of instruments included in the orchestra specified by the user, thus having variable-sized orchestrations from a single instrument to the whole set (this propriety is called \emph{sparsity} of the solution).
+
+In order to have a more fair comparison, we did not allow Orchidea to use any of the advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it forced to use all 10 instruments in each solution.
+This creates a more fair comparison, since our models are unable to create constrained or sparse solutions.
+
+Qualitative evaluation was done through an acoustic inspection of the solution, paying close attention to timbre and pitch.
+For targets that had harmonic content, it was noted if the partials present in the target were also represented in the orchestrated solution.
+For example, one of the samples of a bell had partials that loosely represented a C minor chord, so we checked if the orchestration contained the notes of the chord.
+If a target had specific notes, we identified if the note or its partials were present.
+For example there is target of an Oboe playing an A4 and a Bassoon playing a C\#3, the solution from ResNet contained the partials of the Bassoon's note: C\#3, G\#4, C\#5, and E6 were all included on various instruments in the solution.
+
+For quantitative evaluation, we used the distance metric defined in Eqn.
+\eqref{distance} to calculate differences in timbre between targets and solutions.
+This metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization.
+The equation takes in the full FFT of the target $x$ and full FFT of the solution $\hat{x}$.
+Then for each bin $k$ of the FFT, it calculates the absolute difference between the values.
+The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.
+
+In the first summation, $\lambda_1$ is multiplied by all the distances calculated when there was more energy in the target than the solution, since $\delta_{k1} = 1$ only when $x_k \ge \tilde{x}_k$.
+Similarly in the second summation, $\lambda_2$ is multiplied by all the distances when the solution provided more energy to a frequency than the target.
+Therefore, the relation between $\lambda_1$ and $\lambda_2$ determines whether a solution is penalized more for undershooting or overshooting the target.
+
+We calculated the distance between the target samples and our orchestrated solutions.
+We then orchestrated the same targets with Orchidea and calculated the distance for Orchidea's solutions.
+A comparison of these results is in Tab.
+\ref{orch_eval}.
+
+\begin{equation}\label{distance}
+d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k| \\
+\end{equation}
+where 
+$
+\delta_{k1} = 
+\begin{cases}
+1, \text{if   } x_k \ge \tilde{x}_k \\
+0, \text{otherwise}
+\end{cases} 
+$, $
+\delta_{k2} = 
+\begin{cases}
+1, \text{if   } x_k < \tilde{x}_k \\
+0, \text{otherwise}
+\end{cases}.
+$\\
+
+While our model is not able to outperform Orchidea, it shows consistent results.
+We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets.
+Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.
+
+\section{Conclusions}
+\label{sec:conclusions}
+
+Target-based computer-assisted orchestration through deep learning models seems a promising path, thanks to the ability of deep networks to classify individual instruments and pitches out of dense combinations of samples.
+This work, however, represents only a preliminary study of the potential of these methods for the task of assisted orchestration.
+
+The first natural extension would be to support sparsity in our models.
+Our current models, indeed, orchestrate all targets using a constant number of instruments and it is not able to drop specific instruments from the solution.
+This does not take into account the density of different targets.
+By using sparse solutions, in which the model decides how many samples should be used to best represent the target, would allow a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds.
+This would generate more meaningful orchestrations that would compare more favourably to the state of the art.
+
+Another important extension would be to create a more powerful embedding spaces for the target and combinations.
+In \cite{Gillick19} the authors propose to use LSTM-based models to predict the embedding features for the combinations used during the optimisation process in assisted orchestration.
+We believe that by combining their prediction model with our classification models could generate more faithful representations and improve the overall quality of generated orchestrations.
 
@@
 
+\bibitem{FerVic2013} J.D. Fern\'andez and F. Vico. {AI Methods in Algorithmic Composition: A Comprehensive Survey}. Journal of Artificial Intelligence Research, vol.48, pp. 513--582, 2013.
+
+\bibitem{Ari2005} C.
+Ariza, {Navigating the landscape of computer aided algorithmic composition systems: A definition, seven descriptors, and a lexicon of systems and research}.
+In Proceedings of the International Computer Music Conference, 2005.
+
+\bibitem{Maresz2003}
+Yan Maresz, {On Computer-Assisted Orchestration}. Contemporary Music Review, vol. 32, n. 1, pg. 99-109, Routledge, 2013.
+
+\bibitem{Carpentier2010}
+Gr\'egoire Carpentier, Damien Tardieu, Jonathan Harvey, G\'erard Assayag and Emmanuel Saint-James.
+{Predicting Timbre Features of Instrument Sound Combinations: Application to Automatic Orchestration}, Journal of New Music Research, vol.
+39, n.
+1, pg.
+47-61, Routlegde, 2010.
+
 \bibitem{Benetos07}
@@
 \newblock Large scale musical instrument identification.
-\newblock 2007.
+\newblock Sound and Music Computing Conference, Greece, 2007.
 
@@
 \newblock Audio-based music classification with densenet and data augmentation.
-\newblock ArXiv (https://arxiv.org/abs/1906.11620v1), 2019.
-
-\bibitem{Cella18}
-Carmine~Emanuele Cella and Philippe Esling.
-\newblock Open-Source Modular Toolbox for Computer-Aided Orchestration.
-\newblock Timbre conference, 2018.
-
-\bibitem{Cella19}
-Carmine~Emanuele Cella, Jon Gillick, and David Bamman.
-\newblock Estimating unobserved audio features for targed-based orchestration.
-\newblock ISMIR, 2019.
+\newblock arXiv preprint, https://arxiv.org/abs/1906.11620v1, 2019.
 
@@
 \newblock Orchidea: a comprehensive framework for target-based assisted orchestration.
-\newblock in preparation; available on request from the author, 2020.
+\newblock In preparation; available on request from the author, 2020.
 
@@
 Nicolas Chetry, Mike Davies, and Mark Sandler.
-\newblock Musical instrument identification using lsf and k-means.
-\newblock 2005.
+\newblock Musical instrument identification using LSF and K-means.
+\newblock Audio Engineering Society Convention 118, 2005.
 
@@
 \newblock Deep residual learning for image recognition.
-\newblock 2015.
+\newblock arXiv preprint, https://arxiv.org/abs/1512.03385, 2015.
 
-\bibitem{Hershey16}
+\bibitem{Hershey17}
 Shawn Hershey, Sourish Chaudhuri, Daniel P.~W. Ellis, Jort~F. Gemmeke, Aren
@@
   Seybold, Malcolm Slaney, Ron~J. Weiss, and Kevin Wilson.
-\newblock Cnn architectures for large-scale audio classification.
-\newblock 2016.
+\newblock CNN architectures for large-scale audio classification.
+\newblock IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP), 2017.
 
@@
 \newblock Pitch-dependent identification of musical instrument sounds.
-\newblock 2005.
+\newblock Applied Intelligence 23, 267–275, Springer, 2005.
 
@@
 \newblock Deep convolutional networks on the pitch spiral for musical
-  instrument recognition, 2016.
+  instrument recognition. ISMIR, 2016.
 
@@
 \newblock librosa: Audio and music signal analysis in python.
-\newblock 2015.
+\newblock {P}roceedings of the 14th {P}ython in {S}cience {C}onference, 2015.
 
@@
 
-\bibitem{Stevens37}
-S.~S. Stevens, J.~Volkmann, , and E.~B. Newman.
-\newblock A scale for the measurement of the psychological magnitude pitch.
-\newblock 1937.
+\bibitem{McKinney2003}
+Martin F. McKinney and Jeroen Breebaart.
+\newblock Features for Audio and Music Classification.
+\newblock ISMIR, 2003.
+
+\bibitem{Cella18}
+Carmine-Emanuele Cella and Philippe Esling, {Open-source modular toolbox for computer-aided orchestration}, Proc. of the Timbre conference, Montreal, Canada, 2018.
+
+\bibitem{Gillick19}
+Jon Gillick, Carmine-Emanuele Cella and David Bamman. {Estimating Unobserved Audio Features for Target-Based Orchestration}, ISMIR, 2019.
 

### 2020-05-16 paper/latex/paper.tex
fixed typos - final version submitted for review

@@
 In this paper we will do a preliminary exploration on how neural networks can be used for the task of target-based computer-assisted musical orchestration.
-We will show how it possible to model this  musical problem as a classification task and we will propose two deep learning models.
-We will show, first, how they perform as classifiers for musical instruments recognition by comparing them with specific baselines.
-We will then show how they perform, both qualitative and quantitative, in the task of computer-assisted orchestration by comparing them with state-of-the-art systems.
+We will show how it is possible to model this  musical problem as a classification task and we will propose two deep learning models.
+We will show, first, how they perform as classifiers for musical instrument recognition by comparing them with specific baselines.
+We will then show how they perform, both qualitatively and quantitatively, in the task of computer-assisted orchestration by comparing them with state-of-the-art systems.
 Finally, we will highlight benefits and problems of neural approaches for assisted orchestration and we will propose possible future steps.
@@
 The development of computational tools to assist and inspire the musical composition process constitutes an import research area known as \emph{Computer-Assisted Composition (CAC)} \cite{FerVic2013, Ari2005}.
-Within CAC, target-based computer-assisted orchestration is an important example how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.
+Within CAC, target-based computer-assisted orchestration is an important example of how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.
 
@@
 
-\subsection{Form orchestration to classification}
+\subsection{From orchestration to classification}
 
@@
 
-Three baseline classifiers have been trained to recognize either the instrument only for a total of 12 classes or the pair instrument-pitch \emph{class} (a pitch class is the pitch of a note without the corresponding octave, e.g.
-C, D\#, G, etc.) for a total of 25 classes.
+Three baseline classifiers have been trained to recognize either the instrument only for a total of 12 classes (the instruments in TinySOL) or the pair instrument-pitch \emph{class} (a pitch class is the pitch of a note without the corresponding octave, e.g.
+C, D\#, G, etc.) for a total of 25 classes (we will classifiy combinations of 2 instruments for 12 pitch classes each, plus one extra \emph{trash} class, see section \ref{sec:baseline}).
 
 Two neural classifiers, instead, have been trained on the more complex problem of recognizing instrument-pitch pairs, for a total of 424 classes (i.e.
-the neural classifiers would classify instruments and pitches with respective octaves).
+the neural classifiers would classify instruments and pitches with respective octaves, hence 12 instruments for 12 pitch classes and three dynamics; note that not all the samples in TinySOL have all dynamics).
 We want to point out, however, that this is still a simplified problem compared to assisted orchestration.
@@
 
-The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an accurate representation of sound and musical signals.
-The Mel spectrogram is generated using an FFT hop length of 2048 samples (e.g.
-the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128.
+The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an appropriate representation of sound and musical signals.
+The Mel spectrogram is generated using an FFT hop length of 2048 samples (the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128.
 Therefore, the Mel features fed to the model were matrices of size $128\times 345$.
@@
 
-\subsection{Baseline}
-The baseline classifiers that we tested are support vector machine (SVM), random forest (RF), and K-nearest neighbours (KNN).
+\subsection{Baselines}
+\label{sec:baseline}
+
+In order to get a betters sense on the complexity of the problem, we tested three baseline classifiers: support vector machine (SVM), random forest (RF), and K-nearest neighbours (KNN).
 We used the implementations provided in the \emph{scikit-learn }library for Python \cite{scikit-learn}.
@@
 Each sample was made by a combination of one to ten instruments and was four seconds in length.
-In this case, differently from the neural models, the features used are the MFCCs of the resulting combination; we chose to use MFCCs for this setting, instead of the Mel spectrogram, to have a number of features that is more appropriate for the parametric classifiers.
+In this case, differently from the neural models, the features used are the MFCCs of the resulting combination; we chose to use MFCCs for this setting, instead of the Mel spectrogram, to have a lower number of features, that is more manageable by parametric classifiers.
 We found SVM to have the highest accuracy of the three classifiers across all experiments.
@@
 If another instrument was present in the input combination that was not Violin or Viola, the classifier would simply identify that an instrument that was not one of the two was present.
-The accuracies from this experiment are outlined in table \ref{tab:baselines}.
-Since RF performed worse than SVM is every experiment, we stopped testing with it and used SVM from that point on.
+The accuracies from this experiment are outlined in Table \ref{tab:baselines}.
+Since RF performed worse than SVM in every experiment, we stopped testing with it and used SVM from that point on.
 
@@
 Flute, Oboe, and Violin reached an accuracy of 11.1\%, and Bass Tuba, Trumpet, and Trombone was 0.5\%.
-As we increased the number of instruments whose pitch classes was being identified, the accuracy continued to drop.
-For classifying the pitch class of four instruments: Oboe, French Horn, Violin, and Flute, the accuracy was 2.7\%.
+As we increased the number of instruments whose pitch classes was being identified, the accuracy continued to drop.Indeed, for classifying the pitch class of four instruments (Oboe, French Horn, Violin, and Flute) the accuracy was 2.7\%.
 
@@
 The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit.
-CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey16}.
+
+%CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey17}.
 LSTM units provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.
@@
 
-Different experiments were made by varying the number $N$ of samples in each mixture sample.
+Different experiments were made by varying the number $N$ of samples in each mixture.
 We used an orchestra of 10 instruments containing Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet in C and Clarinet in Bb.
@@
 After training the neural models for classification, we finally tested them for the task of target-based computer-assisted orchestration.
-For this purpose, we decided to use the models trained on combinations of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet.
-The total number of classes in this setting is 424, each one representing an instrument-pitch pair.
+For this purpose, we decided to use the models trained on combinations of 10 instruments, namely French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet.
+Each class, in this model, represents an instrument-pitch pair.
 
@@
 
-Since we decided not to train our models to classify the dynamics of a sample (despite TinySOL having \emph{pianissimo}, \emph{mezzoforte}, and \emph{fortissimo} recordings), the dynamics is determined by the probability of each sample as output by the model.
+Since we decided not to train our models to classify the dynamics of a sample (despite TinySOL having \emph{pianissimo}, \emph{mezzoforte}, and \emph{fortissimo} recordings), the dynamics are determined by the probability of each sample as output by the model.
 If the model outputs a probability higher than $0.66$ for a sample, the fortissimo version of the sample is used.
 If the probability is between $0.33$ and $0.66$, then the mezzoforte version is selected and if it is less than $0.33$ the pianissimo version is used.
-The idea behind this quantization, is that samples that are the most represented in the target should appear as the loudest in the orchestrated solution.
+The idea behind this quantization is that samples that are the most represented in the target should appear as the loudest in the orchestrated solution.
 
@@
 This leads to a solution whose harmonics are much closer to the target, which is an important part of aural similarity.
-Orchidea's solutions, moreover, can use any number of instruments included in the orchestra specified by the user, thus having variable-sized orchestrations from a single instrument to the whole set (this propriety is called \emph{sparsity} of the solution).
+Orchidea's solutions, moreover, can use any number of instruments included in the orchestra specified by the user, thus having variable-sized orchestrations from a single instrument to the whole set (this property is called \emph{sparsity} of the solution).
 
@@
 We then orchestrated the same targets with Orchidea and calculated the distance for Orchidea's solutions.
-A comparison of these results is in Tab.
-\ref{orch_eval}.
+A comparison of these results is in Table \ref{orch_eval}.
 
@@
 This does not take into account the density of different targets.
-By using sparse solutions, in which the model decides how many samples should be used to best represent the target, would allow a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds.
+Sparse solutions, in which the model decides how many samples should be used to best represent the target, would allow a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds.
 This would generate more meaningful orchestrations that would compare more favourably to the state of the art.
@@
 
-\bibitem{FerVic2013} J.D. Fern\'andez and F. Vico. {AI Methods in Algorithmic Composition: A Comprehensive Survey}. Journal of Artificial Intelligence Research, vol.48, pp. 513--582, 2013.
+\bibitem{FerVic2013} Jose David Fern\'andez and Francisco Vico.
+{AI Methods in Algorithmic Composition: A Comprehensive Survey}.
+Journal of Artificial Intelligence Research, vol.48, pp.
+513--582, 2013.
 
-\bibitem{Ari2005} C.
-Ariza, {Navigating the landscape of computer aided algorithmic composition systems: A definition, seven descriptors, and a lexicon of systems and research}.
+\bibitem{Ari2005} Christopher Ariza, {Navigating the landscape of computer aided algorithmic composition systems: A definition, seven descriptors, and a lexicon of systems and research}.
 In Proceedings of the International Computer Music Conference, 2005.
@@
 \bibitem{Cella2020}
-Carmine~Emanuele Cella.
+Carmine-Emanuele Cella.
 \newblock Orchidea: a comprehensive framework for target-based assisted orchestration.
@@
 \bibitem{Cella2020b}
-Carmine~Emanuele Cella, Daniele Ghisi, Vincent Lostanlen, Fabi\'en Levy, Yan Maresz and Joshua Fineberg.
+Carmine-Emanuele Cella, Daniele Ghisi, Vincent Lostanlen, Fabi\'en Levy, Yan Maresz and Joshua Fineberg.
 \newblock OrchideaSOL: a dataset of extended instrumental techniques for computer-aided orchestration.
@@
 \bibitem{scikit-learn}
-F.~Pedregosa, G.~Varoquaux, A.~Gramfort, V.~Michel, B.~Thirion, O.~Grisel,
-  M.~Blondel, P.~Prettenhofer, R.~Weiss, V.~Dubourg, J.~Vanderplas, A.~Passos,
-  D.~Cournapeau, M.~Brucher, M.~Perrot, and E.~Duchesnay.
+Fabian Pedregosa, Ga\"el Varoquaux, Alexandre
+Gramfort, Vincent Michel, Bertrand Thirion, Olivier
+Grisel, Mathieu Blondel, Peter Prettenhofer, Ron
+Weiss, Vincent Dubourg, Jake Vanderplas, Alexandre
+Passos, David Cournapeau, Matthieu Brucher,
+Matthieu Perrot, \'Edouard Duchesnay.
 \newblock Scikit-learn: Machine learning in {P}ython.

### 2020-05-29 paper/latex/paper.tex
minor modif

@@
 
-In order to get a betters sense on the complexity of the problem, we tested three baseline classifiers: support vector machine (SVM), random forest (RF), and K-nearest neighbours (KNN).
+In order to get a better sense on the complexity of the problem, we tested three baseline classifiers: support vector machine (SVM), random forest (RF), and K-nearest neighbours (KNN).
 We used the implementations provided in the \emph{scikit-learn }library for Python \cite{scikit-learn}.

### 2020-08-17 paper/aimusic2020/CSMC_MUME_LaTeX_Template.tex
Add aimusic2020 paper

%%%%%%%%%%%%%%%%%%%% file CSMC_MUME_LaTeX_Template.tex %%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Journal of Creative Music Systems.
% Copyright: http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{url}
\usepackage{apacite}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\pagestyle{headings}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{\LaTeX{} Template for \textit{CSMC-MuMe 2020}}

% a short form should be given in case it is too long for the running head
\titlerunning{Running Title}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{AuthorA\inst{1}\and AuthorB\inst{2} \thanks{Please place acknowledgement here.}}
%
% if the names of the authors are too long for the running head, please use the format: AuthorA et al.
\authorrunning{AuthorA and AuthorB (or AuthorA et al. if too long)}

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{InstituteA \and InstituteB\\ \email{author@phdsrus.ac.uk}}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

\maketitle

\begin{abstract}
The abstract should summarize the contents of the paper and should
contain at least 70 and at most 150 words. It should be written using the
\textit{abstract} environment.
\keywords{Please list your keywords in this section}
\end{abstract}

\section{Introduction}

This is a \LaTeX{} template for submission to the \textit{2020 Joint Conference on AI Music Creativity (CSMC + MuMe)}.

\section{Paper Preparation}

When you use \LaTeX\ together with our document class file,
\texttt{llncs.cls},
your text is typeset automatically in Computer Modern Roman (CM) fonts.
Please do
\emph{not} change the preset fonts. If you have to use fonts other
than the preset fonts, kindly submit these with your files.

Please use the commands \verb+\label+ and \verb+\ref+ for
cross-references and the commands \verb+\bibitem+ and \verb+\cite+ for
references to the bibliography, to enable us to create hyperlinks at
these places.

For preparing your figures electronically and integrating them into
your source file we recommend using the standard \LaTeX{} \verb+graphics+ or
\verb+graphicx+ package. These provide the \verb+\includegraphics+ command.
In general, please refrain from using the \verb+\special+ command.

Remember to submit any further style files and
fonts you have used together with your source files.

\subsubsection{Headings.}

Headings should be capitalized
(i.e., nouns, verbs, and all other words
except articles, prepositions, and conjunctions should be set with an
initial capital) and should,
with the exception of the title, be aligned to the left.
Words joined by a hyphen are subject to a special rule. If the first
word can stand alone, the second word should be capitalized.

Here are some examples of headings: ``Criteria to Disprove
Context-Freeness of Collage Language'', ``On Correcting the Intrusion of
Tracing Non-deterministic Programs by Software'', ``A User-Friendly and
Extendable Data Distribution System'', ``Multi-flip Networks:
Parallelizing GenSAT'', ``Self-determinations of Man''.

\subsubsection{Lemmas, Propositions, and Theorems.}

The numbers accorded to lemmas, propositions, and theorems, etc. should
appear in consecutive order, starting with Lemma 1, and not, for
example, with Lemma 11.

\subsection{Figures}

For \LaTeX\ users, we recommend using the \emph{graphics} or \emph{graphicx}
package and the \verb+\includegraphics+ command.

Please check that the lines in line drawings are not
interrupted and are of a constant width. Grids and details within the
figures must be clearly legible and may not be written one on top of
the other. Line drawings should have a resolution of at least 800 dpi
(preferably 1200 dpi). The lettering in figures should have a height of
2~mm (10-point type). Figures should be numbered and should have a
caption which should always be positioned \emph{under} the figures, in
contrast to the caption belonging to a table, which should always appear
\emph{above} the table; this is simply achieved as matter of sequence in
your source.

Please centre the figures or your tabular material by using the \verb+\centering+
declaration. Short captions are centred by default between the margins
and typeset in 9-point type (Fig.~\ref{fig:example} shows an example).
The distance between text and figure is preset to be about 8~mm, the
distance between figure and caption about 6~mm.

To ensure that the reproduction of your illustrations is of a reasonable
quality, we advise against the use of shading. The contrast should be as
pronounced as possible.

If screenshots are necessary, please make sure that you are happy with
the print quality before you send the files.

\begin{figure}
	\centering
	\includegraphics[height=6.2cm]{graphic.pdf}
	\caption{One kernel at $x_s$ (\emph{dotted kernel}) or two kernels at
	$x_i$ and $x_j$ (\textit{left and right}) lead to the same summed estimate
	at $x_s$. This shows a figure consisting of different types of
	lines. Elements of the figure described in the caption should be set in
	italics, in parentheses, as shown in this sample caption.}
	\label{fig:example}
\end{figure}

Please define figures (and tables) as floating objects. Please avoid
using optional location parameters like ``\verb+[h]+'' for ``here''.

\subsection{Formulas}

Displayed equations or formulas are centred and set on a separate
line (with an extra line or halfline space above and below). Displayed
expressions should be numbered for reference. The numbers should be
consecutive within each section or within the contribution,
with numbers enclosed in parentheses and set on the right margin --
which is the default if you use the \emph{equation} environment, e.g.,

\begin{equation}
  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt \;  .
\end{equation}

Equations should be punctuated in the same way as ordinary
text but with a small space before the end punctuation mark.

\subsection{Footnotes}

The superscript numeral used to refer to a footnote appears in the text
either directly after the word to be discussed or -- in relation to a
phrase or a sentence -- following the punctuation sign (comma,
semicolon, or period). Footnotes should appear at the bottom of
the normal text area, with a line of about 2~cm set
immediately above them.\footnote{The footnote numeral is set flush left
and the text follows with the usual word spacing.}

\subsection{Program Code}

Program listings or program commands in the text are normally set in
typewriter font, e.g., CMTT10 or Courier.

\medskip

\noindent
{\it Example of a Computer Program}
\begin{verbatim}
program Inflation (Output)
  {Assuming annual inflation rates of 7%, 8%, and 10%,...
   years};
   const
     MaxYears = 10;
   var
     Year: 0..MaxYears;
     Factor1, Factor2, Factor3: Real;
   begin
     Year := 0;
     Factor1 := 1.0; Factor2 := 1.0; Factor3 := 1.0;
     WriteLn('Year  7% 8% 10%'); WriteLn;
     repeat
       Year := Year + 1;
       Factor1 := Factor1 * 1.07;
       Factor2 := Factor2 * 1.08;
       Factor3 := Factor3 * 1.10;
       WriteLn(Year:5,Factor1:7:3,Factor2:7:3,Factor3:7:3)
     until Year = MaxYears
end.
\end{verbatim}
%
\noindent
{\small (Example from Jensen K., Wirth N. (1991). \textit{Pascal user manual and
report}. New York: Springer)}

\subsection{Citations}

For citations in the text please use the APA6 system (\url{http://www.apastyle.org/}).

Examples:
\cite{einstein}
\cite<e.g.,>{knuthwebsite}
\cite[p.~10]{latexcompanion}
\citeA{latexcompanion}
(\citeNP{einstein}; \citeNP{latexcompanion})
\cite[pp.~10--11]{einstein}

\subsection{Page Numbering and Running Heads}

There is no need to include page numbers. If your paper title is too
long to serve as a running head, it will be shortened. Your suggestion
as to how to shorten it would be most welcome.

\subsubsection*{Acknowledgements} The heading should be treated as a
subsubsection heading and should not be assigned a number.

\section{The References Section}\label{references}

Please use the APA6 style for references.
The reference section must be complete. You may not omit references.
Instructions as to where to find a fuller version of the references are
not permissible.

We only accept references written using the Latin alphabet. If the title
of the book you are referring to is in Russian or Chinese, then please write
(in Russian) or (in Chinese) at the end of the transcript or translation
of the title.

The following section shows a sample reference list.

\bibliographystyle{apacite}
% Place your own .bib file here:
\bibliography{CSMC_MUME_LaTeX_Template.bib}

\end{document}

### 2020-08-17 paper/aimusic2020/paper.tex
Add aimusic2020 paper

--- 2020-05-29 paper/latex/paper.tex
@@
-% -----------------------------------------------
-% Template for ISMIR Papers
-% 2020 version, based on previous ISMIR templates
+%%%%%%%%%%%%%%%%%%%% file CSMC_MUME_LaTeX_Template.tex %%%%%%%%%%%%%%%%%%%%%
+%
+% This is the LaTeX source for the instructions to authors using
+% the LaTeX document class 'llncs.cls' for contributions to
+% the Journal of Creative Music Systems.
+% Copyright: http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
+%
+% It may be used as a template for your own input - copy it
+% to a new file with a new name and use it as the basis
+% for your article.
+%
+% NB: the document class 'llncs' has its own and detailed documentation, see
+% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Requirements :
-% * 6+n page length maximum
-% * 4MB maximum file size
-% * Copyright note must appear in the bottom left corner of first page
-% * Clearer statement about citing own work in anonymized submission
-% (see conference website for additional details)
-% -----------------------------------------------
+\documentclass[runningheads,a4paper]{llncs}
 
-\documentclass{article}
-\usepackage[T1]{fontenc} % add special characters (e.g., umlaute)
-\usepackage[utf8]{inputenc} % set utf-8 as default input encoding
-\usepackage{ismir,amsmath,cite,url}
+\usepackage{amssymb}
+\setcounter{tocdepth}{3}
 \usepackage{graphicx}
-\usepackage{color}
+\usepackage{url}
+\usepackage{apacite}
+\newcommand{\keywords}[1]{\par\addvspace\baselineskip
+\noindent\keywordname\enspace\ignorespaces#1}
 
-% Optional: To use hyperref, uncomment the following.
-\usepackage[bookmarks=false,hidelinks]{hyperref}
-% Mind the bookmarks=false option; bookmarks are incompatible with ismir.sty.
-
-\usepackage{lineno}
-%\linenumbers
-\newcommand{\carmine}[1]{\textbf{\color{red} Carmine => #1}}
-\renewcommand{\baselinestretch}{.98}
-
-% Title.
-% ------
-\title{Neural models for target-based computer-assisted musical orchestration: a preliminary study}
-
-% Note: Please do NOT use \thanks or a \footnote in any of the author markup
-
-% Single address
-% To use with only one author or several with the same address
-% ---------------
-%\oneauthor
-% {Names should be omitted for double-blind reviewing}
-% {Affiliations should be omitted for double-blind reviewing}
-
-% Two addresses
-% --------------
-%\twoauthors
-%  {First author} {School \\ Department}
-%  {Second author} {Company \\ Address}
-
-%% To make customize author list in Creative Common license, uncomment and customize the next line
-%  \def\authorname{First Author, Second Author}
-
-% Three addresses
-% --------------
-%\threeauthors
-%  {First Author} {Affiliation1 \\ {\tt author1@ismir.edu}}
-%  {Second Author} {\bf Retain these fake authors in\\\bf submission to preserve the formatting}
-%  {Third Author} {Affiliation3 \\ {\tt author3@ismir.edu}}
-
-%% To make customize author list in Creative Common license, uncomment and customize the next line
-%  \def\authorname{First Author, Second Author, Third Author}
-
-% Four or more addresses
-% OR alternative format for large number of co-authors
-% ------------
-\multauthor
-{First author$^1$ \hspace{1cm} Second author$^1$ \hspace{1cm} Third author$^2$} { \bfseries{Fourth author$^3$ \hspace{1cm} Fifth author$^2$}\\
-  $^1$ Department of Computer Science, University , Country\\
-$^2$ International Laboratories, City, Country\\
-$^3$  Company, Address\\
-{\tt\small CorrespondenceAuthor@ismir.edu, PossibleOtherAuthor@ismir.edu}
-}
-\def\authorname{First author, Second author, Third author, Fourth author, Fifth author}
-
-\sloppy % please retain sloppy command for improved formatting
+\pagestyle{headings}
 
@@
 
+\mainmatter  % start of an individual contribution
+
+% first the title is needed
+\title{Neural Models for Target-Based Computer-Assisted Musical Orchestration}
+
+% a short form should be given in case it is too long for the running head
+\titlerunning{Running Title}
+
+% the name(s) of the author(s) follow(s) next
 %
+% NB: Chinese authors should write their first names(s) in front of
+% their surnames. This ensures that the names appear correctly in
+% the running heads and the author index.
+%
+\author{AuthorA\inst{1}\and AuthorB\inst{2} \thanks{Please place acknowledgement here.}}
+%
+% if the names of the authors are too long for the running head, please use the format: AuthorA et al.
+\authorrunning{AuthorA and AuthorB (or AuthorA et al. if too long)}
+
+% the affiliations are given next; don't give your e-mail address
+% unless you accept that it will be published
+\institute{InstituteA \and InstituteB\\ \email{author@phdsrus.ac.uk}}
+
+%
+% NB: a more complex sample for affiliations and the mapping to the
+% corresponding authors can be found in the file "llncs.dem"
+% (search for the string "\mainmatter" where a contribution starts).
+% "llncs.dem" accompanies the document class "llncs.cls".
+%
+
 \maketitle
-%
+
 \begin{abstract}
@@
 \begin{figure}
-\includegraphics[scale=0.17]{figs/method.png}
+\includegraphics[scale=0.17]{../latex/figs/method.png}
 \caption{An overview of the proposed method for assisted orchestration with neural models.
@@
 \begin{center}
-\includegraphics[scale=0.22]{figs/CNN_LSTM_diagram.png}
+\includegraphics[scale=0.22]{../latex/figs/CNN_LSTM_diagram.png}
 \caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
@@
 \center{
-\includegraphics[scale=0.4]{figs/CNN_vs_ResNet.png}}
+\includegraphics[scale=0.4]{../latex/figs/CNN_vs_ResNet.png}}
 \caption{Best overall accuracy for CNN with LSTM and ResNet depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
@@
 \begin{center}
-\includegraphics[scale=0.4]{figs/Acc_vs_N_CNN.png}
+\includegraphics[scale=0.4]{../latex/figs/Acc_vs_N_CNN.png}
 \caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM depending on the number of instruments in the combinations. \label{best_acc_cnn}}
@@
 
-\begin{equation}
-M_N(\hat{X})_i = \left\{\begin{array}{ll}
-1 \text{ if } i \in I_N(\hat{X})\\
-0 \text{ otherwise}
-\end{array}\right.
-\label{NMax}
-\end{equation}
+%\begin{equation}
+%M_N(\hat{X})_i = \left\{\begin{array}{ll}
+%1 \text{ if } i \in I_N(\hat{X})\\
+%0 \text{ otherwise}
+%\end{array}\right.
+%\label{NMax}
+%\end{equation}
 and $I_N(\hat{X})$ is the set of indices containing the $N$ first maximums of the vector $\hat{X}$.
@@
 
-Fig.
+%Fig.
 \ref{best_acc_cnn} and Fig.
@@
 \begin{center}
-\includegraphics[scale=0.40]{figs/Acc_vs_N_ResNet.png}
+\includegraphics[scale=0.40]{../latex/figs/Acc_vs_N_ResNet.png}
 \caption{Best overall accuracy and for each instrument obtained by the ResNet depending on the number of instruments in the combinations. \label{best_acc_resnet}}
@@
   \end{center}
-  \caption{Quantitative comparison of orchestrations as ratios to Orchidea.
+%  \caption{Quantitative comparison of orchestrations as ratios to Orchidea.
 Eqn.
@@
 
-For quantitative evaluation, we used the distance metric defined in Eqn.
+%For quantitative evaluation, we used the distance metric defined in Eqn.
 \eqref{distance} to calculate differences in timbre between targets and solutions.
@@
 where 
-$
-\delta_{k1} = 
-\begin{cases}
-1, \text{if   } x_k \ge \tilde{x}_k \\
-0, \text{otherwise}
-\end{cases} 
-$, $
-\delta_{k2} = 
-\begin{cases}
-1, \text{if   } x_k < \tilde{x}_k \\
-0, \text{otherwise}
-\end{cases}.
-$\\
+%$
+%\delta_{k1} = 
+%\begin{cases}
+%1, \text{if   } x_k \ge \tilde{x}_k \\
+%0, \text{otherwise}
+%\end{cases} 
+%$, $
+%\delta_{k2} = 
+%\begin{cases}
+%1, \text{if   } x_k < \tilde{x}_k \\
+%0, \text{otherwise}
+%\end{cases}.
+%$\\
 
@@
 
+\bibliographystyle{apacite}
+\bibliography{paper}
 % For bibtex users:
 %\bibliography{paper}
-
-\begin{thebibliography}{99}
-
-\bibitem{FerVic2013} Jose David Fern\'andez and Francisco Vico.
+%\bibliographystyle{apalike}
+%\begin{thebibliography}{99}
+%
+%\bibitem{FerVic2013} Jose David Fern\'andez and Francisco Vico.
 {AI Methods in Algorithmic Composition: A Comprehensive Survey}.
@@
 513--582, 2013.
-
-\bibitem{Ari2005} Christopher Ariza, {Navigating the landscape of computer aided algorithmic composition systems: A definition, seven descriptors, and a lexicon of systems and research}.
+%
+%\bibitem{Ari2005} Christopher Ariza, {Navigating the landscape of computer aided algorithmic composition systems: A definition, seven descriptors, and a lexicon of systems and research}.
 In Proceedings of the International Computer Music Conference, 2005.
-
-\bibitem{Maresz2003}
-Yan Maresz, {On Computer-Assisted Orchestration}. Contemporary Music Review, vol. 32, n. 1, pg. 99-109, Routledge, 2013.
-
-\bibitem{Carpentier2010}
-Gr\'egoire Carpentier, Damien Tardieu, Jonathan Harvey, G\'erard Assayag and Emmanuel Saint-James.
+%
+%\bibitem{Maresz2003}
+%Yan Maresz, {On Computer-Assisted Orchestration}. Contemporary Music Review, vol. 32, n. 1, pg. 99-109, Routledge, 2013.
+%
+%\bibitem{Carpentier2010}
+%Gr\'egoire Carpentier, Damien Tardieu, Jonathan Harvey, G\'erard Assayag and Emmanuel Saint-James.
 {Predicting Timbre Features of Instrument Sound Combinations: Application to Automatic Orchestration}, Journal of New Music Research, vol.
@@
 47-61, Routlegde, 2010.
-
-\bibitem{Benetos07}
-Emmanouil Benetos, Margarita Kotti, and Constantine Kotropoulos.
-\newblock Large scale musical instrument identification.
-\newblock Sound and Music Computing Conference, Greece, 2007.
-
-\bibitem{Bhardwaj17}
-Siddharth Bhardwaj.
-\newblock Audio data augmentation with respect to musical instrument
-  recognition.
-\newblock Master's thesis, 2017.
-
-\bibitem{Bian19}
-Wenhao Bian, Jie Wang, Bojin Zhuang, Shaojun~Wang Jiankui~Yang, and Jing Xiao.
-\newblock Audio-based music classification with densenet and data augmentation.
-\newblock arXiv preprint, https://arxiv.org/abs/1906.11620v1, 2019.
-
-\bibitem{Cella2020}
-Carmine-Emanuele Cella.
-\newblock Orchidea: a comprehensive framework for target-based assisted orchestration.
-\newblock In preparation; available on request from the author, 2020.
-
-\bibitem{Cella2020b}
-Carmine-Emanuele Cella, Daniele Ghisi, Vincent Lostanlen, Fabi\'en Levy, Yan Maresz and Joshua Fineberg.
-\newblock OrchideaSOL: a dataset of extended instrumental techniques for computer-aided orchestration.
-\newblock ICMC (under review), 2020.
-
-\bibitem{Chetry05}
-Nicolas Chetry, Mike Davies, and Mark Sandler.
-\newblock Musical instrument identification using LSF and K-means.
-\newblock Audio Engineering Society Convention 118, 2005.
-
-\bibitem{He15}
-Kaiming He, Xiangyu Zhang, Shaoqing Ren, and Jian Sun.
-\newblock Deep residual learning for image recognition.
-\newblock arXiv preprint, https://arxiv.org/abs/1512.03385, 2015.
-
-\bibitem{Hershey17}
-Shawn Hershey, Sourish Chaudhuri, Daniel P.~W. Ellis, Jort~F. Gemmeke, Aren
-  Jansen, R.~Channing Moore, Manoj Plakal, Devin Platt, Rif~A. Saurous, Bryan
-  Seybold, Malcolm Slaney, Ron~J. Weiss, and Kevin Wilson.
-\newblock CNN architectures for large-scale audio classification.
-\newblock IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP), 2017.
-
-\bibitem{Hochreiter97}
-Sepp Hochreiter and J\"{u}rgen Schmidhuber.
-\newblock Long short-term memory.
-\newblock {\em Neural Comput.}, 9(8):1735–1780, November 1997.
-
-\bibitem{Kitahara05}
-Tetsuro Kitahara, Masataka Goto, and Hiroshi~G. Okuno.
-\newblock Pitch-dependent identification of musical instrument sounds.
-\newblock Applied Intelligence 23, 267–275, Springer, 2005.
-
-\bibitem{lostanlen16}
-Vincent Lostanlen and Carmine-Emanuele Cella.
-\newblock Deep convolutional networks on the pitch spiral for musical
-  instrument recognition. ISMIR, 2016.
-
-\bibitem{mcfee15}
-Brian McFee, Colin Raffel, Dawen Liang, Daniel~P.W. Ellis, Matt McVicar, Eric
-  Battenberg, and Oriol Nieto.
-\newblock librosa: Audio and music signal analysis in python.
-\newblock {P}roceedings of the 14th {P}ython in {S}cience {C}onference, 2015.
-
-\bibitem{scikit-learn}
-Fabian Pedregosa, Ga\"el Varoquaux, Alexandre
-Gramfort, Vincent Michel, Bertrand Thirion, Olivier
-Grisel, Mathieu Blondel, Peter Prettenhofer, Ron
-Weiss, Vincent Dubourg, Jake Vanderplas, Alexandre
-Passos, David Cournapeau, Matthieu Brucher,
-Matthieu Perrot, \'Edouard Duchesnay.
-\newblock Scikit-learn: Machine learning in {P}ython.
-\newblock {\em Journal of Machine Learning Research}, 12:2825--2830, 2011.
-
-\bibitem{Salamon17}
-Justin Salamon and Juan~Pablo Bello.
-\newblock Deep convolutional neural networks and data augmentation for
-  environmental sound classification.
-\newblock {\em IEEE Signal Processing Letters}, 24(3):279–283, Mar 2017.
-
-\bibitem{McKinney2003}
-Martin F. McKinney and Jeroen Breebaart.
-\newblock Features for Audio and Music Classification.
-\newblock ISMIR, 2003.
-
-\bibitem{Cella18}
-Carmine-Emanuele Cella and Philippe Esling, {Open-source modular toolbox for computer-aided orchestration}, Proc. of the Timbre conference, Montreal, Canada, 2018.
-
-\bibitem{Gillick19}
-Jon Gillick, Carmine-Emanuele Cella and David Bamman. {Estimating Unobserved Audio Features for Target-Based Orchestration}, ISMIR, 2019.
-
-\end{thebibliography}
+%
+%
+%\bibitem{Benetos07}
+%Emmanouil Benetos, Margarita Kotti, and Constantine Kotropoulos.
+%\newblock Large scale musical instrument identification.
+%\newblock Sound and Music Computing Conference, Greece, 2007.
+%
+%\bibitem{Bhardwaj17}
+%Siddharth Bhardwaj.
+%\newblock Audio data augmentation with respect to musical instrument
+%  recognition.
+%\newblock Master's thesis, 2017.
+%
+%\bibitem{Bian19}
+%Wenhao Bian, Jie Wang, Bojin Zhuang, Shaojun~Wang Jiankui~Yang, and Jing Xiao.
+%\newblock Audio-based music classification with densenet and data augmentation.
+%\newblock arXiv preprint, https://arxiv.org/abs/1906.11620v1, 2019.
+%
+%\bibitem{Cella2020}
+%Carmine-Emanuele Cella.
+%\newblock Orchidea: a comprehensive framework for target-based assisted orchestration.
+%\newblock In preparation; available on request from the author, 2020.
+%
+%\bibitem{Cella2020b}
+%Carmine-Emanuele Cella, Daniele Ghisi, Vincent Lostanlen, Fabi\'en Levy, Yan Maresz and Joshua Fineberg.
+%\newblock OrchideaSOL: a dataset of extended instrumental techniques for computer-aided orchestration.
+%\newblock ICMC (under review), 2020.
+%
+%\bibitem{Chetry05}
+%Nicolas Chetry, Mike Davies, and Mark Sandler.
+%\newblock Musical instrument identification using LSF and K-means.
+%\newblock Audio Engineering Society Convention 118, 2005.
+%
+%\bibitem{He15}
+%Kaiming He, Xiangyu Zhang, Shaoqing Ren, and Jian Sun.
+%\newblock Deep residual learning for image recognition.
+%\newblock arXiv preprint, https://arxiv.org/abs/1512.03385, 2015.
+%
+%\bibitem{Hershey17}
+%Shawn Hershey, Sourish Chaudhuri, Daniel P.~W. Ellis, Jort~F. Gemmeke, Aren
+%  Jansen, R.~Channing Moore, Manoj Plakal, Devin Platt, Rif~A. Saurous, Bryan
+%  Seybold, Malcolm Slaney, Ron~J. Weiss, and Kevin Wilson.
+%\newblock CNN architectures for large-scale audio classification.
+%\newblock IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP), 2017.
+%
+%\bibitem{Hochreiter97}
+%Sepp Hochreiter and J\"{u}rgen Schmidhuber.
+%\newblock Long short-term memory.
+%\newblock {\em Neural Comput.}, 9(8):1735–1780, November 1997.
+%
+%\bibitem{Kitahara05}
+%Tetsuro Kitahara, Masataka Goto, and Hiroshi~G. Okuno.
+%\newblock Pitch-dependent identification of musical instrument sounds.
+%\newblock Applied Intelligence 23, 267–275, Springer, 2005.
+%
+%\bibitem{lostanlen16}
+%Vincent Lostanlen and Carmine-Emanuele Cella.
+%\newblock Deep convolutional networks on the pitch spiral for musical
+%  instrument recognition. ISMIR, 2016.
+%
+%\bibitem{mcfee15}
+%Brian McFee, Colin Raffel, Dawen Liang, Daniel~P.W. Ellis, Matt McVicar, Eric
+%  Battenberg, and Oriol Nieto.
+%\newblock librosa: Audio and music signal analysis in python.
+%\newblock {P}roceedings of the 14th {P}ython in {S}cience {C}onference, 2015.
+%
+%\bibitem{scikit-learn}
+%Fabian Pedregosa, Ga\"el Varoquaux, Alexandre
+%Gramfort, Vincent Michel, Bertrand Thirion, Olivier
+%Grisel, Mathieu Blondel, Peter Prettenhofer, Ron
+%Weiss, Vincent Dubourg, Jake Vanderplas, Alexandre
+%Passos, David Cournapeau, Matthieu Brucher,
+%Matthieu Perrot, \'Edouard Duchesnay.
+%\newblock Scikit-learn: Machine learning in {P}ython.
+%\newblock {\em Journal of Machine Learning Research}, 12:2825--2830, 2011.
+%
+%\bibitem{Salamon17}
+%Justin Salamon and Juan~Pablo Bello.
+%\newblock Deep convolutional neural networks and data augmentation for
+%  environmental sound classification.
+%\newblock {\em IEEE Signal Processing Letters}, 24(3):279–283, Mar 2017.
+%
+%\bibitem{McKinney2003}
+%Martin F. McKinney and Jeroen Breebaart.
+%\newblock Features for Audio and Music Classification.
+%\newblock ISMIR, 2003.
+%
+%\bibitem{Cella18}
+%Carmine-Emanuele Cella and Philippe Esling, {Open-source modular toolbox for computer-aided orchestration}, Proc. of the Timbre conference, Montreal, Canada, 2018.
+%
+%\bibitem{Gillick19}
+%Jon Gillick, Carmine-Emanuele Cella and David Bamman. {Estimating Unobserved Audio Features for Target-Based Orchestration}, ISMIR, 2019.
+%
+%\end{thebibliography}
 

### 2020-08-18 paper/aimusic2020/paper.tex
Fixed errors and merged figures to reduce length

@@
 \usepackage{amssymb}
+\usepackage{amsmath}
 \setcounter{tocdepth}{3}
 \usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 \usepackage{url}
@@
 
-%The instruments and ranges over which they were recorded are summarized in table \ref{tab:ranges}.
+The instruments and ranges over which they were recorded are summarized in table \ref{tab:ranges}.
 
@@
 
-%CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey17}.
+CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey17}.
 LSTM units provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.4]{../latex/figs/Acc_vs_N_CNN.png}
-\caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM depending on the number of instruments in the combinations. \label{best_acc_cnn}}
-\end{center}
-\end{figure}
-
 During training, the loss function used to optimize the inner parameters of the model was binary cross entropy, as it is the common choice for multiclass multilabel classification frameworks.
@@
 
-%\begin{equation}
-%M_N(\hat{X})_i = \left\{\begin{array}{ll}
-%1 \text{ if } i \in I_N(\hat{X})\\
-%0 \text{ otherwise}
-%\end{array}\right.
-%\label{NMax}
-%\end{equation}
+\begin{equation}
+M_N(\hat{X})_i = \left\{\begin{array}{ll}
+1 \text{ if } i \in I_N(\hat{X})\\
+0 \text{ otherwise}
+\end{array}\right.
+\label{NMax}
+\end{equation}
 and $I_N(\hat{X})$ is the set of indices containing the $N$ first maximums of the vector $\hat{X}$.
@@
 
-%Fig.
+Fig.
 \ref{best_acc_cnn} and Fig.
 \ref{best_acc_resnet} show the maximum accuracy computed using the function $f$ from Eqn.
-\eqref{eval}.
+\ref{eval}.
 For ResNet, the variance in accuracy is much smaller until $N$ reaches 5, at which point it becomes similar to the CNN.
@@
 \begin{figure}
-\begin{center}
-\includegraphics[scale=0.40]{../latex/figs/Acc_vs_N_ResNet.png}
-\caption{Best overall accuracy and for each instrument obtained by the ResNet depending on the number of instruments in the combinations. \label{best_acc_resnet}}
-\end{center}
+\centering
+\begin{subfigure}{.65\textwidth}
+  \centering
+  \includegraphics[width=0.9\linewidth]{../latex/figs/Acc_vs_N_CNN.png}
+  \caption{CNN with LSTM}
+  \label{best_acc_cnn}
+\end{subfigure}%
+\begin{subfigure}{.65\textwidth}
+  \centering
+  \includegraphics[width=0.9\linewidth]{../latex/figs/Acc_vs_N_ResNet.png}
+  \caption{ResNet}
+  \label{best_acc_resnet}
+\end{subfigure}
+\caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM and the ResNet depending on the number of instruments in the combinations}
 \end{figure}
@@
   \end{center}
-%  \caption{Quantitative comparison of orchestrations as ratios to Orchidea.
+  \caption{Quantitative comparison of orchestrations as ratios to Orchidea.
 Eqn.
-\eqref{distance} was used with $\lambda_1 = 0.5$ and $\lambda_2 = 10$ to compute distances between orchestrations and targets.
+\ref{distance} was used with $\lambda_1 = 0.5$ and $\lambda_2 = 10$ to compute distances between orchestrations and targets.
 What is shown is the ratio between the distance calculated and Orchidea's distance for the same target.
@@
 
-%For quantitative evaluation, we used the distance metric defined in Eqn.
-\eqref{distance} to calculate differences in timbre between targets and solutions.
+For quantitative evaluation, we used the distance metric defined in Eqn.
+\ref{distance} to calculate differences in timbre between targets and solutions.
 This metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization.
@@
 where 
-%$
-%\delta_{k1} = 
-%\begin{cases}
-%1, \text{if   } x_k \ge \tilde{x}_k \\
-%0, \text{otherwise}
-%\end{cases} 
-%$, $
-%\delta_{k2} = 
-%\begin{cases}
-%1, \text{if   } x_k < \tilde{x}_k \\
-%0, \text{otherwise}
-%\end{cases}.
-%$\\
+$
+\delta_{k1} = 
+\begin{cases}
+1, \text{if   } x_k \ge \tilde{x}_k \\
+0, \text{otherwise}
+\end{cases} 
+$, $
+\delta_{k2} = 
+\begin{cases}
+1, \text{if   } x_k < \tilde{x}_k \\
+0, \text{otherwise}
+\end{cases}.
+$\\
 

### 2020-08-18 paper/aimusic2020/paper.tex
Add authors, titles and minor formatting changes

@@
 \usepackage{subcaption}
+\captionsetup{compatibility=false}
 \usepackage{url}
@@
 % first the title is needed
-\title{Neural Models for Target-Based Computer-Assisted Musical Orchestration}
+\title{Neural Models for Target-Based Computer-Assisted Musical Orchestration: A Preliminary Study}
 
 % a short form should be given in case it is too long for the running head
-\titlerunning{Running Title}
+\titlerunning{Neural Models for Target-Based Computer-Assisted Musical Orchestration}
 
@@
 %
-\author{AuthorA\inst{1}\and AuthorB\inst{2} \thanks{Please place acknowledgement here.}}
+\author{Carmine Emanuele Cella\inst{1}\and Luke Dzwonczyk\inst{1}\and Alejandro Saldarriaga-Fuertes\inst{1}\and Hongfu Liu\inst{2}\and Helene-Camille Crayencour\inst{3} \thanks{Please place acknowledgement here.}}
 %
 % if the names of the authors are too long for the running head, please use the format: AuthorA et al.
-\authorrunning{AuthorA and AuthorB (or AuthorA et al. if too long)}
+\authorrunning{Carmine Emanuele Cella et al.}
 
@@
 % unless you accept that it will be published
-\institute{InstituteA \and InstituteB\\ \email{author@phdsrus.ac.uk}}
+\institute{Unviersity of California, Berkeley \and InstituteB\\ \email{author@phdsrus.ac.uk}}
 
@@
 \begin{figure}
+	\centering
 	\includegraphics[scale=0.17]{../latex/figs/method.png}
@@
 \begin{figure}
-\begin{center}
+	\centering
 	\includegraphics[scale=0.22]{../latex/figs/CNN_LSTM_diagram.png}
 	\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
-\end{center}
 \end{figure}
@@
 
-%\begin{table}
-%\begin{center}
-%\begin{tabular}{|c|c|c|}
-%\hline
-%\textbf{Instrument} & \textbf{Abbreviation} & \textbf{Range} \\
-%\hline
-%Violin & Vn & G3-E7\\
-%Cello & Vc & C2-C6\\
-%Viola & Va & C3-C7\\
-%Trumpet C & TpC & F\#3-C6\\
-%Trombone & Tbn & A\#0;A\#1-C5\\
-%Oboe & Ob & A\#3-A6\\
-%Horn & Hn & G1-F5\\
-%Flute & Fl & B3-D7\\
-%Clarinet Bb & ClBb & D3-G6\\
-%Contrabass & Cb & E1-C5\\
-%Bass Tuba & BTb & F\#1-F4\\
-%Bassoon & Bn & A\#1-D\#5\\
-%\hline
-%\end{tabular}
-%\caption{Table showing the pitch ranges present in the TinySOL database.\label{tab:ranges}}
-%\end{center}
-%\end{table}
-
 \subsection{Data augmentation}
@@
 \begin{table}
-  \begin{center}
+  \centering
+  \caption{Comparison of accuracies between SVM and RF.
+Each data point is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment.
+For the samples drawn from one of the two instruments, the pitch class of that sample is identified.
+For a sample not from one of the two instruments, the classifier simply attempted to identify that a sample from one of the non-specified instruments was present.
+In this setting, there are 25 classes: 24 from the 12 pitch classes from 2 instruments, and 1 class that specified whether an instrument that is not one of the two specified is present.
+\label{tab:baselines}}
     \begin{tabular}{|c|c|c|c|} 
@@
     \end{tabular}
-  \end{center}
-  \caption{Comparison of accuracies between SVM and RF.
-Each data point is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment.
-For the samples drawn from one of the two instruments, the pitch class of that sample is identified.
-For a sample not from one of the two instruments, the classifier simply attempted to identify that a sample from one of the non-specified instruments was present.
-In this setting, there are 25 classes: 24 from the 12 pitch classes from 2 instruments, and 1 class that specified whether an instrument that is not one of the two specified is present.
-\label{tab:baselines}}
 \end{table}
@@
 \begin{table*}
-  \begin{center}
-    \label{orch_eval}
+  \centering
+     \caption{Quantitative comparison of orchestrations as ratios to Orchidea.
+Eqn.
+\ref{distance} was used with $\lambda_1 = 0.5$ and $\lambda_2 = 10$ to compute distances between orchestrations and targets.
+What is shown is the ratio between the distance calculated and Orchidea's distance for the same target.
+A value less than 1 means that the model performed worse (i.e.
+had a larger distance), and a value greater than 1 means the model performed better than Orchidea.
+The last column shows the ratio of the average distances for the model across all targets.}\label{orch_eval}
     \begin{tabular}{|c|c|c|c|c|c|c|c|c|c}
@@
     \end{tabular}
-  \end{center}
-  \caption{Quantitative comparison of orchestrations as ratios to Orchidea.
-Eqn.
-\ref{distance} was used with $\lambda_1 = 0.5$ and $\lambda_2 = 10$ to compute distances between orchestrations and targets.
-What is shown is the ratio between the distance calculated and Orchidea's distance for the same target.
-A value less than 1 means that the model performed worse (i.e.
-had a larger distance), and a value greater than 1 means the model performed better than Orchidea.
-The last column shows the ratio of the average distances for the model across all targets.}
+
 \end{table*}

### 2020-08-18 paper/aimusic2020/paper.tex
remove old bibliography
Cutting out some things

@@
 
+\renewcommand{\baselinestretch}{0.96}
 \pagestyle{headings}
@@
 
-The codebase for this paper and some audio examples can be found at: \url{https://github.com/dzluke/DeepOrchestration}.
+The codebase for this paper can be found at: \url{https://github.com/dzluke/DeepOrchestration}.
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[scale=0.22]{../latex/figs/CNN_LSTM_diagram.png}
-	\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
-\end{figure}
+%\begin{figure}
+%	\centering
+%	\includegraphics[scale=0.22]{../latex/figs/CNN_LSTM_diagram.png}
+%	\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
+%\end{figure}
 
@@
 
-For a given number of instruments $N$, each input to our models is a combination of $N$ TinySOL samples chosen among an orchestra of 10 instruments. The data is generated in the following way:
-
-\begin{itemize}
-\item $N$ random TinySOL samples are selected, leading to a variety of instruments, pitches, and dynamics; we did not allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures;
-\item the chosen samples are combined to be played simultaneously and normalized by the number of instruments;
-\item the Mel frequency cepstral coefficients (MFCCs) and the Mel spectrogram of the mixture are computed to be used as the input features for the baseline classifiers and the neural classifiers respectively.
-\end{itemize}
-
+For a given number of instruments $N$, each input to our models is a combination of $N$ TinySOL samples chosen among an orchestra of 10 instruments.
+The data is generated in by selecting $N$ random TinySOL samples, leading to a variety of instruments, pitches, and dynamics; we did not allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures.
+Then the chosen samples are combined to be played simultaneously and normalized by the number of instruments and the Mel frequency cepstral coefficients (MFCCs) and the Mel spectrogram of the mixture are computed to be used as the input features for the baseline classifiers and the neural classifiers respectively.
 The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly 4 seconds long.
@@
 
-The complete problem of classification of the pairs instrument-pitch has 424 total classes.
-This problem is very hard for parametric classifiers, so we initially switched to the simpler problem of classifying only the instruments and not the pitch.
-This had the benefit of both reducing the number of classes and increasing the number of samples per class.
-We found that SVM was able to very accurately identify the instrument given an input that had only one instrument present; for this case the accuracy was 99.8\%.
-However, as soon as multiple instruments were present in the input, the accuracy dropped significantly.
-With two instruments, accuracy was 55.4\%, with three it was 19.6\%.
-KNN performed significantly worse than SVM, so we did not attempt any further testing with it.
-
-To better approximate our original problem of identifying instrument and pitch, however, we performed experiments in which two instruments were selected and for input data that contained samples from one of those two instruments, both the instrument and pitch \emph{class} of the sample would be classified.
+We started with a simplification of our problem.
+We performed experiments in which two instruments were selected, and the classifier attempted to identify instrument and pitch \emph{class} of the samples.
 The pitch class is the note name without the octave, e.g.
 C, D\#, G, etc.
-The input was a combination of two instruments drawn from a possible twelve instruments, and the classifier attempted to identify which instruments were present and for the specified instruments, say Violin and Viola, which pitch classes were present.
-If another instrument was present in the input combination that was not Violin or Viola, the classifier would simply identify that an instrument that was not one of the two was present.
-The accuracies from this experiment are outlined in Table \ref{tab:baselines}.
+The best accuracies from these experiments were from a combination of Oboe and French Horn, in which SVM had an accuracy of 39.9\% and RF had an accuracy of 17.5\%.
 Since RF performed worse than SVM in every experiment, we stopped testing with it and used SVM from that point on.
@@
 
-\begin{table}
-  \centering
-  \caption{Comparison of accuracies between SVM and RF.
+%\begin{table}
+%  \centering
+%  \caption{Comparison of accuracies between SVM and RF.
 Each data point is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment.
@@
 \label{tab:baselines}}
-    \begin{tabular}{|c|c|c|c|} 
-    	  \hline
-      \textbf{Instr. 1} & \textbf{Instr. 2} & \textbf{SVM Acc.} & \textbf{RF Acc.}\\
-      \hline
-      Violin & Flute & 38.8\% & 9.8\% \\
-      \hline
-      Violin & Trumpet & 33.8\% & 9.1\% \\
-      \hline
-      Violin & Cello & 34.8\% & 6.3\% \\
-      \hline
-      Cello & Viola & 32.1\% & 5.8\% \\
-      \hline
-      Oboe & French Horn & \textbf{39.9\%} & \textbf{17.5\%} \\
-      \hline
-    \end{tabular}
-\end{table}
+%    \begin{tabular}{|c|c|c|c|} 
+%    	  \hline
+%      \textbf{Instr. 1} & \textbf{Instr. 2} & \textbf{SVM Acc.} & \textbf{RF Acc.}\\
+%      \hline
+%      Violin & Flute & 38.8\% & 9.8\% \\
+%      \hline
+%      Violin & Trumpet & 33.8\% & 9.1\% \\
+%      \hline
+%      Violin & Cello & 34.8\% & 6.3\% \\
+%      \hline
+%      Cello & Viola & 32.1\% & 5.8\% \\
+%      \hline
+%      Oboe & French Horn & \textbf{39.9\%} & \textbf{17.5\%} \\
+%      \hline
+%    \end{tabular}
+%\end{table}
 
@@
 Fig.
-\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$.\\
+\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$.
 
@@
 The last column shows the ratio of the average distances for the model across all targets.}\label{orch_eval}
+	\centering    
     \begin{tabular}{|c|c|c|c|c|c|c|c|c|c}
@@
 For example, it is able to apply symbolic constraints to the search, hence allowing only specific instrumental combinations or playing styles.
-It is also able to reduce the search space by applying harmonic analysis on the target sound.
-The dominant harmonic partials of the target are identified, and the search space is limited to only include samples of those pitches.
-For example, if the target is a recording of an instrument playing a C4, then the partials identified may be C4, C5, G5, and E6.
-The model would then only consider samples of these pitches to be used in the solution.
-This leads to a solution whose harmonics are much closer to the target, which is an important part of aural similarity.
-Orchidea's solutions, moreover, can use any number of instruments included in the orchestra specified by the user, thus having variable-sized orchestrations from a single instrument to the whole set (this property is called \emph{sparsity} of the solution).
+It is also able to reduce the search space by applying harmonic analysis on the target sound, and create sparse solutions that do not use every instrument specified.
 
@@
 For example, one of the samples of a bell had partials that loosely represented a C minor chord, so we checked if the orchestration contained the notes of the chord.
-If a target had specific notes, we identified if the note or its partials were present.
-For example there is target of an Oboe playing an A4 and a Bassoon playing a C\#3, the solution from ResNet contained the partials of the Bassoon's note: C\#3, G\#4, C\#5, and E6 were all included on various instruments in the solution.
 
@@
 
-In the first summation, $\lambda_1$ is multiplied by all the distances calculated when there was more energy in the target than the solution, since $\delta_{k1} = 1$ only when $x_k \ge \tilde{x}_k$.
-Similarly in the second summation, $\lambda_2$ is multiplied by all the distances when the solution provided more energy to a frequency than the target.
-Therefore, the relation between $\lambda_1$ and $\lambda_2$ determines whether a solution is penalized more for undershooting or overshooting the target.
-
 We calculated the distance between the target samples and our orchestrated solutions.
@@
 \end{equation}
-where 
-$
-\delta_{k1} = 
-\begin{cases}
-1, \text{if   } x_k \ge \tilde{x}_k \\
-0, \text{otherwise}
-\end{cases} 
-$, $
-\delta_{k2} = 
-\begin{cases}
-1, \text{if   } x_k < \tilde{x}_k \\
-0, \text{otherwise}
-\end{cases}.
-$\\
+where $\delta_{k1} = 1 \text{  if  } x_k \ge \tilde{x}_k, 0 \text{  otherwise}$; and $\delta_{k2} = 1 \text{  if  } x_k < \tilde{x}_k, 0 \text{  otherwise}$.
 
@@
 Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.
+You can listen to the targets and orchestrated solutions from Orchidea, ResNet, and the CNN with LSTM at \url{https://dzluke.github.io/DeepOrchestration/}.
 
@@
 
-\iffalse
-\begin{figure}
-\includegraphics[scale=0.6]{figs/CNN10.png}
-\end{figure}
-\fi
-
 \bibliographystyle{apacite}
 \bibliography{paper}
-% For bibtex users:
-%\bibliography{paper}
-%\bibliographystyle{apalike}
-%\begin{thebibliography}{99}
-%
-%\bibitem{FerVic2013} Jose David Fern\'andez and Francisco Vico.
-{AI Methods in Algorithmic Composition: A Comprehensive Survey}.
-Journal of Artificial Intelligence Research, vol.48, pp.
-513--582, 2013.
-%
-%\bibitem{Ari2005} Christopher Ariza, {Navigating the landscape of computer aided algorithmic composition systems: A definition, seven descriptors, and a lexicon of systems and research}.
-In Proceedings of the International Computer Music Conference, 2005.
-%
-%\bibitem{Maresz2003}
-%Yan Maresz, {On Computer-Assisted Orchestration}. Contemporary Music Review, vol. 32, n. 1, pg. 99-109, Routledge, 2013.
-%
-%\bibitem{Carpentier2010}
-%Gr\'egoire Carpentier, Damien Tardieu, Jonathan Harvey, G\'erard Assayag and Emmanuel Saint-James.
-{Predicting Timbre Features of Instrument Sound Combinations: Application to Automatic Orchestration}, Journal of New Music Research, vol.
-39, n.
-1, pg.
-47-61, Routlegde, 2010.
-%
-%
-%\bibitem{Benetos07}
-%Emmanouil Benetos, Margarita Kotti, and Constantine Kotropoulos.
-%\newblock Large scale musical instrument identification.
-%\newblock Sound and Music Computing Conference, Greece, 2007.
-%
-%\bibitem{Bhardwaj17}
-%Siddharth Bhardwaj.
-%\newblock Audio data augmentation with respect to musical instrument
-%  recognition.
-%\newblock Master's thesis, 2017.
-%
-%\bibitem{Bian19}
-%Wenhao Bian, Jie Wang, Bojin Zhuang, Shaojun~Wang Jiankui~Yang, and Jing Xiao.
-%\newblock Audio-based music classification with densenet and data augmentation.
-%\newblock arXiv preprint, https://arxiv.org/abs/1906.11620v1, 2019.
-%
-%\bibitem{Cella2020}
-%Carmine-Emanuele Cella.
-%\newblock Orchidea: a comprehensive framework for target-based assisted orchestration.
-%\newblock In preparation; available on request from the author, 2020.
-%
-%\bibitem{Cella2020b}
-%Carmine-Emanuele Cella, Daniele Ghisi, Vincent Lostanlen, Fabi\'en Levy, Yan Maresz and Joshua Fineberg.
-%\newblock OrchideaSOL: a dataset of extended instrumental techniques for computer-aided orchestration.
-%\newblock ICMC (under review), 2020.
-%
-%\bibitem{Chetry05}
-%Nicolas Chetry, Mike Davies, and Mark Sandler.
-%\newblock Musical instrument identification using LSF and K-means.
-%\newblock Audio Engineering Society Convention 118, 2005.
-%
-%\bibitem{He15}
-%Kaiming He, Xiangyu Zhang, Shaoqing Ren, and Jian Sun.
-%\newblock Deep residual learning for image recognition.
-%\newblock arXiv preprint, https://arxiv.org/abs/1512.03385, 2015.
-%
-%\bibitem{Hershey17}
-%Shawn Hershey, Sourish Chaudhuri, Daniel P.~W. Ellis, Jort~F. Gemmeke, Aren
-%  Jansen, R.~Channing Moore, Manoj Plakal, Devin Platt, Rif~A. Saurous, Bryan
-%  Seybold, Malcolm Slaney, Ron~J. Weiss, and Kevin Wilson.
-%\newblock CNN architectures for large-scale audio classification.
-%\newblock IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP), 2017.
-%
-%\bibitem{Hochreiter97}
-%Sepp Hochreiter and J\"{u}rgen Schmidhuber.
-%\newblock Long short-term memory.
-%\newblock {\em Neural Comput.}, 9(8):1735–1780, November 1997.
-%
-%\bibitem{Kitahara05}
-%Tetsuro Kitahara, Masataka Goto, and Hiroshi~G. Okuno.
-%\newblock Pitch-dependent identification of musical instrument sounds.
-%\newblock Applied Intelligence 23, 267–275, Springer, 2005.
-%
-%\bibitem{lostanlen16}
-%Vincent Lostanlen and Carmine-Emanuele Cella.
-%\newblock Deep convolutional networks on the pitch spiral for musical
-%  instrument recognition. ISMIR, 2016.
-%
-%\bibitem{mcfee15}
-%Brian McFee, Colin Raffel, Dawen Liang, Daniel~P.W. Ellis, Matt McVicar, Eric
-%  Battenberg, and Oriol Nieto.
-%\newblock librosa: Audio and music signal analysis in python.
-%\newblock {P}roceedings of the 14th {P}ython in {S}cience {C}onference, 2015.
-%
-%\bibitem{scikit-learn}
-%Fabian Pedregosa, Ga\"el Varoquaux, Alexandre
-%Gramfort, Vincent Michel, Bertrand Thirion, Olivier
-%Grisel, Mathieu Blondel, Peter Prettenhofer, Ron
-%Weiss, Vincent Dubourg, Jake Vanderplas, Alexandre
-%Passos, David Cournapeau, Matthieu Brucher,
-%Matthieu Perrot, \'Edouard Duchesnay.
-%\newblock Scikit-learn: Machine learning in {P}ython.
-%\newblock {\em Journal of Machine Learning Research}, 12:2825--2830, 2011.
-%
-%\bibitem{Salamon17}
-%Justin Salamon and Juan~Pablo Bello.
-%\newblock Deep convolutional neural networks and data augmentation for
-%  environmental sound classification.
-%\newblock {\em IEEE Signal Processing Letters}, 24(3):279–283, Mar 2017.
-%
-%\bibitem{McKinney2003}
-%Martin F. McKinney and Jeroen Breebaart.
-%\newblock Features for Audio and Music Classification.
-%\newblock ISMIR, 2003.
-%
-%\bibitem{Cella18}
-%Carmine-Emanuele Cella and Philippe Esling, {Open-source modular toolbox for computer-aided orchestration}, Proc. of the Timbre conference, Montreal, Canada, 2018.
-%
-%\bibitem{Gillick19}
-%Jon Gillick, Carmine-Emanuele Cella and David Bamman. {Estimating Unobserved Audio Features for Target-Based Orchestration}, ISMIR, 2019.
-%
-%\end{thebibliography}
 

### 2020-08-19 paper/aimusic2020/paper.tex
Fix small typos

@@
 
-The second and more deep model that we trained as classifier was the well known deep residual network \emph{ResNet}.
+The second and deeper model that we trained as classifier was the well known deep residual network \emph{ResNet}.
 Specifically, we used 18-layer ResNet, which allows information to pass directly from the input to the final layer of each block.
@@
 
-In order to have a more fair comparison, we did not allow Orchidea to use any of the advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it forced to use all 10 instruments in each solution.
+In order to have a fairer comparison, we did not allow Orchidea to use any of the advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it forced to use all 10 instruments in each solution.
 This creates a more fair comparison, since our models are unable to create constrained or sparse solutions.

### 2020-08-19 paper/aimusic2020/paper.tex
Fixed some typos and added justification for structure and data augmentation choices

@@
 % unless you accept that it will be published
-\institute{Unviersity of California, Berkeley \and InstituteB\\ \email{author@phdsrus.ac.uk}}
+\institute{University of California, Berkeley \and InstituteB\\ \email{author@phdsrus.ac.uk}}
 
@@
 We then compared our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means.
-For quantitative comparison, we computed a specific distance metric on the FFTs of the target and solution, comparing the distance between Orchidea's solution and the target and our model's solution and the target.
+For quantitative comparison, we computed a specific distance metric on the amplitude spectrum of the target and solution, comparing the distance between Orchidea's solution and the target and our model's solution and the target.
 
@@
 
-The instruments and ranges over which they were recorded are summarized in table \ref{tab:ranges}.
+%The instruments and ranges over which they were recorded are summarized in table \ref{tab:ranges}.
 
@@
 The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed.
+Such setup proved to be successful in \cite{Salamon17}.
 
@@
 
-The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit.
+The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit, whose structure is inspired by the one successfully used in \cite{Salamon17}.
+The LSTM unit was added in order to provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.
 
-CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey17}.
+%CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey17}.
 LSTM units provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.
@@
 
-Following the first three convolutional layers, there is an LSTM layer with 32 outputs.
+Following the first three convolutional layers, there is an LSTM layer which outputs 32 matrices.
 After the LSTM layer, there is a final convolutional layer yielding a tensor of dimensions $32\times 8 \times 21$.
@@
 
-Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class. The model architecture is shown in Figure \ref{cnn_lstm}.
+Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class. %The model architecture is shown in Figure \ref{cnn_lstm}.
 
@@
 \end{equation}
-
-where
-
 \begin{equation}
@@
 This metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization.
-The equation takes in the full FFT of the target $x$ and full FFT of the solution $\hat{x}$.
-Then for each bin $k$ of the FFT, it calculates the absolute difference between the values.
+The equation takes in the full amplitude spectrum of the target $x$ and full amplitude spectrum of the solution $\hat{x}$.
+Then for each bin $k$ of the amplitude spectrum, it calculates the absolute difference between the values.
 The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.

### 2020-08-19 paper/aimusic2020/paper.tex
Reduced length of paper

@@
 
-\begin{figure}
-\center{
-\includegraphics[scale=0.4]{../latex/figs/CNN_vs_ResNet.png}}
-\caption{Best overall accuracy for CNN with LSTM and ResNet depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
-\end{figure}
-
 In other words, the classifiers learn how to take a complex combination of pitches and timbres and deconstruct it into its original parts.
@@
 We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}.
-The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed.
+%The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed.
 Such setup proved to be successful in \cite{Salamon17}.
@@
 After the LSTM layer, there is a final convolutional layer yielding a tensor of dimensions $32\times 8 \times 21$.
-We flatten the outputs and feed them into fully a connected layer with Dropout, then another another fully connected layer.
+We flatten the outputs and feed them into a fully connected layer with Dropout, then another another fully connected layer.
 Finally, the sigmoid function is applied to the final layer.
@@
 However, the value of the loss function alone is difficult to interpret.
-For this reason we created a complementary function $f$ to be used for evaluation only.
+%For this reason we created a complementary function $f$ to be used for evaluation only.
 This function compares a vector of expected outputs $\overline{X}$ with the estimated output from the model $\hat{X}$ by using the following function
-
-\begin{equation}
-f(\overline{X}, \hat{X}) = \frac{1}{N}<\overline{X}, M_N(\hat{X})>
-\label{eval}
-\end{equation}
-\begin{equation}
-M_N(\hat{X})_i = \left\{\begin{array}{ll}
-1 \text{ if } i \in I_N(\hat{X})\\
-0 \text{ otherwise}
-\end{array}\right.
-\label{NMax}
-\end{equation}
-and $I_N(\hat{X})$ is the set of indices containing the $N$ first maximums of the vector $\hat{X}$.
-
+%
+%\begin{equation}
+%f(\overline{X}, \hat{X}) = \frac{1}{N}<\overline{X}, M_N(\hat{X})>
+%\label{eval}
+%\end{equation}
+%\begin{equation}
+%M_N(\hat{X})_i = \left\{\begin{array}{ll}
+%1 \text{ if } i \in I_N(\hat{X})\\
+%0 \text{ otherwise}
+%\end{array}\right.
+%\label{NMax}
+%\end{equation}
+%and $I_N(\hat{X})$ is the set of indices containing the $N$ first maximums of the vector $\hat{X}$.
 More specifically, the function $M_N(\hat{X})$ takes as an input a vector of probabilities and outputs a vector where only the positions of the $N$ first maxima are set to 1.
@@
 Thus, the function $f$ simply outputs the proportion of the estimated orchestration that matches the expected one.
+For this reason, the evaluation was done by comparing the proportion of estimated orchestration samples, chosen among the samples that output the highest probability, that matched the expected orchestration.
 
@@
 Then, for both CNN with LSTM and ResNet, we computed the maximum accuracy over the epochs. 
-The CNN was trained on 200,000 generated samples over 50 epochs, and ResNet was 	trained on 400,000 generated samples over 20 epochs.
+%The CNN was trained on 200,000 generated samples over 50 epochs, and ResNet was 	trained on 400,000 generated samples over 20 epochs.
 Fig.
 \ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$.
-
 ResNet outperforms the CNN regardless of the number of samples used in the combination.
 This result is consistent with previous research \cite{He15}, as residual networks usually perform well in classification problems.
+
+\begin{figure}
+\center{
+\includegraphics[scale=0.4]{../latex/figs/CNN_vs_ResNet.png}}
+\caption{Best overall accuracy for CNN with LSTM (50 epochs, 200k samples per epoch) and ResNet (20 epochs, 400k samples per epoch) depending on the number of instruments in the combinations used for training.
+\label{cnn_vs_resnet}}
+\end{figure}
 
@@
 For example, flute notes tend to exhibit a steep spectral rolloff, with most of the energy captured by the first few partials.
-Moreover, the noisy nature of the transient portions of these note is not well represented by frequency-based descriptions such as Mel spectra.
+Moreover, the noisy nature of the transient portions of these notes is not well represented by frequency-based descriptions such as Mel spectra.
 These two factors combined, could make the disentanglement of the flute from the analyzed combination more difficult.
@@
 
-Orchidea implements many advanced features that are not supported by our models.
+%Orchidea implements many advanced features that are not supported by our models.
 For example, it is able to apply symbolic constraints to the search, hence allowing only specific instrumental combinations or playing styles.
@@
 
-In order to have a fairer comparison, we did not allow Orchidea to use any of the advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it forced to use all 10 instruments in each solution.
-This creates a more fair comparison, since our models are unable to create constrained or sparse solutions.
+In order to have a fairer comparison, we did not allow Orchidea to use any of its advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it forced to use all 10 instruments in each solution.
+%This creates a fairer comparison, since our models are unable to create constrained or sparse solutions.
 

### 2020-08-19 paper/aimusic2020/paper.tex
Minor fixes and formatting
Small change

@@
 
-\subsection{From orchestration to classification}
+\subsection{From Orchestration to Classification}
 
@@
 
-\subsection{Classification results}
+\subsection{Classification Results}
 
@@
 \ref{best_acc_cnn} and Fig.
-\ref{best_acc_resnet} show the maximum accuracy computed using the function $f$ from Eqn.
-\ref{eval}.
+\ref{best_acc_resnet} show the maximum test accuracy for each model.
 For ResNet, the variance in accuracy is much smaller until $N$ reaches 5, at which point it becomes similar to the CNN.
@@
 
-\section{Orchestration experiments}
+\section{Orchestration Experiments}
 \label{sec:orchestration}
@@
 After training the neural models for classification, we finally tested them for the task of target-based computer-assisted orchestration.
-For this purpose, we decided to use the models trained on combinations of 10 instruments, namely French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet.
+For this purpose, we decided to use the models trained on combinations of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet.
 Each class, in this model, represents an instrument-pitch pair.

### 2020-08-21 paper/aimusic2020/paper.tex
Reduce to 8 pages

@@
 \renewcommand{\baselinestretch}{0.96}
+\raggedbottom
 \pagestyle{headings}
@@
 %
-\author{Carmine Emanuele Cella\inst{1}\and Luke Dzwonczyk\inst{1}\and Alejandro Saldarriaga-Fuertes\inst{1}\and Hongfu Liu\inst{2}\and Helene-Camille Crayencour\inst{3} \thanks{Please place acknowledgement here.}}
+\author{Carmine Emanuele Cella\inst{1}\and Luke Dzwonczyk\inst{1}\and Alejandro Saldarriaga-Fuertes\inst{1}\and Hongfu Liu\inst{2}\and Helene-Camille Crayencour\inst{3}}
 %
@@
 % unless you accept that it will be published
-\institute{University of California, Berkeley \and InstituteB\\ \email{author@phdsrus.ac.uk}}
+\institute{University of California, Berkeley \and InstituteB\\ \email{ , dz.luke@berkeley.edu, }}
 
@@
 \begin{abstract}
-In this paper we will do a preliminary exploration on how neural networks can be used for the task of target-based computer-assisted musical orchestration.
+In this paper we will perform a preliminary exploration on how neural networks can be used for the task of target-based computer-assisted musical orchestration.
 We will show how it is possible to model this  musical problem as a classification task and we will propose two deep learning models.
@@
 A solution to this problem is a set of orchestral scores that represent the mixtures of audio samples in the database, ranked by similarity with the target sound.
-Valid orchestral scores may contain several instruments sounding simultaneously, selected out of a large number of possible combinations of sounds from the database.
 
@@
 In this paper, we try a different approach to this problem by experimenting with deep neural architectures.
-The main idea is to train a model to classify combinations of real instruments and use it then for orchestration.
+The main idea is to train a model to classify combinations of real instruments and then use it for orchestration.
 A typical solution for assisted orchestration is a set of triples \emph{instrument-pitch-dynamics} such as \{\texttt{Flute C6 pp, Bassoon C4 mf, Bassoon G4 ff}\}.
@@
 Thus, when an unknown target sound is given as input, the network will identify which are the best instruments to match the target sound, and it will be able to  deconstruct a complex mixture of timbres into individual instrument notes.
-
-%\footnote{The dynamics is retrived using the probability distributions generated as output by the network; more details on this will be given in the next sections.}.
-
 This method is motivated by the good results obtained in previous research on musical instruments identification \cite{Benetos07, Kitahara05} and the more recent use of deep neural networks for musical classification \cite{lostanlen16, Bian19}.
 
-In this paper we do preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit in the middle and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}.
+In this paper we perform preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}.
 We chose to use a CNN because of its success in audio classification \cite{Hershey17} and we decided to include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.
-
-After training these models on varying combinations of 10 instruments, we used them for orchestration.
-More specifically, we orchestrated 15 target sounds by feeding them to the networks and synthesizing a solution from the most likely samples output from the system.
-We then compared our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means.
-For quantitative comparison, we computed a specific distance metric on the amplitude spectrum of the target and solution, comparing the distance between Orchidea's solution and the target and our model's solution and the target.
-
-Next sections will be as follows: section \ref{sec:models} will describe the dataset we used for experiments and the models we trained as classifiers, showing the classification results.
-Sections \ref{sec:orchestration} and \ref{sec:conclusions} will present, respectively, the results of orchestration experiments using the trained models and the conclusions of our preliminary study on neural approaches for target-based computer-assisted orchestration.
 
@@
 
-A complete orchestration solution, however, would normally be made by triples of instrument-pitch-dynamics.
+A complete orchestration solution, however, would normally be made by triples of \textit{instrument-pitch-dynamics}.
 It is difficult to frame this problem as classification, since we would need to have a very high number of classes.
 Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in \ref{sec:dataset}), each class would be represented by a single sample.
-For these reasons, our models have been trained on simplified versions of the original problem and we used different strategies to provide the remaining information needed to generate orchestrations.
-
-In order to have a baseline to compare our classification results and to get a sense on the complexity of the problem, we used various parametric classifiers along with the neural models.
-
-Three baseline classifiers have been trained to recognize either the instrument only for a total of 12 classes (the instruments in TinySOL) or the pair instrument-pitch \emph{class} (a pitch class is the pitch of a note without the corresponding octave, e.g.
-C, D\#, G, etc.) for a total of 25 classes (we will classifiy combinations of 2 instruments for 12 pitch classes each, plus one extra \emph{trash} class, see section \ref{sec:baseline}).
-
-Two neural classifiers, instead, have been trained on the more complex problem of recognizing instrument-pitch pairs, for a total of 424 classes (i.e.
-the neural classifiers would classify instruments and pitches with respective octaves, hence 12 instruments for 12 pitch classes and three dynamics; note that not all the samples in TinySOL have all dynamics).
-We want to point out, however, that this is still a simplified problem compared to assisted orchestration.
-Even in the case of instrument-pitch pair classification, indeed, we would miss the dynamics of the note.
-To cope with this problem, in the orchestration phase, we quantize the probability functions generated by the models.
-When a network is given a target sound as input, it will attempt to apply the same classification rules, outputting a vector of the probabilities of each sample being in the target sound.
-Since the classes only encode instrument-pitch pairs and since we also need the dynamics, we quantize the probability function generated by the network into three echelons coresponding, respectively, to the dynamic levels \emph{pianissimo}, \emph{mezzoforte} and \emph{fortissimo} (see section \ref{sec:orchestration}).
-By taking the instruments and pitches that have the highest probability and by using quantized probabilities to retrieve dynamics, a full orchestrated solution can be created.
-Figure \ref{fig:method} illustrates the described ideas.
+For this reasons, our models were not trained to determine \textit{instrument-pitch-dynamics} triples but instead \textit{instrument-pitch} pairs.
 
@@
 
-For a given number of instruments $N$, each input to our models is a combination of $N$ TinySOL samples chosen among an orchestra of 10 instruments.
-The data is generated in by selecting $N$ random TinySOL samples, leading to a variety of instruments, pitches, and dynamics; we did not allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures.
-Then the chosen samples are combined to be played simultaneously and normalized by the number of instruments and the Mel frequency cepstral coefficients (MFCCs) and the Mel spectrogram of the mixture are computed to be used as the input features for the baseline classifiers and the neural classifiers respectively.
+For a given number of instruments $N$, each input to our model is a combination of $N$ TinySOL samples chosen among an orchestra of 10 instruments.
+The data is generated by selecting $N$ random TinySOL samples, leading to a variety of instruments, pitches, and dynamics; we did not allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures.
+Then the chosen samples are combined to be played simultaneously and normalized by the number of instruments.
+The Mel spectrogram of the mixture is computed to be used as the input features for the classifiers.
 The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly 4 seconds long.
 
-The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an appropriate representation of sound and musical signals.
-The Mel spectrogram is generated using an FFT hop length of 2048 samples (the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128.
+The choice of using the Mel spectrogram as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an appropriate representation of sound and musical signals.
+%The Mel spectrogram is generated using an FFT hop length of 2048 samples (the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128.
 Therefore, the Mel features fed to the model were matrices of size $128\times 345$.
@@
 
-The former has been directly applied on the TinySOL samples each time they are selected to generate a new combination.
+Pitch shifting was applied on the TinySOL samples each time they were selected to generate a new combination.
 We performed a small pitch shift by reading the samples with different sample rates: a small difference in sample rate will slightly modify the duration and the perceived pitch if played at the normal sample rate.
@@
 
-The latter was performed on the feature matrix itself of input samples, the Mel spectrogram.
+Partial feature dropout was performed on the feature matrix itself of input samples, the Mel spectrogram.
 We chose random columns and rows of the matrix to zero out.
-For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$  columns and $3.45$ rows being zero-ed out.
+%For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$  columns and $3.45$ rows being zero-ed out.
 This method of data augmentation aimed to be more resilient to the possible variations in the recording of the instruments.
@@
 We used the implementations provided in the \emph{scikit-learn }library for Python \cite{scikit-learn}.
-More specifically, for SVM we used a non-linear RBF kernel and for RF we set the maximum depth of each tree to be 15.
-All of the baseline experiments used 50,000 generated samples with a train-test split of 60/40.
-Each sample was made by a combination of one to ten instruments and was four seconds in length.
-In this case, differently from the neural models, the features used are the MFCCs of the resulting combination; we chose to use MFCCs for this setting, instead of the Mel spectrogram, to have a lower number of features, that is more manageable by parametric classifiers.
+In this case, differently from the neural models, the features used are the MFCCs of the resulting combination; as the number of features is more manageable for parametric classifiers.
 We found SVM to have the highest accuracy of the three classifiers across all experiments.
@@
 We performed experiments in which two instruments were selected, and the classifier attempted to identify instrument and pitch \emph{class} of the samples.
-The pitch class is the note name without the octave, e.g.
-C, D\#, G, etc.
-The best accuracies from these experiments were from a combination of Oboe and French Horn, in which SVM had an accuracy of 39.9\% and RF had an accuracy of 17.5\%.
-Since RF performed worse than SVM in every experiment, we stopped testing with it and used SVM from that point on.
-
-We then performed this same experiment with three instruments having their pitch class identified.
-Flute, Oboe, and Violin reached an accuracy of 11.1\%, and Bass Tuba, Trumpet, and Trombone was 0.5\%.
-As we increased the number of instruments whose pitch classes was being identified, the accuracy continued to drop.Indeed, for classifying the pitch class of four instruments (Oboe, French Horn, Violin, and Flute) the accuracy was 2.7\%.
-
-This was still a simplified version of the problem, as we were identifying only the pitch class of a few instruments.
-However, the parametric classifiers were unable to achieve accurate results as the number of instruments increased.
-Therefore, we did not attempt, with parametric classifiers, the full setting of the problem in which individual pitches are classified for all instruments.
+For two instruments, SVM had an accuracy of 39.9\% and RF had an accuracy of 17.5\%.
+As the number of instruments used in combination increased, the accuracy dropped sharply.
+With three instruments, SVM accuracy was 11.1\%, with four instruments it was 2.7\%.
+Clearly, these classifiers were not going to be able to achieve meaningful results with the full setting of our problem.
 
@@
 
-In the rest of this section, we will show the classification results of the two neural models.
+%In the rest of this section, we will show the classification results of the two neural models.
 
@@
 
-The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit, whose structure is inspired by the one successfully used in \cite{Salamon17}.
+The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit, whose structure is inspired by the success in \cite{Salamon17}.
 The LSTM unit was added in order to provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.
@@
 Finally, the sigmoid function is applied to the final layer.
-
-Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class. %The model architecture is shown in Figure \ref{cnn_lstm}.
+Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class.
+%The model architecture is shown in Figure \ref{cnn_lstm}.
 
@@
 Specifically, we used 18-layer ResNet, which allows information to pass directly from the input to the final layer of each block.
-Besides, to make the model more suitable to our problem, we decided to reduce the number of parameters and to reset the output channel numbers of each block with 32, 64, 32, 32 respectively.
+To make the model more suitable to our problem, we decided to reduce the number of parameters and to reset the output channel numbers of each block with 32, 64, 32, 32 respectively.
 
@@
 Different experiments were made by varying the number $N$ of samples in each mixture.
-We used an orchestra of 10 instruments containing Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet in C and Clarinet in Bb.
-
+We used an orchestra of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet in C and Clarinet in Bb.
 Then, for both CNN with LSTM and ResNet, we computed the maximum accuracy over the epochs.
@@
 After training the neural models for classification, we finally tested them for the task of target-based computer-assisted orchestration.
-For this purpose, we decided to use the models trained on combinations of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet.
-Each class, in this model, represents an instrument-pitch pair.
 
@@
 
-Since we decided not to train our models to classify the dynamics of a sample (despite TinySOL having \emph{pianissimo}, \emph{mezzoforte}, and \emph{fortissimo} recordings), the dynamics are determined by the probability of each sample as output by the model.
+Since we decided not to train our models to classify the dynamics of a sample, the dynamics are determined by the probability of each sample as output by the model.
 If the model outputs a probability higher than $0.66$ for a sample, the fortissimo version of the sample is used.
-If the probability is between $0.33$ and $0.66$, then the mezzoforte version is selected and if it is less than $0.33$ the pianissimo version is used.
+A probability between $0.33$ and $0.66$, is mezzoforte and less than $0.33$ is pianissimo.
 The idea behind this quantization is that samples that are the most represented in the target should appear as the loudest in the orchestrated solution.
@@
 These targets represent a variety of signal types but are mostly static, in the sense that they do not change sensibly over time.
-
-Some of the targets were made of instrumental samples: a sum of oboe and bassoon notes, single bassoon and bass clarinet notes, two multiphonics of bassoon and two symphonic chords.
-The other targets there are: the recordings of two bells, of a car horn, of a gong, of a screaming voice, of a wind harp, of the attack of a brass instrument and the recording of a boat docking.
-
-Although the data we trained the model on was fixed to four seconds of audio, the target samples can be of arbitrary length since we adapt the representation to the size of the training data by changing the hop length during the computation of the Mel spectrogram.
+Some of the targets were made of instrumental samples and chords, others are bells or gongs, and some do not feature any musical instruments
 
@@
 Eqn.
-\ref{distance} was used with $\lambda_1 = 0.5$ and $\lambda_2 = 10$ to compute distances between orchestrations and targets.
+\ref{distance} was used to compute distances between orchestrations and targets.
 What is shown is the ratio between the distance calculated and Orchidea's distance for the same target.
@@
 We evaluated our orchestrations both qualitatively and quantitatively by comparing our solutions to the solutions generated by Orchidea, the state-of-the-art system for computer-assisted orchestration.
+In order to have a fairer comparison, we did not allow Orchidea to use any of its advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it to use all 10 instruments in each solution.
 
@@
 It is also able to reduce the search space by applying harmonic analysis on the target sound, and create sparse solutions that do not use every instrument specified.
-
-In order to have a fairer comparison, we did not allow Orchidea to use any of its advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it forced to use all 10 instruments in each solution.
 %This creates a fairer comparison, since our models are unable to create constrained or sparse solutions.
@@
 For targets that had harmonic content, it was noted if the partials present in the target were also represented in the orchestrated solution.
-For example, one of the samples of a bell had partials that loosely represented a C minor chord, so we checked if the orchestration contained the notes of the chord.
 
@@
 This metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization.
-The equation takes in the full amplitude spectrum of the target $x$ and full amplitude spectrum of the solution $\hat{x}$.
-Then for each bin $k$ of the amplitude spectrum, it calculates the absolute difference between the values.
-The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.
-
-We calculated the distance between the target samples and our orchestrated solutions.
-We then orchestrated the same targets with Orchidea and calculated the distance for Orchidea's solutions.
-A comparison of these results is in Table \ref{orch_eval}.
 
@@
 
+A comparison of distances between our solutions and targets and Orchidea's solutions and targets is in Table \ref{orch_eval}.
 While our model is not able to outperform Orchidea, it shows consistent results.
@@
 The first natural extension would be to support sparsity in our models.
-Our current models, indeed, orchestrate all targets using a constant number of instruments and it is not able to drop specific instruments from the solution.
+Our current models orchestrate all targets using a constant number of instruments and are not able to drop specific instruments from the solution.
 This does not take into account the density of different targets.
 Sparse solutions, in which the model decides how many samples should be used to best represent the target, would allow a small number of samples to be used for sonically sparse sounds and many to be used for sonically dense sounds.
-This would generate more meaningful orchestrations that would compare more favourably to the state of the art.
+%This would generate more meaningful orchestrations that would compare more favourably to the state of the art.
 

### 2020-08-21 paper/aimusic2020/paper.tex
minor modif for submission

@@
 % first the title is needed
-\title{Neural Models for Target-Based Computer-Assisted Musical Orchestration: A Preliminary Study}
+\title{A Study on Neural Models for Target-Based Computer-Assisted Musical Orchestration}
 
@@
 %
-\author{Carmine Emanuele Cella\inst{1}\and Luke Dzwonczyk\inst{1}\and Alejandro Saldarriaga-Fuertes\inst{1}\and Hongfu Liu\inst{2}\and Helene-Camille Crayencour\inst{3}}
+\author{Carmine Emanuele Cella\inst{1}\and Luke Dzwonczyk\inst{1}\and Alejandro Saldarriaga-Fuertes\inst{1}\and Hongfu Liu\inst{1}\and Helene-Camille Crayencour\inst{2}}
 %
@@
 % unless you accept that it will be published
-\institute{University of California, Berkeley \and InstituteB\\ \email{ , dz.luke@berkeley.edu, }}
+\institute{University of California, Berkeley \and Centrale Sup\'elec, L2S, Univ. Paris-Saclay, CNRS \\ \email{carmine.cella@berkeley.edu, dz.luke@berkeley.edu }}
 
@@
 	\centering
-	\includegraphics[scale=0.17]{../latex/figs/method.png}
+	\includegraphics[scale=0.20]{../latex/figs/method.png}
 	\caption{An overview of the proposed method for assisted orchestration with neural models.
@@
 \centering
-\begin{subfigure}{.65\textwidth}
+\begin{subfigure}{.52\textwidth}
   \centering
@@
 \end{subfigure}%
-\begin{subfigure}{.65\textwidth}
+\begin{subfigure}{.52\textwidth}
   \centering
@@
 	\centering    
-     \caption{Quantitative comparison of orchestrations as ratios to Orchidea.
-Eqn.
-\ref{distance} was used to compute distances between orchestrations and targets.
-What is shown is the ratio between the distance calculated and Orchidea's distance for the same target.
-A value less than 1 means that the model performed worse (i.e.
-had a larger distance), and a value greater than 1 means the model performed better than Orchidea.
-The last column shows the ratio of the average distances for the model across all targets.}\label{orch_eval}
-	\centering    
-    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c}
+  \resizebox{\columnwidth}{!}{\begin{tabular}{|c|c|c|c|c|c|c|c|c|c}
      \hline
@@
       \hline
-    \end{tabular}
-
+    \end{tabular}}
+    \caption{Quantitative comparison of orchestrations as ratios to Orchidea.
+Eqn.
+\ref{distance} was used to compute distances between orchestrations and targets.
+What is shown is the ratio between the distance calculated and Orchidea's distance for the same target.
+A value less than 1 means that the model performed worse (i.e.
+had a larger distance), and a value greater than 1 means the model performed better than Orchidea.
+The last column shows the ratio of the average distances for the model across all targets.}\label{orch_eval}
 \end{table*}

### 2020-08-21 paper/aimusic2020/paper.tex
biblio and typos

@@
 %
-\author{Carmine Emanuele Cella\inst{1}\and Luke Dzwonczyk\inst{1}\and Alejandro Saldarriaga-Fuertes\inst{1}\and Hongfu Liu\inst{1}\and Helene-Camille Crayencour\inst{2}}
+\author{Carmine Emanuele Cella\inst{1}\and Luke Dzwonczyk\inst{1}\and Alejandro Saldarriaga-Fuertes\inst{1}\and Hongfu Liu\inst{1}\and H\'el\`ene-Camille Crayencour\inst{2}}
 %
@@
 % unless you accept that it will be published
-\institute{University of California, Berkeley \and Centrale Sup\'elec, L2S, Univ. Paris-Saclay, CNRS \\ \email{carmine.cella@berkeley.edu, dz.luke@berkeley.edu }}
+\institute{CNMAT - University of California, Berkeley \and Centrale Sup\'elec, L2S, Univ. Paris-Saclay, CNRS \\ \email{carmine.cella@berkeley.edu, dz.luke@berkeley.edu }}
 
@@
 
-The development of computational tools to assist and inspire the musical composition process constitutes an import research area known as \emph{Computer-Assisted Composition (CAC)} \cite{FerVic2013, Ari2005}.
-Within CAC, target-based computer-assisted orchestration is an important example of how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.
+The development of computational tools to assist and inspire the musical composition process constitutes an important research area known as \emph{Computer-Assisted Composition (CAC)} \cite{FerVic2013, Ari2005}.
+Within CAC, target-based computer-assisted orchestration is a compelling case of how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.
 
@@
 
-The second and deeper model that we trained as classifier was the well known deep residual network \emph{ResNet}.
+The second and deeper model that we trained as classifier was the well known deep residual network \emph{ResNet} \cite{He15}.
 Specifically, we used 18-layer ResNet, which allows information to pass directly from the input to the final layer of each block.
@@
 However, the value of the loss function alone is difficult to interpret.
+
+\begin{figure}
+  \center{
+  \includegraphics[scale=0.4]{../latex/figs/CNN_vs_ResNet.png}}
+  \caption{Best overall accuracy for CNN with LSTM (50 epochs, 200k samples per epoch) and ResNet (20 epochs, 400k samples per epoch) depending on the number of instruments in the combinations used for training.
+\label{cnn_vs_resnet}}
+  \end{figure}
+
 %For this reason we created a complementary function $f$ to be used for evaluation only.
@@
 \begin{figure}
-\center{
-\includegraphics[scale=0.4]{../latex/figs/CNN_vs_ResNet.png}}
-\caption{Best overall accuracy for CNN with LSTM (50 epochs, 200k samples per epoch) and ResNet (20 epochs, 400k samples per epoch) depending on the number of instruments in the combinations used for training.
-\label{cnn_vs_resnet}}
+  \centering
+  \begin{subfigure}{.52\textwidth}
+    \centering
+    \includegraphics[width=0.9\linewidth]{../latex/figs/Acc_vs_N_CNN.png}
+    \caption{CNN with LSTM}
+    \label{best_acc_cnn}
+  \end{subfigure}%
+  \begin{subfigure}{.52\textwidth}
+    \centering
+    \includegraphics[width=0.9\linewidth]{../latex/figs/Acc_vs_N_ResNet.png}
+    \caption{ResNet}
+    \label{best_acc_resnet}
+  \end{subfigure}
+  \caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM and the ResNet depending on the number of instruments in the combinations}
   \end{figure}
@@
 This could indicate that there is an optimal spectral shape that maximizes the probability of being detected in such classification framework.
-
-\begin{figure}
-\centering
-\begin{subfigure}{.52\textwidth}
-  \centering
-  \includegraphics[width=0.9\linewidth]{../latex/figs/Acc_vs_N_CNN.png}
-  \caption{CNN with LSTM}
-  \label{best_acc_cnn}
-\end{subfigure}%
-\begin{subfigure}{.52\textwidth}
-  \centering
-  \includegraphics[width=0.9\linewidth]{../latex/figs/Acc_vs_N_ResNet.png}
-  \caption{ResNet}
-  \label{best_acc_resnet}
-\end{subfigure}
-\caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM and the ResNet depending on the number of instruments in the combinations}
-\end{figure}
 
@@
 In \cite{Gillick19} the authors propose to use LSTM-based models to predict the embedding features for the combinations used during the optimisation process in assisted orchestration.
-We believe that by combining their prediction model with our classification models could generate more faithful representations and improve the overall quality of generated orchestrations.
+We believe that by combining their prediction model with our classification models we could generate more faithful representations and improve the overall quality of generated orchestrations.
 

### 2020-10-06 paper/aimusic2020/paper.tex
Add mel spectrogram details

@@
 Then the chosen samples are combined to be played simultaneously and normalized by the number of instruments.
-The Mel spectrogram of the mixture is computed to be used as the input features for the classifiers.
 The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly 4 seconds long.
+The Mel spectrogram of the mixture is computed using an FFT hop length of 2048 samples (the window of each FFT is $46$ms wide) and a 128 Mel bins.
+Therefore, the Mel features fed to the model are matrices of size $128\times 345$.
 

### 2020-10-06 paper/aimusic2020/paper.tex
Rephrase CAC quote

@@
 
-Target-based computer-assisted orchestration can be thought of as the process of searching for the best combinations of sounds in a database of instrumental samples, to match a target sound under a specific similarity metric and specific sets of constraints.
+Target-based computer-assisted orchestration takes a target sound as an input and attempts to find instrumental samples that best match the target given a specific similarity metric and a set of constraints.
 A solution to this problem is a set of orchestral scores that represent the mixtures of audio samples in the database, ranked by similarity with the target sound.

### 2020-10-06 paper/aimusic2020/paper.tex
Correct table description

@@
 \ref{distance} was used to compute distances between orchestrations and targets.
-What is shown is the ratio between the distance calculated and Orchidea's distance for the same target.
-A value less than 1 means that the model performed worse (i.e.
-had a larger distance), and a value greater than 1 means the model performed better than Orchidea.
+What is shown is the ratio between the distance of Orchidea's solution to the target and our solution's distance to the same target.
+A value less than 1 means that our model performed worse (i.e.
+had a larger distance), and a value greater than 1 means our model performed better than Orchidea.
 The last column shows the ratio of the average distances for the model across all targets.}\label{orch_eval}

### 2020-10-07 paper/latex/paper.tex
Fix baselinestretch
Added clarification on test accuracy and samples per epoch

@@
 \newcommand{\carmine}[1]{\textbf{\color{red} Carmine => #1}}
-\renewcommand{\baselinestretch}{.98}
+\renewcommand{\baselinestretch}{1}
 
@@
 \item $N$ random TinySOL samples are selected, leading to a variety of instruments, pitches, and dynamics; we did not allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures;
-\item the chosen samples are combined to be played simultaneously and normalized by the number of instruments;
+\item the chosen samples are combined to be played simultaneously, and the resulting sum is normalized by the number of instruments;
 \item the Mel frequency cepstral coefficients (MFCCs) and the Mel spectrogram of the mixture are computed to be used as the input features for the baseline classifiers and the neural classifiers respectively.
@@
 
-Then, for both CNN with LSTM and ResNet, we computed the maximum accuracy over the epochs. 
+Then, for both CNN with LSTM and ResNet, we computed the accuracy on the test sets across epochs, and kept the best one as an indicator of the performance of the models. 
 The CNN was trained on 200,000 generated samples over 50 epochs, and ResNet was trained on 400,000 generated samples over 20 epochs.
 Fig.
-\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$.\\
+\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$ per mixture.\\
 

### 2020-10-09 paper/aimusic2020/paper.tex
Add explanation of terms in eqn 1
Minor fixes

@@
 The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly 4 seconds long.
-The Mel spectrogram of the mixture is computed using an FFT hop length of 2048 samples (the window of each FFT is $46$ms wide) and a 128 Mel bins.
+The Mel spectrogram of the mixture is computed using an FFT hop length of 2048 samples (the window of each FFT is $46$ms wide) and 128 Mel bins.
 Therefore, the Mel features fed to the model are matrices of size $128\times 345$.
@@
 This metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization.
+The equation takes in the full FFT of the target $x$ and full FFT of the solution $\tilde{x}$.
+Then for each bin $k$ of the FFT, it calculates the absolute difference between the values.
+The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.
 

### 2020-10-09 paper/aimusic2020/paper.tex
Add back Alejandro's changes that git overwrote

@@
 Specifically, we used 18-layer ResNet, which allows information to pass directly from the input to the final layer of each block.
-To make the model more suitable to our problem, we decided to reduce the number of parameters and to reset the output channel numbers of each block with 32, 64, 32, 32 respectively.
+To make the model more suitable to our problem, we decided to use an architecture with 4 blocks whose outputs are of size 32, 64, 32 and 32 respectively.
 

### 2020-11-09 paper/ISMIR_2020/paper.tex
Rename paper folders to ISMIR and csmc mume

--- 2020-10-07 paper/latex/paper.tex
@@
 Specifically, we used 18-layer ResNet, which allows information to pass directly from the input to the final layer of each block.
-Besides, to make the model more suitable to our problem, we decided to reduce the number of parameters and to reset the output channel numbers of each block with 32, 64, 32, 32 respectively.
+To make the model more suitable to our problem, we decided to use an architecture with 4 blocks whose outputs are of size 32, 64, 32 and 32 respectively.
 
@@
 Fig.
-\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$ per mixture.\\
+\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of TinySOL samples $N$ per mixture.\\
 

### 2020-11-09 CSMC + MuMe 2020 /OrchDataset.py
Create new folder structure

@@ -1,0 +1,295 @@
+import itertools
+from functools import reduce
+from generateDB import generateDBLabels
+import librosa.display
+import random
+import math
+from torch.utils import data
+import torch
+import numpy as np
+import os
+import pickle
+import random
+import hashlib
+
+from parameters import GLOBAL_PARAMS
+
+def showSample(s):
+    print(s[1])
+    librosa.display.specshow(s[0], x_axis='time',y_axis='mel',sr=44100)
+
+class IndexIncrementor:
+    '''
+        This class provides a way to iterate over the set of features contained in 
+        a RawDatabase object to ensure that all pitch bins are uniformly chosen.
+        This helps generating a homogeneous database that doesn't miss some dynamics.
+        
+        A certain degree of randomness is added in the sense that the step and starting point
+        are randomized.
+    '''
+    def __init__(self, N, base, offset=None):
+        self.N = N
+        if type(base) == int:
+            self.base = [base]*N
+        elif type(base) == list and len(base) == N:
+            self.base = base
+        else:
+            raise Exception("Wrong argument for IndexIncrementor (must be int or list of size N)")
+        self.max_count = reduce((lambda x,y:x*y), self.base)
+        self.restart(offset)
+    
+    def findStep(self):
+        self.step = int(random.randint(int(0.05*self.max_count),int(0.95*self.max_count)))
+        while math.gcd(self.step, self.max_count) > 1:
+            self.step += 1
+    
+    def add(self, e):
+        r = e
+        i = 0
+        while i < self.N and r != 0:
+            r,self.index[i] = divmod(self.index[i]+r,self.base[i])
+            i += 1
+            
+    def incr(self):
+        self.add(self.step)
+        self.count += 1
+        
+    def restart(self, offset = None):
+        self.count = 0
+        self.index = [0]*self.N
+        self.findStep()
+        if offset is None:
+            self.add(random.randint(0,self.max_count-1))
+        elif offset >= 0:
+            self.add(offset)
+        else:
+            raise Exception("Wrong offset")
+    
+    def isOver(self):
+        return self.count >= self.max_count
+            
+class HashTable:
+    def __init__(self, max_depth, depth=0):
+        self.depth = depth
+        self.max_depth = max_depth
+        if max_depth == depth+1:
+            self.data = []
+        else:
+            self.data = {}
+            for i in '0123456789abcdef':
+                self.data[i] = HashTable(max_depth,depth+1)
+            
[…]

### 2020-11-09 CSMC + MuMe 2020 /augment.py

@@ -1,0 +1,81 @@
+import librosa
+import numpy as np
+import copy
+import random
+from sklearn.utils import shuffle
+
+
+def set_length(new_y, y):
+    if len(new_y) < len(y):
+        new_y = librosa.util.fix_length(new_y, len(y))
+    elif len(new_y) > len(y):
+        new_y = new_y[0:len(y)]
+    return new_y
+
+
+def random_stretch(y, sr, prob=1., min_stretch=0.9, max_stretch=1.1):
+    if np.random.uniform(0, 1) < prob:
+        factor = np.random.uniform(min_stretch, max_stretch)
+        new_y = librosa.effects.time_stretch(y, factor)
+        return set_length(new_y, y)
+    else:
+        return y
+
+
+def random_pitch(y, sr, prob=1., min_shift=-0.3, max_shift=0.3, bins_per_octave=12):
+    if np.random.uniform(0, 1) < prob:
+        steps = np.random.uniform(min_shift, max_shift)
+        new_y = librosa.effects.pitch_shift(
+            y, sr, n_steps=steps, bins_per_octave=bins_per_octave)
+        return set_length(new_y, y)
+    else:
+        return y
+
+
+def wav_augment(y, sr):
+    functions = shuffle([random_pitch, random_stretch])
+    for fn in functions:
+        y = fn(y, sr=sr)
+    return y
+
+
+def freq_mask(spec, F=30, num_masks=1, replace_with_zero=False):
+    cloned = copy.deepcopy(spec)
+    num_mel_channels = cloned.shape[0]
+    for i in range(0, num_masks):
+        f = random.randrange(0, F)
+        f_zero = random.randrange(0, num_mel_channels - f)
+        # avoids randrange error if values are equal and range is empty
+        if (f_zero == f_zero + f):
+            return cloned
+        mask_end = random.randrange(f_zero, f_zero + f)
+        if (replace_with_zero):
+            cloned[0][f_zero:mask_end] = 0
+        else:
+            cloned[0][f_zero:mask_end] = cloned.mean()
+    return cloned
+
+
+def time_mask(spec, T=40, num_masks=1, replace_with_zero=False):
+    cloned = copy.deepcopy(spec)
+    len_spectro = cloned.shape[1]
+    for i in range(0, num_masks):
+        t = random.randrange(0, T)
+        t_zero = random.randrange(0, len_spectro - t)
+        # avoids randrange error if values are equal and range is empty
[…]

### 2020-11-09 CSMC + MuMe 2020 /baseline_classification.py

@@ -1,0 +1,293 @@
+# Adapted from code written by Carmine E. Cella, 2020
+
+
+import numpy as np
+import pandas as pd
+import matplotlib
+import matplotlib.pyplot as plt
+from sklearn.preprocessing import LabelEncoder, StandardScaler
+from sklearn.linear_model import LogisticRegression
+from sklearn.pipeline import Pipeline
+from sklearn.model_selection import train_test_split
+from sklearn.metrics import accuracy_score
+from sklearn.decomposition import PCA
+from sklearn.model_selection import cross_val_score
+from sklearn.svm import SVC
+from sklearn.neighbors import KNeighborsClassifier
+from sklearn.model_selection import KFold
+from sklearn.ensemble import RandomForestClassifier
+from sklearn.multioutput import MultiOutputClassifier
+from collections import Counter
+from matplotlib import pyplot as plt
+
+import librosa
+import librosa.display
+import json
+import os
+import random
+import sys
+from time import process_time
+
+# raw dataset
+# file hierarchy: (note that folders Brass, Winds, Strings are not present)
+# ----TinySOL
+#   ----Bn
+#   ----Cb
+#   ----Va
+#   etc...
+path = './TinySOL_0.6/TinySOL'
+
+# time duration
+time = 4
+
+# number of samples between successive frames in librosa.melspectrogram
+mel_hop_length = 44100
+
+pitch_classes = ['A', 'A#', 'B', 'C', 'C#', 'D', 'D#', 'E', 'F', 'F#', 'G', 'G#']
+
+# maps an instrument or instrument-pitch to a class #
+label_mapping = {}
+
+def mix(fa, fb):
+    diff = len(fa) - len(fb)
+
+    if diff > 0:
+        add = np.zeros((1, diff), dtype=np.float32)
+        fb = np.append(fb, add)
+    elif diff < 0:
+        add = np.zeros((1, -diff), dtype=np.float32)
+        fa = np.append(fa, add)
+
+    return fa+fb
+    
+def combine_sounds(soundlist):
+    mixed_file = np.zeros((1, 1))
+    sr = 0
+    for sound in soundlist:
+        sound_path = os
+        sfile, sr = librosa.load(sound, sr=None)
+        if len(sfile) > time*sr:
+            # randomly select one part of the raw audio
+            n = np.random.randint(0, len(sfile)-time*sr)
+            sfile = sfile[n:n+time*sr]
+        # add augment
[…]

### 2020-11-09 CSMC + MuMe 2020 /dataset.py

@@ -1,0 +1,59 @@
+from torch.utils import data
+import torchvision.transforms as transforms
+import torch
+import numpy as np
+import pickle
+import librosa
+import os
+from augment import spec_augment
+import librosa.display
+
+
+class OrchDataSet(data.Dataset):
+    def __init__(self, root, mode, transform):
+        if not os.path.exists(root):
+            print("[Error] root does not exist")
+            return
+        self.mode = mode
+        self.audio_feature = []
+        self.labels = []
+        self.mix = []
+
+        for dir in os.listdir(root):
+            if dir.startswith(mode):
+                new_path = os.path.join(root, dir)
+                f = open(new_path, 'rb')
+                inp = pickle.load(f)
+                for i, x in enumerate(inp):
+                    self.audio_feature.append(x[0])
+                    self.labels.append(x[1])
+
+        if transform is not None:
+            self.transform = transform
+
+    def __len__(self):
+        return len(self.audio_feature)
+
+    def __getitem__(self, idx):
+        audio = self.audio_feature[idx]
+
+        if self.mode.startswith('train'):
+            a = np.array(audio[0])
+            a = spec_augment(a)
+            audio = np.array([a])
+
+            audio = torch.tensor(audio)
+
+        return audio, self.labels[idx]
+
+
+if __name__ == '__main__':
+    root = './featurized_data'
+    training = OrchDataSet(root, 'training', transforms.ToTensor())
+    testing = OrchDataSet(root, 'test', transforms.ToTensor())
+    print("training set size: {}".format(len(training)))
+    print("testing set size: {}".format(len(testing)))
+
+    print("training set sample feature shape: {}".format(training[0][0].shape))
+    # print("testing set  feature shape: {}".format(testing[0].shape))
+

### 2020-11-09 CSMC + MuMe 2020 /generateDB.py

@@ -1,0 +1,90 @@
+import os
+from functools import reduce
+import librosa
+import numpy as np
+import random
+
+from parameters import GLOBAL_PARAMS
+
+def getSampleMetaData(path,name):
+    s = {}
+    s['path'] = path
+    s['stft'] = {}
+    X = np.linspace((1-GLOBAL_PARAMS.coeff_freq_shift_data_augment)*GLOBAL_PARAMS.RATE, (1+GLOBAL_PARAMS.coeff_freq_shift_data_augment)*GLOBAL_PARAMS.RATE, 5)
+    for freq in X:
+        y,sr=librosa.load(path,sr=int(freq))
+        nb_samples = int(GLOBAL_PARAMS.TIME_LENGTH*GLOBAL_PARAMS.RATE)
+        if len(y) < nb_samples:
+            y = np.append(y, np.zeros((1,(nb_samples-len(y))), dtype=np.float32))
+        else:
+            y = y[:nb_samples]
+        s['stft'][int(freq)] = librosa.stft(y=y,hop_length=GLOBAL_PARAMS.MEL_HOP_LENGTH,n_fft=GLOBAL_PARAMS.N_FFT)
+    t = name.split('.')[0].split('-')
+    s['instrument'] = t[0]
+    s['style'] = t[1]
+    s['pitch_name'] = t[2]
+    s['nuance'] = t[3]
+    if len(t) == 5:
+        s['other'] = t[4]
+    
+    return s
+    
+
+def recursiveSearch(path, i=0):
+    l = []
+    s = i
+    for x in os.listdir(path):
+        p = os.path.join(path,x)
+        if os.path.isdir(p):
+            l.extend(recursiveSearch(p,s))
+        elif os.path.isfile(p) and p.split('.')[-1] == 'wav':
+            print("Sample {}".format(s))
+            l.append(getSampleMetaData(p,x))
+            s += 1
+    return l
+
+def getPitchRange(d):
+    pr = set()
+    for i in d:
+        pr.add(i['pitch_name'])
+    pr = list(pr)
+    pr.sort(key = lambda x : 10000*ord(x[-1]) + 100*((ord(x[0])-ord('C'))%7) + (len(x)==3))
+    return pr
+
+
+def generateDBLabels(path, nb_pitch_range = 1, instr_filter=None):
+    '''
+        Generates the short term fourier transforms for each sample of the data base.
+        In order to ensure a good repartition of the pitches when generating the combinations,
[…]

### 2020-11-09 CSMC + MuMe 2020 /main.py

@@ -1,0 +1,430 @@
+import torch
+import torch.nn as nn
+import torchvision.transforms as transforms
+import torch.nn.functional as F
+from torch.autograd import Variable
+import numpy as np
+import argparse
+import time
+import json
+import os
+
+from model import OrchMatchNet
+from process_OrchDB import instruments, decode, N, out_num, stat_test_db
+from dataset import OrchDataSet
+
+
+epoch_num = 200
+batch_size = 16
+my_model_path = './model'
+server_model_path = './model'
+my_data_path = './data/five'
+featurized_data_path = './featurized_data/'
+
+TEST_MODEL = True
+
+# db = {'Vc': 3560.0, 'Fl': 2831.0, 'Va': 3469.0, 'Vn': 3328.0, 'Ob': 2668.0, 'BTb': 2565.0,
+#       'Cb': 3305.0, 'ClBb': 2960.0, 'Hn': 3271.0, 'TpC': 2245.0, 'Bn': 2944.0, 'Tbn': 2854.0}
+
+db = stat_test_db()
+
+class Timer:
+    def __init__(self, size_buffer):
+        self.measuring = False
+        self.begin = []
+        self.end = []
+        self.size_buffer = size_buffer
+        
+    def start(self):
+        if not self.measuring:
+            if len(self.begin) == self.size_buffer:
+                self.begin.pop(0)
+            self.begin.append(time.time())
+            self.measuring = True
+            
+    def stop(self):
+        if self.measuring:
+            if len(self.end) == self.size_buffer:
+                self.end.pop(0)
+            self.end.append(time.time())
+            self.measuring = False
+    
+    def estimate(self, remaining):
+        if len(self.end) > 0 and not self.measuring:
+            diff = np.array(self.end)-np.array(self.begin)
+            diff = np.mean(diff)
+            h,r=divmod(diff*remaining,3600)
+            m,s=divmod(r,60)
+            print(diff)
+            print("Estimated remaining time : {}h{}m{}s".format(int(h), int(m), int(s)))
+            
+
+timer = Timer(10)
+
+def arg_parse():
+    parser = argparse.ArgumentParser(description='combination of orch')
+    parser.add_argument('--model', default='resnet')
+    parser.add_argument('--is_resume', default='False',
+                        choices=['True', 'False'])
+
+    return parser.parse_args()
+
+
+def main():
+    print("Starting parsing -----")
+    arg = arg_parse()
+    print("End parsing")
+
[…]

### 2020-11-09 CSMC + MuMe 2020 /model.py

@@ -1,0 +1,121 @@
+import torch
+import numpy as np
+import torch.nn as nn
+import torch.nn.functional as F
+
+from resnet import ResNet, count_parameters, init_weights, device
+
+
+class CNN(nn.Module):
+    def __init__(self, out_num, features_dim):
+        super(CNN, self).__init__()
+
+        shape_input = features_dim
+        self.layer1 = nn.Sequential(
+            nn.Conv2d(in_channels=1, out_channels=8,
+                      kernel_size=3, stride=1, padding=1),
+            nn.BatchNorm2d(num_features=8),
+            nn.ReLU(inplace=True),
+            nn.MaxPool2d(kernel_size=2)
+        )
+        shape_input = (shape_input[0]//2, shape_input[1]//2)
+
+        self.layer2 = nn.Sequential(
+            nn.Conv2d(in_channels=8, out_channels=16,
+                      kernel_size=3, stride=1, padding=1),
+            nn.BatchNorm2d(num_features=16),
+            nn.ReLU(inplace=True),
+            nn.MaxPool2d(kernel_size=2)
+        )
+        shape_input = (shape_input[0]//2, shape_input[1]//2)
+
+        self.layer3 = nn.Sequential(
+            nn.Conv2d(in_channels=16, out_channels=32,
+                      kernel_size=3, stride=1, padding=1),
+            nn.BatchNorm2d(num_features=32),
+            nn.ReLU(inplace=True),
+            nn.MaxPool2d(kernel_size=2)
+        )
+        shape_input = (shape_input[0]//2, shape_input[1]//2)
+        self.lstm_shape = (shape_input[0], shape_input[1])
+
+        self.lstm = nn.LSTM(input_size=shape_input[1], hidden_size=shape_input[1], batch_first=True)
+
+        self.layer4 = nn.Sequential(
+            nn.Conv2d(in_channels=32, out_channels=32,
+                      kernel_size=3, stride=1, padding=1),
+            nn.ReLU(inplace=True),
+            nn.MaxPool2d(kernel_size=2),
+            nn.Dropout(0.5)
+        )
+        shape_input = (shape_input[0]//2, shape_input[1]//2)
+
+        self.fc1 = nn.Sequential(
+            nn.Linear(in_features=32*shape_input[0]*shape_input[1], out_features=2048),
+            nn.ReLU(inplace=True),
+            nn.Dropout(0.5)
+        )
+
+        self.fc3 = nn.Linear(in_features=2048, out_features=out_num)
+
[…]

### 2020-11-09 CSMC + MuMe 2020 /parameters.py

@@ -1,0 +1,46 @@
+import pickle
+
+class SimParams:
+    def __init__(self):
+        self.path = "./TinySOL"
+        self.MEL_HOP_LENGTH = 2048
+        self.RATE = 44100
+        self.N_FFT = 2048
+        self.N_MELS = 128
+        self.TIME_LENGTH = 4
+        self.FEATURE_TYPE = 'mel'
+        
+        self.N = 2
+        self.nb_samples = 400000
+        self.rdm_granularity = 10
+        self.nb_pitch_range = 8
+        self.instr_filter = ['Hn','Ob','Vn','Va', 'Vc', 'Fl', 'Tbn', 'Bn', 'TpC', 'ClBb'][:10]
+        self.batch_size = 16
+        self.model_type = 'resnet'
+        self.nb_epoch = 100
+        self.train_proportion = 0.8
+        
+        self.coeff_freq_shift_data_augment = 0.008 # For data augmentation, proportional change to sampling rate
+        self.delay_offset_avg = 0.005 # Average of offset applied to delay filters (ie the position of the first impulse)
+        self.delay_period_avg = 0.002 # Average of the period of the impulses for delaying
+        self.delay_feedback_avg = 0.5 # Average of the feedback factor of delay impulses
+        self.prop_zero_row = 0.01
+        self.prop_zero_col = 0.01
+        self.noise_kernel_var = 0.0001
+        
+        self.lab_class = {} # Contains the labels
+        
+    def load_parameters(self, path):
+        f = open(path + '/params.pkl', 'rb')
+        p = pickle.load(f)
+        f.close()
+        
+        for attr in p.__dict__:
+            setattr(self, attr, getattr(p, attr))
+    
+    def save_parameters(self, path):
+        f = open(path + '/params.pkl', 'wb')
+        pickle.dump(self, f)
+        f.close()
+
+GLOBAL_PARAMS = SimParams()

### 2020-11-09 CSMC + MuMe 2020 /plot.py

@@ -1,0 +1,138 @@
+import matplotlib.pyplot as plt
+import pickle
+import os
+import parameters
+
+def f(l):
+    Y = [0,0,0,0,0,0]
+    for i in l:
+        Y[X.index(i[0])] = i[1]['acc']
+    return Y
+
+order = ['acc', 'Fl', 'ClBb', 'Ob', 'Bn', 'Va', 'Vn', 'Vc', 'TpC', 'Tbn', 'Hn']
+
+def plot_best_acc(path):
+    
+    list_accs = []
+    for i in os.listdir(path):
+        accs = {}
+        MM = len([x for x in os.listdir('{}/{}'.format(path, i)) if 'result' in x])
+        epochs = list(range(4,MM*5,5))
+        max_epoch = -1
+        
+        parameters.GLOBAL_PARAMS.load_parameters('{}/{}'.format(path, i))
+    
+        for j in epochs:
+            f = open('{}/{}/result{}.pkl'.format(path, i, j), 'rb')
+            r = pickle.load(f)
+            if len(accs.keys()) == 0:
+                accs['acc'] = 0
+                for k in r['pitch_acc']:
+                    accs[k] = 0
+            
+            if r['acc'] > accs['acc']:
+                max_epoch = j
+                accs['acc'] = r['acc']
+                for k in r['pitch_acc']:
+                    accs[k] = r['pitch_acc'][k]
+            f.close()
+            
+        list_accs.append((parameters.GLOBAL_PARAMS.N, accs, max_epoch, epochs[-1]))
+    plt.figure(figsize=(6,4))
+    for i in order:
+        accs = []
+        for x in list_accs:
+            accs.append((x[0], x[1][i]))
+            
+        accs.sort(key = lambda x : x[0])
+        
+        if i == 'acc':
+            plt.plot([x[0] for x in accs], [x[1] for x in accs], '-o', label = "Overall accuracy")
+        else:
+            plt.plot([x[0] for x in accs], [x[1] for x in accs], '--x', label = "{} accuracy".format(i))
+    plt.grid()
+    plt.xlabel("Number of samples in the combinations")
+    plt.ylabel("Accuracy")
+    plt.title("Maximum test accuracy achieved vs number of samples N".format(i))
+        
+    return list_accs
+
+def plot_run(path, run_nb):
+    res = {'acc' : []}
+    
+    MM = len([x for x in os.listdir('{}/run{}'.format(path, run_nb)) if 'result' in x])
+        
+        
+    epochs = list(range(4,MM*5,5))
[…]

### 2020-11-09 CSMC + MuMe 2020 /process_OrchDB.py

@@ -1,0 +1,419 @@
+import numpy as np
+import torch
+from torchvision.transforms import transforms
+import matplotlib.pyplot as plt
+import random
+import json
+import copy
+import librosa
+import pickle
+import os
+
+from dataset import OrchDataSet
+from augment import wav_augment
+
+# raw dataset
+# file hierarchy: (note that folders Brass, Winds, Strings are not present)
+# ----TinySOL
+#   ----Bn
+#   ----Cb
+#   ----Va
+#   etc...
+path = './TinySOL'
+
+# generated dataset were picklized files will be stored
+featurized_data_path = './featurized_data/'
+
+instruments = ['Vc', 'Fl', 'Va', 'Vn', 'Ob', 'BTb',
+       'Cb', 'ClBb', 'Hn', 'TpC', 'Bn', 'Tbn']
+
+# number of mixture
+N = 2
+# time duration
+time = 4
+# max sample number
+MAX_NUM = 5040
+# class number
+out_num = 505
+# number of samples between successive frames in librosa.melspectrogram
+mel_hop_length = 512
+
+
+
+
+def random_combine():
+    '''
+        call this function to generate dataset and divide it into training set and test set
+    '''
+    all_samples = []
+    all_mixture = []
+    inx = json.load(open('class.index', 'r'))
+
+    for instrument in instruments:
+        instrument_path = os.path.join(os.path.join(path, instrument), "ordinario")
+        for sample in os.listdir(instrument_path):
+            sample_path = os.path.join(instrument_path, sample)
+            all_samples.append(sample_path)
+
+    print(all_samples)
+    # combine
+    all_selects = []
+    all_mix = []
+    init = 0
+    while init < MAX_NUM:
+        # select N files randomly
+        selects = random.sample(range(len(all_samples)), N)
+        flag = False
+        soundlist = []
+        labellist = []
+
+        for num in selects:
+            f = os.path.split(all_samples[num])[1].split('-')[:3]
+            f = f[0]+'-'+f[1]+'-'+f[2]
+            idx = inx[f]
+
+            if str(idx) in labellist:
+                # ignore the difference of dynamics
+                flag = True
+
+            labellist.append(str(idx))
+            soundlist.append(all_samples[num])
+
+        if flag or set(labellist) in all_selects:
+            continue
+
+        mixed = combine(soundlist, labellist)
[…]

### 2020-11-09 CSMC + MuMe 2020 /process_TinySOL.py

@@ -1,0 +1,411 @@
+##################################################
+# Process the data of orchestration combinations
+# and make preparation for training, the steps include:
+#   1. combine N instruments with various pitches
+#   2. extract the melspectrugram features of these combinations as input while training
+#   3. save the corresponding label(index)
+#   4. make the train and test dataset
+
+import librosa
+import numpy as np
+import torch
+from torchvision.transforms import transforms
+import matplotlib.pyplot as plt
+import random
+import json
+import copy
+import pickle
+import os
+
+from dataset import OrchDataSet
+
+
+data_path = './data/'
+brass = '/Brass'
+strings = '/Strings'
+winds = '/Winds'
+instru_type = [brass, strings, winds]
+instruments = ['BTb', 'TpC', 'Hn', 'Tbn', 'Va', 'Vn',
+       'Vc', 'Cb', 'Ob', 'Fl', 'Bn', 'ClBb']
+N = 2
+time = 4
+MAX_NUM = 200
+featurized_data_path = './featurized_data/'
+
+"""
+Creates .wav files of combinations of instruments
+"""
+def random_combine():
+    all_instrument = []
+    all_mixture = []
+
+    for instruments in instru_type:
+        for instrument in os.listdir(data_path + instruments):
+            if instrument.startswith('.'):
+                continue
+            newpath = os.path.join(data_path + instruments, instrument)
+            all_instrument.append(newpath)
+
+    # combine
+    all_selects = []
+    all_mix = []
+    init = 0
+    while init < MAX_NUM:
+        # select N instruments randomly
+        selects = random.sample(range(len(all_instrument)), N)
+
+        soundlist = []
+        labellist = []
+
+        for instrument in selects:
+            instr = all_instrument[instrument]
+            instr_file_list = os.listdir(instr)
+
+            # select a file from the instrument
+            num = random.randint(0, len(instr_file_list)-1)
+            soundlist.append(os.path.join(instr, instr_file_list[num]))
+            labellist.append(instr.split('/')[-1]+'!'+str(num))
+
+        if set(labellist) in all_selects:
+            continue
+
+        combine(soundlist, labellist)
+        # mix = deal_mix(mix)
+
+        all_selects.append(set(labellist))
+        # all_mix.append(mix)
+
+        init += 1
+        if init % 100 == 0:
+            print(
[…]

### 2020-11-09 CSMC + MuMe 2020 /sampleRNN/OrchDataset.py

@@ -1,0 +1,287 @@
+import itertools
+from functools import reduce
+from generateDB import generateDBLabels
+import librosa.display
+import random
+import math
+from torch.utils import data
+import torch
+import numpy as np
+import os
+import pickle
+import random
+
+from parameters import TIME_LENGTH, RATE, N_FFT, MEL_HOP_LENGTH, N_MELS, rdm_granularity, instr_filter
+from parameters import coeff_freq_shift_data_augment, prop_zero_col, prop_zero_row, noise_kernel_var, delay_feedback_avg, delay_offset_avg, delay_period_avg
+
+def showSample(s):
+    print(s[1])
+    librosa.display.specshow(s[0], x_axis='time',y_axis='mel',sr=44100)
+
+
+def mix(fa, fb):
+    diff = len(fa) - len(fb)
+
+    if diff > 0:
+        add = np.zeros((1, diff), dtype=np.float32)
+        fb = np.append(fb, add)
+    elif diff < 0:
+        add = np.zeros((1, -diff), dtype=np.float32)
+        fa = np.append(fa, add)
+
+    return fa+fb
+
+def combine(soundlist):
+    mixed_file = np.zeros((1, 1))
+    sr = 0
+    for sound in soundlist:
+        sound_path = os
+        sfile, sr = librosa.load(sound, sr=None)
+        if len(sfile) > TIME_LENGTH*sr:
+            # randomly select one part of the raw audio
+            n = np.random.randint(0, len(sfile)-TIME_LENGTH*sr)
+            sfile = sfile[n:n+TIME_LENGTH*sr]
+        mixed_file = mix(mixed_file, sfile)
+    mixed_file = mixed_file/len(soundlist)
+
+    # librosa.output.write_wav(name, y=mixed_file, sr=sr)
+    return mixed_file
+
+class IndexIncrementor:
+    '''
+        This class provides a way to iterate over the set of features contained in 
+        a RawDatabase object to ensure that all pitch bins are uniformly chosen.
+        This helps generating a homogeneous database that doesn't miss some dynamics.
+        
+        A certain degree of randomness is added in the sense that the step and starting point
+        are randomized.
+    '''
+    def __init__(self, N, base, offset=None):
+        self.N = N
+        if type(base) == int:
+            self.base = [base]*N
+        elif type(base) == list and len(base) == N:
+            self.base = base
+        else:
+            raise Exception("Wrong argument for IndexIncrementor (must be int or list of size N)")
[…]

### 2020-11-09 CSMC + MuMe 2020 /sampleRNN/generateDB.py

@@ -1,0 +1,86 @@
+import os
+import itertools
+import argparse
+from functools import reduce
+import librosa
+import numpy as np
+import random
+
+from parameters import MEL_HOP_LENGTH, N_FFT, TIME_LENGTH
+
+def getSampleMetaData(path,name):
+    s = {}
+    s['path'] = path
+    y,sr=librosa.load(path,sr=None)
+    nb_samples = int(TIME_LENGTH*sr)
+    if len(y) < nb_samples:
+        y = np.append(y, np.zeros((1,(nb_samples-len(y))), dtype=np.float32))
+    else:
+        y = y[:nb_samples]
+    s['stft'] = librosa.stft(y=y,hop_length=MEL_HOP_LENGTH,n_fft=N_FFT)
+    t = name.split('.')[0].split('-')
+    s['instrument'] = t[0]
+    s['style'] = t[1]
+    s['pitch_name'] = t[2]
+    s['nuance'] = t[3]
+    if len(t) == 5:
+        s['other'] = t[4]
+    
+    return s
+    
+
+def recursiveSearch(path):
+    l = []
+    for x in os.listdir(path):
+        p = os.path.join(path,x)
+        if os.path.isdir(p):
+            l.extend(recursiveSearch(p))
+        elif os.path.isfile(p) and p.split('.')[-1] == 'wav':
+            l.append(getSampleMetaData(p,x))
+    return l
+
+def getPitchRange(d):
+    pr = set()
+    for i in d:
+        pr.add(i['pitch_name'])
+    pr = list(pr)
+    pr.sort(key = lambda x : 10000*ord(x[-1]) + 100*((ord(x[0])-ord('C'))%7) + (len(x)==3))
+    return pr
+
+
+def generateDBLabels(path, nb_pitch_range = 1, instr_filter=None):
+    '''
+        Generates the short term fourier transforms for each sample of the data base.
+        In order to ensure a good repartition of the pitches when generating the combinations,
+        the set is divided into pitch_bins (the total number of bins amond the pitch range is defined
+        by nb_pitch_range).
+        If instr_filter is not None, the generated set will contain only samples of the instruments
+        specified by instr_filter.
+        
+        Note that if the global paramaters MEL_HOP_LENGTH, N_FFT or TIME_LENGTH are modified,
+        this function needs to be called again
+    '''
+    l = recursiveSearch(path)
+    if not instr_filter is None:
[…]

### 2020-11-09 CSMC + MuMe 2020 /sampleRNN/model.py

@@ -1,0 +1,318 @@
+import nn
+import utils
+
+import torch
+from torch.nn import functional as F
+from torch.nn import init
+
+import numpy as np
+
+
+class SampleRNN(torch.nn.Module):
+
+    def __init__(self, frame_sizes, n_rnn, dim, learn_h0, q_levels,
+                 weight_norm):
+        super().__init__()
+
+        self.dim = dim
+        self.q_levels = q_levels
+
+        ns_frame_samples = map(int, np.cumprod(frame_sizes))
+        self.frame_level_rnns = torch.nn.ModuleList([
+            FrameLevelRNN(
+                frame_size, n_frame_samples, n_rnn, dim, learn_h0, weight_norm
+            )
+            for (frame_size, n_frame_samples) in zip(
+                frame_sizes, ns_frame_samples
+            )
+        ])
+
+        self.sample_level_mlp = SampleLevelMLP(
+            frame_sizes[0], dim, q_levels, weight_norm
+        )
+
+    @property
+    def lookback(self):
+        return self.frame_level_rnns[-1].n_frame_samples
+
+
+class FrameLevelRNN(torch.nn.Module):
+
+    def __init__(self, frame_size, n_frame_samples, n_rnn, dim,
+                 learn_h0, weight_norm):
+        super().__init__()
+
+        self.frame_size = frame_size
+        self.n_frame_samples = n_frame_samples
+        self.dim = dim
+
+        h0 = torch.zeros(n_rnn, dim)
+        if learn_h0:
+            self.h0 = torch.nn.Parameter(h0)
+        else:
+            self.register_buffer('h0', torch.autograd.Variable(h0))
+
+        self.input_expand = torch.nn.Conv1d(
+            in_channels=n_frame_samples,
+            out_channels=dim,
+            kernel_size=1
+        )
+        init.kaiming_uniform_(self.input_expand.weight)
+        init.constant_(self.input_expand.bias, 0)
+        if weight_norm:
+            self.input_expand = torch.nn.utils.weight_norm(self.input_expand)
+
+        self.rnn = torch.nn.GRU(
+            input_size=dim,
+            hidden_size=dim,
+            num_layers=n_rnn,
+            batch_first=True
+        )
+        for i in range(n_rnn):
+            nn.concat_init(
+                getattr(self.rnn, 'weight_ih_l{}'.format(i)),
[…]

### 2020-11-09 CSMC + MuMe 2020 /sampleRNN/parameters.py

@@ -1,0 +1,133 @@
+import pickle
+
+path = "./TinySOL"
+MEL_HOP_LENGTH = 2048
+RATE = 44100
+N_FFT = 2048
+N_MELS = 128
+TIME_LENGTH = 4
+FEATURE_TYPE = 'mel'
+
+N = 4
+nb_samples = 100000
+rdm_granularity = 10
+nb_pitch_range = 8
+instr_filter = ['Hn','Ob','Vn','Va', 'Vc']
+batch_size = 16
+model_type = 'cnn'
+nb_epoch = 50
+train_proportion = 0.8
+
+coeff_freq_shift_data_augment = 0.005 # For data augmentation, proportional change to sampling rate
+delay_offset_avg = 0.005 # Average of offset applied to delay filters (ie the position of the first impulse)
+delay_period_avg = 0.002 # Average of the period of the impulses for delaying
+delay_feedback_avg = 0.5 # Average of the feedback factor of delay impulses
+prop_zero_row = 0.01
+prop_zero_col = 0.01
+noise_kernel_var = 0.0001
+
+
+
+resume_model = False
+model_path = './model'
+model_run_resume = 4
+model_epoch_resume = 23
+
+
+def load_parameters(path):
+    global MEL_HOP_LENGTH
+    global RATE
+    global N_FFT
+    global N_MELS
+    global TIME_LENGTH
+    global FEATURE_TYPE
+    global PITCH_REGROUP
+    
+    global N
+    global nb_samples
+    global rdm_granularity
+    global nb_pitch_range
+    global instr_filter
+    global batch_size
+    global model_type
+    global nb_epoch
+    global train_proportion
+    
+    global coeff_freq_shift_data_augment
+    global delay_offset_avg
+    global delay_period_avg
+    global delay_feedback_avg
+    global noise_kernel_var
+    global prop_zero_row
+    global prop_zero_col
+    
+    f = open(path + '/params.pkl', 'rb')
+    params = pickle.load(f)
+    f.close()
+    
+    MEL_HOP_LENGTH = params['MEL_HOP_LENGTH']
+    RATE = params['RATE']
+    N_FFT = params['N_FFT']
+    N_MELS = params['N_MELS']
+    TIME_LENGTH = params['TIME_LENGTH']
+    FEATURE_TYPE = params['FEATURE_TYPE']
+    
+    N = params['N']
+    nb_samples = params['nb_samples']
+    rdm_granularity = params['rdm_granularity']
+    nb_pitch_range = params['nb_pitch_range']
[…]

### 2020-11-09 CSMC + MuMe 2020 /sampleRNN/train.py

@@ -1,0 +1,393 @@
+# CometML needs to be imported first.
+try:
+    import comet_ml
+except ImportError:
+    pass
+
+from model import SampleRNN, Predictor
+from optim import gradient_clipping
+from nn import sequence_nll_loss_bits
+from trainer import Trainer
+from trainer.plugins import (
+    TrainingLossMonitor, ValidationPlugin, AbsoluteTimeMonitor, SaverPlugin,
+    GeneratorPlugin, StatsPlugin
+)
+from dataset import FolderDataset, DataLoader
+
+import torch
+from torch.utils.trainer.plugins import Logger
+
+from natsort import natsorted
+
+from functools import reduce
+import os
+import shutil
+import sys
+from glob import glob
+import re
+import argparse
+
+import numpy as np
+from OrchDataset import OrchDataSet, RawDatabase
+
+default_params = {
+    # model parameters
+    'n_rnn': 1,
+    'dim': 1024,
+    'learn_h0': True,
+    'q_levels': 256,
+    'seq_len': 1024,
+    'weight_norm': True,
+    'batch_size': 128,
+    'val_frac': 0.1,
+    'test_frac': 0.1,
+
+    # training parameters
+    'keep_old_checkpoints': False,
+    'datasets_path': 'datasets',
+    'results_path': 'results',
+    'epoch_limit': 1000,  # default: 1000
+    'resume': True,
+    'sample_rate': 16000,
+    'n_samples': 1,
+    'sample_length': 30000,  # 8s -> 80000
+    'loss_smoothing': 0.99,
+    'cuda': True,
+    'comet_key': None
+}
+
+tag_params = [
+    'exp', 'frame_sizes', 'n_rnn', 'dim', 'learn_h0', 'q_levels', 'seq_len',
+    'batch_size', 'dataset', 'val_frac', 'test_frac'
+]
+
+def param_to_string(value):
+    if isinstance(value, bool):
+        return 'T' if value else 'F'
+    elif isinstance(value, list):
+        return ','.join(map(param_to_string, value))
+    else:
+        return str(value)
+
+def make_tag(params):
+    return '-'.join(
+        key + ':' + param_to_string(params[key])
+        for key in tag_params
+        if key not in default_params or params[key] != default_params[key]
+    )
+
+def setup_results_dir(params):
+    def ensure_dir_exists(path):
+        if not os.path.exists(path):
+            os.makedirs(path)
+
+    tag = make_tag(params)
[…]

### 2020-11-09 CSMC + MuMe 2020 /sampleRNN/trainer/plugins.py

@@ -1,0 +1,291 @@
+import matplotlib
+matplotlib.use('Agg')
+
+# from model import Generator
+
+import torch
+from torch.autograd import Variable
+# from torch.utils.trainer.plugins.plugin import Plugin
+# from torch.utils.trainer.plugins.monitor import Monitor
+# from torch.utils.trainer.plugins import LossMonitor
+
+from librosa.output import write_wav
+from matplotlib import pyplot
+
+from glob import glob
+import os
+import pickle
+import time
+
+
+# class TrainingLossMonitor(LossMonitor):
+
+#     stat_name = 'training_loss'
+
+
+# class ValidationPlugin(Plugin):
+
+#     def __init__(self, val_dataset, test_dataset):
+#         super().__init__([(1, 'epoch')])
+#         self.val_dataset = val_dataset
+#         self.test_dataset = test_dataset
+
+#     def register(self, trainer):
+#         self.trainer = trainer
+#         val_stats = self.trainer.stats.setdefault('validation_loss', {})
+#         val_stats['log_epoch_fields'] = ['{last:.4f}']
+#         test_stats = self.trainer.stats.setdefault('test_loss', {})
+#         test_stats['log_epoch_fields'] = ['{last:.4f}']
+
+#     def epoch(self, idx):
+#         self.trainer.model.eval()
+
+#         val_stats = self.trainer.stats.setdefault('validation_loss', {})
+#         val_stats['last'] = self._evaluate(self.val_dataset)
+#         test_stats = self.trainer.stats.setdefault('test_loss', {})
+#         test_stats['last'] = self._evaluate(self.test_dataset)
+
+#         self.trainer.model.train()
+
+#     def _evaluate(self, dataset):
+#         loss_sum = 0
+#         n_examples = 0
+#         for data in dataset:
+#             batch_inputs = data[: -1]
+#             batch_target = data[-1]
+#             batch_size = batch_target.size()[0]
+
+#             def wrap(input):
+#                 if torch.is_tensor(input):
+#                     input = Variable(input, volatile=True)
+#                     if self.trainer.cuda:
+#                         input = input.cuda()
+#                 return input
+#             batch_inputs = list(map(wrap, batch_inputs))
+
+#             batch_target = Variable(batch_target, volatile=True)
+#             if self.trainer.cuda:
+#                 batch_target = batch_target.cuda()
+
+#             batch_output = self.trainer.model(*batch_inputs)
+#             loss_sum += self.trainer.criterion(batch_output, batch_target) \
[…]

### 2020-11-09 CSMC + MuMe 2020 /script_generate_rdb.py

@@ -1,0 +1,10 @@
+import parameters
+from OrchDataset_2 import RawDatabase
+
+print("Generating raw database")
+rdb = RawDatabase('../rdb.pkl',
+                  random_granularity=parameters.GLOBAL_PARAMS.rdm_granularity,
+                  instr_filter=parameters.GLOBAL_PARAMS.instr_filter)
+print("Saving raw database")
+rdb.save()
+print("Done")

### 2020-11-09 CSMC + MuMe 2020 /showOutputs.py

@@ -1,0 +1,79 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.animation import FuncAnimation
+from matplotlib import cm
+from matplotlib.colors import ListedColormap
+
+from train import getPosNMax, prediction
+
+path = './model/run4/outputs.npy'
+
+viridis = cm.get_cmap('viridis')
+prec = 0.2
+cmap = ListedColormap([viridis(x**prec) for x in np.linspace(0,1,256)])
+
+fig = plt.figure()
+ax1 = fig.add_subplot(2,3,1)
+ax3 = fig.add_subplot(2,3,2)
+ax4 = fig.add_subplot(2,3,3)
+ax2 = fig.add_subplot(2,1,2)
+
+loss_array = []
+pred_loss = []
+avg_pred_loss = []
+avg_length = 100
+
+global prev
+prev = (-1,-1)
+
+def load(i):
+    global prev
+    r = np.load(path, allow_pickle=True).item()
+    epoch = r.get('epoch')
+    step = r.get('step')
+    outputs = r.get('outputs')
+    labels = r.get('labels')
+    
+    
+    if (epoch, step) != prev:
+        prev = (epoch, step)
+        loss_array.append(r.get('loss'))
+        ax2.cla()
+        ax1.cla()
+        ax3.cla()
+        ax4.cla()
+        plt.title('Epoch {}; Step {}'.format(epoch, step))
+        line = np.ones((1,outputs.shape[1]))
+        line[0,0] = 0.0
+        line[0,-1] = 0.0
+        N = len([x for x in labels[0] if x > 0.0])
+        pred = prediction(outputs, N)
+        pred_loss.append(np.sum(pred*labels)/max(1.0,np.sum(labels)))
+        wdw = min(len(pred_loss), avg_length)
+        ax1.plot(loss_array)
+        ax1.plot([len(loss_array)-1, len(loss_array)-1], [r.get('loss_min'), loss_array[-1]], marker='o')
+        ax1.grid()
+        ax1.set_title('BCE with logits loss')
+        avg_pred_loss.append(sum(pred_loss[-wdw:])/wdw)
+        ax2.imshow(np.vstack([outputs, line, labels, line, pred]), cmap=cmap)
+        ax3.plot(pred_loss)
+        ax3.plot(avg_pred_loss)
+        ax3.set_title('Accuracy over batch')
+        ax3.grid()
[…]

### 2020-11-09 CSMC + MuMe 2020 /test.py

@@ -1,0 +1,399 @@
+import torch
+import numpy as np
+import librosa
+import librosa.display
+import os
+import random
+import soundfile as sf
+
+from model import OrchMatchNet
+from OrchDataset import RawDatabase
+from train import getPosNMax
+from parameters import GLOBAL_PARAMS
+
+""" 
+To run this file:
+- You need a folder of targets to be orchestrated, and set 'target_path' to point to this
+- You need to create a folder to store the orchestrated solutions, and set 'solutions_path'
+to be the path to this folder
+- You need a trained model saved as a .pth file, and set 'state_path' to point to this
+
+- Set 'tinysol_path' to point to your TinySOL database
+- Set model_type, instr_filter, and n
+
+"""
+
+
+# path to TinySOL data
+tinysol_path = './TinySOL'
+
+# path to target samples
+target_path = './targets'
+
+# number of samples to be used in solution
+n = 10
+
+# instruments to be used (all instruments will be used)
+instr_filter = ['Hn', 'Ob', 'Vn', 'Va',
+                'Vc', 'Fl', 'Tbn', 'Bn', 'TpC', 'ClBb']
+
+# if true, doesn't orchestrate targets but instead creates combinations of TinySOL samples to input to the model
+sanity_check = False
+
+
+def test(model, state_path, data, targets):
+    device = torch.device('cpu')
+    state = torch.load(state_path, map_location=device)
+    model.load_state_dict(state['state_dict'])
+    model.eval()
+
+    outputs = model(data)
+    outputs = outputs.detach().cpu().clone().numpy()
+
+    if sanity_check:
+        accuracy = evaluate(outputs, labels)
+        print(accuracy)
+
+    # text file to write solutions to
+    f = open(solutions_path + '/orchestration_results.txt', 'w+')
+
+    for i in range(len(outputs)):
+        output = outputs[i]
+        target = targets[i]
+        # get indices of top n probabilities
+        indices = getPosNMax(output, n)
+        # turn indices into [instr, pitch]
+        classes = get_classes(indices)
+        # get top n probabiltiies
+        probs = [output[i] for i in indices]
+        # turn probabilities into dynamic markings (pp, mf, ff)
+        dynamics = prob_to_dynamic(probs)
+        # combine into (instr, pitch, dynamic)
+        for j in range(len(classes)):
+            classes[j].append(dynamics[j])
+        
[…]

### 2020-11-09 CSMC + MuMe 2020 /train.py

@@ -1,0 +1,431 @@
+import torch
+import torch.nn as nn
+import torchvision.transforms as transforms
+import torch.nn.functional as F
+from torch.autograd import Variable
+from torch.utils import data
+import numpy as np
+import argparse
+import time
+import pickle
+import os
+from functools import reduce
+import librosa.display
+import matplotlib.pyplot as plt
+from matplotlib.animation import FuncAnimation
+
+from model import OrchMatchNet
+from parameters import GLOBAL_PARAMS, SimParams
+from OrchDataset import OrchDataSet,RawDatabase
+
+class Timer:
+    def __init__(self, size_buffer):
+        self.measuring = False
+        self.begin = []
+        self.end = []
+        self.size_buffer = size_buffer
+        
+    def start(self):
+        if not self.measuring:
+            if len(self.begin) == self.size_buffer:
+                self.begin.pop(0)
+            self.begin.append(time.time())
+            self.measuring = True
+            
+    def stop(self):
+        if self.measuring:
+            if len(self.end) == self.size_buffer:
+                self.end.pop(0)
+            self.end.append(time.time())
+            self.measuring = False
+    
+    def estimate(self, remaining):
+        if len(self.end) > 0 and not self.measuring:
+            diff = np.array(self.end)-np.array(self.begin)
+            diff = np.mean(diff)
+            h,r=divmod(diff*remaining,3600)
+            m,s=divmod(r,60)
+            print(diff)
+            print("Estimated remaining time : {}h{}m{}s".format(int(h), int(m), int(s)))
+            
+
+class DataSaver:
+    '''
+        Class used to write data for visualization (see showOutputs.py)
+    '''
+    def __init__(self, path_file):
+        self.path = path_file
+        self.count = 0
+        
+    def save(self, outs, labels, loss, loss_min, epoch, step):
+        r = {}
+        r['epoch'] = epoch
+        r['step'] = step
+        r['outputs'] = outs
+        r['labels'] = labels
+        r['loss'] = loss
+        r['loss_min'] = loss_min
+        np.save(self.path, r)
+        self.count += 1
+
+timer = Timer(1000)
+
+def main(resume_model, rdb = None):
+    #print("Starting parsing -----")
+    #arg = arg_parse()
+    #print("End parsing")
+
[…]

### 2020-11-10 OrchDataset.py
Create new file structure

@@ -1,295 +1,0 @@
-import itertools
-from functools import reduce
-from generateDB import generateDBLabels
-import librosa.display
-import random
-import math
-from torch.utils import data
-import torch
-import numpy as np
-import os
-import pickle
-import random
-import hashlib
-
-from parameters import GLOBAL_PARAMS
-
-def showSample(s):
-    print(s[1])
-    librosa.display.specshow(s[0], x_axis='time',y_axis='mel',sr=44100)
-
-class IndexIncrementor:
-    '''
-        This class provides a way to iterate over the set of features contained in 
-        a RawDatabase object to ensure that all pitch bins are uniformly chosen.
-        This helps generating a homogeneous database that doesn't miss some dynamics.
-        
-        A certain degree of randomness is added in the sense that the step and starting point
-        are randomized.
-    '''
-    def __init__(self, N, base, offset=None):
-        self.N = N
-        if type(base) == int:
-            self.base = [base]*N
-        elif type(base) == list and len(base) == N:
-            self.base = base
-        else:
-            raise Exception("Wrong argument for IndexIncrementor (must be int or list of size N)")
-        self.max_count = reduce((lambda x,y:x*y), self.base)
-        self.restart(offset)
-    
-    def findStep(self):
-        self.step = int(random.randint(int(0.05*self.max_count),int(0.95*self.max_count)))
-        while math.gcd(self.step, self.max_count) > 1:
-            self.step += 1
-    
-    def add(self, e):
-        r = e
-        i = 0
-        while i < self.N and r != 0:
-            r,self.index[i] = divmod(self.index[i]+r,self.base[i])
-            i += 1
-            
-    def incr(self):
-        self.add(self.step)
-        self.count += 1
-        
-    def restart(self, offset = None):
-        self.count = 0
-        self.index = [0]*self.N
-        self.findStep()
-        if offset is None:
-            self.add(random.randint(0,self.max_count-1))
-        elif offset >= 0:
-            self.add(offset)
-        else:
-            raise Exception("Wrong offset")
-    
-    def isOver(self):
-        return self.count >= self.max_count
-            
-class HashTable:
-    def __init__(self, max_depth, depth=0):
-        self.depth = depth
-        self.max_depth = max_depth
-        if max_depth == depth+1:
-            self.data = []
-        else:
-            self.data = {}
-            for i in '0123456789abcdef':
-                self.data[i] = HashTable(max_depth,depth+1)
-            
[…]

### 2020-11-10 augment.py

@@ -1,81 +1,0 @@
-import librosa
-import numpy as np
-import copy
-import random
-from sklearn.utils import shuffle
-
-
-def set_length(new_y, y):
-    if len(new_y) < len(y):
-        new_y = librosa.util.fix_length(new_y, len(y))
-    elif len(new_y) > len(y):
-        new_y = new_y[0:len(y)]
-    return new_y
-
-
-def random_stretch(y, sr, prob=1., min_stretch=0.9, max_stretch=1.1):
-    if np.random.uniform(0, 1) < prob:
-        factor = np.random.uniform(min_stretch, max_stretch)
-        new_y = librosa.effects.time_stretch(y, factor)
-        return set_length(new_y, y)
-    else:
-        return y
-
-
-def random_pitch(y, sr, prob=1., min_shift=-0.3, max_shift=0.3, bins_per_octave=12):
-    if np.random.uniform(0, 1) < prob:
-        steps = np.random.uniform(min_shift, max_shift)
-        new_y = librosa.effects.pitch_shift(
-            y, sr, n_steps=steps, bins_per_octave=bins_per_octave)
-        return set_length(new_y, y)
-    else:
-        return y
-
-
-def wav_augment(y, sr):
-    functions = shuffle([random_pitch, random_stretch])
-    for fn in functions:
-        y = fn(y, sr=sr)
-    return y
-
-
-def freq_mask(spec, F=30, num_masks=1, replace_with_zero=False):
-    cloned = copy.deepcopy(spec)
-    num_mel_channels = cloned.shape[0]
-    for i in range(0, num_masks):
-        f = random.randrange(0, F)
-        f_zero = random.randrange(0, num_mel_channels - f)
-        # avoids randrange error if values are equal and range is empty
-        if (f_zero == f_zero + f):
-            return cloned
-        mask_end = random.randrange(f_zero, f_zero + f)
-        if (replace_with_zero):
-            cloned[0][f_zero:mask_end] = 0
-        else:
-            cloned[0][f_zero:mask_end] = cloned.mean()
-    return cloned
-
-
-def time_mask(spec, T=40, num_masks=1, replace_with_zero=False):
-    cloned = copy.deepcopy(spec)
-    len_spectro = cloned.shape[1]
-    for i in range(0, num_masks):
-        t = random.randrange(0, T)
-        t_zero = random.randrange(0, len_spectro - t)
-        # avoids randrange error if values are equal and range is empty
[…]

### 2020-11-10 baseline_classification.py

@@ -1,293 +1,0 @@
-# Adapted from code written by Carmine E. Cella, 2020
-
-
-import numpy as np
-import pandas as pd
-import matplotlib
-import matplotlib.pyplot as plt
-from sklearn.preprocessing import LabelEncoder, StandardScaler
-from sklearn.linear_model import LogisticRegression
-from sklearn.pipeline import Pipeline
-from sklearn.model_selection import train_test_split
-from sklearn.metrics import accuracy_score
-from sklearn.decomposition import PCA
-from sklearn.model_selection import cross_val_score
-from sklearn.svm import SVC
-from sklearn.neighbors import KNeighborsClassifier
-from sklearn.model_selection import KFold
-from sklearn.ensemble import RandomForestClassifier
-from sklearn.multioutput import MultiOutputClassifier
-from collections import Counter
-from matplotlib import pyplot as plt
-
-import librosa
-import librosa.display
-import json
-import os
-import random
-import sys
-from time import process_time
-
-# raw dataset
-# file hierarchy: (note that folders Brass, Winds, Strings are not present)
-# ----TinySOL
-#   ----Bn
-#   ----Cb
-#   ----Va
-#   etc...
-path = './TinySOL_0.6/TinySOL'
-
-# time duration
-time = 4
-
-# number of samples between successive frames in librosa.melspectrogram
-mel_hop_length = 44100
-
-pitch_classes = ['A', 'A#', 'B', 'C', 'C#', 'D', 'D#', 'E', 'F', 'F#', 'G', 'G#']
-
-# maps an instrument or instrument-pitch to a class #
-label_mapping = {}
-
-def mix(fa, fb):
-    diff = len(fa) - len(fb)
-
-    if diff > 0:
-        add = np.zeros((1, diff), dtype=np.float32)
-        fb = np.append(fb, add)
-    elif diff < 0:
-        add = np.zeros((1, -diff), dtype=np.float32)
-        fa = np.append(fa, add)
-
-    return fa+fb
-    
-def combine_sounds(soundlist):
-    mixed_file = np.zeros((1, 1))
-    sr = 0
-    for sound in soundlist:
-        sound_path = os
-        sfile, sr = librosa.load(sound, sr=None)
-        if len(sfile) > time*sr:
-            # randomly select one part of the raw audio
-            n = np.random.randint(0, len(sfile)-time*sr)
-            sfile = sfile[n:n+time*sr]
-        # add augment
[…]

### 2020-11-10 latent_space_vis.py

@@ -1,75 +1,0 @@
-import librosa
-import torch
-import torch.nn as nn
-import torchvision.transforms as transforms
-import torch.nn.functional as F
-from torch.autograd import Variable
-from torch.utils import data
-import numpy as np
-import matplotlib.pyplot as plt
-import pickle
-
-from model import OrchMatchNet
-from parameters import GLOBAL_PARAMS, SimParams
-from OrchDataset_2 import RawDatabase, OrchDataSet
-
-path = "D:/DeepOrchestration/New_tests/CNN/run0/epoch_39.pth"
-
-print("Loading rdb")
-try:
-    rdb
-except:
-    with open("D:/DeepOrchestration/rdb.pkl", 'rb') as f:
-        rdb = pickle.load(f)
-    
-GLOBAL_PARAMS.lab_class = {}
-GLOBAL_PARAMS.N = 10
-GLOBAL_PARAMS.nb_samples = 100
-
-tot_size = 0
-for i in rdb.db:
-    GLOBAL_PARAMS.lab_class[i] = {}
-    a = set()
-    for k in rdb.db[i]:
-        for j in k:
-            a.add(j['pitch_name'])
-    for x in a:
-        GLOBAL_PARAMS.lab_class[i][x] = tot_size
-        tot_size += 1
-    
-def class_encoder(list_samp):
-    label = [0 for i in range(tot_size)]
-    for s in list_samp:
-        label[GLOBAL_PARAMS.lab_class[s['instrument']][s['pitch_name']]] = 1
-    return np.array(label).astype(np.float32)
-    
-train_dataset = OrchDataSet(rdb,class_encoder, GLOBAL_PARAMS.FEATURE_TYPE)
-train_dataset.generate(GLOBAL_PARAMS.N, GLOBAL_PARAMS.nb_samples)
-
-print("Loading model")
-print("Shape : {}".format(train_dataset[0][0].shape[1:]))
-model = OrchMatchNet(424, 'cnn', train_dataset[0][0].shape[1:])
-state = torch.load(path)
-model.load_state_dict(state['state_dict'])
-    
-print("Generating sample")
-samp = train_dataset[0][0].view([1,1,128,87])
-
-res = [samp]
-res.extend(model.getLatentSpace(samp))
-
-def plot_latent_space(res, layer):
-    nrows = 2**(int(np.log2(res[layer].shape[1]))//2)
-    ncols = res[layer].shape[1]/nrows
-    
-    plt.figure()
[…]

### 2020-11-10 model.py

@@ -1,121 +1,0 @@
-import torch
-import numpy as np
-import torch.nn as nn
-import torch.nn.functional as F
-
-from resnet import ResNet, count_parameters, init_weights, device
-
-
-class CNN(nn.Module):
-    def __init__(self, out_num, features_dim):
-        super(CNN, self).__init__()
-
-        shape_input = features_dim
-        self.layer1 = nn.Sequential(
-            nn.Conv2d(in_channels=1, out_channels=8,
-                      kernel_size=3, stride=1, padding=1),
-            nn.BatchNorm2d(num_features=8),
-            nn.ReLU(inplace=True),
-            nn.MaxPool2d(kernel_size=2)
-        )
-        shape_input = (shape_input[0]//2, shape_input[1]//2)
-
-        self.layer2 = nn.Sequential(
-            nn.Conv2d(in_channels=8, out_channels=16,
-                      kernel_size=3, stride=1, padding=1),
-            nn.BatchNorm2d(num_features=16),
-            nn.ReLU(inplace=True),
-            nn.MaxPool2d(kernel_size=2)
-        )
-        shape_input = (shape_input[0]//2, shape_input[1]//2)
-
-        self.layer3 = nn.Sequential(
-            nn.Conv2d(in_channels=16, out_channels=32,
-                      kernel_size=3, stride=1, padding=1),
-            nn.BatchNorm2d(num_features=32),
-            nn.ReLU(inplace=True),
-            nn.MaxPool2d(kernel_size=2)
-        )
-        shape_input = (shape_input[0]//2, shape_input[1]//2)
-        self.lstm_shape = (shape_input[0], shape_input[1])
-
-        self.lstm = nn.LSTM(input_size=shape_input[1], hidden_size=shape_input[1], batch_first=True)
-
-        self.layer4 = nn.Sequential(
-            nn.Conv2d(in_channels=32, out_channels=32,
-                      kernel_size=3, stride=1, padding=1),
-            nn.ReLU(inplace=True),
-            nn.MaxPool2d(kernel_size=2),
-            nn.Dropout(0.5)
-        )
-        shape_input = (shape_input[0]//2, shape_input[1]//2)
-
-        self.fc1 = nn.Sequential(
-            nn.Linear(in_features=32*shape_input[0]*shape_input[1], out_features=2048),
-            nn.ReLU(inplace=True),
-            nn.Dropout(0.5)
-        )
-
-        self.fc3 = nn.Linear(in_features=2048, out_features=out_num)
-
[…]

### 2020-11-10 parameters.py

@@ -1,46 +1,0 @@
-import pickle
-
-class SimParams:
-    def __init__(self):
-        self.path = "./TinySOL"
-        self.MEL_HOP_LENGTH = 2048
-        self.RATE = 44100
-        self.N_FFT = 2048
-        self.N_MELS = 128
-        self.TIME_LENGTH = 4
-        self.FEATURE_TYPE = 'mel'
-        
-        self.N = 2
-        self.nb_samples = 400000
-        self.rdm_granularity = 10
-        self.nb_pitch_range = 8
-        self.instr_filter = ['Hn','Ob','Vn','Va', 'Vc', 'Fl', 'Tbn', 'Bn', 'TpC', 'ClBb'][:10]
-        self.batch_size = 16
-        self.model_type = 'resnet'
-        self.nb_epoch = 100
-        self.train_proportion = 0.8
-        
-        self.coeff_freq_shift_data_augment = 0.008 # For data augmentation, proportional change to sampling rate
-        self.delay_offset_avg = 0.005 # Average of offset applied to delay filters (ie the position of the first impulse)
-        self.delay_period_avg = 0.002 # Average of the period of the impulses for delaying
-        self.delay_feedback_avg = 0.5 # Average of the feedback factor of delay impulses
-        self.prop_zero_row = 0.01
-        self.prop_zero_col = 0.01
-        self.noise_kernel_var = 0.0001
-        
-        self.lab_class = {} # Contains the labels
-        
-    def load_parameters(self, path):
-        f = open(path + '/params.pkl', 'rb')
-        p = pickle.load(f)
-        f.close()
-        
-        for attr in p.__dict__:
-            setattr(self, attr, getattr(p, attr))
-    
-    def save_parameters(self, path):
-        f = open(path + '/params.pkl', 'wb')
-        pickle.dump(self, f)
-        f.close()
-
-GLOBAL_PARAMS = SimParams()

### 2020-11-10 plot.py

@@ -1,138 +1,0 @@
-import matplotlib.pyplot as plt
-import pickle
-import os
-import parameters
-
-def f(l):
-    Y = [0,0,0,0,0,0]
-    for i in l:
-        Y[X.index(i[0])] = i[1]['acc']
-    return Y
-
-order = ['acc', 'Fl', 'ClBb', 'Ob', 'Bn', 'Va', 'Vn', 'Vc', 'TpC', 'Tbn', 'Hn']
-
-def plot_best_acc(path):
-    
-    list_accs = []
-    for i in os.listdir(path):
-        accs = {}
-        MM = len([x for x in os.listdir('{}/{}'.format(path, i)) if 'result' in x])
-        epochs = list(range(4,MM*5,5))
-        max_epoch = -1
-        
-        parameters.GLOBAL_PARAMS.load_parameters('{}/{}'.format(path, i))
-    
-        for j in epochs:
-            f = open('{}/{}/result{}.pkl'.format(path, i, j), 'rb')
-            r = pickle.load(f)
-            if len(accs.keys()) == 0:
-                accs['acc'] = 0
-                for k in r['pitch_acc']:
-                    accs[k] = 0
-            
-            if r['acc'] > accs['acc']:
-                max_epoch = j
-                accs['acc'] = r['acc']
-                for k in r['pitch_acc']:
-                    accs[k] = r['pitch_acc'][k]
-            f.close()
-            
-        list_accs.append((parameters.GLOBAL_PARAMS.N, accs, max_epoch, epochs[-1]))
-    plt.figure(figsize=(6,4))
-    for i in order:
-        accs = []
-        for x in list_accs:
-            accs.append((x[0], x[1][i]))
-            
-        accs.sort(key = lambda x : x[0])
-        
-        if i == 'acc':
-            plt.plot([x[0] for x in accs], [x[1] for x in accs], '-o', label = "Overall accuracy")
-        else:
-            plt.plot([x[0] for x in accs], [x[1] for x in accs], '--x', label = "{} accuracy".format(i))
-    plt.grid()
-    plt.xlabel("Number of samples in the combinations")
-    plt.ylabel("Accuracy")
-    plt.title("Maximum test accuracy achieved vs number of samples N".format(i))
-        
-    return list_accs
-
-def plot_run(path, run_nb):
-    res = {'acc' : []}
-    
-    MM = len([x for x in os.listdir('{}/run{}'.format(path, run_nb)) if 'result' in x])
-        
-        
-    epochs = list(range(4,MM*5,5))
[…]

### 2020-11-10 resnet.py

@@ -1,159 +1,0 @@
-import librosa
-import os
-import sys
-import pickle
-import time
-import librosa
-import torch
-import numpy as np
-from torch import nn
-import torch.nn.functional as F
-from torch import optim
-from torch.distributions.categorical import Categorical
-
-
-device = torch.device('cuda' if torch.cuda.is_available() else 'cpu')
-
-
-def count_parameters(model):
-    counts = sum(p.numel() for p in model.parameters() if p.requires_grad)
-    print('The model has {:,} trainable parameters'.format(counts))
-
-
-def init_weights(model):
-    for name, param in model.named_parameters():
-        nn.init.normal_(param.data, mean=0, std=0.01)
-
-
-class ResidualBlock(nn.Module):
-    def __init__(self, in_channels, out_channels, stride=1):
-        super(ResidualBlock, self).__init__()
-
-        # Conv Layer 1
-        self.conv1 = nn.Conv2d(
-            in_channels=in_channels, out_channels=out_channels,
-            kernel_size=(3, 3), stride=stride, padding=1, bias=True
-        )
-        self.bn1 = nn.BatchNorm2d(out_channels)
-
-        # Conv Layer 2
-        self.conv2 = nn.Conv2d(
-            in_channels=out_channels, out_channels=out_channels,
-            kernel_size=(3, 3), stride=1, padding=1, bias=True
-        )
-        self.bn2 = nn.BatchNorm2d(out_channels)
-
-        self.shortcut = nn.Sequential()
-        if stride != 1 or in_channels != out_channels:
-            self.shortcut = nn.Sequential(
-                nn.Conv2d(
-                    in_channels=in_channels, out_channels=out_channels,
-                    kernel_size=(1, 1), stride=stride, bias=False
-                ),
-                nn.BatchNorm2d(out_channels)
-            )
-
-    def forward(self, x):
-        out = nn.ReLU()(self.bn1(self.conv1(x)))
-        out = self.bn2(self.conv2(out))
-        out += self.shortcut(x)
-        out = nn.ReLU()(out)
-        return out
-
-
-class ResNet(nn.Module):
-    def __init__(self, num_classes, features_dim, dropout_rate=0.5, output_nonlinearity='sigmoid'):
-        super(ResNet, self).__init__()
-        # Initial input conv
-        shape_input = features_dim
-        self.conv1 = nn.Conv2d(
[…]

### 2020-11-10 script_generate_rdb.py

@@ -1,10 +1,0 @@
-import parameters
-from OrchDataset_2 import RawDatabase
-
-print("Generating raw database")
-rdb = RawDatabase('../rdb.pkl',
-                  random_granularity=parameters.GLOBAL_PARAMS.rdm_granularity,
-                  instr_filter=parameters.GLOBAL_PARAMS.instr_filter)
-print("Saving raw database")
-rdb.save()
-print("Done")

### 2020-11-10 showOutputs.py

@@ -1,79 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib.animation import FuncAnimation
-from matplotlib import cm
-from matplotlib.colors import ListedColormap
-
-from train import getPosNMax, prediction
-
-path = './model/run4/outputs.npy'
-
-viridis = cm.get_cmap('viridis')
-prec = 0.2
-cmap = ListedColormap([viridis(x**prec) for x in np.linspace(0,1,256)])
-
-fig = plt.figure()
-ax1 = fig.add_subplot(2,3,1)
-ax3 = fig.add_subplot(2,3,2)
-ax4 = fig.add_subplot(2,3,3)
-ax2 = fig.add_subplot(2,1,2)
-
-loss_array = []
-pred_loss = []
-avg_pred_loss = []
-avg_length = 100
-
-global prev
-prev = (-1,-1)
-
-def load(i):
-    global prev
-    r = np.load(path, allow_pickle=True).item()
-    epoch = r.get('epoch')
-    step = r.get('step')
-    outputs = r.get('outputs')
-    labels = r.get('labels')
-    
-    
-    if (epoch, step) != prev:
-        prev = (epoch, step)
-        loss_array.append(r.get('loss'))
-        ax2.cla()
-        ax1.cla()
-        ax3.cla()
-        ax4.cla()
-        plt.title('Epoch {}; Step {}'.format(epoch, step))
-        line = np.ones((1,outputs.shape[1]))
-        line[0,0] = 0.0
-        line[0,-1] = 0.0
-        N = len([x for x in labels[0] if x > 0.0])
-        pred = prediction(outputs, N)
-        pred_loss.append(np.sum(pred*labels)/max(1.0,np.sum(labels)))
-        wdw = min(len(pred_loss), avg_length)
-        ax1.plot(loss_array)
-        ax1.plot([len(loss_array)-1, len(loss_array)-1], [r.get('loss_min'), loss_array[-1]], marker='o')
-        ax1.grid()
-        ax1.set_title('BCE with logits loss')
-        avg_pred_loss.append(sum(pred_loss[-wdw:])/wdw)
-        ax2.imshow(np.vstack([outputs, line, labels, line, pred]), cmap=cmap)
-        ax3.plot(pred_loss)
-        ax3.plot(avg_pred_loss)
-        ax3.set_title('Accuracy over batch')
-        ax3.grid()
[…]

### 2020-11-10 test.py

@@ -1,399 +1,0 @@
-import torch
-import numpy as np
-import librosa
-import librosa.display
-import os
-import random
-import soundfile as sf
-
-from model import OrchMatchNet
-from OrchDataset import RawDatabase
-from train import getPosNMax
-from parameters import GLOBAL_PARAMS
-
-""" 
-To run this file:
-- You need a folder of targets to be orchestrated, and set 'target_path' to point to this
-- You need to create a folder to store the orchestrated solutions, and set 'solutions_path'
-to be the path to this folder
-- You need a trained model saved as a .pth file, and set 'state_path' to point to this
-
-- Set 'tinysol_path' to point to your TinySOL database
-- Set model_type, instr_filter, and n
-
-"""
-
-
-# path to TinySOL data
-tinysol_path = './TinySOL'
-
-# path to target samples
-target_path = './targets'
-
-# number of samples to be used in solution
-n = 10
-
-# instruments to be used (all instruments will be used)
-instr_filter = ['Hn', 'Ob', 'Vn', 'Va',
-                'Vc', 'Fl', 'Tbn', 'Bn', 'TpC', 'ClBb']
-
-# if true, doesn't orchestrate targets but instead creates combinations of TinySOL samples to input to the model
-sanity_check = False
-
-
-def test(model, state_path, data, targets):
-    device = torch.device('cpu')
-    state = torch.load(state_path, map_location=device)
-    model.load_state_dict(state['state_dict'])
-    model.eval()
-
-    outputs = model(data)
-    outputs = outputs.detach().cpu().clone().numpy()
-
-    if sanity_check:
-        accuracy = evaluate(outputs, labels)
-        print(accuracy)
-
-    # text file to write solutions to
-    f = open(solutions_path + '/orchestration_results.txt', 'w+')
-
-    for i in range(len(outputs)):
-        output = outputs[i]
-        target = targets[i]
-        # get indices of top n probabilities
-        indices = getPosNMax(output, n)
-        # turn indices into [instr, pitch]
-        classes = get_classes(indices)
-        # get top n probabiltiies
-        probs = [output[i] for i in indices]
-        # turn probabilities into dynamic markings (pp, mf, ff)
-        dynamics = prob_to_dynamic(probs)
-        # combine into (instr, pitch, dynamic)
-        for j in range(len(classes)):
-            classes[j].append(dynamics[j])
-        
[…]

### 2020-11-10 train.py

@@ -1,431 +1,0 @@
-import torch
-import torch.nn as nn
-import torchvision.transforms as transforms
-import torch.nn.functional as F
-from torch.autograd import Variable
-from torch.utils import data
-import numpy as np
-import argparse
-import time
-import pickle
-import os
-from functools import reduce
-import librosa.display
-import matplotlib.pyplot as plt
-from matplotlib.animation import FuncAnimation
-
-from model import OrchMatchNet
-from parameters import GLOBAL_PARAMS, SimParams
-from OrchDataset import OrchDataSet,RawDatabase
-
-class Timer:
-    def __init__(self, size_buffer):
-        self.measuring = False
-        self.begin = []
-        self.end = []
-        self.size_buffer = size_buffer
-        
-    def start(self):
-        if not self.measuring:
-            if len(self.begin) == self.size_buffer:
-                self.begin.pop(0)
-            self.begin.append(time.time())
-            self.measuring = True
-            
-    def stop(self):
-        if self.measuring:
-            if len(self.end) == self.size_buffer:
-                self.end.pop(0)
-            self.end.append(time.time())
-            self.measuring = False
-    
-    def estimate(self, remaining):
-        if len(self.end) > 0 and not self.measuring:
-            diff = np.array(self.end)-np.array(self.begin)
-            diff = np.mean(diff)
-            h,r=divmod(diff*remaining,3600)
-            m,s=divmod(r,60)
-            print(diff)
-            print("Estimated remaining time : {}h{}m{}s".format(int(h), int(m), int(s)))
-            
-
-class DataSaver:
-    '''
-        Class used to write data for visualization (see showOutputs.py)
-    '''
-    def __init__(self, path_file):
-        self.path = path_file
-        self.count = 0
-        
-    def save(self, outs, labels, loss, loss_min, epoch, step):
-        r = {}
-        r['epoch'] = epoch
-        r['step'] = step
-        r['outputs'] = outs
-        r['labels'] = labels
-        r['loss'] = loss
-        r['loss_min'] = loss_min
-        np.save(self.path, r)
-        self.count += 1
-
-timer = Timer(1000)
-
-def main(resume_model, rdb = None):
-    #print("Starting parsing -----")
-    #arg = arg_parse()
-    #print("End parsing")
-
[…]

### 2020-11-19 constraint_filtering/BiLSTM-CRF-Gxzzz/OrchDataset.py
Properly loaded data, working on forward pass shapes

@@ -164,7 +164,7 @@ class OrchDataSet(data.Dataset):
             feature_type defines whether the features generated are mel spectrograms or mfcc.
             It must be either 'mel' or 'mfcc'.
         '''
-        
+
         # For data augmentation, a set of mel basis is generated, using slightly shifted
         # sample rates. This ensures frequency variation of the selected samples.
         self.mel_basis = librosa.filters.mel(GLOBAL_PARAMS.RATE, GLOBAL_PARAMS.N_FFT, n_mels=GLOBAL_PARAMS.N_MELS)
@@ -210,6 +210,7 @@ class OrchDataSet(data.Dataset):
         self.class_encoder = class_encoder
 
 
+
     
     def generate(self, N, nb_samples=None):
         '''
@@ -226,7 +227,7 @@ class OrchDataSet(data.Dataset):
         if not nb_samples is None:
             self.nb_samples = min(self.nb_samples,nb_samples)
         
-        self.list_instr = list(self.db.keys())
+        self.list_instr = GLOBAL_PARAMS.instr_filter
         self.list_sr = np.linspace((1-GLOBAL_PARAMS.coeff_freq_shift_data_augment)*GLOBAL_PARAMS.RATE, (1+GLOBAL_PARAMS.coeff_freq_shift_data_augment)*GLOBAL_PARAMS.RATE, 5)
         self.list_sr = [int(x) for x in self.list_sr]
         
@@ -362,9 +363,9 @@ class OrchDataSet(data.Dataset):
         if self.feature_type == 'mfcc':
             pow_db = librosa.power_to_db(mel_spec)
             mfcc = librosa.feature.mfcc(S=pow_db)
-            return torch.tensor(np.array([mfcc])), self.class_encoder(list_samp)
+            return torch.tensor(mfcc), self.class_encoder(list_samp)
         elif self.feature_type == 'mel':
-            return torch.tensor(np.array([mel_spec])), self.class_encoder(list_samp)
+            return torch.tensor(mel_spec), self.class_encoder(list_samp)
 
 
 class IndexIncrementor:

### 2020-11-19 constraint_filtering/BiLSTM-CRF-Gxzzz/bilstm_crf.py

@@ -6,12 +6,11 @@ import utils
 
 
 class BiLSTMCRF(nn.Module):
-    def __init__(self, num_classes, dropout_rate=0.5, embed_size=256, hidden_size=256):
+    def __init__(self, num_classes, feature_size, dropout_rate=0.5, hidden_size=256):
         """ Initialize the model
         Args:
-            sent_vocab (Vocab): vocabulary of words
             num_classes:
-            embed_size (int): embedding size (# of features per word)
+            feature_size (int): embedding size, # of features
             hidden_size (int): hidden state size
 
             len: how many discrete inputs come from the data point (number of words)
@@ -21,12 +20,12 @@ class BiLSTMCRF(nn.Module):
         """
         super(BiLSTMCRF, self).__init__()
         self.dropout_rate = dropout_rate
-        self.embed_size = embed_size
+        self.feature_size = feature_size
         self.hidden_size = hidden_size  # "the number of features in the hidden state" of the LSTM
         self.num_classes = num_classes  # labels
         self.dropout = nn.Dropout(dropout_rate)
 
-        self.LSTM = nn.LSTM(input_size=embed_size, hidden_size=hidden_size, bidirectional=True)
+        self.LSTM = nn.LSTM(input_size=feature_size, hidden_size=hidden_size, bidirectional=True)
 
         # a linear layer from hidden size * 2 to # of possible labels? or # of samples?
         # I think len(self.tag_vocab) = K, meaning K is the number of classes
@@ -36,7 +35,7 @@ class BiLSTMCRF(nn.Module):
         # CRF transition matrix, shape: (K, K)
         self.transition = nn.Parameter(torch.randn(self.num_classes, self.num_classes))
 
-    def forward(self, input_batch, labels, sen_lengths):
+    def forward(self, input_batch, labels):
         """
         1. takes in sentences, featurizes them (embeds them in feature space)
         2. calls encode on them, which puts them through the BiLSTM
@@ -54,13 +53,15 @@ class BiLSTMCRF(nn.Module):
         # mask = (input_batch != self.sent_vocab[self.sent_vocab.PAD]).to(self.device)  # shape: (b, len)
         # for now, let's have mask be an entire array of True
         # later we can try removing mask entirely, since I don't think it is necessary
+
+        # input shape (b, s, N_MELS, t)
         mask = torch.full(input_batch.shape, True, dtype=bool)
[…]

### 2020-11-19 constraint_filtering/BiLSTM-CRF-Gxzzz/run.py

@@ -1,7 +1,7 @@
 """
 Usage:
-    run.py train TRAIN SENT_VOCAB TAG_VOCAB [options]
-    run.py test TEST RESULT SENT_VOCAB TAG_VOCAB MODEL [options]
+    run.py train [options]
+    run.py test [options]
 
 Options:
     --dropout-rate=<float>              dropout rate [default: 0.5]
@@ -61,13 +61,15 @@ def train(args):
     num_test_samples = GLOBAL_PARAMS.nb_samples - num_training_samples
     test_data.generate(GLOBAL_PARAMS.N, num_test_samples)
 
-    train_data = torch.utils.data.DataLoader(dataset=train_data,
+    train_dataset = torch.utils.data.DataLoader(dataset=train_data,
                                              batch_size=GLOBAL_PARAMS.batch_size,
                                              shuffle=True)
 
-    test_data = torch.utils.data.DataLoader(dataset=test_data,
+    test_dataset = torch.utils.data.DataLoader(dataset=test_data,
                                             batch_size=GLOBAL_PARAMS.batch_size,
                                             shuffle=False)
+
+    feature_shape = train_data[0][0].shape
 
     max_epoch = int(args['--max-epoch'])
     log_every = int(args['--log-every'])
@@ -78,7 +80,7 @@ def train(args):
     device = torch.device('cuda' if args['--cuda'] else 'cpu')
     patience, decay_num = 0, 0
 
-    model = bilstm_crf.BiLSTMCRF(train_data.num_classes, float(args['--dropout-rate']), int(args['--embed-size']),
+    model = bilstm_crf.BiLSTMCRF(train_data.num_classes, feature_shape[1], float(args['--dropout-rate']),
                                  int(args['--hidden-size'])).to(device)
     for name, param in model.named_parameters():
         if 'weight' in name:
@@ -94,15 +96,17 @@ def train(args):
 
     print('start training...')
     for epoch in range(max_epoch):
-        for sentences, tags in utils.batch_iter(train_data, batch_size=int(args['--batch-size'])):
+        for i, (data, labels) in enumerate(train_dataset):
             train_iter += 1
-            current_batch_size = len(sentences)
-            sentences, sent_lengths = utils.pad(sentences, sent_vocab[sent_vocab.PAD], device)
-            tags, _ = utils.pad(tags, tag_vocab[tag_vocab.PAD], device)
+            current_batch_size = data.shape[0]
[…]

### 2020-11-20 source_separation/MMDenseLSTM.py
Added MMDenseLSTM code

@@ -1,0 +1,486 @@
+import torch
+# import encoding ## pip install torch-encoding . For synchnonized Batch norm in pytorch 1.0.0
+import torch.nn as nn
+import numpy as np
+from torch.nn import functional as F
+from torch import Tensor
+
+import time
+
+'''
+    MMDenseLSTM is comprised of several instances of MDenseLSTM, one for each frequency band + one for the whole frequency range.
+    
+    The flow of data is as follows:
+        - First convolution to generate [first_channel] channels
+        - Succession of [scale+1] dense blocksused to downsample the feature matrices. Each dense block has [l] layers.
+            * Layer number [i] takes as input [k0+i*k] channels and outputs [k] channels.
+            * Layer [i] has a first Norm-ReLU-1x1Conv layer to reduce input features to [bn_size*k] channels
+            * After bottleneck, another layer of Norm-ReLU-3x3Conv is applied to obtain [k] channels.
+        - 
+'''
+global DLC
+global DBC
+global LLC
+global MBC
+DLC = 0
+DBC = 0
+LLC = 0
+MBC = 0
+
+CUMULATIVE_DENSE_BLOCK_OUTPUT = False
+VERBOSE = False
+
+class MMDenseLSTMParams:
+    def __init__(self, init_nb_channels, band_splits, band_growth_rates, scales, input_shape, final_conv_channels, final_dense_block_layers, final_dense_block_growth_rate, drop_rate = 0.1, bn_size=4):
+        self.N = len(band_splits) + 2
+        self.drop_rate = drop_rate
+        self.init_nb_channels = init_nb_channels
+        self.bn_size = bn_size
+        self.final_conv_channels = final_conv_channels
+        self.final_dense_block_layers = final_dense_block_layers
+        self.final_dense_block_growth_rate = final_dense_block_growth_rate
+        if len(band_growth_rates) != self.N:
+            raise Exception("band_growth_rates must contain {} elements : 1 for each band and 1 for the full range".format(self.N))
+        if len(scales) != self.N:
+            raise Exception("scales must contain {} elements : 1 for each band and 1 for the full range".format(self.N))
+        self.layer_shapes = [[input_shape]]
+        ext_band_splits = [0] + band_splits + [input_shape[1]]
+        for i in range(len(ext_band_splits)-1):
+            self.layer_shapes.append([(input_shape[0], ext_band_splits[i+1]-ext_band_splits[i])])
[…]

### 2020-11-20 source_separation/MMDenseLSTM.py
Added MMDenseLSTM code

@@ -27,8 +27,8 @@ DBC = 0
 LLC = 0
 MBC = 0
 
-CUMULATIVE_DENSE_BLOCK_OUTPUT = False
-VERBOSE = False
+CUMULATIVE_DENSE_BLOCK_OUTPUT = True
+VERBOSE = True
 
 class MMDenseLSTMParams:
     def __init__(self, init_nb_channels, band_splits, band_growth_rates, scales, input_shape, final_conv_channels, final_dense_block_layers, final_dense_block_growth_rate, drop_rate = 0.1, bn_size=4):
@@ -447,7 +447,7 @@ class MMDenseLSTM(nn.Module):
         return output
 
 # Split at 4100Hz and 11000Hz
-B,C,T,Fr = 15,1,75,128
+B,C,T,Fr = 16,1,50,100
 params = MMDenseLSTMParams(1, [43,86],[7,14,4,2],[4,3,3,2], (T,Fr), 32, 3, 12)
 
 params.set_downsample_dense_layers(0,[3,3,4,5,5])
@@ -483,4 +483,209 @@ for i in range(200):
     t = time.time()
     r=m(l)
     dt.append(time.time()-t)
+    l = None
+    r = None
+    torch.cuda.empty_cache()
 print("Computation time for 1 batch : {}".format(sum(x for x in dt)/len(dt)))
+
+'''
+dt = [1.2316265106201172,
+ 0.9567806720733643,
+ 1.0095221996307373,
+ 1.069530725479126,
+ 0.9404206275939941,
+ 0.9825644493103027,
+ 0.943284273147583,
+ 0.9432389736175537,
+ 1.3145031929016113,
+ 0.9436233043670654,
+ 0.9450030326843262,
+ 1.050917387008667,
+ 1.0431499481201172,
+ 0.9439489841461182,
+ 0.9484851360321045,
+ 0.9486229419708252,
+ 1.2562696933746338,
+ 0.9581849575042725,
+ 0.9514966011047363,
+ 0.9680142402648926,
+ 1.2484300136566162,
+ 0.9650261402130127,
+ 1.4151840209960938,
+ 1.220715045928955,
+ 0.9658191204071045,
+ 1.3943183422088623,
+ 1.1896495819091797,
+ 0.9695940017700195,
+ 1.3323705196380615,
+ 1.1225497722625732,
+ 1.0147838592529297,
+ 1.1678142547607422,
+ 0.9433512687683105,
[…]

### 2021-02-27 source_separation/pipeline.py
Add new distance metric

@@ -26,7 +26,7 @@ TDCNNpp_model_path = "./trained_TDCNNpp"  # path to pretrained TDCNNpp model
 
 # If you want to save orchestrations, set the following variables
 SAVED_ORCHESTRATIONS_PATH = "./saved_orchestrations"
-num_orchestrations_to_save = 5
+num_orchestrations_to_save = 0
 
 ########################################################
 # You shouldn't need to change the following variables #
@@ -280,34 +280,39 @@ def spectral_distance(target, solution):
 
 def frame_spectral_distance(target, solution):
     """
-
-    :param target:
-    :param solution:
-    :return:
+    Calculates a weighted sum of the spectral difference between frames of target and solution
+    :param target: audio as numpy array
+    :param solution: audio as numpy array
+    :return: distance as float
     """
     # target and solution should be the same length
     length = max(target.size, solution.size)
     target = librosa.util.fix_length(target, length)
     solution = librosa.util.fix_length(solution, length)
 
-    num_frames = 100
-    frame_length = length // num_frames
+    frame_length = 4000  # 4000 samples ~ 10 ms
     target_frames = librosa.util.frame(target, frame_length, frame_length, axis=0)
-    last_frame = target[num_frames * frame_length:]
-    last_frame = librosa.util.fix_length(last_frame, frame_length)
-    target_frames = np.append(target_frames, [last_frame], axis=0)
+    # last_frame = target[num_frames * frame_length:]
+    # last_frame = librosa.util.fix_length(last_frame, frame_length)
+    # target_frames = np.append(target_frames, [last_frame], axis=0)
 
     solution_frames = librosa.util.frame(solution, frame_length, frame_length, axis=0)
-    last_frame = solution[num_frames * frame_length:]
-    last_frame = librosa.util.fix_length(last_frame, frame_length)
-    solution_frames = np.append(solution_frames, [last_frame], axis=0)
+    # last_frame = solution[num_frames * frame_length:]
+    # last_frame = librosa.util.fix_length(last_frame, frame_length)
+    # solution_frames = np.append(solution_frames, [last_frame], axis=0)
 
     distances = []
+    target_energy = []
     assert target_frames.size == solution_frames.size
     for i in range(target_frames.shape[0]):
         distance = spectral_distance(target_frames[i], solution_frames[i])
         distances.append(distance)
-    return mean(distances)
+        energy = np.sum(np.sqrt(np.square(target_frames[i])))
[…]

### 2021-02-27 source_separation/create_targets.py
Allow different targets to use same samples

@@ -5,9 +5,11 @@ import json
 import os
 import librosa
 import soundfile as sf
+from itertools import product
 
 from pipeline import NUM_SUBTARGETS, SAMPLE_DATABASE_PATH, TARGETS_PATH, TARGET_METADATA_PATH, SAMPLING_RATE, clear_directory, combine_with_offset
 
+NUM_TARGETS = 100  # number of targets to create
 
 def create_targets(paths):
     """
@@ -67,15 +69,12 @@ def create_targets(paths):
 if __name__ == "__main__":
     clear_directory(TARGETS_PATH)
     samples = librosa.util.find_files(SAMPLE_DATABASE_PATH)
-    random.shuffle(samples)
-    sample_paths = []
-    i = 0
-    while i < len(samples):
-        if i % NUM_SUBTARGETS == 0:
-            sample_paths.append([])
-        sample_paths[-1].append(samples[i])
-        i += 1
-    sample_paths = [l for l in sample_paths if len(l) == NUM_SUBTARGETS]
+    sample_paths = set()
+    while len(sample_paths) < NUM_TARGETS:
+        paths = random.sample(samples, NUM_SUBTARGETS)
+        paths = sorted(paths)
+        paths = tuple(paths)  # you can't add lists to a set
+        sample_paths.add(paths)
     print("Creating {} targets".format(len(sample_paths)))
     create_targets(sample_paths)
     print("Done.")

### 2021-02-27 source_separation/paper/paper.tex
Add outline of paper

% -----------------------------------------------
% Template for SMC 2021
% Adapted from previous SMC paper templates
% -----------------------------------------------
\documentclass{article}
\usepackage{smc2021}
%%%%%%%%%%%%%%%%%%%%%%%% Some useful packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%% See related documentation %%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[caption=false, font=footnotesize]{subfig}% Modern replacement for subfigure package
\usepackage{paralist}% extended list environments
\usepackage[figure,table]{hypcap}% hyperref companion
% Enable for Review only, remove for Camera Ready version
\pagewiselinenumbers

% Use this if english is the only language/alphabet used in the document
\usepackage[english]{babel}

% Title.
% ------
\def\papertitle{Paper Title}

% Authors
% Please note that submissions are NOT anonymous, therefore 
% authors' names have to be VISIBLE in your manuscript. 
% Authors are entered as an ordered list, each one can be linked to multiple affiliations using the correct index.
% Available tags for authors are: \firstname \middlename \lastname \generation \originalname \email \orcid
% Available tags for affiliations are: \unit \department \institution \streetaddress \city \state \postcode \country \type
% type can take as value: University, Company, Music, Independent, Other
%
% \author[]{\mbox{\firstname{}\middlename{}\lastname{}\originalname{}\generation{}\email{}\orcid{}}}
% mbox force an author not to be split over multiple lines
\author[1]{\mbox{\firstname{Carmine}\lastname{Cella}}}

%%Affiliations
\affil[1]{\department{Center for New Music and Audio Technologies}\institution{University of California, Berkeley}\city{Berkeley}\state{California}\country{USA}\affiliationtype{University}}

% Complete setup stage
\completesetup

% Title.
% ------
\title{\papertitle}
% ***************************************** the document starts here ***************
\begin{document}
	%
	\capstartfalse
	\maketitle
	\capstarttrue
	%

	\begin{abstract}
		The abstract should be placed at the top left column and should contain about 150-200 words.
	\end{abstract}
	%

	\section{Introduction}\label{sec:introduction}
	\begin{itemize}
		\item what is CAO 
		\item why we need source separation 
		\item which type of sounds are useful for us (targets)
	\end{itemize}

	\section{Methodology}\label{sec:methodology}
	We compare a number of source separation methods and our metrics depend on the orchestration
	Distance metric and formula here

	\section{Experiments}\label{sec:experiments}
	\begin{itemize}
		\item Full orchestration
		\item Ground truth
		\item NMF
		\item 4 neural models: TDCNN++, TDCNN, Demucs, OpenUnmix
		\item table with results
	\end{itemize}

	\section{Conclusions}\label{sec:conclusions}
	We think that adding source separation improves orchestration. Unsupervised methods work better because of the data that supervised methods are trained on \cite{Someone:00}

	\section{Future Work}\label{sec:futurework}
	Implementing these methods in Orchidea. We could improve the supervised methods by training them ourselves with data that fits our problem better.

	\begin{acknowledgments}
		At the end of the Conclusions, acknowledgements to people, projects, funding agencies, etc. can be included after the second-level heading  ``Acknowledgments'' (with no numbering).
	\end{acknowledgments}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%bibliography here
	\bibliography{references}

\end{document}

### 2021-02-28 source_separation/paper/paper.tex
Add distances table
Add distance metric

@@
 		\item which type of sounds are useful for us (targets)
+		\item What is Orchidea?
 	\end{itemize}
 
+	\textit{The following is directly copy/pasted from our last paper, it will need to be changed in order to avoid self-plagiarism}
+
+	The development of computational tools to assist and inspire the musical composition process constitutes an important research area known as \emph{Computer-Assisted Composition (CAC)} \cite{FerVic2013, Ari2005}.
+Within CAC, target-based computer-assisted orchestration is a compelling case of how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.
+
+	Target-based computer-assisted orchestration takes a target sound as an input and attempts to find instrumental samples that best match the target given a specific similarity metric and a set of constraints.
+A solution to this problem is a set of orchestral scores that represent the mixtures of audio samples in the database, ranked by similarity with the target sound.
+
+	The approach studied in \cite{Carpentier2010} consists in finding a good orchestration for any given sound by searching combinations of sounds from a database with a multi-objective optimization heuristics and a constraint solver that are jointly optimized.
+Both the target sound and the sounds in the database are embedded in a feature space defined by a fixed feature function and each generated combination of sounds is evaluated by using a specific metric.
+This method has been substantially improved in \cite{Cella18, Cella2020} and is implemented in the \emph{Orchidea} toolbox for assisted orchestration (\url{www.orch-idea.org}), currently considered the state-of-the-art system for assisted orchestration.
+
+	In this paper, we try a different approach to this problem by experimenting with deep neural architectures.
+The main idea is to train a model to classify combinations of real instruments and then use it for orchestration.
+A typical solution for assisted orchestration is a set of triples \emph{instrument-pitch-dynamics} such as \{\texttt{Flute C6 pp, Bassoon C4 mf, Bassoon G4 ff}\}.
+By training a neural network with real combinations of instrumental notes, it will acquire the ability to identify the presence of each instrument and its associated pitch by building the appropriate latent representation.
+Thus, when an unknown target sound is given as input, the network will identify which are the best instruments to match the target sound, and it will be able to  deconstruct a complex mixture of timbres into individual instrument notes.
+This method is motivated by the good results obtained in previous research on musical instruments identification \cite{Benetos07, Kitahara05} and the more recent use of deep neural networks for musical classification \cite{lostanlen16, Bian19}.
+
+	In this paper we perform preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}.
+We chose to use a CNN because of its success in audio classification \cite{Hershey17} and we decided to include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.
+The codebase for this paper can be found at: \url{}.
+
 	\section{Methodology}\label{sec:methodology}
-	We compare a number of source separation methods and our metrics depend on the orchestration
-	Distance metric and formula here
+	\begin{itemize}
+		\item compare a number of source separation methods
+		\item we use the final orchestration to compare the separation methods
+		\item include distance metric and formula here
+	\end{itemize}
+
+	Targets are separated into four sub-targets.
+Each sub-target is independently orchestrated with a randomly assigned "sub-orchestra." These orchestrations are then combined to play simultaneously, creating a final orchestrated solution.
+Then the distance between the target and solution is calculated, giving us a metric to compare the various separation methods.
+
+		\subsection{Separation Methods}
+
+			\subsubsection{Non-negative Matrix Factorization}
+			@Max
+
+			\subsubsection{Demucs}
+			@Leo/Alejandro
+
+			\subsubsection{OpenUnmix}
+			@Leo/Alejandro
+
+			\subsubsection{TDCNN (use full name instead of abbreviation)}
+			@Leo/Alejandro
+
+			\subsubsection{TDCNN++}
+			@Leo/Alejandro
+
+		\subsection{Evaluation}
+		We compare the effectiveness of different separation methods by comparing how well they work for orchestration.
+The output of a method is orchestrated, and these orchestrations are compared.
+A quantitative evaluation is performed through the use of a distance metric that measures the spectral distance between target and solution.
+
+		The distance metric cuts the target and solution into successive frames that are 4,000 samples in length, then calculates the spectral distance as defined in Eqn.
+\ref{eqn:distance} between corresponding frames.
+This metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization.
+The equation takes in the full FFT of the target $x$ and full FFT of the solution $\tilde{x}$.
+Then for each bin $k$ of the FFT, it calculates the absolute difference between the values.
+The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.
+
+		\begin{equation}\label{eqn:distance}
+d(x, \tilde{x}) =\lambda_1 \sum_k \delta_{k1}(x_k - \tilde{x}_k) + \lambda_2 \sum_k \delta_{k2}|x_k - \tilde{x	}_k| \\
+\end{equation}
+where $\delta_{k1} = 1 \text{  if  } x_k \ge \tilde{x}_k, 0 \text{  otherwise}$; and $\delta_{k2} = 1 \text{  if  } x_k < \tilde{x}_k, 0 \text{  otherwise}$.
 
@@
 	\end{itemize}
+
+		\subsection{Data}
+		We created our own targets as combinations of four sub-targets.
+The sub-targets come from the NIGENS and BBC databases \cite{} and freesound \cite{}.
+We selected ?
+sub-targets from these databases, choosing sub-targets that fit the following criteria:
+
+		\begin{enumerate}
+			\item Static sounds in which the harmonic average across time is a fitting representation of the sound
+			\item Sounds in which there is at least some pitched content and not only noise
+		\end{enumerate}			
+		During the creation of the targets, the sub-targets were randomly offset in time from the beginning of the target, so that the various sub-targets begin playing at different times.
+
+	\begin{table}[t]
+		\begin{center}
+			\begin{tabular}{|c|c|}
+				\hline
+				& Average distance \\
+				\hline
+				Full target & 25.73 \\
+				\hline
+				TDCNN++ & 25.44 \\
+				\hline
+				Demucs & 27.62 \\
+				\hline
+				NMF & 22.32 \\
+				\hline
+				Ground truth & 24.27 \\
+				\hline
+			\end{tabular}
+		\end{center}
+		\caption{Average distance between target and orchestration for various methods. "Full target" means no separation.}
+		\label{tab:distances}
+	\end{table}
 

### 2021-03-01 source_separation/paper/paper.tex
Move some things from methodology to experiments

@@
 		\item we use the final orchestration to compare the separation methods
-		\item include distance metric and formula here
+	\end{itemize}
+
+	We compare the effectiveness of different source separation methods for the task of orchestration by applying the separation methods to various targets, orchestrating the separated output of the method, and finally comparing this orchestration to the orchestration of the target if no separation was performed \textbf{Diagram needed}.
+For example, consider a target sound that is a combination of a low droning sound and high-pitched whistle.
+With separation, these two sounds would be disentangled into two separate sub-targets and each would be individually orchestrated.
+The separated orchestrations would then be combined to create the solution.
+This would then be compared to the orchestration if no separation had been performed.
+
+	\section{Experiments}\label{sec:experiments}
+	\begin{itemize}
+		\item Full orchestration
+		\item Ground truth
+		\item NMF
+		\item 4 neural models: TDCNN++, TDCNN, Demucs, OpenUnmix
+		\item table with results
 	\end{itemize}
@@
 Then the distance between the target and solution is calculated, giving us a metric to compare the various separation methods.
+
+		\subsection{Data}
+		We created our own targets as combinations of four sub-targets.
+The sub-targets come from the NIGENS and BBC databases \cite{} and freesound \cite{}.
+We selected ?
+sub-targets from these databases, choosing sub-targets that fit the following criteria:
+
+		\begin{enumerate}
+			\item Static sounds in which the harmonic average across time is a fitting representation of the sound
+			\item Sounds in which there is at least some pitched content and not only noise
+		\end{enumerate}			
+		During the creation of the targets, the sub-targets were randomly offset in time from the beginning of the target, so that the various sub-targets begin playing at different times.
 
@@
 where $\delta_{k1} = 1 \text{  if  } x_k \ge \tilde{x}_k, 0 \text{  otherwise}$; and $\delta_{k2} = 1 \text{  if  } x_k < \tilde{x}_k, 0 \text{  otherwise}$.
-
-	\section{Experiments}\label{sec:experiments}
-	\begin{itemize}
-		\item Full orchestration
-		\item Ground truth
-		\item NMF
-		\item 4 neural models: TDCNN++, TDCNN, Demucs, OpenUnmix
-		\item table with results
-	\end{itemize}
-
-		\subsection{Data}
-		We created our own targets as combinations of four sub-targets.
-The sub-targets come from the NIGENS and BBC databases \cite{} and freesound \cite{}.
-We selected ?
-sub-targets from these databases, choosing sub-targets that fit the following criteria:
-
-		\begin{enumerate}
-			\item Static sounds in which the harmonic average across time is a fitting representation of the sound
-			\item Sounds in which there is at least some pitched content and not only noise
-		\end{enumerate}			
-		During the creation of the targets, the sub-targets were randomly offset in time from the beginning of the target, so that the various sub-targets begin playing at different times.
 

### 2021-03-01 source_separation/paper/paper.tex
Add introduction references

@@
 
-	In this paper, we try a different approach to this problem by experimenting with deep neural architectures.
-The main idea is to train a model to classify combinations of real instruments and then use it for orchestration.
-A typical solution for assisted orchestration is a set of triples \emph{instrument-pitch-dynamics} such as \{\texttt{Flute C6 pp, Bassoon C4 mf, Bassoon G4 ff}\}.
-By training a neural network with real combinations of instrumental notes, it will acquire the ability to identify the presence of each instrument and its associated pitch by building the appropriate latent representation.
-Thus, when an unknown target sound is given as input, the network will identify which are the best instruments to match the target sound, and it will be able to  deconstruct a complex mixture of timbres into individual instrument notes.
-This method is motivated by the good results obtained in previous research on musical instruments identification \cite{Benetos07, Kitahara05} and the more recent use of deep neural networks for musical classification \cite{lostanlen16, Bian19}.
-
-	In this paper we perform preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}.
-We chose to use a CNN because of its success in audio classification \cite{Hershey17} and we decided to include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.
 	 The codebase for this paper can be found at: \url{}.
@@
 	\section{Conclusions}\label{sec:conclusions}
-	We think that adding source separation improves orchestration. Unsupervised methods work better because of the data that supervised methods are trained on \cite{Someone:00}
+	We think that adding source separation improves orchestration. Unsupervised methods work better because of the data that supervised methods are trained on
 

### 2021-03-01 source_separation/paper/paper.tex
Add testing section

@@
 The sub-targets come from the NIGENS and BBC databases \cite{} and freesound \cite{}.
-We selected ?
-sub-targets from these databases, choosing sub-targets that fit the following criteria:
+We selected 90 samples from these databases, choosing sounds that fit the following criteria:
 
@@
 		\end{enumerate}			
-		During the creation of the targets, the sub-targets were randomly offset in time from the beginning of the target, so that the various sub-targets begin playing at different times.
+		Each target that we used for testing was a combination of 4 randomly chosen source sounds from the group of 90 sounds.
+During the creation of the targets, the sources were randomly offset in time from the beginning of the target, so that they began playing at different times.
 
@@
 			@Leo/Alejandro
+
+		\subsection{Testing}
+		In order to compare the effectiveness of the separation methods, a full target orchestration and a ground truth orchestration are done for each target.
+The full target orchestration is the orchestration of the target without any separation performed.
+The ground truth orchestration takes the four sources that make up the target and orchestrates each one separately, then combines them.
+This ground truth represents the orchestration that could take place if the separation method was perfect and separated a target exactly into its constituent sources.
+
+		The procedure we used for testing is as follows:
+		\begin{enumerate}
+			\item The target is created as a combination of four randomly chosen sources, which are offset to begin playing at different times
+			\item The full target, without any separation performed, is orchestrated using the entire orchestra
+			\item A given separation method is performed on the target, splitting the target into 4 sub-targets
+			\item Each sub-target is separately orchestrated using a randomly chosen sub-orchestra that is one quarter the size of the full orchestra, and then each orchestration is combined to play simultaneously
+			\item The ground truth orchestration is created by orchestrating the four sources, each with a different sub-orchestra, and then combined to play simultaneously
+			\item The distances between target and orchestration are calculated for the full target orchestration, the separated orchestration, and the ground truth orchestration 
+		\end{enumerate}
+
+		The orchestrations performed were static orchestrations, meaning a single onset of notes is created for each target, no matter if the target itself has multiple onsets.
 

### 2021-03-02 source_separation/paper/paper.tex
Add my name
Add some references

@@
 \author[1]{\mbox{\firstname{Carmine}\lastname{Cella}}}
+\author[1]{\mbox{\firstname{Luke}\lastname{Dzwonczyk}}}
 
@@
 
-	Target-based computer-assisted orchestration takes a target sound as an input and attempts to find instrumental samples that best match the target given a specific similarity metric and a set of constraints.
+	takes a target sound as an input and attempts to find instrumental samples that best match the target given a specific similarity metric and a set of constraints.
 A solution to this problem is a set of orchestral scores that represent the mixtures of audio samples in the database, ranked by similarity with the target sound.
@@
 	 The codebase for this paper can be found at: \url{}.
+
+	 ----------------------------------------------------------
+
+	 \textit{Everything below this line in the introduction is newly written for this paper}
+
+	 Target-based computer-assisted orchestration is the process of creating an orchestral score that best matches an arbitrary target sound given a similarity metric and constraints.
+It recreates the timbre and harmonics of the target sound using orchestral instruments.
 
@@
 		We created our own targets as combinations of four sub-targets.
-The sub-targets come from the NIGENS and BBC databases \cite{} and freesound \cite{}.
-We selected 90 samples from these databases, choosing sounds that fit the following criteria:
+The sub-targets come from the NIGENS \cite{NIGENS} and BBC \cite{} databases, and freesound.org \cite{freesound}.
+We selected a total of 90 samples from these databases, choosing sounds that fit the following criteria:
 
@@
 \ref{eqn:distance} between corresponding frames.
-This metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization.
+The spectral distance metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization, and used in \cite{Cella2020b} \textbf{(is it weird/bad to cite our last paper?)}to compare accuracies of orchestrations.
 The equation takes in the full FFT of the target $x$ and full FFT of the solution $\tilde{x}$.
@@
 		\end{center}
-		\caption{Average distance between target and orchestration for various methods. "Full target" means no separation.}
+		\caption{Average across \textbf{X} targets showing the distance between target and orchestration for various methods. "Full target" means no separation was performed. }
 		\label{tab:distances}

### 2021-03-03 source_separation/paper/paper.tex
Add to introduction

@@
 
-	\textit{The following is directly copy/pasted from our last paper, it will need to be changed in order to avoid self-plagiarism}
+	\textit{The following (from here until the line) is directly copy/pasted from our last paper, it will need to be changed in order to avoid self-plagiarism}
 
@@
 Within CAC, target-based computer-assisted orchestration is a compelling case of how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.
-
-	takes a target sound as an input and attempts to find instrumental samples that best match the target given a specific similarity metric and a set of constraints.
-A solution to this problem is a set of orchestral scores that represent the mixtures of audio samples in the database, ranked by similarity with the target sound.
 
@@
 It recreates the timbre and harmonics of the target sound using orchestral instruments.
+
+		Source separation is an important addition to improve orchestration as it allows for the orchestration of more complex sounds in a way that makes sense in an orchestral context.
+Orchestral music often uses multiple layers, so by separating the layers of a target and individually orchestrating them, the resulting solution is improved.
+\textbf{(@Carmine, can you make sure this paragraph makes sense/is correct?)}
+
+		For example, consider an example target that has a continuous drone sound throughout and a bird song playing above this drone.
+Without source separation, Orchidea would detect each note of the bird song as a new onset and cut the target in time at each new note.
+However, this segmentation in time would also effect the drone, forcing it to have a new onset each time the bird song changes notes.
+The resulting orchestration would having multiple onsets for the drone, when really only one onset is needed at the beginning of the drone.
+When source separation is applied to this target, the drone and bird song would be orchestrated separately.
+Therefore, the drone could be orchestrated by a single onset of note(s) that lasts for the duration of the drone.
+At the same time, the bird song could have as many new onsets as needed without affecting the drone.
+The orchestration would be greatly improved by removing the unnecessary onsets in the orchestration of the drone.
+This is the effect we hope to achieve through implementing source separation in computer-assisted orchestration.
 
@@
 \ref{eqn:distance} between corresponding frames.
-The spectral distance metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization, and used in \cite{Cella2020b} \textbf{(is it weird/bad to cite our last paper?)}to compare accuracies of orchestrations.
+The spectral distance metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization, and used in \cite{Cella2020b} \textbf{(is it weird/bad to cite our last paper?)} to compare accuracies of orchestrations.
 The equation takes in the full FFT of the target $x$ and full FFT of the solution $\tilde{x}$.
@@
 	\section{Future Work}\label{sec:futurework}
-	Implementing these methods in Orchidea. We could improve the supervised methods by training them ourselves with data that fits our problem better.
+	Implementing these methods in Orchidea.
+We could improve the supervised methods by training them ourselves with data that fits our problem better.
+Adding a joint optimization to determine which instruments should be used in which sources.
 

### 2021-03-04 source_separation/paper/paper.tex
Add to abstract

@@
 		The abstract should be placed at the top left column and should contain about 150-200 words.
+
+		We identify that sound source separation is a pre-processing method that can improve computer-assisted musical orchestration.
+In this paper, we compare various state-of-the-art models for source separation, including supervised and unsupervised methods, in their effectiveness at improving orchestration.
+By applying these methods to hand picked target sounds, orchestrating the separated outputs, and comparing them to the orchestration of the same target without separation, we can determine which methods improve orchestration the most.
+The target sounds we test with are not songs or musical phrases but a collection of atmospheric recordings, sound effects, bells, alarms, and more.
+These sounds more closely match common targets used in computer-assisted orchestration.
 	\end{abstract}

### 2021-03-04 source_separation/paper/paper.tex
Add to table
Add BBC reference

@@
 \author[1]{\mbox{\firstname{Carmine}\lastname{Cella}}}
-\author[1]{\mbox{\firstname{Luke}\lastname{Dzwonczyk}}}
+\author[1]{\mbox{\firstname{Luke}\lastname{Dzwonczyk}\email{dz.luke@berkeley.edu}}}
 
@@
 		We created our own targets as combinations of four sub-targets.
-The sub-targets come from the NIGENS \cite{NIGENS} and BBC \cite{} databases, and freesound.org \cite{freesound}.
+The sub-targets come from the NIGENS \cite{NIGENS} and BBC \cite{BBC} databases, and freesound.org \cite{freesound}.
 We selected a total of 90 samples from these databases, choosing sounds that fit the following criteria:
@@
 				\hline
+				TDCNN & \\
+				\hline
 				TDCNN++ & 25.44 \\
@@
 				Demucs & 27.62 \\
+				\hline
+				OpenUnmix & \\
 				\hline

### 2021-03-04 source_separation/paper/paper.tex
Wrote conclusion and added references to separation methods

--- 2021-03-03 source_separation/paper/paper.tex
@@
 
+			Demucs \cite{demucs}
+
 			\subsubsection{OpenUnmix}
@@
 	\section{Conclusions}\label{sec:conclusions}
-	We think that adding source separation improves orchestration. Unsupervised methods work better because of the data that supervised methods are trained on
+	In this paper, we addressed the problem of orchestrating dynamic targets using source separation as a preprocessing step.
+The results showed that performing source separation followed by a static orchestration of each individual sub targets improved the spectral distance between the original sample and its orchestrated version.\\
+
+	Furthermore, it seems that unsupervised methods such as NMF outperform all supervised methods.
+However, it is important to notice that the different neural models (Demucs, ConvTasNet, ConvTasNetUniversal and OpenUnmix) were used based on pre-trained versions.
+Those versions were trained on musical data, except for TDCNN++ which used selected generic sound samples, which opens a large path for improvement.
+More specifically, we may infer from our results that models trained not only on musical data can generate a better separation for orchestration, as TDCNN++ outperforms other supervised models.\\
+
+	Since unsupervised methods perform well, it is reasonable to think that supervised methods trained on a dataset well suited for source separation could yield better overall performances on dynamic orchestration.
 

### 2021-03-04 source_separation/paper/paper.tex
Added ConvTasNet description

--- 2021-03-04 source_separation/paper/paper.tex
@@
 
-			\subsubsection{TDCNN (use full name instead of abbreviation)}
-			@Leo/Alejandro
+			\subsubsection{ConvTasNet}
+			ConvTasNet \cite{tdcnn} (TDCNN) is a model based on a temporal convolutional encoder-decoder.
+It uses the audio waveform as input and performs a series of 1-D convolutions at different resolutions in order to generate a pool of masks that are then applied on the embedding of the initial waveform to output the separated targets.
+A final 1-D convolutional layer is used as a decoder to recover the waveforms of each sub target.
+This type of architecture replaces RNNs in modelling sequential data, as series of dilateds convolutions are a good way to capture temporal patterns.\\
+
+			The purpose of the model was initially to separate multiple speakers from a single-channel audio, and was thus trained on speech datasets.
+Because of its good performance at speech separation, a pre-trained version of the model based on the settings of the paper was added to the benchmark.
 

### 2021-03-04 source_separation/paper/paper.tex
Added TDCNN++ description

@@
 			ConvTasNet \cite{tdcnn} (TDCNN) is a model based on a temporal convolutional encoder-decoder.
-It uses the audio waveform as input and performs a series of 1-D convolutions at different resolutions in order to generate a pool of masks that are then applied on the embedding of the initial waveform to output the separated targets.
+It uses the audio waveform as input and performs a series of 1-D convolutions at different resolutions with residual connections in order to generate a pool of masks that are then applied on the embedding of the initial waveform to output the separated targets.
 A final 1-D convolutional layer is used as a decoder to recover the waveforms of each sub target.
@@
 
-			\subsubsection{TDCNN++}
-			@Leo/Alejandro
+			\subsubsection{Universal Sound Separation ConvTasNet}
+			This version of ConvTasNet, called TDCNN++, was improved by \cite{tdcnnpp} in order to be used for general sound separation, and not only for speech.
+The main architecture remains the same as in \cite{tdcnn}, and the changes only affected the masking network.
+More specifically, feature-wise normalization replaces global normalization, residual connections ranges are extended to improve the flow of information, and learnable scale parameters are introduced to weight the outputs of each residual layer.
+Those modifications help the model to better extract features without drastically increasing its size.\\
+
+			TDCNN++ presents the interesting property of having almost the same architecture as TDCNN but trained with generic sounds.
+Thus, it should naturally be better suited for our task, and the comparison between those two models in particular can give good insights about the impact of the data type used for training.
 

### 2021-03-04 source_separation/paper/paper.tex
-
Add back changes that were overwritten

@@
 			@Leo/Alejandro
+
+			OpenUnmix \cite{open-unmix}
 

### 2021-03-04 source_separation/paper/paper.tex
Add Leo's abstract

@@
 		The abstract should be placed at the top left column and should contain about 150-200 words.
+
+		In the field of computer-assisted composition, musical orchestration is probably the most empirical one and it has been traditionally taught in an intuitive and non formalized way.
+Computer-assisted orchestration seeks to help composers by accelerating certain creative processes.
+In particular, target-based computer-assisted orchestration helps composers to explore new timbral possibilities by associating instrumental samples, which, combined, will sound perceptually similar to a target sound.
+In this paper, we will study the possibility of adding a source separation step to the target-assisted orchestration process.
+We first discuss the interest of this addition to improve the quality of orchestrations of multi-layered sounds.
+Second, we select several state of the art models from music source separation (separation of instruments) and universal sound separation (separation of arbitrary sounds of different types), and we compare their efficiency for the task of orchestration.
+Our experiments show that the quality of orchestrations improved, both qualitatively and quantitatively, indicating that our approach is promising.
+Finally, we discuss the interest of unsupervised methods compared to supervised methods in the case of assisted orchestration, where data are relatively scarce.
 

### 2021-03-04 source_separation/paper/paper.tex
Edit abstract

--- 2021-03-04 source_separation/paper/paper.tex
@@
 	\begin{abstract}
-		The abstract should be placed at the top left column and should contain about 150-200 words.
+		In this paper, we will study the possibility of adding source separation as a pre-processing step to the orchestration process.
+We first discuss the motivation of this addition and its effect on the quality of orchestrations of multi-layered sounds.
+Second, we select several state-of-the-art models for both music source separation (separation of instruments) and universal sound separation (separation of arbitrary sounds of different types), and compare their effectiveness for the task of orchestration.
+By applying these methods to hand picked target sounds, then orchestrating the separated outputs, and finally comparing them to the orchestration of the same target without separation, we can determine which methods best suit the needs of orchestration.
+Our experiments show that the quality of orchestrations improved, both qualitatively and quantitatively, indicating that our approach is promising.
+Finally, we discuss the interest of unsupervised methods compared to supervised methods in the case of assisted orchestration, where data is relatively scarce.
 	\end{abstract}
@@
 	 \textit{Everything below this line in the introduction is newly written for this paper}
+
+	 In the field of computer-assisted composition, musical orchestration is probably the most empirical one and it has been traditionally taught in an intuitive and non formalized way.
+Computer-assisted orchestration seeks to help composers by accelerating certain creative processes.
+In particular, target-based computer-assisted orchestration helps composers to explore new timbral possibilities by associating instrumental samples, which, combined, will sound perceptually similar to a target sound.
+\textbf{(Note: This paragraph was moved to the intro from the abstract.
+We can move it back to the abstract if that would be better)}
 

### 2021-03-05 source_separation/paper/paper.tex
Add examples of source sounds

@@
 		\subsection{Data}
-		We created our own targets as combinations of four sub-targets.
-The sub-targets come from the NIGENS \cite{NIGENS} and BBC \cite{BBC} databases, and freesound.org \cite{freesound}.
+		We created our own targets as combinations of four source sounds.
+The sources come from the NIGENS \cite{NIGENS} and BBC \cite{BBC} databases, and freesound.org \cite{freesound}.
 We selected a total of 90 samples from these databases, choosing sounds that fit the following criteria:
@@
 		\end{enumerate}			
+		The sounds chosen include alarms, bells, engine noises, animals, soundscapes, synthesizer chords, and sound effects.
 Each target that we used for testing was a combination of 4 randomly chosen source sounds from the group of 90 sounds.

### 2021-03-05 source_separation/paper/paper.tex
add short description to demucs and open-unmix

@@
 			\subsubsection{Demucs}
-			@Leo/Alejandro
-
-			Demucs \cite{demucs}
+			The Demucs~\cite{demucs} architecture is a deep learning model for musical source separation.
+It takes as input the stereo mixture, and outputs 4 stereo waveforms corresponding to the stems (vocals, bass, drums and others).
+Demucs is based on a convolutional encoder, a bidirectional LSTM and a convolutional decoder.
+The encoder and the decoder are linked with skip U-Net connections.
+Both the encoder and the decoder are made of 6  blocks, each made of a combination of a convolutional layer with kernel size 8 followed by ReLU activation and a convolution with kernel size 1 followed by gated linear units (GLU).
+The U-network structure allows the decoder to directly access the input signal and to transfer its phase, improving results.
 
 			\subsubsection{OpenUnmix}
-			@Leo/Alejandro
-
-			OpenUnmix \cite{open-unmix}
+			Open-Unmix~\cite{open-unmix} is based on a bidirectional LSTM model from~\cite{Uhlich2017}.
+It is an open-source implementation of a musical source separation system.
+It takes as input the Short-Time Fourier Transform of the input signal.
 

### 2021-03-05 source_separation/paper/paper.tex
Minor changes to abstract

@@
 Second, we select several state-of-the-art models for both music source separation (separation of instruments) and universal sound separation (separation of arbitrary sounds of different types), and compare their effectiveness for the task of orchestration.
-By applying these methods to hand picked target sounds, then orchestrating the separated outputs, and finally comparing them to the orchestration of the same target without separation, we can determine which methods best suit the needs of orchestration.
+We assessed which methods best suit the needs of orchestration by applying them on hand picked target sounds, orchestrating the separated outputs, and finally comparing them to the orchestration of the same target without separation.
 Our experiments show that the quality of orchestrations improved, both qualitatively and quantitatively, indicating that our approach is promising.

### 2021-03-05 source_separation/paper/paper.tex
added title
typos in abstract

@@
 % ------
-\def\papertitle{Paper Title}
+\def\papertitle{Source separation methods for computer-assisted orchestration}
 
@@
 % mbox force an author not to be split over multiple lines
-\author[1]{\mbox{\firstname{Carmine}\lastname{Cella}}}
 \author[1]{\mbox{\firstname{Luke}\lastname{Dzwonczyk}}}
+\author[1, 2]{\mbox{\firstname{L\'eo}\lastname{Chedin}}}
+\author[1]{\mbox{\firstname{Alejandro}\lastname{Saldarriaga-Fuertes}}}
+\author[1]{\mbox{\firstname{Max}\lastname{Sherr}}}
+\author[1]{\mbox{\firstname{H\'el\`ene-Camille}\lastname{Crayencour}}}
+\author[1]{\mbox{\firstname{Carmine-Emanuele}\lastname{Cella}}}
 
@@
 \affil[1]{\department{Center for New Music and Audio Technologies}\institution{University of California, Berkeley}\city{Berkeley}\state{California}\country{USA}\affiliationtype{University}}
+\affil[2]{\department{ENS}\institution{}\city{Paris}\state{}\country{France}\affiliationtype{University}}
 
@@
 	\begin{abstract}
-		In this paper, we will study the possibility of adding source separation as a pre-processing step to the orchestration process.
+		In this paper, we will study the possibility of adding source separation as a pre-processing step to the computer-assisted orchestration process.
 We first discuss the motivation of this addition and its effect on the quality of orchestrations of multi-layered sounds.

### 2021-03-05 source_separation/paper/paper.tex
Finalize abstract:

@@
 	\begin{abstract}
-		In this paper, we will study the possibility of adding source separation as a pre-processing step to the computer-assisted orchestration process.
-We first discuss the motivation of this addition and its effect on the quality of orchestrations of multi-layered sounds.
+		In this paper, we will study the possibility of adding source separation as a pre-processing step to the orchestration process.
+We first discuss the motivation of this addition and its effects on the quality of orchestrations of multi-layered sounds.
 Second, we select several state-of-the-art models for both music source separation (separation of instruments) and universal sound separation (separation of arbitrary sounds of different types), and compare their effectiveness for the task of orchestration.
-We assessed which methods best suit the needs of orchestration by applying them on hand picked target sounds, orchestrating the separated outputs, and finally comparing them to the orchestration of the same target without separation.
-Our experiments show that the quality of orchestrations improved, both qualitatively and quantitatively, indicating that our approach is promising.
-Finally, we discuss the interest of unsupervised methods compared to supervised methods in the case of assisted orchestration, where data is relatively scarce.
+We assess which methods best suit the needs of orchestration by applying them on hand picked target sounds, orchestrating the separated outputs, and finally comparing them to the orchestration of the same target without separation.
+Our experiments show that the quality of orchestrations improves, both qualitatively and quantitatively, indicating that our approach is promising.
+Finally, we compare unsupervised methods to supervised methods in the case of assisted orchestration, where data is relatively scarce.
 	\end{abstract}

### 2021-03-05 source_separation/paper/paper.tex

@@
 	\begin{abstract}
-		In this paper, we will study the possibility of adding source separation as a pre-processing step to the orchestration process.
+		In this paper, we will study the possibility of adding source separation as a pre-processing step to the computer-assisted orchestration process.
 We first discuss the motivation of this addition and its effects on the quality of orchestrations of multi-layered sounds.

### 2021-03-05 source_separation/paper/paper.tex
Add descr of offset in orchestrations

@@
 
-		\subsection{Data}
+		\subsection{Data}\label{subsec:data}
 		We created our own targets as combinations of four source sounds.
@@
 		\end{enumerate}			
-		The sounds chosen include alarms, bells, engine noises, animals, soundscapes, synthesizer chords, and sound effects.
-Each target that we used for testing was a combination of 4 randomly chosen source sounds from the group of 90 sounds.
-During the creation of the targets, the sources were randomly offset in time from the beginning of the target, so that they began playing at different times.
+		The sounds chosen include alarms, bells, engine noises, soundscapes, synthesizer chords, and sound effects.
+Each target that we used for testing was a combination of four randomly chosen source sounds from the group of 90 sounds.
+		When a target is created, the longest of the four sources is chosen to start playing at the beginning of the target.
+The other three sources are then randomly assigned different times to begin playing such that they will start between the beginning of the target and half-way through the longest source.
 
@@
 		The orchestrations performed were static orchestrations, meaning a single onset of notes is created for each target, no matter if the target itself has multiple onsets.
+
+		As described in Section \ref{subsec:data}, each target is comprised of four sources that are offset in time to begin playing at different times.
+\textbf{Figure needed} illustrates the effect this offset has on the full target, separated, and ground truth orchestrations.
+The full target orchestration places all onsets at the beginning of the orchestration, since it is unable to differentiate between different sources, it cannot place the different orchestrations at the appropriate offsets.
+However, the separated orchestrations are able to correctly place the onsets of the sub-target orchestrations so far as the method has correctly separated out the sources.
+In the ground truth orchestration, the sources are offset by the same amount as they are in the creation of the target, leading to the ground truth having each source orchestration beginning at the correct time.
 

### 2021-03-05 source_separation/paper/paper.tex
Add paragraph on sound source separation

@@
 	\section{Methodology}\label{sec:methodology}
-	\begin{itemize}
-		\item compare a number of source separation methods
-		\item we use the final orchestration to compare the separation methods
-	\end{itemize}
+	Sound source separation is the process by which a single audio file is separated into multiple sound sources.
+A perfect separation is able to dissect a sound exactly into its constituent parts, in which each part is an independent sound event.
+Music source separation is specific application of sound source in which the input audio is expected to be music that is comprised of a subset of instruments or sounds.
+For example, one music source separation method could attempt to split a song into three parts: voice, bass, and drums.
+Another aimed at orchestral music may try to separate the input into families of instruments: woodwinds, brass, strings, and percussion.
+In contrast, universal sound separation does not operate under the assumption that the input is of a musical nature, and can be applied to arbitrary sounds.
+In this case, the number of sources expected is defined, and the method will attempt to divide the input into the given number of sources.
 

### 2021-03-05 source_separation/paper/paper.tex
Work on introduction

@@
 % mbox force an author not to be split over multiple lines
-\author[1]{\mbox{\firstname{Luke}\lastname{Dzwonczyk}}}
+\author[1]{\mbox{\firstname{Luke}\lastname{Dzwonczyk}\email{dz.luke@berkeley.edu}}}
 \author[1, 2]{\mbox{\firstname{L\'eo}\lastname{Chedin}}}
@@
 	\section{Introduction}\label{sec:introduction}
-	\begin{itemize}
-		\item what is CAO 
-		\item why we need source separation 
-		\item which type of sounds are useful for us (targets)
-		\item What is Orchidea?
-	\end{itemize}
 
@@
 Within CAC, target-based computer-assisted orchestration is a compelling case of how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.
-
-	The approach studied in \cite{Carpentier2010} consists in finding a good orchestration for any given sound by searching combinations of sounds from a database with a multi-objective optimization heuristics and a constraint solver that are jointly optimized.
-Both the target sound and the sounds in the database are embedded in a feature space defined by a fixed feature function and each generated combination of sounds is evaluated by using a specific metric.
-This method has been substantially improved in \cite{Cella18, Cella2020} and is implemented in the \emph{Orchidea} toolbox for assisted orchestration (\url{www.orch-idea.org}), currently considered the state-of-the-art system for assisted orchestration.
-
-	 The codebase for this paper can be found at: \url{}.
 
@@
 Computer-assisted orchestration seeks to help composers by accelerating certain creative processes.
-In particular, target-based computer-assisted orchestration helps composers to explore new timbral possibilities by associating instrumental samples, which, combined, will sound perceptually similar to a target sound.
-\textbf{(Note: This paragraph was moved to the intro from the abstract.
-We can move it back to the abstract if that would be better)}
+In particular, target-based computer-assisted orchestration helps composers explore new timbral possibilities by combining instrumental samples in such a way as to mimic the timbre of a given target sound.
+Formally, it is the process of creating an orchestral score that best matches an arbitrary target sound given a similarity metric and constraints~\cite{Maresz2003}.
 
-	 Target-based computer-assisted orchestration is the process of creating an orchestral score that best matches an arbitrary target sound given a similarity metric and constraints.
-It recreates the timbre and harmonics of the target sound using orchestral instruments.
+	 Orchidea is a framework and set of tools that is currently considered the state-of-the-art for computer-assisted orchestration.
+Orchidea embeds the target in a feature space and uses a jointly-optimized multi-objective optimization heuristic and constraint solver to find a combination of samples that best match the target~\cite{Cella18, Cella2020}.
+
+	Sound source separation is the process by which a single audio file is separated into multiple sound sources.
+A perfect separation is able to dissect a sound exactly into its constituent parts, in which each part is an independent sound event.
+Music source separation is specific application of sound source in which the input audio is expected to be music that is comprised of a subset of instruments or sounds.
+For example, one music source separation method could attempt to split a song into three parts: voice, bass, and drums.
+Another aimed at orchestral music may try to separate the input into families of instruments: woodwinds, brass, strings, and percussion.
+In contrast, universal sound separation does not operate under the assumption that the input is of a musical nature, and can be applied to arbitrary sounds.
+In this case, the number of sources expected is specified, and the method will attempt to divide the input into the given number of sources.
 
@@
 Orchestral music often uses multiple layers, so by separating the layers of a target and individually orchestrating them, the resulting solution is improved.
-\textbf{(@Carmine, can you make sure this paragraph makes sense/is correct?)}
 
@@
 
+	The codebase for this paper can be found at: \url{url needed}, and you can listen to our targets and orchestrations here: \url{url needed}.
+
 	\section{Methodology}\label{sec:methodology}
-	Sound source separation is the process by which a single audio file is separated into multiple sound sources.
-A perfect separation is able to dissect a sound exactly into its constituent parts, in which each part is an independent sound event.
-Music source separation is specific application of sound source in which the input audio is expected to be music that is comprised of a subset of instruments or sounds.
-For example, one music source separation method could attempt to split a song into three parts: voice, bass, and drums.
-Another aimed at orchestral music may try to separate the input into families of instruments: woodwinds, brass, strings, and percussion.
-In contrast, universal sound separation does not operate under the assumption that the input is of a musical nature, and can be applied to arbitrary sounds.
-In this case, the number of sources expected is defined, and the method will attempt to divide the input into the given number of sources.
 
@@
 	\section{Experiments}\label{sec:experiments}
-	\begin{itemize}
-		\item Full orchestration
-		\item Ground truth
-		\item NMF
-		\item 4 neural models: TDCNN++, TDCNN, Demucs, OpenUnmix
-		\item table with results
-	\end{itemize}
 
@@
 A final 1-D convolutional layer is used as a decoder to recover the waveforms of each sub target.
-This type of architecture replaces RNNs in modelling sequential data, as series of dilateds convolutions are a good way to capture temporal patterns.\\
+This type of architecture replaces RNNs in modelling sequential data, as series of dilateds convolutions are a good way to capture temporal patterns.
 
@@
 More specifically, feature-wise normalization replaces global normalization, residual connections ranges are extended to improve the flow of information, and learnable scale parameters are introduced to weight the outputs of each residual layer.
-Those modifications help the model to better extract features without drastically increasing its size.\\
+Those modifications help the model to better extract features without drastically increasing its size.
 
@@
 	In this paper, we addressed the problem of orchestrating dynamic targets using source separation as a preprocessing step.
-The results showed that performing source separation followed by a static orchestration of each individual sub targets improved the spectral distance between the original sample and its orchestrated version.\\
+The results showed that performing source separation followed by a static orchestration of each individual sub targets improved the spectral distance between the original sample and its orchestrated version.
 
@@
 Those versions were trained on musical data, except for TDCNN++ which used selected generic sound samples, which opens a large path for improvement.
-More specifically, we may infer from our results that models trained not only on musical data can generate a better separation for orchestration, as TDCNN++ outperforms other supervised models.\\
+More specifically, we may infer from our results that models trained not only on musical data can generate a better separation for orchestration, as TDCNN++ outperforms other supervised models.
 

### 2021-03-08 source_separation/paper/paper.tex
Add names and affiliations

@@
 \author[1]{\mbox{\firstname{Luke}\lastname{Dzwonczyk}\email{dz.luke@berkeley.edu}}}
-\author[1, 2]{\mbox{\firstname{L\'eo}\lastname{Chedin}}}
+\author[1, 2, 3]{\mbox{\firstname{L\'eo}\lastname{Chedin}}}
 \author[1]{\mbox{\firstname{Alejandro}\lastname{Saldarriaga-Fuertes}}}
 \author[1]{\mbox{\firstname{Max}\lastname{Sherr}}}
-\author[1]{\mbox{\firstname{H\'el\`ene-Camille}\lastname{Crayencour}}}
+\author[3]{\mbox{\firstname{H\'el\`ene-Camille}\lastname{Crayencour}}}
 \author[1]{\mbox{\firstname{Carmine-Emanuele}\lastname{Cella}}}
@@
 \affil[1]{\department{Center for New Music and Audio Technologies}\institution{University of California, Berkeley}\city{Berkeley}\state{California}\country{USA}\affiliationtype{University}}
-\affil[2]{\department{ENS}\institution{}\city{Paris}\state{}\country{France}\affiliationtype{University}}
+\affil[2]{\department{\'Ecole Normale Sup\'erieure Paris-Saclay}\institution{Universi\'e Paris-Saclay}\city{Paris}\state{}\country{France}\affiliationtype{University}}
+\affil[3]{\department{Laboratoire des Signaux et Syst\`emes}\institution{Centrale Supélec, CNRS, Université Paris-Saclay}\city{Paris}\state{}\country{France}\affiliationtype{University}}
 

### 2021-03-08 source_separation/paper/paper.tex
Add results

@@
 \ref{eqn:distance} between corresponding frames.
-The spectral distance metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization, and used in \cite{Cella2020b} \textbf{(is it weird/bad to cite our last paper?)} to compare accuracies of orchestrations.
+The spectral distance metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization, and used in \cite{Cella2020b} to compare accuracies of orchestrations.
 The equation takes in the full FFT of the target $x$ and full FFT of the solution $\tilde{x}$.
@@
 The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.
+For our purposes, we used $\lambda_1 = 0.5$ and $\lambda_2 = 10$, which penalizes a solution that overshoots the harmonic energy of the target.
 
@@
 				\hline
-				Full target & 25.73 \\
+				Full target & 24.96 \\
 				\hline
-				TDCNN++ & 25.44 \\
+				NMF & 22.44 \\
 				\hline
-				Demucs & 27.62 \\
+				Demucs & 25.11 \\
 				\hline
-				NMF & 22.32 \\
+				OpenUnmix & 24.48\\
 				\hline
-				Ground truth & 24.27 \\
+				TDCNN & 26.44\\
+				\hline
+				TDCNN++ & 24.38 \\
+				\hline
+				Ground truth & 24.19 \\
 				\hline
@@
 		\end{center}
-		\caption{Average across \textbf{X} targets showing the distance between target and orchestration for various methods. "Full target" means no separation was performed. }
+		\caption{Average across 300 targets showing the distance between target and orchestration for various methods. "Full target" means no separation was performed.}
 		\label{tab:distances}
 	\end{table}
+
+		\subsection{Results}
+		The average distances for full target, separated, and ground truth orchestrations are displayed in Table \ref{tab:distances}.
+These results are the average distance between target and orchestration for 300 different targets.
+We found that NMF performed the best out of all separation methods, and even outperformed the ground truth orchestrations.
+TDCNN++, OpenUnmix, and NMF also performed better than the full target orchestrations, meaning that separating targets using these methods leads to a better orchestration than if no separation is applied.
 

### 2021-03-08 source_separation/paper/paper.tex
Edit intro

@@
 
-	\textit{The following (from here until the line) is directly copy/pasted from our last paper, it will need to be changed in order to avoid self-plagiarism}
-
-	The development of computational tools to assist and inspire the musical composition process constitutes an important research area known as \emph{Computer-Assisted Composition (CAC)} \cite{FerVic2013, Ari2005}.
-Within CAC, target-based computer-assisted orchestration is a compelling case of how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.
-
-	 ----------------------------------------------------------
-
-	 \textit{Everything below this line in the introduction is newly written for this paper}
-
-	 In the field of computer-assisted composition, musical orchestration is probably the most empirical one and it has been traditionally taught in an intuitive and non formalized way.
-Computer-assisted orchestration seeks to help composers by accelerating certain creative processes.
+	Computer-assisted composition is a field that focuses on the creation of computational tools to aid in the musical composition process \cite{FerVic2013, Ari2005}.
+Within this field, target-based computer-assisted musical orchestration is an example of how musical creativity can be assisted and supported by tools such as machine learning.
+Orchestration has been traditionally taught in an intuitive and non formalized way, and computer-assisted orchestration seeks to help composers by accelerating certain creative processes.
 In particular, target-based computer-assisted orchestration helps composers explore new timbral possibilities by combining instrumental samples in such a way as to mimic the timbre of a given target sound.

### 2021-03-08 source_separation/paper/paper.tex
Edit beginning of Experiments

@@
 
-	Targets are separated into four sub-targets.
-Each sub-target is independently orchestrated with a randomly assigned "sub-orchestra." These orchestrations are then combined to play simultaneously, creating a final orchestrated solution.
+	We test the effectiveness of five difference source separation methods, both supervised and unsupervised, on 300 custom made targets that are combinations of sounds from different databases.
+During testing, targets input to each source separation method, which outputs four sub-targets.
+Then each sub-target is independently orchestrated with a randomly assigned subset of the full orchestra.
+These orchestrations are then combined to play simultaneously, creating a final orchestrated solution.
 Then the distance between the target and solution is calculated, giving us a metric to compare the various separation methods.

### 2021-03-08 source_separation/create_targets.py
Small changes to pipeline

@@ -9,7 +9,7 @@ from itertools import product
 
 from pipeline import NUM_SUBTARGETS, SAMPLE_DATABASE_PATH, TARGETS_PATH, TARGET_METADATA_PATH, SAMPLING_RATE, clear_directory, combine_with_offset
 
-NUM_TARGETS = 100  # number of targets to create
+NUM_TARGETS = 300  # number of targets to create
 
 def create_targets(paths):
     """

### 2021-03-08 source_separation/pipeline.py

@@ -22,7 +22,7 @@ SOL_PATH = "../OrchideaSOL2020"  # path to the SOL database
 ANALYSIS_DB_PATH = "../OrchideaSOL2020.spectrum.db"  # path to the analysis file of SOL, ex: TinySOL.spectrum.db
 TARGETS_PATH = "./targets"  # path to targets to be orchestrated
 SAMPLE_DATABASE_PATH = "./subtargets"  # path to samples that will be combined to create targets
-TDCNNpp_model_path = "./trained_TDCNNpp"  # path to pretrained TDCNNpp model
+TDCNNpp_model_path = "./trained_TDCNNpp"  # path to pretrained TDCNN++ model
 
 # If you want to save orchestrations, set the following variables
 SAVED_ORCHESTRATIONS_PATH = "./saved_orchestrations"
@@ -43,7 +43,7 @@ full_orchestra = ['Fl', 'Fl', 'Ob', 'Ob', 'ClBb', 'ClBb', 'Bn', 'Bn', 'Tr', 'Tr'
                   'Vn', 'Vn', 'Vn', 'Vn', 'Vn', 'Vn', 'Vn', 'Vn', 'Va', 'Va', 'Va', 'Va', 'Vc', 'Vc', 'Vc', 'Cb']
 
 separation_models = ["TDCNN++", "TDCNN", "Demucs", "OpenUnmix", "NMF"]
-thresholds = [2]  # onset thresholds for dynamic orchestration
+thresholds = [0.1]  # onset thresholds for dynamic orchestration
 
 
 def clear_temp():
@@ -290,7 +290,7 @@ def frame_spectral_distance(target, solution):
     target = librosa.util.fix_length(target, length)
     solution = librosa.util.fix_length(solution, length)
 
-    frame_length = 4000  # 4000 samples ~ 10 ms
+    frame_length = 4000  # 4000 samples ~ 90 ms
     target_frames = librosa.util.frame(target, frame_length, frame_length, axis=0)
     # last_frame = target[num_frames * frame_length:]
     # last_frame = librosa.util.fix_length(last_frame, frame_length)
@@ -618,10 +618,10 @@ if __name__ == "__main__":
             json.dump(results, f, indent='\t')
 
         # remove temp files created during separation
-        # for model_name in separation_models:
-        #     path = os.path.join(TEMP_OUTPUT_PATH, model_name, target_name)
-        #     if os.path.exists(path):
-        #         remove_directory(path)
+        for model_name in separation_models:
[…]

### 2021-03-08 source_separation/results.py

@@ -24,7 +24,7 @@ if __name__ == "__main__":
     separated_target_distances = results['separated_target_distances']
     ground_truth_distances = results['ground_truth_distances']
 
-    aggregate_data(lambda x: x[0], median, 'Median')
+    # aggregate_data(lambda x: x[0], median, 'Median')
     aggregate_data(lambda x: x[0], mean, 'Mean')
 
     x = 0  # num times separate is better than full
@@ -46,8 +46,6 @@ if __name__ == "__main__":
             y += 1
         if ground < full:
             z += 1
-
-        ground_increase += (full - ground)
 
     print("Separate is better than full:", x / num_completed * 100, "%")
     print("Ground is better than separate:", y / num_completed * 100, "%")

### 2021-03-08 source_separation/paper/paper.tex
Add future work; add ref to SOL and orchestra used

@@
 		The orchestrations performed were static orchestrations, meaning a single onset of notes is created for each target, no matter if the target itself has multiple onsets.
+The OrchideaSOL database of orchestral samples was used with Orchidea to create the orchestrations~\cite{Cella2020c}.
+This database contains recordings of extended playing techniques, which better fits the often noisy or inharmonic nature of the targets we used.
+The full orchestra, of which non-overlapping subsets were selected to orchestrate the subtargets, contained the following instruments: 8 violins, 4 violas, 3 cellos, 1 bass, 2 oboes, 2 flutes, 2 clarinets in B$\flat$, 2 bassoons, 2 trumpets, 2 trombones, and 2 French horns.
 
@@
 	\section{Future Work}\label{sec:futurework}
-	Implementing these methods in Orchidea.
-We could improve the supervised methods by training them ourselves with data that fits our problem better.
-Adding a joint optimization to determine which instruments should be used in which sources.
+
+	As we stated in Section \ref{sec:conclusions}, we believe that one of the reasons the supervised methods did not perform as well is because of the data they were trained on.
+Therefore, a natural next step is to train these models ourselves with sounds that are more applicable to computer-assisted orchestration.
+
+	Another aspect we would like to work on is the assignment of the orchestras.
+Currently, the subset of the orchestra that is assigned to each sub-target is randomly selected.
+However, it is possible that this random assignment could reduce the quality of the final orchestration.
+If, for example, one of the sub-targets contains mostly low pitch content, but the orchestra assigned to it contains mostly violins, violas, flutes, and oboes, the resulting orchestration will suffer.
+To solve this problem, a step could be added that jointly-optimizes the the orchestras assigned to each sub-target, ultimately choosing an assignment that leads to the best orchestration of each sub-target.
+
+	Finally, one or multiple of the separation methods we tested should be implemented in Orchidea in order to improve the orchestrations Orchidea is capable of creating.
+NMF performed the best across the targets we tested on, and therefore makes sense to be added to Orchidea.
+However, it could also make sense to add one of the musical source spearation methods, such as OpenUnmix, to allow users to choose different methods based on the targets they are orchestrating.
 

### 2021-03-09 source_separation/paper/paper.tex
Correct TDCN name
It should be TDCN and not TDCNN for Time-Dilated Convolutional Network (Network is already in the acronym!)

@@
 			\subsubsection{ConvTasNet}
-			ConvTasNet \cite{tdcnn} (TDCNN) is a model based on a temporal convolutional encoder-decoder.
+			ConvTasNet \cite{tdcn} (TDCN) is a model based on a temporal convolutional encoder-decoder.
 It uses the audio waveform as input and performs a series of 1-D convolutions at different resolutions with residual connections in order to generate a pool of masks that are then applied on the embedding of the initial waveform to output the separated targets.
@@
 			\subsubsection{Universal Sound Separation ConvTasNet}
-			This version of ConvTasNet, called TDCNN++, was improved by \cite{tdcnnpp} in order to be used for general sound separation, and not only for speech.
-The main architecture remains the same as in \cite{tdcnn}, and the changes only affected the masking network.
+			This version of ConvTasNet, called TDCN++, was improved by \cite{tdcnpp} in order to be used for general sound separation, and not only for speech.
+The main architecture remains the same as in \cite{tdcn}, and the changes only affected the masking network.
 More specifically, feature-wise normalization replaces global normalization, residual connections ranges are extended to improve the flow of information, and learnable scale parameters are introduced to weight the outputs of each residual layer.
@@
 
-			TDCNN++ presents the interesting property of having almost the same architecture as TDCNN but trained with generic sounds.
+			TDCN++ presents the interesting property of having almost the same architecture as TDCN but trained with generic sounds.
 Thus, it should naturally be better suited for our task, and the comparison between those two models in particular can give good insights about the impact of the data type used for training.
@@
 				\hline
-				TDCNN & 26.44\\
+				TDCN & 26.44\\
 				\hline
-				TDCNN++ & 24.38 \\
+				TDCN++ & 24.38 \\
 				\hline
@@
 We found that NMF performed the best out of all separation methods, and even outperformed the ground truth orchestrations.
-TDCNN++, OpenUnmix, and NMF also performed better than the full target orchestrations, meaning that separating targets using these methods leads to a better orchestration than if no separation is applied.
+TDCN++, OpenUnmix, and NMF also performed better than the full target orchestrations, meaning that separating targets using these methods leads to a better orchestration than if no separation is applied.
 
@@
 However, it is important to notice that the different neural models (Demucs, ConvTasNet, ConvTasNetUniversal and OpenUnmix) were used based on pre-trained versions.
-Those versions were trained on musical data, except for TDCNN++ which used selected generic sound samples, which opens a large path for improvement.
-More specifically, we may infer from our results that models trained not only on musical data can generate a better separation for orchestration, as TDCNN++ outperforms other supervised models.
+Those versions were trained on musical data, except for TDCN++ which used selected generic sound samples, which opens a large path for improvement.
+More specifically, we may infer from our results that models trained not only on musical data can generate a better separation for orchestration, as TDCN++ outperforms other supervised models.
 

### 2021-03-09 source_separation/paper/paper.tex
Edit conclusion

@@
 	\section{Conclusions}\label{sec:conclusions}
-	In this paper, we addressed the problem of orchestrating dynamic targets using source separation as a preprocessing step.
-The results showed that performing source separation followed by a static orchestration of each individual sub targets improved the spectral distance between the original sample and its orchestrated version.
+	In this paper, we addressed the problem of using source separation as a pre-processing step for computer-assisted orchestration.
+The results show that performing source separation followed by a static orchestration of each individual sub-target improved the spectral distance between the original target and its orchestrated version in three out of five of the methods tested.
 
-	Furthermore, it seems that unsupervised methods such as NMF outperform all supervised methods.
-However, it is important to notice that the different neural models (Demucs, ConvTasNet, ConvTasNetUniversal and OpenUnmix) were used based on pre-trained versions.
-Those versions were trained on musical data, except for TDCN++ which used selected generic sound samples, which opens a large path for improvement.
-More specifically, we may infer from our results that models trained not only on musical data can generate a better separation for orchestration, as TDCN++ outperforms other supervised models.
+	Furthermore, it seems that unsupervised methods, such as NMF, outperform all supervised methods tested.
+However, it is important to notice that the pre-trained version of the neural models (Demucs, TDCN, TDCN++ and OpenUnmix) were used.
+Except for TDCN++, these models were trained on musical data, yet we tested them on a wide range of targets, many of which were not strictly musical.
+More specifically, we may infer from our results that models trained on generic sounds generate a better separation for orchestration, as TDCN++ outperformed all other supervised models.
 
-	Since unsupervised methods perform well, it is reasonable to think that supervised methods trained on a dataset well suited for source separation could yield better overall performances on dynamic orchestration.
+	Since unsupervised methods perform well, it is reasonable to think that supervised methods trained on a dataset of arbitrary sounds could yield better overall performance for orchestration.
 

### 2021-03-09 source_separation/paper/paper.tex
Edit separation methods

@@
 			The Demucs~\cite{demucs} architecture is a deep learning model for musical source separation.
-It takes as input the stereo mixture, and outputs 4 stereo waveforms corresponding to the stems (vocals, bass, drums and others).
-Demucs is based on a convolutional encoder, a bidirectional LSTM and a convolutional decoder.
+It takes as input the stereo mixture, and outputs 4 stereo waveforms corresponding to the stems (vocals, bass, drums and other).
+Demucs is based on a convolutional encoder, a bidirectional LSTM, and a convolutional decoder.
 The encoder and the decoder are linked with skip U-Net connections.
-Both the encoder and the decoder are made of 6  blocks, each made of a combination of a convolutional layer with kernel size 8 followed by ReLU activation and a convolution with kernel size 1 followed by gated linear units (GLU).
+Both the encoder and the decoder are made of 6 blocks, each made of a combination of a convolutional layer with kernel size 8, followed by ReLU activation.
+Finally there is another convolution with kernel size 1, followed by gated linear units (GLU).
 The U-network structure allows the decoder to directly access the input signal and to transfer its phase, improving results.
@@
 A final 1-D convolutional layer is used as a decoder to recover the waveforms of each sub target.
-This type of architecture replaces RNNs in modelling sequential data, as series of dilateds convolutions are a good way to capture temporal patterns.
+This type of architecture replaces RNNs in modeling sequential data, as series of dilated convolutions are a good way to capture temporal patterns.
 
@@
 The main architecture remains the same as in \cite{tdcn}, and the changes only affected the masking network.
-More specifically, feature-wise normalization replaces global normalization, residual connections ranges are extended to improve the flow of information, and learnable scale parameters are introduced to weight the outputs of each residual layer.
+More specifically, feature-wise normalization replaces global normalization, residual connection ranges are extended to improve the flow of information, and learnable scale parameters are introduced to weight the outputs of each residual layer.
 Those modifications help the model to better extract features without drastically increasing its size.
@@
 			TDCN++ presents the interesting property of having almost the same architecture as TDCN but trained with generic sounds.
-Thus, it should naturally be better suited for our task, and the comparison between those two models in particular can give good insights about the impact of the data type used for training.
+Thus, it should naturally be better suited for our task, and the comparison between those two models in particular can give good insights about the impact of the data used for training.
 
@@
 We found that NMF performed the best out of all separation methods, and even outperformed the ground truth orchestrations.
-TDCN++, OpenUnmix, and NMF also performed better than the full target orchestrations, meaning that separating targets using these methods leads to a better orchestration than if no separation is applied.
+TDCN++, OpenUnmix, and NMF performed better than the full target orchestrations, meaning that separating targets using these methods leads to a better orchestration than if no separation is applied.
 

### 2021-03-09 source_separation/paper/paper.tex
Reword conclusion section

@@
 
-	Furthermore, it seems that unsupervised methods, such as NMF, outperform all supervised methods tested.
-However, it is important to notice that the pre-trained version of the neural models (Demucs, TDCN, TDCN++ and OpenUnmix) were used.
-Except for TDCN++, these models were trained on musical data, yet we tested them on a wide range of targets, many of which were not strictly musical.
-More specifically, we may infer from our results that models trained on generic sounds generate a better separation for orchestration, as TDCN++ outperformed all other supervised models.
+	Furthermore, the one unsupervised method we tested, NMF, outperformed all the supervised methods.
+It is important to notice that for the supervised methods, we used pre-trained versions of the neural models.
+All of these models, except for TDCN++, were trained on musical data.
+However, we tested them on a wide range of targets, many of which were not strictly musical.
+This could explain why Demucs, OpenUnmix, and TDCN performed worse than NMF, and why TDCN++, which was trained on arbitrary sounds, outperformed the other supervised methods.
+Therefore, we may infer from our results that models trained on generic sounds generate a better separation for orchestration when compared to models trained only on musical sounds.
 

### 2021-03-09 source_separation/paper/paper.tex
Improve OpenUnmix description

@@
 
+			\subsubsection{OpenUnmix}
+			Open-Unmix~\cite{open-unmix} is an open source implementation of a deep music source separation model, based on \cite{Uhlich207}.
+It is actually composed of multiple models that are trained for each target (such as 'vocals', 'drums', ...).
+Each of these models is trained on the specific target using the same architecture, based on a three layer bidirectionnal LSTM network.
+Open-Unmix operates on the Short Time Fourier Transform of the input mixture, which is first standardized.
+It predicts the target by multiplying the output of the LSTM network with the magnitude spectogram of the input, essentially applying a mask.
+Open-Unmix also use Wiener filtering as a last processing step.
+
 			\subsubsection{Demucs}
@@
 The U-network structure allows the decoder to directly access the input signal and to transfer its phase, improving results.
-
-			\subsubsection{OpenUnmix}
-			Open-Unmix~\cite{open-unmix} is based on a bidirectional LSTM model from~\cite{Uhlich2017}.
-It is an open-source implementation of a musical source separation system.
-It takes as input the Short-Time Fourier Transform of the input signal.
 

### 2021-03-09 source_separation/paper/paper.tex
Add the accent to my name
Add orch comparison figure

@@
 \author[1]{\mbox{\firstname{Luke}\lastname{Dzwonczyk}\email{dz.luke@berkeley.edu}}}
-\author[1, 2, 3]{\mbox{\firstname{L\'eo}\lastname{Chedin}}}
+\author[1, 2, 3]{\mbox{\firstname{L\'eo}\lastname{Ch\'edin}}}
 \author[1]{\mbox{\firstname{Alejandro}\lastname{Saldarriaga-Fuertes}}}
@@
 
+		\begin{figure}[t]
+		\centering
+			\includegraphics[width=\columnwidth]{figures/orch.jpg}
+			\caption{Examples waveforms showing onsets of four different sources in the target, and possible onsets in the full, separated, and ground truth orchestrations.
+In this example, the separation method could not distinguish between sources 2 and 3, and therefore identified them as a single sub-target.
+Therefore, the separated orchestration has a single orchestration that combines sources 2 and 3.
+Note that this is a simplified representation, in the actual targets sources can overlap in time and are not as discrete as this diagram shows.}\label{fig:orchestrations}
+		\end{figure}
+
 		As described in Section \ref{subsec:data}, each target is comprised of four sources that are offset in time to begin playing at different times.
-\textbf{Figure needed} illustrates the effect this offset has on the full target, separated, and ground truth orchestrations.
+Figure \ref{fig:orchestrations} illustrates the effect this offset has on the full target, separated, and ground truth orchestrations.
 The full target orchestration places all onsets at the beginning of the orchestration, since it is unable to differentiate between different sources, it cannot place the different orchestrations at the appropriate offsets.

### 2021-03-09 source_separation/paper/paper.tex
Update fig caption

--- 2021-03-09 source_separation/paper/paper.tex
@@
 
+		\begin{figure}[t]
+		\centering
+			\includegraphics[width=\columnwidth]{figures/orch.jpg}
+			\caption{Example waveforms showing onsets of the four sources in the target, and possible onsets in the full, separated, and ground truth orchestrations.
+In this example, the separation method could not distinguish between sources 2 and 3 and identified them as a single sub-target.
+Therefore, the separated orchestration has only 3 onsets, one of them being an orchestration that combines sources 2 and 3.
+Note that this is a simplified representation, in the actual targets sources overlap in time and are not as discrete as this diagram shows.}\label{fig:orchestrations}
+		\end{figure}
+
 		As described in Section \ref{subsec:data}, each target is comprised of four sources that are offset in time to begin playing at different times.
-\textbf{Figure needed} illustrates the effect this offset has on the full target, separated, and ground truth orchestrations.
+Figure \ref{fig:orchestrations} illustrates the effect this offset has on the full target, separated, and ground truth orchestrations.
 The full target orchestration places all onsets at the beginning of the orchestration, since it is unable to differentiate between different sources, it cannot place the different orchestrations at the appropriate offsets.

### 2021-03-09 source_separation/paper/paper.tex
Added diagram

@@
 
-	We compare the effectiveness of different source separation methods for the task of orchestration by applying the separation methods to various targets, orchestrating the separated output of the method, and finally comparing this orchestration to the orchestration of the target if no separation was performed \textbf{Diagram needed}.
+	We compare the effectiveness of different source separation methods for the task of orchestration by applying the separation methods to various targets, orchestrating the separated output of the method, and finally comparing this orchestration to the orchestration of the target if no separation was performed (Fig.
+\ref{fig:full_diagram}).
 For example, consider a target sound that is a combination of a low droning sound and high-pitched whistle.
@@
 This would then be compared to the orchestration if no separation had been performed.
+
+			\begin{figure}[t]
+		\centering
+			\includegraphics[width=\columnwidth]{figures/diagram.png}
+			\caption{Diagram showing the pipeline of the orchestration with (bottom) and without (top) separation.
+The pipeline using separation divides the target into subtargets that will be orchestrated separately using specific subsets of the orchestra, and then recombined to obtain the full orchestration.}\label{fig:full_diagram}
+		\end{figure}
 

### 2021-03-10 source_separation/paper/paper.tex
add references for NMF
fix accent in institutions

--- 2021-03-09 source_separation/paper/paper.tex
@@
 \affil[2]{\department{\'Ecole Normale Sup\'erieure Paris-Saclay}\institution{Universi\'e Paris-Saclay}\city{Paris}\state{}\country{France}\affiliationtype{University}}
-\affil[3]{\department{Laboratoire des Signaux et Syst\`emes}\institution{Centrale Supélec, CNRS, Université Paris-Saclay}\city{Paris}\state{}\country{France}\affiliationtype{University}}
+\affil[3]{\department{Laboratoire des Signaux et Syst\`emes}\institution{Centrale Sup\'elec, CNRS, Universit\'e Paris-Saclay}\city{Paris}\state{}\country{France}\affiliationtype{University}}
 
@@
 			@Max 
+			We use scikit learn implementation. Related papers:
+			\cite{Cichocki_NMF} \cite{Fevotte_NMF}
 

### 2021-03-10 source_separation/paper/paper.tex
Added NMF section, fixed a broken citation, some grammar

@@
 			\subsubsection{Non-negative Matrix Factorization} 
-			@Max 
-			We use scikit learn implementation. Related papers:
-			\cite{Cichocki_NMF} \cite{Fevotte_NMF}
+			Non-negative matrix factorization \cite{Cichocki_NMF} \cite{Fevotte_NMF} is an \`a priori method, which takes an input non-negative matrix, in this case a spectrogram $S$, in $\mathbb{R} ^{N\times M}$ and decomposes it into a basis matrix $A$ in $\mathbb{R}^{N\times J}$ and a component matrix $X$ in $\mathbb{R}^{J\times M}$ such that $S\approx A\times X$.
+We use the coordinate descent algorithm from the scikit-learn Python package based on \cite{Cichocki_NMF} with $J$ set to $4$, which selects for 4 features in the dataset.
+We then output 4 mono waveforms containing each isolated feature $F_i$, where $F_i=A_i\times X_i$ for $i = 1, 2, 3$, or $4$.
 
 			\subsubsection{OpenUnmix}
-			Open-Unmix~\cite{open-unmix} is an open source implementation of a deep music source separation model, based on \cite{Uhlich207}.
+			Open-Unmix~\cite{open-unmix} is an open source implementation of a deep music source separation model, based on \cite{Uhlich2017}.
 It is actually composed of multiple models that are trained for each target (such as 'vocals', 'drums', ...).
@@
 It predicts the target by multiplying the output of the LSTM network with the magnitude spectogram of the input, essentially applying a mask.
-Open-Unmix also use Wiener filtering as a last processing step.
+Open-Unmix also uses Wiener filtering as a last processing step.
 

### 2021-03-10 source_separation/paper/paper.tex
Add short introductive text to separation methods

@@
 		\subsection{Separation Methods}
-
+	We consider source separation models trained to tackle two distinct separation problems.
+The first is music source separation (Open-Unmix, Demucs, ConvTasnet).
+For these architectures, we used models that were pre-trained on the MUSDB18 dataset~\cite{MUSDB18}.
+Thus, these models outputs four stereo tracks, that correspond to broad instrumental categories defined in the SiSEC 2018 Mus evaluation campaign~\cite{Stoter_SiSEC}, i.e.
+\textit{vocals, drums, bass} and \textit{other}.
+The audio sources we consider in this work are often non-musical, but we hope that these source separation methods can still improve the orchestration by separating other types of sound in a musically meaningful way.
+	The second one is universal sound separation (TDCN++, NMF), which \cite{tdcnpp} defines as "separating mixtures of arbitrary sounds of different types". 
 			\subsubsection{Non-negative Matrix Factorization} 
@@
 The main architecture remains the same as in \cite{tdcn}, and the changes only affected the masking network.
-More specifically, feature-wise normalization replaces global normalization, residual connection ranges are extended to improve the flow of information, and learnable scale parameters are introduced to weight the outputs of each residual layer.
+More specifically, feature-wise normalizatio	n replaces global normalization, residual connection ranges are extended to improve the flow of information, and learnable scale parameters are introduced to weight the outputs of each residual layer.
 Those modifications help the model to better extract features without drastically increasing its size.

### 2021-03-10 source_separation/paper/paper.tex
fix typo in affiliations

@@
 \affil[1]{\department{Center for New Music and Audio Technologies}\institution{University of California, Berkeley}\city{Berkeley}\state{California}\country{USA}\affiliationtype{University}}
-\affil[2]{\department{\'Ecole Normale Sup\'erieure Paris-Saclay}\institution{Universi\'e Paris-Saclay}\city{Paris}\state{}\country{France}\affiliationtype{University}}
+\affil[2]{\department{\'Ecole Normale Sup\'erieure Paris-Saclay}\institution{Universit\'e Paris-Saclay}\city{Paris}\state{}\country{France}\affiliationtype{University}}
 \affil[3]{\department{Laboratoire des Signaux et Syst\`emes}\institution{Centrale Sup\'elec, CNRS, Universit\'e Paris-Saclay}\city{Paris}\state{}\country{France}\affiliationtype{University}}

### 2021-03-10 source_separation/paper/paper.tex
correct mistakes regarding models based on Conv-Tasnet

@@
 For these architectures, we used models that were pre-trained on the MUSDB18 dataset~\cite{MUSDB18}.
-Thus, these models outputs four stereo tracks, that correspond to broad instrumental categories defined in the SiSEC 2018 Mus evaluation campaign~\cite{Stoter_SiSEC}, i.e.
+Thus, these models outputs 4 stereo tracks, that correspond to broad instrumental categories defined in the SiSEC 2018 Mus evaluation campaign~\cite{Stoter_SiSEC}, i.e.
 \textit{vocals, drums, bass} and \textit{other}.
 The audio sources we consider in this work are often non-musical, but we hope that these source separation methods can still improve the orchestration by separating other types of sound in a musically meaningful way.
-	The second one is universal sound separation (TDCN++, NMF), which \cite{tdcnpp} defines as "separating mixtures of arbitrary sounds of different types". 
+	The second one is universal sound separation (TDCN++, NMF), which \cite{tdcnpp} defines as \textit{separating mixtures of arbitrary sounds of different types}.
+
 			\subsubsection{Non-negative Matrix Factorization} 
@@
 			Open-Unmix~\cite{open-unmix} is an open source implementation of a deep music source separation model, based on \cite{Uhlich2017}.
-It is actually composed of multiple models that are trained for each target (such as 'vocals', 'drums', ...).
+It is actually composed of multiple models that are trained for each instrumental target.
 Each of these models is trained on the specific target using the same architecture, based on a three layer bidirectionnal LSTM network.
-Open-Unmix operates on the Short Time Fourier Transform of the input mixture, which is first standardized.
+Open-Unmix operates on the Short Time Fourier Transform (STFT) of the input mixture, which is first standardized.
 It predicts the target by multiplying the output of the LSTM network with the magnitude spectogram of the input, essentially applying a mask.
@@
 
-			\subsubsection{ConvTasNet}
-			ConvTasNet \cite{tdcn} (TDCN) is a model based on a temporal convolutional encoder-decoder.
+			\subsubsection{Conv-TasNet}
+			We also used the Conv-Tasnet\cite{tdcn} implementation of \cite{demucs}, who adapted the architecture, originally designed for speech separation at 8 kHz, for the task of music source separation. 
+			Conv-Tasnet uses  a combination of learned time-domain analysis and synthesis transforms with a time-dilated convolutional network (TDCN)
+			Conv-Tasnet (TDCN) is a model based on a temporal convolutional encoder-decoder.
 It uses the audio waveform as input and performs a series of 1-D convolutions at different resolutions with residual connections in order to generate a pool of masks that are then applied on the embedding of the initial waveform to output the separated targets.
@@
 
-			The purpose of the model was initially to separate multiple speakers from a single-channel audio, and was thus trained on speech datasets.
-Because of its good performance at speech separation, a pre-trained version of the model based on the settings of the paper was added to the benchmark.
-
-			\subsubsection{Universal Sound Separation ConvTasNet}
-			This version of ConvTasNet, called TDCN++, was improved by \cite{tdcnpp} in order to be used for general sound separation, and not only for speech.
+			\subsubsection{Universal Sound Separation Conv-Tasnet}
+			This version of Conv-Tasnet, called TDCN++, was improved by \cite{tdcnpp} in order to be used for general sound separation, and not only for speech.
 The main architecture remains the same as in \cite{tdcn}, and the changes only affected the masking network.
-More specifically, feature-wise normalizatio	n replaces global normalization, residual connection ranges are extended to improve the flow of information, and learnable scale parameters are introduced to weight the outputs of each residual layer.
+More specifically, feature-wise normalization replaces global normalization, residual connection ranges are extended to improve the flow of information, and learnable scale parameters are introduced to weight the outputs of each residual layer.
 Those modifications help the model to better extract features without drastically increasing its size.

### 2021-03-10 source_separation/paper/paper.tex
Add to future work, small edits to sep methods

@@
 	We consider source separation models trained to tackle two distinct separation problems.
-The first is music source separation (Open-Unmix, Demucs, ConvTasnet).
+The first is music source separation (OpenUnmix, Demucs, ConvTasnet).
 For these architectures, we used models that were pre-trained on the MUSDB18 dataset~\cite{MUSDB18}.
-Thus, these models outputs 4 stereo tracks, that correspond to broad instrumental categories defined in the SiSEC 2018 Mus evaluation campaign~\cite{Stoter_SiSEC}, i.e.
-\textit{vocals, drums, bass} and \textit{other}.
-The audio sources we consider in this work are often non-musical, but we hope that these source separation methods can still improve the orchestration by separating other types of sound in a musically meaningful way.
-	The second one is universal sound separation (TDCN++, NMF), which \cite{tdcnpp} defines as \textit{separating mixtures of arbitrary sounds of different types}.
+Thus, these models output 4 stereo tracks that correspond to broad instrumental categories defined in the SiSEC 2018 Mus evaluation campaign: \textit{vocals, drums, bass} and \textit{other}~\cite{Stoter_SiSEC}.
+Although the audio sources we consider in this work are often non-musical, we hope that these source separation methods can still improve the orchestration by separating other types of sound in a musically meaningful way.
+	The second type of model we tested are ones trained for universal sound separation (TDCN++, NMF), which \cite{tdcnpp} defines as \textit{separating mixtures of arbitrary sounds of different types}.
 
@@
 Therefore, a natural next step is to train these models ourselves with sounds that are more applicable to computer-assisted orchestration.
+This would be a dataset similar to the samples that make up our targets, instead of the music tracks that the models were trained on.
 

### 2021-03-10 source_separation/paper/paper.tex
Add back Alejandro's diagram

@@
 
-	The codebase for this paper can be found at: \url{url needed}, and you can listen to our targets and orchestrations here: \url{url needed}.
+	The codebase for this paper can be found at: \url{url needed}, and you can listen to a selection of targets and orchestrations here: \url{url needed}.
 
@@
 
-	We compare the effectiveness of different source separation methods for the task of orchestration by applying the separation methods to various targets, orchestrating the separated output of the method, and finally comparing this orchestration to the orchestration of the target if no separation was performed \textbf{Diagram needed}.
+	We compare the effectiveness of different source separation methods for the task of orchestration by applying the separation methods to various targets, orchestrating the separated output of the method, and finally comparing this orchestration to the orchestration of the target if no separation was performed (see Fig.
+\ref{fig:full_diagram}).
 For example, consider a target sound that is a combination of a low droning sound and high-pitched whistle.
@@
 This would then be compared to the orchestration if no separation had been performed.
+
+	\begin{figure}[t]
+		\centering
+			\includegraphics[width=\columnwidth]{figures/diagram.png}
+			\caption{Diagram showing the pipeline of the orchestration with (bottom) and without (top) separation.
+The pipeline using separation divides the target into subtargets that will be orchestrated separately using specific subsets of the orchestra, and then recombined to obtain the full orchestration.}\label{fig:full_diagram}
+	\end{figure}
 

### 2021-03-10 source_separation/paper/paper.tex
Edits throughout all sections

--- 2021-03-10 source_separation/paper/paper.tex
@@
 		In this paper, we will study the possibility of adding source separation as a pre-processing step to the computer-assisted orchestration process.
-We first discuss the motivation of this addition and its effects on the quality of orchestrations of multi-layered sounds.
+We first discuss the motivation of this addition and its potential to increase the quality of orchestrations of multi-layered sounds.
 Second, we select several state-of-the-art models for both music source separation (separation of instruments) and universal sound separation (separation of arbitrary sounds of different types), and compare their effectiveness for the task of orchestration.
@@
 Our experiments show that the quality of orchestrations improves, both qualitatively and quantitatively, indicating that our approach is promising.
-Finally, we compare unsupervised methods to supervised methods in the case of assisted orchestration, where data is relatively scarce.
+Finally, we compare unsupervised methods to supervised methods for separation, and the effect of training data selection on performance for supervised methods.
 	\end{abstract}
@@
 	Computer-assisted composition is a field that focuses on the creation of computational tools to aid in the musical composition process \cite{FerVic2013, Ari2005}.
-Within this field, target-based computer-assisted musical orchestration is an example of how musical creativity can be assisted and supported by tools such as machine learning.
-Orchestration has been traditionally taught in an intuitive and non formalized way, and computer-assisted orchestration seeks to help composers by accelerating certain creative processes.
+Within this field, target-based computer-assisted musical orchestration is an example of how musical creativity can be supported by tools such as machine learning.
+Orchestration has been traditionally taught in an intuitive and non-formalized way, and computer-assisted orchestration seeks to help composers by accelerating certain creative processes.
 In particular, target-based computer-assisted orchestration helps composers explore new timbral possibilities by combining instrumental samples in such a way as to mimic the timbre of a given target sound.
@@
 A perfect separation is able to dissect a sound exactly into its constituent parts, in which each part is an independent sound event.
-Music source separation is specific application of sound source in which the input audio is expected to be music that is comprised of a subset of instruments or sounds.
+Music source separation is a specific application of source separation in which the input audio is expected to be music that is comprised of a subset of instruments or sounds.
 For example, one music source separation method could attempt to split a song into three parts: voice, bass, and drums.
@@
 
-	For example, consider an example target that has a continuous drone sound throughout and a bird song playing above this drone.
-Without source separation, Orchidea would detect each note of the bird song as a new onset and cut the target in time at each new note.
-However, this segmentation in time would also effect the drone, forcing it to have a new onset each time the bird song changes notes.
-The resulting orchestration would having multiple onsets for the drone, when really only one onset is needed at the beginning of the drone.
-When source separation is applied to this target, the drone and bird song would be orchestrated separately.
-Therefore, the drone could be orchestrated by a single onset of note(s) that lasts for the duration of the drone.
-At the same time, the bird song could have as many new onsets as needed without affecting the drone.
+	For example, consider an example target that has a continuous drone sound throughout and a melody playing above this drone.
+Without source separation, Orchidea would detect each note of the melody as a new onset and cut the target in time at each new note.
+However, this segmentation in time would also effect the drone, cutting it in time and forcing it to have a new onset each time the melody changes notes.
+The resulting orchestration would have multiple onsets for the drone, when really only one onset is needed at the beginning of the drone.
+When source separation is applied to this target, the drone and melody would be orchestrated separately.
+Therefore, the drone could be orchestrated by a single onset that lasts for the duration of the drone.
+At the same time, the melody could have as many new onsets as needed without affecting the drone.
 The orchestration would be greatly improved by removing the unnecessary onsets in the orchestration of the drone.
@@
 
-	The codebase for this paper can be found at: \url{url needed}, and you can listen to our targets and orchestrations here: \url{url needed}.
+	The codebase for this paper can be found at: \mbox{\url{https://github.com/dzluke/smc2021}}, and you can listen to a selection of targets and orchestrations here: \mbox{\url{https://dzluke.github.io/smc2021/}}.
 
@@
 
-	We compare the effectiveness of different source separation methods for the task of orchestration by applying the separation methods to various targets, orchestrating the separated output of the method, and finally comparing this orchestration to the orchestration of the target if no separation was performed \textbf{Diagram needed}.
+	We compare the effectiveness of different source separation methods for the task of orchestration by applying the separation methods to various targets, orchestrating the separated output of the method, and finally comparing this to the orchestration of the target if no separation was performed (see Fig.
+\ref{fig:full_diagram}).
 For example, consider a target sound that is a combination of a low droning sound and high-pitched whistle.
@@
 
+	\begin{figure}[t]
+		\centering
+			\includegraphics[width=\columnwidth]{figures/diagram.png}
+			\caption{Diagram comparing the process of orchestration with and without separation.
+At the top, the full target orchestration is created simply by orchestrating the target.
+At the bottom, the separation process is performed, and the resulting sub-targets are individually orchestrated with subsets of the orchestra before being recombined to obtain the full orchestration.}\label{fig:full_diagram}
+	\end{figure}
+
 	\section{Experiments}\label{sec:experiments}
@@
 	We test the effectiveness of five difference source separation methods, both supervised and unsupervised, on 300 custom made targets that are combinations of sounds from different databases.
-During testing, targets input to each source separation method, which outputs four sub-targets.
+During testing, a target is input to a source separation method, which outputs four sub-targets.
 Then each sub-target is independently orchestrated with a randomly assigned subset of the full orchestra.
@@
 Then the distance between the target and solution is calculated, giving us a metric to compare the various separation methods.
+The random split of the orchestra is the same for each separation method of the same target.
 
@@
 		When a target is created, the longest of the four sources is chosen to start playing at the beginning of the target.
-The other three sources are then randomly assigned different times to begin playing such that they will start between the beginning of the target and half-way through the longest source.
+The other three sources are then randomly assigned different times to begin playing such that they will start between the beginning of the target and half-way through the longest source sound.
 
@@
 	We consider source separation models trained to tackle two distinct separation problems.
-The first is music source separation (OpenUnmix, Demucs, ConvTasnet).
+The first is music source separation (Open-Unmix, Demucs, ConvTasnet).
 For these architectures, we used models that were pre-trained on the MUSDB18 dataset~\cite{MUSDB18}.
@@
 
-			\subsubsection{OpenUnmix}
+			\subsubsection{Open-Unmix}
 			Open-Unmix~\cite{open-unmix} is an open source implementation of a deep music source separation model, based on \cite{Uhlich2017}.
 It is actually composed of multiple models that are trained for each instrumental target.
-Each of these models is trained on the specific target using the same architecture, based on a three layer bidirectionnal LSTM network.
+Each of these models is trained on the specific target using the same architecture, based on a three layer bidirectional LSTM network.
 Open-Unmix operates on the Short Time Fourier Transform (STFT) of the input mixture, which is first standardized.
-It predicts the target by multiplying the output of the LSTM network with the magnitude spectogram of the input, essentially applying a mask.
+It predicts the target by multiplying the output of the LSTM network with the magnitude spectrogram of the input, essentially applying a mask.
 Open-Unmix also uses Wiener filtering as a last processing step.
@@
 			The Demucs~\cite{demucs} architecture is a deep learning model for musical source separation.
-It takes as input the stereo mixture, and outputs 4 stereo waveforms corresponding to the stems (vocals, bass, drums and other).
+It takes as input the stereo mixture, and outputs 4 stereo waveforms corresponding to the four categories: vocals, bass, drums and other.
 Demucs is based on a convolutional encoder, a bidirectional LSTM, and a convolutional decoder.
@@
 			We also used the Conv-Tasnet\cite{tdcn} implementation of \cite{demucs}, who adapted the architecture, originally designed for speech separation at 8 kHz, for the task of music source separation.
-			Conv-Tasnet uses  a combination of learned time-domain analysis and synthesis transforms with a time-dilated convolutional network (TDCN)
+Conv-Tasnet uses  a combination of learned time-domain analysis and synthesis transforms with a time-dilated convolutional network (TDCN).
 			Conv-Tasnet (TDCN) is a model based on a temporal convolutional encoder-decoder.
@@
 			This version of Conv-Tasnet, called TDCN++, was improved by \cite{tdcnpp} in order to be used for general sound separation, and not only for speech.
-The main architecture remains the same as in \cite{tdcn}, and the changes only affected the masking network.
+The main architecture remains the same as in \cite{tdcn}, and the changes only affect the masking network.
 More specifically, feature-wise normalization replaces global normalization, residual connection ranges are extended to improve the flow of information, and learnable scale parameters are introduced to weight the outputs of each residual layer.
@@
 			\item The full target, without any separation performed, is orchestrated using the entire orchestra
-			\item A given separation method is performed on the target, splitting the target into 4 sub-targets
-			\item Each sub-target is separately orchestrated using a randomly chosen sub-orchestra that is one quarter the size of the full orchestra, and then each orchestration is combined to play simultaneously
-			\item The ground truth orchestration is created by orchestrating the four sources, each with a different sub-orchestra, and then combined to play simultaneously
+			\item Each separation method is performed on the target, each splitting the target into 4 sub-targets 
+			\item The full orchestra is randomly split into four equal sized subsets
+			\item For each method, the four sub-targets are separately orchestrated, each using a different subset of the orchestra, and then each orchestration is combined to play simultaneously
+			\item The ground truth orchestration is created by orchestrating the four sources, using the same orchestra splits from step 4, and then combined to play simultaneously
 			\item The distances between target and orchestration are calculated for the full target orchestration, the separated orchestration, and the ground truth orchestration 
@@
 
-		The orchestrations performed were static orchestrations, meaning a single onset of notes is created for each target, no matter if the target itself has multiple onsets.
-The OrchideaSOL database of orchestral samples was used with Orchidea to create the orchestrations~\cite{Cella2020c}.
-This database contains recordings of extended playing techniques, which better fits the often noisy or inharmonic nature of the targets we used.
-The full orchestra, of which non-overlapping subsets were selected to orchestrate the subtargets, contained the following instruments: 8 violins, 4 violas, 3 cellos, 1 bass, 2 oboes, 2 flutes, 2 clarinets in B$\flat$, 2 bassoons, 2 trumpets, 2 trombones, and 2 French horns.
+		The orchestrations performed are static orchestrations, meaning a single onset of notes is created for each target, no matter if the target itself has multiple onsets.
+The OrchideaSOL database of orchestral samples is used with Orchidea to create the orchestrations~\cite{Cella2020c}.
+This database contains recordings of extended playing techniques, which better fits the often noisy or inharmonic nature of our targets.
+The full orchestra, of which non-overlapping subsets were selected to orchestrate the subtargets, contains the following instruments: 8 violins, 4 violas, 3 cellos, 1 bass, 2 oboes, 2 flutes, 2 clarinets in B$\flat$, 2 bassoons, 2 trumpets, 2 trombones, and 2 French horns.
 
@@
 
-		The distance metric cuts the target and solution into successive frames that are 4,000 samples in length, then calculates the spectral distance as defined in Eqn.
-\ref{eqn:distance} between corresponding frames.
+		The distance metric cuts the target and solution into successive frames that are 4,000 samples in length, then calculates the spectral distance, as defined in Eqn.
+\ref{eqn:distance}, between corresponding frames.
 The spectral distance metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization, and used in \cite{Cella2020b} to compare accuracies of orchestrations.
@@
 				\hline
-				OpenUnmix & 24.48\\
+				Open-Unmix & 24.48\\
 				\hline
@@
 We found that NMF performed the best out of all separation methods, and even outperformed the ground truth orchestrations.
-TDCN++, OpenUnmix, and NMF performed better than the full target orchestrations, meaning that separating targets using these methods leads to a better orchestration than if no separation is applied.
+TDCN++, Open-Unmix, and NMF performed better than the full target orchestrations, meaning that separating targets using these methods leads to a better orchestration than if no separation is applied.
 
@@
 However, we tested them on a wide range of targets, many of which were not strictly musical.
-This could explain why Demucs, OpenUnmix, and TDCN performed worse than NMF, and why TDCN++, which was trained on arbitrary sounds, outperformed the other supervised methods.
+This could explain why Demucs, Open-Unmix, and TDCN performed worse than NMF, and why TDCN++, which was trained on arbitrary sounds, outperformed the other supervised methods.
 Therefore, we may infer from our results that models trained on generic sounds generate a better separation for orchestration when compared to models trained only on musical sounds.
@@
 NMF performed the best across the targets we tested on, and therefore makes sense to be added to Orchidea.
-However, it could also make sense to add one of the musical source spearation methods, such as OpenUnmix, to allow users to choose different methods based on the targets they are orchestrating.
+However, it could also make sense to add one of the musical source spearation methods, such as Open-Unmix, to allow users to choose different methods based on the targets they are orchestrating.
 
 	\begin{acknowledgments}
-		At the end of the Conclusions, acknowledgements to people, projects, funding agencies, etc. can be included after the second-level heading  ``Acknowledgments'' (with no numbering).
+		Do we have any acknowledgments or should we remove this section?
 	\end{acknowledgments}

### 2021-03-11 source_separation/paper/paper.tex
made some of the suggested changes, added TODO in Discussion section

@@
 	 Orchidea is a framework and set of tools that is currently considered the state-of-the-art for computer-assisted orchestration.
-Orchidea embeds the target in a feature space and uses a jointly-optimized multi-objective optimization heuristic and constraint solver to find a combination of samples that best match the target~\cite{Cella18, Cella2020}.
+Orchidea embeds the target in a feature space and uses a jointly-optimized optimization heuristic and constraint solver to find a combination of samples that best match the target~\cite{Cella18, Cella2020}.
 
@@
 \ref{fig:full_diagram}).
-For example, consider a target sound that is a combination of a low droning sound and high-pitched whistle.
+Following the previous example, consider a target sound that is a combination of a low droning sound and high-pitched whistle.
 With separation, these two sounds would be disentangled into two separate sub-targets and each would be individually orchestrated.
@@
 These results are the average distance between target and orchestration for 300 different targets.
-We found that NMF performed the best out of all separation methods, and even outperformed the ground truth orchestrations.
+We found that NMF performed the best out of all separation methods, and even outperformed the ground truth orchestrations on average.
 TDCN++, Open-Unmix, and NMF performed better than the full target orchestrations, meaning that separating targets using these methods leads to a better orchestration than if no separation is applied.
 
-	\section{Conclusions}\label{sec:conclusions}
+	\section{Discussion}\label{sec:conclusions}
 	In this paper, we addressed the problem of using source separation as a pre-processing step for computer-assisted orchestration.
 The results show that performing source separation followed by a static orchestration of each individual sub-target improved the spectral distance between the original target and its orchestrated version in three out of five of the methods tested.
+This shows that source separation is a useful pre-processing stage for computer assisted orchestration.
 
@@
 
+	TODO: say that unsupervised methods can still do well in particular cases where there are not many examples
+
 	Since unsupervised methods perform well, it is reasonable to think that supervised methods trained on a dataset of arbitrary sounds could yield better overall performance for orchestration.
 
-	\section{Future Work}\label{sec:futurework}
+		\subsection{Future Work}\label{sec:futurework}
 
 		As we stated in Section \ref{sec:conclusions}, we believe that one of the reasons the supervised methods did not perform as well is because of the data they were trained on.
-Therefore, a natural next step is to train these models ourselves with sounds that are more applicable to computer-assisted orchestration.
-This would be a dataset similar to the samples that make up our targets, instead of the music tracks that the models were trained on.
+Therefore, a natural next step is to train these models with sounds that are more appropriate for computer-assisted orchestration.
 
@@
 		Finally, one or multiple of the separation methods we tested should be implemented in Orchidea in order to improve the orchestrations Orchidea is capable of creating.
-NMF performed the best across the targets we tested on, and therefore makes sense to be added to Orchidea.
-However, it could also make sense to add one of the musical source spearation methods, such as Open-Unmix, to allow users to choose different methods based on the targets they are orchestrating.
-
-	\begin{acknowledgments}
-		Do we have any acknowledgments or should we remove this section?
-	\end{acknowledgments}
 

### 2021-03-11 source_separation/paper/paper.tex
added a few more changes I missed

@@
 The orchestration would be greatly improved by removing the unnecessary onsets in the orchestration of the drone.
-This is the effect we hope to achieve through implementing source separation in computer-assisted orchestration.
+This is the effect we achieve through implementing source separation in computer-assisted orchestration.
 
@@
 
+		Suggestion: A block-diagram of the algo would be nice to have
+
 		The procedure we used for testing is as follows:
@@
 		\end{center}
-		\caption{Average across 300 targets showing the distance between target and orchestration for various methods. "Full target" means no separation was performed.}
+		\caption{Average across 300 targets showing the distance between target and orchestration for various methods. "Full target" means no separation was performed. Lower values are better.}
 		\label{tab:distances}

### 2021-03-11 source_separation/paper/paper.tex
Carmine's changes

@@
 Our experiments show that the quality of orchestrations improves, both qualitatively and quantitatively, indicating that our approach is promising.
-Finally, we compare unsupervised methods to supervised methods for separation, and the effect of training data selection on performance for supervised methods.
+Finally, we compare unsupervised methods to supervised methods for separation, and comment on the effect of training data selection on performance of supervised methods.
 	\end{abstract}
@@
 	 Orchidea is a framework and set of tools that is currently considered the state-of-the-art for computer-assisted orchestration.
-Orchidea embeds the target in a feature space and uses a jointly-optimized optimization heuristic and constraint solver to find a combination of samples that best match the target~\cite{Cella18, Cella2020}.
+Orchidea embeds the target in a feature space and uses a jointly-optimized heuristic and constraint solver to find a combination of samples that best match the target~\cite{Cella18, Cella2020}.
 
@@
 
-		Suggestion: A block-diagram of the algo would be nice to have
-
 		The procedure we used for testing is as follows:
@@
 		\end{center}
-		\caption{Average across 300 targets showing the distance between target and orchestration for various methods. "Full target" means no separation was performed. Lower values are better.}
+		\caption{Average across 300 targets showing the distance between target and orchestration for various methods. ``Full target'' means no separation was performed. Lower values are better.}
 		\label{tab:distances}
@@
 These results are the average distance between target and orchestration for 300 different targets.
-We found that NMF performed the best out of all separation methods, and even outperformed the ground truth orchestrations on average.
+We found that NMF performed the best out of all separation methods, and even outperformed the ground truth orchestrations, on average.
 TDCN++, Open-Unmix, and NMF performed better than the full target orchestrations, meaning that separating targets using these methods leads to a better orchestration than if no separation is applied.
 
-	\section{Discussion}\label{sec:conclusions}
+	\section{Discussion}\label{sec:discussion}
 	In this paper, we addressed the problem of using source separation as a pre-processing step for computer-assisted orchestration.
 The results show that performing source separation followed by a static orchestration of each individual sub-target improved the spectral distance between the original target and its orchestrated version in three out of five of the methods tested.
-This shows that source separation is a useful pre-processing stage for computer assisted orchestration.
+This shows that source separation is a useful pre-processing stage for computer-assisted orchestration.
 
 	Furthermore, the one unsupervised method we tested, NMF, outperformed all the supervised methods.
+When training data for a particular problem is scarce, it is logical that unsupervised methods will perform well, as they do not rely on training data.
 It is important to notice that for the supervised methods, we used pre-trained versions of the neural models.
@@
 
-	TODO: say that unsupervised methods can still do well in particular cases where there are not many examples
-
 	Since unsupervised methods perform well, it is reasonable to think that supervised methods trained on a dataset of arbitrary sounds could yield better overall performance for orchestration.
@@
 
-		As we stated in Section \ref{sec:conclusions}, we believe that one of the reasons the supervised methods did not perform as well is because of the data they were trained on.
+		As we stated previously, we believe that one of the reasons the supervised methods did not perform as well is because of the data they were trained on.
 Therefore, a natural next step is to train these models with sounds that are more appropriate for computer-assisted orchestration.

### 2021-03-11 source_separation/paper/paper.tex

@@
 		\subsection{Future Work}\label{sec:futurework}	
-
 		As we stated previously, we believe that one of the reasons the supervised methods did not perform as well is because of the data they were trained on.

### 2021-03-11 source_separation/paper/paper.tex
Add Helene-Camille's paragraph

@@
 In this case, the number of sources expected is specified, and the method will attempt to divide the input into the given number of sources.
+
+	Using source separation techniques as a pre-processing step has been shown to improve results in several music processing tasks.
+Source separation can be used to remove part of the spectrum in order to enhance the clarity of the features of interest for the task at hand.
+For instance, in the task of automatic chord estimation from audio, some authors have chosen to remove part of the spectrum related to percussive sounds to improve chord accuracy \cite{Reed_al2009}.
+Similarly some authors have investigated how beat tracking can be improved by using source separation as a pre-processing step for difficult songs with highly predominant expressive vocals \cite{ZapGom2012}.
+Source separation can also be applied to separate different components of the signal and work on them separately to reduce the complexity of the task.
+For instance, in the context of tempo detection~\cite{ChoRae2009}, the authors decompose the signal into sources to reduce rhythmic complexity, under the idea that some layers may be more rhythmically regular than the overall mix.
+Another example can be found in~\cite{GomAbecan2018} where harmonic/percussive and solo/accompaniment source separation techniques are investigated as a pre-processing step to improve predominant instrument recognition in jazz music.
 

### 2021-03-11 source_separation/paper/paper.tex
Final edits

@@
 
-	For example, consider an example target that has a continuous drone sound throughout and a melody playing above this drone.
+	For example, consider a target that has a continuous drone sound throughout and a melody playing above this drone.
 Without source separation, Orchidea would detect each note of the melody as a new onset and cut the target in time at each new note.
 However, this segmentation in time would also effect the drone, cutting it in time and forcing it to have a new onset each time the melody changes notes.
-The resulting orchestration would have multiple onsets for the drone, when really only one onset is needed at the beginning of the drone.
+The resulting orchestration would have multiple onsets for the drone, even though only one onset is needed at the beginning of the drone.
 When source separation is applied to this target, the drone and melody would be orchestrated separately.
@@
 			\item The full target, without any separation performed, is orchestrated using the entire orchestra
-			\item Each separation method is performed on the target, each splitting the target into 4 sub-targets 
+			\item Each separation method is performed on the target, each splitting the target into four sub-targets 
 			\item The full orchestra is randomly split into four equal sized subsets
@@
 
-		The distance metric cuts the target and solution into successive frames that are 4,000 samples in length, then calculates the spectral distance, as defined in Eqn.
-\ref{eqn:distance}, between corresponding frames.
+		The distance metric cuts the target and solution into successive frames that are 4,000 samples in length, which is approximately 90 milliseconds at a sampling rate of 44.1 kHz.
+The spectral distance, as defined in Eqn.
+\ref{eqn:distance}, is calculated between corresponding frames of the target and solution.
 The spectral distance metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization, and used in \cite{Cella2020b} to compare accuracies of orchestrations.
@@
 If, for example, one of the sub-targets contains mostly low pitch content, but the orchestra assigned to it contains mostly violins, violas, flutes, and oboes, the resulting orchestration will suffer.
-To solve this problem, a step could be added that jointly-optimizes the the orchestras assigned to each sub-target, ultimately choosing an assignment that leads to the best orchestration of each sub-target.
+To solve this problem, a step could be added that jointly-optimizes the orchestras assigned to each sub-target, ultimately choosing an assignment that leads to the best orchestration of each sub-target.
 

### 2021-03-12 source_separation/paper/paper.tex
Change order of separation methods and results table

@@
 \usepackage[figure,table]{hypcap}% hyperref companion
+\usepackage{multirow}
 % Enable for Review only, remove for Camera Ready version
@@
 
-			\subsubsection{Non-negative Matrix Factorization} 
-			Non-negative matrix factorization \cite{Cichocki_NMF} \cite{Fevotte_NMF} is an \`a priori method, which takes an input non-negative matrix, in this case a spectrogram $S$, in $\mathbb{R} ^{N\times M}$ and decomposes it into a basis matrix $A$ in $\mathbb{R}^{N\times J}$ and a component matrix $X$ in $\mathbb{R}^{J\times M}$ such that $S\approx A\times X$.
-We use the coordinate descent algorithm from the scikit-learn Python package based on \cite{Cichocki_NMF} with $J$ set to $4$, which selects for 4 features in the dataset.
-We then output 4 mono waveforms containing each isolated feature $F_i$, where $F_i=A_i\times X_i$ for $i = 1, 2, 3$, or $4$.
-
-			\subsubsection{Open-Unmix}
+			\subsubsection{Music source separation}
+			\textbf{Open-Unmix}\\
 			Open-Unmix~\cite{open-unmix} is an open source implementation of a deep music source separation model, based on \cite{Uhlich2017}.
@@
 
-			\subsubsection{Demucs}
+			\textbf{Demucs}\\
 			The Demucs~\cite{demucs} architecture is a deep learning model for musical source separation.
@@
 
-			\subsubsection{Conv-TasNet}
+			\textbf{Conv-TasNet}\\
 			We also used the Conv-Tasnet\cite{tdcn} implementation of \cite{demucs}, who adapted the architecture, originally designed for speech separation at 8 kHz, for the task of music source separation.
@@
 
-			\subsubsection{Universal Sound Separation Conv-Tasnet}
+			\subsubsection{Universal sound separation}
+			\textbf{Non-negative Matrix Factorization}\\
+			Non-negative matrix factorization \cite{Cichocki_NMF} \cite{Fevotte_NMF} is an \`a priori method, which takes an input non-negative matrix, in this case a spectrogram $S$, in $\mathbb{R} ^{N\times M}$ and decomposes it into a basis matrix $A$ in $\mathbb{R}^{N\times J}$ and a component matrix $X$ in $\mathbb{R}^{J\times M}$ such that $S\approx A\times X$.
+We use the coordinate descent algorithm from the scikit-learn Python package based on \cite{Cichocki_NMF} with $J$ set to $4$, which selects for 4 features in the dataset.
+We then output 4 mono waveforms containing each isolated feature $F_i$, where $F_i=A_i\times X_i$ for $i = 1, 2, 3$, or $4$.
+
+			\textbf{Universal Sound Separation Conv-Tasnet}\\
 			This version of Conv-Tasnet, called TDCN++, was improved by \cite{tdcnpp} in order to be used for general sound separation, and not only for speech.
@@
 Those modifications help the model to better extract features without drastically increasing its size.
-
 			TDCN++ presents the interesting property of having almost the same architecture as TDCN but trained with generic sounds.
@@
 				\hline
-				Full target & 24.96 \\
-				\hline
-				NMF & 22.44 \\
-				\hline
 				Demucs & 25.11 \\
@@
 				\hline
+				NMF & 22.44 \\
+				\hline
 				TDCN++ & 24.38 \\
+				\hline
+				Full target & 24.96 \\
 				\hline
@@
 		\label{tab:distances}
+	\end{table}
+
+	\begin{table}[t]
+		\begin{center}
+		\begin{tabular}{|c|c|c|c|}
+			\hline
+			& Model & Average distance & Variance\\
+			\hline
+			\multirow{3}{4em}{Music source separation} & Open-Unmix & 24.48 & \\
+			& Demucs & 25.11 & \\
+			& TDCN & 26.44 & \\
+			\hline
+			\multirow{2}{4em}{Universal separation} & NMF & 22.44 & \\
+			& TDCN+++ & 24.38 & \\
+			\hline
+			\multirow{2}{4em}{} & Full target & 24.96 & \\
+			\hline
+			& Ground truth & 24.19 & \\
+			\hline
+		\end{tabular}
+		\end{center}
+		\caption{Average across 300 targets showing the distance between target and orchestration for various methods. ``Full target'' means no separation was performed. Lower values are better.}
+		\label{tab:test}
 	\end{table}

### 2021-03-12 source_separation/results.py
Add plot with avg and stddev

@@ -1,5 +1,6 @@
 import json
-from statistics import median
+from statistics import median, stdev
+import matplotlib.pyplot as plt
 
 from pipeline import thresholds, mean, RESULTS_PATH
 
@@ -24,7 +25,37 @@ if __name__ == "__main__":
     separated_target_distances = results['separated_target_distances']
     ground_truth_distances = results['ground_truth_distances']
 
-    # aggregate_data(lambda x: x[0], median, 'Median')
+    y = []
+    e = []
+    x_labels = []
+
+    # order: Open-Unmix, Demucs, TDCN, NMF, TDCN++, Full, Ground
+
+    def make_plot_data(distances, label):
+        distances = [x[0] for x in distances]
+        y.append(mean(distances))
+        e.append(stdev(distances))
+        x_labels.append(label)
+
+    make_plot_data(separated_target_distances['OpenUnmix'], 'Open-Unmix')
+    make_plot_data(separated_target_distances['Demucs'], 'Demucs')
+    make_plot_data(separated_target_distances['TDCNN'], 'TDCN')
+    make_plot_data(separated_target_distances['NMF'], 'NMF')
+    make_plot_data(separated_target_distances['TDCNN++'], 'TDCN++')
+    make_plot_data(full_target_distances, 'Full target')
+    make_plot_data(ground_truth_distances, 'Ground truth')
+
+    x = list(range(len(y)))
+    plt.title("Average distance between target and solution with standard deviation")
+    plt.errorbar(x, y, e, linestyle='None', marker='.', markersize=10, capsize=3, color='black')
+    plt.xticks(x, x_labels, rotation=15)
+    plt.xlabel("Orchestration type")
+    plt.ylabel("Distance")
+    plt.ylim(0, 50)
+    plt.tight_layout()
+    plt.show()
+
+    aggregate_data(lambda x: x[0], median, 'Median')
     aggregate_data(lambda x: x[0], mean, 'Mean')
 
     x = 0  # num times separate is better than full

### 2021-03-12 source_separation/paper/paper.tex
Add plot with avg and stddev
Capitalize headings
Add to future work

@@
 
-			\subsubsection{Music source separation}
+			\subsubsection{Music Source Separation}
 			\textbf{Open-Unmix}\\
@@
 
-			\subsubsection{Universal sound separation}
+			\subsubsection{Universal Sound Separation}
 			\textbf{Non-negative Matrix Factorization}\\
@@
 
+	\begin{figure}[t]
+		\centering
+			\includegraphics[width=\columnwidth]{figures/graph.png}
+			\caption{Plot showing the average distance and standard deviation between target and solution for each orchestration method.}\label{fig:plot}
+	\end{figure}
+
 		\subsection{Results}
@@
 
+		As we stated in Section \ref{sec:introduction}, we chose static orchestration instead of dynamic orchestration to avoid entangling the problems of time segmentation and source separation.
+Future work could include applying source separation to dynamic targets and performing dynamic orchestration.
+
 		Finally, one or multiple of the separation methods we tested should be implemented in Orchidea in order to improve the orchestrations Orchidea is capable of creating.

### 2021-03-12 source_separation/paper/paper.tex
Added evaluation diagram
Fix level of headings problem and improve clarity of TDCN description

--- 2021-03-12 source_separation/paper/paper.tex
@@
 			\subsubsection{Music source separation}
-			\textbf{Open-Unmix}\\
-			Open-Unmix~\cite{open-unmix} is an open source implementation of a deep music source separation model, based on \cite{Uhlich2017}.
+			\textbf{Open-Unmix}~\cite{open-unmix} is an open source implementation of a deep music source separation model, based on \cite{Uhlich2017}.
 It is actually composed of multiple models that are trained for each instrumental target.
@@
 
-			\textbf{Demucs}\\
-			The Demucs~\cite{demucs} architecture is a deep learning model for musical source separation.
+			The \textbf{Demucs}~\cite{demucs} architecture is a deep learning model for musical source separation.
 It takes as input the stereo mixture, and outputs 4 stereo waveforms corresponding to the four categories: vocals, bass, drums and other.
@@
 
-			\textbf{Conv-TasNet}\\
-			We also used the Conv-Tasnet\cite{tdcn} implementation of \cite{demucs}, who adapted the architecture, originally designed for speech separation at 8 kHz, for the task of music source separation.
-Conv-Tasnet uses  a combination of learned time-domain analysis and synthesis transforms with a time-dilated convolutional network (TDCN).
-			Conv-Tasnet (TDCN) is a model based on a temporal convolutional encoder-decoder.
+			We also used the Conv-Tasnet\cite{tdcn} implementation of~\cite{demucs}, who adapted the architecture, originally designed for speech separation at 8 kHz, for the task of music source separation.
+Conv-Tasnet is a model based on a convolutional encoder-decoder, with a time-dilated convolutional network separation module, estimating masks based on the encoder output.
+This is why we refer to it as \textbf{TDCN}, in accordance with~\cite{tdcnpp}.
 It uses the audio waveform as input and performs a series of 1-D convolutions at different resolutions with residual connections in order to generate a pool of masks that are then applied on the embedding of the initial waveform to output the separated targets.
@@
 			\subsubsection{Universal sound separation}
-			\textbf{Non-negative Matrix Factorization}\\
-			Non-negative matrix factorization \cite{Cichocki_NMF} \cite{Fevotte_NMF} is an \`a priori method, which takes an input non-negative matrix, in this case a spectrogram $S$, in $\mathbb{R} ^{N\times M}$ and decomposes it into a basis matrix $A$ in $\mathbb{R}^{N\times J}$ and a component matrix $X$ in $\mathbb{R}^{J\times M}$ such that $S\approx A\times X$.
+			\textbf{Non-negative matrix factorization}~\cite{Cichocki_NMF, Fevotte_NMF} is an \`a priori method, which takes an input non-negative matrix, in this case a spectrogram $S$, in $\mathbb{R} ^{N\times M}$ and decomposes it into a basis matrix $A$ in $\mathbb{R}^{N\times J}$ and a component matrix $X$ in $\mathbb{R}^{J\times M}$ such that $S\approx A\times X$.
 We use the coordinate descent algorithm from the scikit-learn Python package based on \cite{Cichocki_NMF} with $J$ set to $4$, which selects for 4 features in the dataset.
@@
 
-			\textbf{Universal Sound Separation Conv-Tasnet}\\
-			This version of Conv-Tasnet, called TDCN++, was improved by \cite{tdcnpp} in order to be used for general sound separation, and not only for speech.
+			This version of Conv-Tasnet, called \textbf{TDCN++}, was improved by \cite{tdcnpp} in order to be used for general sound separation, and not only for speech.
 The main architecture remains the same as in \cite{tdcn}, and the changes only affect the masking network.
@@
 		\end{enumerate}
+
+		\begin{figure}[t]
+			\centering
+				\includegraphics[width=\columnwidth]{figures/evaluation_diagram.png}
+				\caption{Diagram showing the evaluation method. Note that in both orchestrations with split orchestra, the split is the same.}\label{fig:full_diagram}
+		\end{figure}
 

### 2021-03-12 source_separation/paper/paper.tex
Add Leo's paragraph on static vs dynamic

@@
 
+	Computer-assisted orchestrations can be divided into two categories: static and dynamic.
+Static orchestration considers the target as a single vector of timbral features and creates a single onset of note(s) as the solution.
+Dynamic orchestration considers a time series of features, which allows it to create an orchestration that evolves over time with multiple onsets.
+
 	Orchidea is a framework and set of tools that is currently considered the state-of-the-art for computer-assisted orchestration.
-Orchidea embeds the target in a feature space and uses a jointly-optimized heuristic and constraint solver to find a combination of samples that best match the target~\cite{Cella18, Cella2020}.
+It embeds the target in a feature space and uses a jointly-optimized heuristic and constraint solver to find a combination of samples that best match the target~\cite{Cella18, Cella2020}.
+Orchidea is able to do both static and dynamic orchestration.
+In order to better observe the effects of source separation, we perform only static orchestrations in this paper.
+The reasoning for this is that we wish to disentangle the two difficult problems of source separation and time segmentation.
+By only performing static orchestrations, we reduce the complexity of the problem by removing time segmentation and can better observe the effects of source separation.
 

### 2021-03-12 source_separation/paper/paper.tex
Fix sentence about unsupervised methods not needing data

@@
 	Furthermore, the one unsupervised method we tested, NMF, outperformed all the supervised methods.
-When training data for a particular problem is scarce, it is logical that unsupervised methods will perform well, as they do not rely on training data.
+When training data for a particular problem is scarce, it is logical that unsupervised methods will perform well, as they require less data than supervised methods.
 It is important to notice that for the supervised methods, we used pre-trained versions of the neural models.

### 2021-03-12 source_separation/paper/paper.tex
Minor fixes

@@
 The random split of the orchestra is the same for each separation method of the same target.
+See Figure \ref{fig:eval_diagram} for a diagram of this process.
 
@@
 
-			This version of Conv-Tasnet, called \textbf{TDCN++}, was improved by \cite{tdcnpp} in order to be used for general sound separation, and not only for speech.
+			\textbf{TDCN++}, a version of Conv-Tasnet, was improved by \cite{tdcnpp} in order to be used for general sound separation, and not only for speech.
 The main architecture remains the same as in \cite{tdcn}, and the changes only affect the masking network.
@@
 				\includegraphics[width=\columnwidth]{figures/evaluation_diagram.png}
-				\caption{Diagram showing the evaluation method. Note that in both orchestrations with split orchestra, the split is the same.}\label{fig:full_diagram}
+				\caption{Diagram showing the testing method. Note that in both orchestrations that use a split orchestra, the split is the same.}\label{fig:eval_diagram}
 		\end{figure}

### 2021-03-12 source_separation/paper/paper.tex
Edits from reading through, add back future work paragraph and plot

@@
 Dynamic orchestration considers a time series of features, which allows it to create an orchestration that evolves over time with multiple onsets.
+It uses time segmentation to cut the target in time into multiple segments, and orchestrates each segment.
 
@@
 The random split of the orchestra is the same for each separation method of the same target.
-See Figure \ref{fig:eval_diagram} for a diagram of this process.
+See Fig.
+\ref{fig:eval_diagram} for a diagram of this process.
 
@@
 
-		The procedure we used for testing is as follows:
+		The procedure we use for testing is as follows:
 		\begin{enumerate}
@@
 				\includegraphics[width=\columnwidth]{figures/evaluation_diagram.png}
-				\caption{Diagram showing the testing method. Note that in both orchestrations that use a split orchestra, the split is the same.}\label{fig:eval_diagram}
+				\caption{Diagram showing the testing process. Note that in both orchestrations that use a split orchestra, the split is the same.}\label{fig:eval_diagram}
 		\end{figure}
@@
 In the ground truth orchestration, the sources are offset by the same amount as they are in the creation of the target, leading to the ground truth having each source orchestration beginning at the correct time.
+For this reason, despite only performing static orchestrations, the resulting separated and ground truth orchestrations are offset in time compared to the full target orchestration.
 

### 2021-03-12 source_separation/paper/paper.tex
added items 3 and 4 from TODO

@@
 			\subsubsection{Universal sound separation}
-			\textbf{Non-negative matrix factorization}~\cite{Cichocki_NMF, Fevotte_NMF} is an \`a priori method, which takes an input non-negative matrix, in this case a spectrogram $S$, in $\mathbb{R} ^{N\times M}$ and decomposes it into a basis matrix $A$ in $\mathbb{R}^{N\times J}$ and a component matrix $X$ in $\mathbb{R}^{J\times M}$ such that $S\approx A\times X$.
+			\textbf{Non-negative matrix factorization}~\cite{Cichocki_NMF, Fevotte_NMF} is an unsupervised learning method, which takes an input non-negative matrix, in this case a spectrogram $S$, in $\mathbb{R} ^{N\times M}$ and decomposes it into a basis matrix $A$ in $\mathbb{R}^{N\times J}$ and a component matrix $X$ in $\mathbb{R}^{J\times M}$ such that $S\approx A\times X$.
 We use the coordinate descent algorithm from the scikit-learn Python package based on \cite{Cichocki_NMF} with $J$ set to $4$, which selects for 4 features in the dataset.
@@
 	Furthermore, the one unsupervised method we tested, NMF, outperformed all the supervised methods.
+This suggests that methods like NMF can still be useful for specific problems when available data is limited.
 When training data for a particular problem is scarce, it is logical that unsupervised methods will perform well, as they require less data than supervised methods.
+The NMF also outperformed the ground truth orchestrations, which could be due to the relative simplicity of NMF's output subtargets making subsequent orchestration easier for Orchidea, but more experimentation would need to be done in order to confirm this.
 It is important to notice that for the supervised methods, we used pre-trained versions of the neural models.

### 2021-03-12 source_separation/paper/paper.tex
Made 2nd paragraph a little less awkward

@@
 This suggests that methods like NMF can still be useful for specific problems when available data is limited.
-When training data for a particular problem is scarce, it is logical that unsupervised methods will perform well, as they require less data than supervised methods.
+It is logical that unsupervised methods will perform well for problems in which there is limited data, as they require less data than supervised methods.
 The NMF also outperformed the ground truth orchestrations, which could be due to the relative simplicity of NMF's output subtargets making subsequent orchestration easier for Orchidea, but more experimentation would need to be done in order to confirm this.

### 2021-03-13 source_separation/paper/paper.tex
Last edits

@@
 The orchestration would be greatly improved by removing the unnecessary onsets in the orchestration of the drone.
-This is the effect we achieve through implementing source separation in computer-assisted orchestration.
+Although this example is a dynamic orchestration, and we perform only static orchestrations, this paper represents a first step toward achieving dynamic orchestration with source separation.
 
@@
 
-	\begin{table}[t]
-		\begin{center}
-		\begin{tabular}{|c|c|c|c|}
-			\hline
-			& Model & Average distance & Variance\\
-			\hline
-			\multirow{3}{4em}{Music source separation} & Open-Unmix & 24.48 & \\
-			& Demucs & 25.11 & \\
-			& TDCN & 26.44 & \\
-			\hline
-			\multirow{2}{4em}{Universal separation} & NMF & 22.44 & \\
-			& TDCN+++ & 24.38 & \\
-			\hline
-			\multirow{2}{4em}{} & Full target & 24.96 & \\
-			\hline
-			& Ground truth & 24.19 & \\
-			\hline
-		\end{tabular}
-		\end{center}
-		\caption{Average across 300 targets showing the distance between target and orchestration for various methods. ``Full target'' means no separation was performed. Lower values are better.}
-		\label{tab:test}
-	\end{table}
+%	\begin{table}[t]
+%		\begin{center}
+%		\begin{tabular}{|c|c|c|c|}
+%			\hline
+%			& Model & Average distance & Variance\\
+%			\hline
+%			\multirow{3}{4em}{Music source separation} & Open-Unmix & 24.48 & \\
+%			& Demucs & 25.11 & \\
+%			& TDCN & 26.44 & \\
+%			\hline
+%			\multirow{2}{4em}{Universal separation} & NMF & 22.44 & \\
+%			& TDCN+++ & 24.38 & \\
+%			\hline
+%			\multirow{2}{4em}{} & Full target & 24.96 & \\
+%			\hline
+%			& Ground truth & 24.19 & \\
+%			\hline
+%		\end{tabular}
+%		\end{center}
+%		\caption{Average across 300 targets showing the distance between target and orchestration for various methods. ``Full target'' means no separation was performed. Lower values are better.}
+%		\label{tab:test}
+%	\end{table}
 
-	\begin{figure}[t]
-		\centering
-			\includegraphics[width=\columnwidth]{figures/graph.png}
-			\caption{Plot showing the average distance and standard deviation between target and solution for each orchestration method.}\label{fig:plot}
-	\end{figure}
+%	\begin{figure}[t]
+%		\centering
+%			\includegraphics[width=\columnwidth]{figures/graph.png}
+%			\caption{Plot showing the average distance and standard deviation between target and solution for each orchestration method.}\label{fig:plot}
+%	\end{figure}
 
@@
 It is logical that unsupervised methods will perform well for problems in which there is limited data, as they require less data than supervised methods.
-The NMF also outperformed the ground truth orchestrations, which could be due to the relative simplicity of NMF's output subtargets making subsequent orchestration easier for Orchidea, but more experimentation would need to be done in order to confirm this.
+NMF also outperformed the ground truth orchestrations, which could be due to the relative simplicity of NMF's output subtargets making subsequent orchestration easier for Orchidea, but more experimentation would need to be done in order to confirm this.
 It is important to notice that for the supervised methods, we used pre-trained versions of the neural models.

### 2021-03-22 source_separation/pipeline.py
Correct name of TDCN

@@ -8,7 +8,7 @@ import soundfile as sf
 import librosa
 import json
 
-import ConvTasNetUniversal.separate as TDCNNpp_separate
+import ConvTasNetUniversal.separate as TDCNpp_separate
 import open_unmix.separate as open_unmix
 import demucs.separate as demucs
 import NMF
@@ -22,7 +22,7 @@ SOL_PATH = "../OrchideaSOL2020"  # path to the SOL database
 ANALYSIS_DB_PATH = "../OrchideaSOL2020.spectrum.db"  # path to the analysis file of SOL, ex: TinySOL.spectrum.db
 TARGETS_PATH = "./targets"  # path to targets to be orchestrated
 SAMPLE_DATABASE_PATH = "./subtargets"  # path to samples that will be combined to create targets
-TDCNNpp_model_path = "./trained_TDCNNpp"  # path to pretrained TDCNN++ model
+TDCNpp_model_path = "./trained_TDCNNpp"  # path to pretrained TDCN++ model
 
 # If you want to save orchestrations, set the following variables
 SAVED_ORCHESTRATIONS_PATH = "./saved_orchestrations"
@@ -36,13 +36,13 @@ CONFIG_PATH = "orch_config.txt"  # path to the Orchidea configuration file (you 
 TEMP_OUTPUT_PATH = "./TEMP"
 RESULTS_PATH = "./results.json"
 TARGET_METADATA_PATH = os.path.join(TARGETS_PATH, 'metadata.json')
-TDCNNpp_nb_sub_targets = 4
+TDCNpp_nb_sub_targets = 4
 SAMPLING_RATE = 44100
 NUM_SUBTARGETS = 4
 full_orchestra = ['Fl', 'Fl', 'Ob', 'Ob', 'ClBb', 'ClBb', 'Bn', 'Bn', 'Tr', 'Tr', 'Tbn', 'Tbn', 'Hn', 'Hn',
                   'Vn', 'Vn', 'Vn', 'Vn', 'Vn', 'Vn', 'Vn', 'Vn', 'Va', 'Va', 'Va', 'Va', 'Vc', 'Vc', 'Vc', 'Cb']
 
-separation_models = ["TDCNN++", "TDCNN", "Demucs", "OpenUnmix", "NMF"]
+separation_models = ["TDCN++", "TDCN", "Demucs", "OpenUnmix", "NMF"]
 thresholds = [0.1]  # onset thresholds for dynamic orchestration
 
 
[…]

### 2021-04-08 constraint_filtering/ConstrainedResNet/train.py
Minor changes to constrained resnet

@@ -153,7 +153,7 @@ def train(model, save_path, optimizer, train_load, test_load, start_epoch, out_n
     device = torch.device('cuda' if torch.cuda.is_available() else 'cpu')
     model = model.to(device)
 
-    criterion = CustomLoss(0.001)
+    criterion = CustomLoss(0.0001)
     sig = F.sigmoid
 
     best_acc = 0
@@ -219,7 +219,7 @@ def train(model, save_path, optimizer, train_load, test_load, start_epoch, out_n
                     epoch + 1, GLOBAL_PARAMS.nb_epoch, i+1, size_train_load, out_num*total_loss/50))
                 total_loss = 0
 
-        # graph_loss(criterion.bce_loss_list, criterion.semantic_loss_list, epoch + 1, criterion.semantic_weight)
+        graph_loss(criterion.bce_loss_list, criterion.semantic_loss_list, epoch + 1, criterion.semantic_weight)
 
         if (epoch+1) % 5 == 0:
             model.eval()
@@ -249,10 +249,6 @@ def train(model, save_path, optimizer, train_load, test_load, start_epoch, out_n
 
                 test_loss += float(loss)
 
-                # result_ins = evaluate(predicts, labels, result_ins)
-
-            # pret_tmp[pret_tmp >= 0.05] = 1
-            # pret_tmp[pret_tmp < 0.05] = 0
             result = evaluate(pret_tmp, grod_tmp)
             f = open(save_path + '/result{}.pkl'.format(epoch), 'wb')
             pickle.dump(result, f)
@@ -384,6 +380,10 @@ def calculateConstraint(audio_combos, labels):
     for i in range(labels.shape[0]):
         audio = audio_combos[i]
         label = labels[i]
+        samples = []
+        for j in range(label.shape[0]):
+            if label[j] != 0:
+                samples.append(GLOBAL_PARAMS.index2sample[j])
         # list of pitches allowed in the orchestration
         pitches_to_include = get_harmonic_peaks(audio, GLOBAL_PARAMS.harmonic_threshold)
         # this part might be slow
@@ -407,7 +407,7 @@ def get_harmonic_peaks(audio, thresh):
     :return: a set of pitches (with octaves)
     """
     nyquist = GLOBAL_PARAMS.RATE / 2
-    n = next_power_of_2(audio.size)
+    n = next_power_of_2(audio.shape[0])
[…]

### 2021-04-08 csmc_mume_2020/paper/JCMS_2021/paper.tex
Add JCMS paper

--- 2020-11-09 paper/ISMIR_2020/paper.tex
@@
-% -----------------------------------------------
-% Template for ISMIR Papers
-% 2020 version, based on previous ISMIR templates
+%%%%%%%%%%%%%%%%%%%% file CSMC_MUME_LaTeX_Template.tex %%%%%%%%%%%%%%%%%%%%%
+%
+% This is the LaTeX source for the instructions to authors using
+% the LaTeX document class 'llncs.cls' for contributions to
+% the Journal of Creative Music Systems.
+% Copyright: http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
+%
+% It may be used as a template for your own input - copy it
+% to a new file with a new name and use it as the basis
+% for your article.
+%
+% NB: the document class 'llncs' has its own and detailed documentation, see
+% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Requirements :
-% * 6+n page length maximum
-% * 4MB maximum file size
-% * Copyright note must appear in the bottom left corner of first page
-% * Clearer statement about citing own work in anonymized submission
-% (see conference website for additional details)
-% -----------------------------------------------
+\documentclass[runningheads,a4paper]{llncs}
 
-\documentclass{article}
-\usepackage[T1]{fontenc} % add special characters (e.g., umlaute)
-\usepackage[utf8]{inputenc} % set utf-8 as default input encoding
-\usepackage{ismir,amsmath,cite,url}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\setcounter{tocdepth}{3}
 \usepackage{graphicx}
-\usepackage{color}
+\usepackage{caption}
+\usepackage{subcaption}
+\captionsetup{labelfont=bf, labelsep=period,compatibility=false}
+\usepackage{apacite}
+\newcommand{\keywords}[1]{\par\addvspace\baselineskip
+\noindent\keywordname\enspace\ignorespaces#1}
 
-% Optional: To use hyperref, uncomment the following.
-\usepackage[bookmarks=false,hidelinks]{hyperref}
-% Mind the bookmarks=false option; bookmarks are incompatible with ismir.sty.
-
-\usepackage{lineno}
-%\linenumbers
-\newcommand{\carmine}[1]{\textbf{\color{red} Carmine => #1}}
-\renewcommand{\baselinestretch}{1}
-
-% Title.
-% ------
-\title{Neural models for target-based computer-assisted musical orchestration: a preliminary study}
-
-% Note: Please do NOT use \thanks or a \footnote in any of the author markup
-
-% Single address
-% To use with only one author or several with the same address
-% ---------------
-%\oneauthor
-% {Names should be omitted for double-blind reviewing}
-% {Affiliations should be omitted for double-blind reviewing}
-
-% Two addresses
-% --------------
-%\twoauthors
-%  {First author} {School \\ Department}
-%  {Second author} {Company \\ Address}
-
-%% To make customize author list in Creative Common license, uncomment and customize the next line
-%  \def\authorname{First Author, Second Author}
-
-% Three addresses
-% --------------
-%\threeauthors
-%  {First Author} {Affiliation1 \\ {\tt author1@ismir.edu}}
-%  {Second Author} {\bf Retain these fake authors in\\\bf submission to preserve the formatting}
-%  {Third Author} {Affiliation3 \\ {\tt author3@ismir.edu}}
-
-%% To make customize author list in Creative Common license, uncomment and customize the next line
-%  \def\authorname{First Author, Second Author, Third Author}
-
-% Four or more addresses
-% OR alternative format for large number of co-authors
-% ------------
-\multauthor
-{First author$^1$ \hspace{1cm} Second author$^1$ \hspace{1cm} Third author$^2$} { \bfseries{Fourth author$^3$ \hspace{1cm} Fifth author$^2$}\\
-  $^1$ Department of Computer Science, University , Country\\
-$^2$ International Laboratories, City, Country\\
-$^3$  Company, Address\\
-{\tt\small CorrespondenceAuthor@ismir.edu, PossibleOtherAuthor@ismir.edu}
-}
-\def\authorname{First author, Second author, Third author, Fourth author, Fifth author}
-
-\sloppy % please retain sloppy command for improved formatting
+\renewcommand{\baselinestretch}{0.96}
+\raggedbottom
+\pagestyle{headings}
 
@@
 
+\mainmatter  % start of an individual contribution
+
+% first the title is needed
+\title{change}
+
+% a short form should be given in case it is too long for the running head
+\titlerunning{change}
+
+% the name(s) of the author(s) follow(s) next
 %
+% NB: Chinese authors should write their first names(s) in front of
+% their surnames. This ensures that the names appear correctly in
+% the running heads and the author index.
+%
+\author{Carmine Emanuele Cella\inst{1}\and Luke Dzwonczyk\inst{1}\and Alejandro Saldarriaga-Fuertes\inst{1}\and Hongfu Liu\inst{1}\and H\'el\`ene-Camille Crayencour\inst{2}}
+%
+% if the names of the authors are too long for the running head, please use the format: AuthorA et al.
+\authorrunning{Carmine Emanuele Cella et al.}
+
+% the affiliations are given next; don't give your e-mail address
+% unless you accept that it will be published
+\institute{CNMAT - University of California, Berkeley \and Centrale Sup\'elec, L2S, Univ. Paris-Saclay, CNRS \\ \email{carmine.cella@berkeley.edu, dz.luke@berkeley.edu }}
+
+%
+% NB: a more complex sample for affiliations and the mapping to the
+% corresponding authors can be found in the file "llncs.dem"
+% (search for the string "\mainmatter" where a contribution starts).
+% "llncs.dem" accompanies the document class "llncs.cls".
+%
+
 \maketitle
-%
+
 \begin{abstract}
-In this paper we will do a preliminary exploration on how neural networks can be used for the task of target-based computer-assisted musical orchestration.
+In this paper we will perform a preliminary exploration on how neural networks can be used for the task of target-based computer-assisted musical orchestration.
 We will show how it is possible to model this  musical problem as a classification task and we will propose two deep learning models.
@@
 Finally, we will highlight benefits and problems of neural approaches for assisted orchestration and we will propose possible future steps.
+
+\keywords{computer-assisted orchestration, CNN, LSTM, ResNet, Orchidea}
+
 \end{abstract}
+
 %
@@
 
-The development of computational tools to assist and inspire the musical composition process constitutes an import research area known as \emph{Computer-Assisted Composition (CAC)} \cite{FerVic2013, Ari2005}.
-Within CAC, target-based computer-assisted orchestration is an important example of how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.
+The development of computational tools to assist and inspire the musical composition process constitutes an important research area known as \emph{Computer-Assisted Composition (CAC)} \cite{FerVic2013, Ari2005}.
+Within CAC, target-based computer-assisted orchestration is a compelling case of how machine learning can be used for {enhancing} and {assisting} music creativity \cite{Maresz2003}.
 
-Target-based computer-assisted orchestration can be thought of as the process of searching for the best combinations of sounds in a database of instrumental samples, to match a target sound under a specific similarity metric and specific sets of constraints.
+Target-based computer-assisted orchestration takes a target sound as an input and attempts to find instrumental samples that best match the target given a specific similarity metric and a set of constraints.
 A solution to this problem is a set of orchestral scores that represent the mixtures of audio samples in the database, ranked by similarity with the target sound.
@@
 \begin{figure}
-\includegraphics[scale=0.17]{figs/method.png}
+	\centering
+	\includegraphics[scale=0.20]{../ISMIR_2020/figs/method.png}
 	\caption{An overview of the proposed method for assisted orchestration with neural models.
@@
 Thus, when an unknown target sound is given as input, the network will identify which are the best instruments to match the target sound, and it will be able to  deconstruct a complex mixture of timbres into individual instrument notes.
-
-%\footnote{The dynamics is retrived using the probability distributions generated as output by the network; more details on this will be given in the next sections.}.
-
 This method is motivated by the good results obtained in previous research on musical instruments identification \cite{Benetos07, Kitahara05} and the more recent use of deep neural networks for musical classification \cite{lostanlen16, Bian19}.
@@
 
+%In this paper we perform preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}.
+We chose to use a CNN because of its success in audio classification \cite{Hershey17} and we decided to include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.
+The codebase for this paper can be found at: \url{https://github.com/dzluke/DeepOrchestration}.
+
 \section{Neural models}
@@
 \begin{figure}
-\begin{center}
-\includegraphics[scale=0.22]{figs/CNN_LSTM_diagram.png}
+	\centering
+	\includegraphics[scale=0.22]{../ISMIR_2020/figs/CNN_LSTM_diagram.png}
 	\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
-\end{center}
 \end{figure}
 
-\subsection{From orchestration to classification}
+\subsection{From Orchestration to Classification}
 
@@
 \end{enumerate}
-
-\begin{figure}
-\center{
-\includegraphics[scale=0.4]{figs/CNN_vs_ResNet.png}}
-\caption{Best overall accuracy for CNN with LSTM and ResNet depending on the number of instruments in the combinations used for training. \label{cnn_vs_resnet}}
-\end{figure}
 
@@
 
+%A complete orchestration solution, however, would normally be made by triples of \textit{instrument-pitch-dynamics}.
+It is difficult to frame this problem as classification, since we would need to have a very high number of classes.
+Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in \ref{sec:dataset}), each class would be represented by a single sample.
+For this reasons, our models were not trained to determine \textit{instrument-pitch-dynamics} triples but instead \textit{instrument-pitch} pairs.
+
 \subsection{Dataset}
@@
 
+For a given number of instruments $N$, each input to our model is a combination of $N$ TinySOL samples chosen among an orchestra of 10 instruments.
+The data is generated by selecting $N$ random TinySOL samples, leading to a variety of instruments, pitches, and dynamics; we did not allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures.
+Then the chosen samples are combined to be played simultaneously and normalized by the number of instruments.
 The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly 4 seconds long.
+The Mel spectrogram of the mixture is computed using an FFT hop length of 2048 samples (the window of each FFT is $46$ms wide) and 128 Mel bins.
+Therefore, the Mel features fed to the model are matrices of size $128\times 345$.
 
-The choice of using the Mel spectrogram and MFCCs as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an appropriate representation of sound and musical signals.
+The choice of using the Mel spectrogram as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an appropriate representation of sound and musical signals.
 The Mel spectrogram is generated using an FFT hop length of 2048 samples (the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128.
@@
 The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed.
+Such setup proved to be successful in \cite{Salamon17}.
 
-%\begin{table}
-%\begin{center}
-%\begin{tabular}{|c|c|c|}
-%\hline
-%\textbf{Instrument} & \textbf{Abbreviation} & \textbf{Range} \\
-%\hline
-%Violin & Vn & G3-E7\\
-%Cello & Vc & C2-C6\\
-%Viola & Va & C3-C7\\
-%Trumpet C & TpC & F\#3-C6\\
-%Trombone & Tbn & A\#0;A\#1-C5\\
-%Oboe & Ob & A\#3-A6\\
-%Horn & Hn & G1-F5\\
-%Flute & Fl & B3-D7\\
-%Clarinet Bb & ClBb & D3-G6\\
-%Contrabass & Cb & E1-C5\\
-%Bass Tuba & BTb & F\#1-F4\\
-%Bassoon & Bn & A\#1-D\#5\\
-%\hline
-%\end{tabular}
-%\caption{Table showing the pitch ranges present in the TinySOL database.\label{tab:ranges}}
-%\end{center}
-%\end{table}
+\begin{table}
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+\textbf{Instrument} & \textbf{Abbreviation} & \textbf{Range} \\
+\hline
+Violin & Vn & G3-E7\\
+Cello & Vc & C2-C6\\
+Viola & Va & C3-C7\\
+Trumpet C & TpC & F\#3-C6\\
+Trombone & Tbn & A\#0;A\#1-C5\\
+Oboe & Ob & A\#3-A6\\
+Horn & Hn & G1-F5\\
+Flute & Fl & B3-D7\\
+Clarinet Bb & ClBb & D3-G6\\
+Contrabass & Cb & E1-C5\\
+Bass Tuba & BTb & F\#1-F4\\
+Bassoon & Bn & A\#1-D\#5\\
+\hline
+\end{tabular}
+\caption{Table showing the pitch ranges present in the TinySOL database.\label{tab:ranges}}
+\end{center}
+\end{table}
 
-\subsection{Data augmentation}
+\subsection{Data Augmentation}
 
@@
 
-The latter was performed on the feature matrix itself of input samples, the Mel spectrogram.
+Partial feature dropout was performed on the feature matrix itself of input samples, the Mel spectrogram.
 We chose random columns and rows of the matrix to zero out.
@@
 
+%We started with a simplification of our problem.
+We performed experiments in which two instruments were selected, and the classifier attempted to identify instrument and pitch \emph{class} of the samples.
+For two instruments, SVM had an accuracy of 39.9\% and RF had an accuracy of 17.5\%.
+As the number of instruments used in combination increased, the accuracy dropped sharply.
+With three instruments, SVM accuracy was 11.1\%, with four instruments it was 2.7\%.
+Clearly, these classifiers were not going to be able to achieve meaningful results with the full setting of our problem.
+
 \begin{table}
-  \begin{center}
+  \centering
+  \caption{Comparison of accuracies between SVM and RF.
+Each data point is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment.
+For the samples drawn from one of the two instruments, the pitch class of that sample is identified.
+For a sample not from one of the two instruments, the classifier simply attempted to identify that a sample from one of the non-specified instruments was present.
+In this setting, there are 25 classes: 24 from the 12 pitch classes from 2 instruments, and 1 class that specified whether an instrument that is not one of the two specified is present.
+\label{tab:baselines}}
     \begin{tabular}{|c|c|c|c|}
@@
     \end{tabular}
-  \end{center}
-  \caption{Comparison of accuracies between SVM and RF.
-Each data point is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment.
-For the samples drawn from one of the two instruments, the pitch class of that sample is identified.
-For a sample not from one of the two instruments, the classifier simply attempted to identify that a sample from one of the non-specified instruments was present.
-In this setting, there are 25 classes: 24 from the 12 pitch classes from 2 instruments, and 1 class that specified whether an instrument that is not one of the two specified is present.
-\label{tab:baselines}}
 \end{table}
 
-In the rest of this section, we will show the classification results of the two neural models.
+%In the rest of this section, we will show the classification results of the two neural models.
 
@@
 
-The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit.
-
-%CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey17}.
+The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit, whose structure is inspired by the success in \cite{Salamon17}.
+CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey17}.
 LSTM units provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.
@@
 
-Following the first three convolutional layers, there is an LSTM layer with 32 outputs.
+Following the first three convolutional layers, there is an LSTM layer which outputs 32 matrices.
 After the LSTM layer, there is a final convolutional layer yielding a tensor of dimensions $32\times 8 \times 21$.
-We flatten the outputs and feed them into fully a connected layer with Dropout, then another another fully connected layer.
+We flatten the outputs and feed them into a fully connected layer with Dropout, then another another fully connected layer.
 Finally, the sigmoid function is applied to the final layer.
-
-Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class. The model architecture is shown in Figure \ref{cnn_lstm}.
+Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class.
+%The model architecture is shown in Figure \ref{cnn_lstm}.
 
@@
 
-The second and more deep model that we trained as classifier was the well known deep residual network \emph{ResNet}.
+The second and deeper model that we trained as classifier was the well known deep residual network \emph{ResNet} \cite{He15}.
 Specifically, we used 18-layer ResNet, which allows information to pass directly from the input to the final layer of each block.
@@
 \begin{center}
-\includegraphics[scale=0.4]{figs/Acc_vs_N_CNN.png}
+\includegraphics[scale=0.4]{../ISMIR_2020/figs/Acc_vs_N_CNN.png}
 \caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM depending on the number of instruments in the combinations. \label{best_acc_cnn}}
 \end{center}
+\end{figure}
+
+\begin{figure}
+  \center{
+  \includegraphics[scale=0.4]{../ISMIR_2020/figs/CNN_vs_ResNet.png}}
+  \caption{Best overall accuracy for CNN with LSTM (50 epochs, 200k samples per epoch) and ResNet (20 epochs, 400k samples per epoch) depending on the number of instruments in the combinations used for training.
+\label{cnn_vs_resnet}}
 \end{figure}
@@
 and $I_N(\hat{X})$ is the set of indices containing the $N$ first maximums of the vector $\hat{X}$.
-
 More specifically, the function $M_N(\hat{X})$ takes as an input a vector of probabilities and outputs a vector where only the positions of the $N$ first maxima are set to 1.
@@
 Thus, the function $f$ simply outputs the proportion of the estimated orchestration that matches the expected one.
+For this reason, the evaluation was done by comparing the proportion of estimated orchestration samples, chosen among the samples that output the highest probability, that matched the expected orchestration.
 
 Different experiments were made by varying the number $N$ of samples in each mixture.
-We used an orchestra of 10 instruments containing Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet in C and Clarinet in Bb.
+We used an orchestra of 10 instruments: French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet in C and Clarinet in Bb.
+Then, for both CNN with LSTM and ResNet, we computed the accuracy on the test sets across epochs, and kept the best one as an indicator of the performance of the models.
 
-Then, for both CNN with LSTM and ResNet, we computed the accuracy on the test sets across epochs, and kept the best one as an indicator of the performance of the models. 
 The CNN was trained on 200,000 generated samples over 50 epochs, and ResNet was trained on 400,000 generated samples over 20 epochs.
 Fig.
-\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of TinySOL samples $N$ per mixture.\\
-
+\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$.
 ResNet outperforms the CNN regardless of the number of samples used in the combination.
@@
 \ref{best_acc_cnn} and Fig.
-\ref{best_acc_resnet} show the maximum accuracy computed using the function $f$ from Eqn.
-\eqref{eval}.
+\ref{best_acc_resnet} show the maximum test accuracy for each model.
 For ResNet, the variance in accuracy is much smaller until $N$ reaches 5, at which point it becomes similar to the CNN.
@@
 \begin{figure}
-\begin{center}
-\includegraphics[scale=0.40]{figs/Acc_vs_N_ResNet.png}
-\caption{Best overall accuracy and for each instrument obtained by the ResNet depending on the number of instruments in the combinations. \label{best_acc_resnet}}
-\end{center}
+  \centering
+  \begin{subfigure}{.52\textwidth}
+    \centering
+    \includegraphics[width=0.9\linewidth]{../ISMIR_2020/figs/Acc_vs_N_CNN.png}
+    \caption{CNN with LSTM}
+    \label{best_acc_cnn}
+  \end{subfigure}%
+  \begin{subfigure}{.52\textwidth}
+    \centering
+    \includegraphics[width=0.9\linewidth]{../ISMIR_2020/figs/Acc_vs_N_ResNet.png}
+    \caption{ResNet}
+    \label{best_acc_resnet}
+  \end{subfigure}
+  \caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM and the ResNet depending on the number of instruments in the combinations}
   \end{figure}
 
-\section{Orchestration experiments}
+\section{Orchestration Experiments}
 \label{sec:orchestration}
@@
 After training the neural models for classification, we finally tested them for the task of target-based computer-assisted orchestration.
-For this purpose, we decided to use the models trained on combinations of 10 instruments, namely French Horn, Oboe, Violin, Viola, Cello, Flute, Trombone, Bassoon, Trumpet, and Clarinet.
-Each class, in this model, represents an instrument-pitch pair.
 
@@
 In order to test our models for orchestration, we used 15 targets from the Orchidea distribution.
-These targets represent a variety of signal types but are mostly static, in the sense that they do not change sensibly over time.
-
+These targets represent a variety of signal types but are mostly static, in the sense that the harmonic average across time is a good representation of the sample.
 Some of the targets were made of instrumental samples: a sum of oboe and bassoon notes, single bassoon and bass clarinet notes, two multiphonics of bassoon and two symphonic chords.
@@
 
-Although the data we trained the model on was fixed to four seconds of audio, the target samples can be of arbitrary length since we adapt the representation to the size of the training data by changing the hop length during the computation of the Mel spectrogram.
+In order to test our models for orchestration, we used 15 targets from the Orchidea distribution.
+These targets represent a variety of signal types but are mostly static, in the sense that they do not change sensibly over time.
+Some of the targets were made of instrumental samples and chords, others are bells or gongs, and some do not feature any musical instruments
 
 \begin{table*}
-  \begin{center}
-    \label{orch_eval}
-    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c}
+	\centering    
+  \resizebox{\columnwidth}{!}{\begin{tabular}{|c|c|c|c|c|c|c|c|c|c}
      \hline
@@
       \hline
-    \end{tabular}
-  \end{center}
+    \end{tabular}}
     \caption{Quantitative comparison of orchestrations as ratios to Orchidea.
 Eqn.
-\eqref{distance} was used with $\lambda_1 = 0.5$ and $\lambda_2 = 10$ to compute distances between orchestrations and targets.
-What is shown is the ratio between the distance calculated and Orchidea's distance for the same target.
-A value less than 1 means that the model performed worse (i.e.
-had a larger distance), and a value greater than 1 means the model performed better than Orchidea.
-The last column shows the ratio of the average distances for the model across all targets.}
+\ref{distance} was used to compute distances between orchestrations and targets.
+What is shown is the ratio between the distance of Orchidea's solution to the target and our solution's distance to the same target.
+A value less than 1 means that our model performed worse (i.e.
+had a larger distance), and a value greater than 1 means our model performed better than Orchidea.
+The last column shows the ratio of the average distances for the model across all targets.}\label{orch_eval}
 \end{table*}
@@
 For quantitative evaluation, we used the distance metric defined in Eqn.
-\eqref{distance} to calculate differences in timbre between targets and solutions.
+\ref{distance} to calculate differences in timbre between targets and solutions.
 This metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization.
-The equation takes in the full FFT of the target $x$ and full FFT of the solution $\hat{x}$.
-Then for each bin $k$ of the FFT, it calculates the absolute difference between the values.
+The equation takes in the full amplitude spectrum of the target $x$ and full amplitude spectrum of the solution $\tilde{x}$.
+Then for each bin $k$ of the amplitude spectrum, it calculates the absolute difference between the values.
 The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.
@@
 \end{equation}
-where 
-$
-\delta_{k1} = 
-\begin{cases}
-1, \text{if   } x_k \ge \tilde{x}_k \\
-0, \text{otherwise}
-\end{cases} 
-$, $
-\delta_{k2} = 
-\begin{cases}
-1, \text{if   } x_k < \tilde{x}_k \\
-0, \text{otherwise}
-\end{cases}.
-$\\
+where $\delta_{k1} = 1 \text{  if  } x_k \ge \tilde{x}_k, 0 \text{  otherwise}$; and $\delta_{k2} = 1 \text{  if  } x_k < \tilde{x}_k, 0 \text{  otherwise}$.
 
+A comparison of distances between our solutions and targets and Orchidea's solutions and targets is in Table \ref{orch_eval}.
 While our model is not able to outperform Orchidea, it shows consistent results.
@@
 The first natural extension would be to support sparsity in our models.
-Our current models, indeed, orchestrate all targets using a constant number of instruments and it is not able to drop specific instruments from the solution.
+Our current models orchestrate all targets using a constant number of instruments and are not able to drop specific instruments from the solution.
 This does not take into account the density of different targets.
@@
 In \cite{Gillick19} the authors propose to use LSTM-based models to predict the embedding features for the combinations used during the optimisation process in assisted orchestration.
-We believe that by combining their prediction model with our classification models could generate more faithful representations and improve the overall quality of generated orchestrations.
+We believe that by combining their prediction model with our classification models we could generate more faithful representations and improve the overall quality of generated orchestrations.
 
-%\subsection{Interpreting the Latent Space}
-%
-%After training the CNN with LSTM for $N=10$ samples per combination, it is possible to visualize the filters applied to the input layer after layer.
-Figs.
-\ref{latent0}-\ref{latent5} show the successive outputs of each layer of the convolutional network.
-\carmine{I put here all the intermediary features of the CNN with LSTM.
-I don't find any particular conclusion other than saying that LSTM seems to focus more on vertical components rather than horizontal, ie time over frequency}
-%
-%\begin{figure}
-%\includegraphics[scale=0.5]{figs/latent_space_layer0.png}
-%\caption{Initial feature matrix used as input of the model \label{latent0}}
-%\end{figure}
-%
-%\begin{figure}
-%\includegraphics[scale=0.5]{figs/latent_space_layer1.png}
-%\caption{Intermediary features after the first convolutional layer of the CNN with LSTM. \label{latent1}}
-%\end{figure}
-%
-%\begin{figure}
-%\includegraphics[scale=0.5]{figs/latent_space_layer2.png}
-%\caption{Intermediary features after the second convolutional layer of the CNN with LSTM. \label{latent2}}
-%\end{figure}
-%
-%\begin{figure}
-%\includegraphics[scale=0.5]{figs/latent_space_layer3.png}
-%\caption{Intermediary features after the third convolutional layer of the CNN with LSTM. \label{latent3}}
-%\end{figure}
-%
-%\begin{figure}
-%\includegraphics[scale=0.5]{figs/latent_space_layer4.png}
-%\caption{Intermediary features after the LSTM layer. \label{latent4}}
-%\end{figure}
-%
-%\begin{figure}
-%\includegraphics[scale=0.5]{figs/latent_space_layer5.png}
-%\caption{Intermediary features after final convolutional layer of the CNN with LSTM. \label{latent5}}
-%\end{figure}
-
-\iffalse
-\begin{figure}
-\includegraphics[scale=0.6]{figs/CNN10.png}
-\end{figure}
-\fi
-
-% For bibtex users:
-%\bibliography{paper}
-
-\begin{thebibliography}{99}
-
-\bibitem{FerVic2013} Jose David Fern\'andez and Francisco Vico.
-{AI Methods in Algorithmic Composition: A Comprehensive Survey}.
-Journal of Artificial Intelligence Research, vol.48, pp.
-513--582, 2013.
-
-\bibitem{Ari2005} Christopher Ariza, {Navigating the landscape of computer aided algorithmic composition systems: A definition, seven descriptors, and a lexicon of systems and research}.
-In Proceedings of the International Computer Music Conference, 2005.
-
-\bibitem{Maresz2003}
-Yan Maresz, {On Computer-Assisted Orchestration}. Contemporary Music Review, vol. 32, n. 1, pg. 99-109, Routledge, 2013.
-
-\bibitem{Carpentier2010}
-Gr\'egoire Carpentier, Damien Tardieu, Jonathan Harvey, G\'erard Assayag and Emmanuel Saint-James.
-{Predicting Timbre Features of Instrument Sound Combinations: Application to Automatic Orchestration}, Journal of New Music Research, vol.
-39, n.
-1, pg.
-47-61, Routlegde, 2010.
-
-\bibitem{Benetos07}
-Emmanouil Benetos, Margarita Kotti, and Constantine Kotropoulos.
-\newblock Large scale musical instrument identification.
-\newblock Sound and Music Computing Conference, Greece, 2007.
-
-\bibitem{Bhardwaj17}
-Siddharth Bhardwaj.
-\newblock Audio data augmentation with respect to musical instrument
-  recognition.
-\newblock Master's thesis, 2017.
-
-\bibitem{Bian19}
-Wenhao Bian, Jie Wang, Bojin Zhuang, Shaojun~Wang Jiankui~Yang, and Jing Xiao.
-\newblock Audio-based music classification with densenet and data augmentation.
-\newblock arXiv preprint, https://arxiv.org/abs/1906.11620v1, 2019.
-
-\bibitem{Cella2020}
-Carmine-Emanuele Cella.
-\newblock Orchidea: a comprehensive framework for target-based assisted orchestration.
-\newblock In preparation; available on request from the author, 2020.
-
-\bibitem{Cella2020b}
-Carmine-Emanuele Cella, Daniele Ghisi, Vincent Lostanlen, Fabi\'en Levy, Yan Maresz and Joshua Fineberg.
-\newblock OrchideaSOL: a dataset of extended instrumental techniques for computer-aided orchestration.
-\newblock ICMC (under review), 2020.
-
-\bibitem{Chetry05}
-Nicolas Chetry, Mike Davies, and Mark Sandler.
-\newblock Musical instrument identification using LSF and K-means.
-\newblock Audio Engineering Society Convention 118, 2005.
-
-\bibitem{He15}
-Kaiming He, Xiangyu Zhang, Shaoqing Ren, and Jian Sun.
-\newblock Deep residual learning for image recognition.
-\newblock arXiv preprint, https://arxiv.org/abs/1512.03385, 2015.
-
-\bibitem{Hershey17}
-Shawn Hershey, Sourish Chaudhuri, Daniel P.~W. Ellis, Jort~F. Gemmeke, Aren
-  Jansen, R.~Channing Moore, Manoj Plakal, Devin Platt, Rif~A. Saurous, Bryan
-  Seybold, Malcolm Slaney, Ron~J. Weiss, and Kevin Wilson.
-\newblock CNN architectures for large-scale audio classification.
-\newblock IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP), 2017.
-
-\bibitem{Hochreiter97}
-Sepp Hochreiter and J\"{u}rgen Schmidhuber.
-\newblock Long short-term memory.
-\newblock {\em Neural Comput.}, 9(8):1735–1780, November 1997.
-
-\bibitem{Kitahara05}
-Tetsuro Kitahara, Masataka Goto, and Hiroshi~G. Okuno.
-\newblock Pitch-dependent identification of musical instrument sounds.
-\newblock Applied Intelligence 23, 267–275, Springer, 2005.
-
-\bibitem{lostanlen16}
-Vincent Lostanlen and Carmine-Emanuele Cella.
-\newblock Deep convolutional networks on the pitch spiral for musical
-  instrument recognition. ISMIR, 2016.
-
-\bibitem{mcfee15}
-Brian McFee, Colin Raffel, Dawen Liang, Daniel~P.W. Ellis, Matt McVicar, Eric
-  Battenberg, and Oriol Nieto.
-\newblock librosa: Audio and music signal analysis in python.
-\newblock {P}roceedings of the 14th {P}ython in {S}cience {C}onference, 2015.
-
-\bibitem{scikit-learn}
-Fabian Pedregosa, Ga\"el Varoquaux, Alexandre
-Gramfort, Vincent Michel, Bertrand Thirion, Olivier
-Grisel, Mathieu Blondel, Peter Prettenhofer, Ron
-Weiss, Vincent Dubourg, Jake Vanderplas, Alexandre
-Passos, David Cournapeau, Matthieu Brucher,
-Matthieu Perrot, \'Edouard Duchesnay.
-\newblock Scikit-learn: Machine learning in {P}ython.
-\newblock {\em Journal of Machine Learning Research}, 12:2825--2830, 2011.
-
-\bibitem{Salamon17}
-Justin Salamon and Juan~Pablo Bello.
-\newblock Deep convolutional neural networks and data augmentation for
-  environmental sound classification.
-\newblock {\em IEEE Signal Processing Letters}, 24(3):279–283, Mar 2017.
-
-\bibitem{McKinney2003}
-Martin F. McKinney and Jeroen Breebaart.
-\newblock Features for Audio and Music Classification.
-\newblock ISMIR, 2003.
-
-\bibitem{Cella18}
-Carmine-Emanuele Cella and Philippe Esling, {Open-source modular toolbox for computer-aided orchestration}, Proc. of the Timbre conference, Montreal, Canada, 2018.
-
-\bibitem{Gillick19}
-Jon Gillick, Carmine-Emanuele Cella and David Bamman. {Estimating Unobserved Audio Features for Target-Based Orchestration}, ISMIR, 2019.
-
-\end{thebibliography}
+\bibliographystyle{apacite}
+\bibliography{references}
 

### 2021-04-08 csmc_mume_2020/paper/JCMS_2021/paper.tex
First round of edits, adjusting figures

@@
 % first the title is needed
-\title{change}
+\title{A Study on Neural Models for Target-Based Computer-Assisted Musical Orchestration}
 
 % a short form should be given in case it is too long for the running head
-\titlerunning{change}
+\titlerunning{Neural Models for Target-Based Computer-Assisted Musical Orchestration}
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[scale=0.20]{../ISMIR_2020/figs/method.png}
-	\caption{An overview of the proposed method for assisted orchestration with neural models.
-Instruments and pitches are determined as peaks of the output probability distribution, while the dynamics are computed by quantizing the probabilities.
-\label{fig:method}}
-\end{figure}
-
 The approach studied in \cite{Carpentier2010} consists in finding a good orchestration for any given sound by searching combinations of sounds from a database with a multi-objective optimization heuristics and a constraint solver that are jointly optimized.
@@
 
-In this paper we do preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit in the middle and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}.
-We chose to use a CNN because of its success in audio classification \cite{Hershey17} and we decided to include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.
+We perform preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit in the middle and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}.
+We choose to use a CNN because of its success in audio classification \cite{Hershey17} and include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.
 
-After training these models on varying combinations of 10 instruments, we used them for orchestration.
-More specifically, we orchestrated 15 target sounds by feeding them to the networks and synthesizing a solution from the most likely samples output from the system.
-We then compared our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means.
-For quantitative comparison, we computed a specific distance metric on the FFTs of the target and solution, comparing the distance between Orchidea's solution and the target and our model's solution and the target.
+After training these models on varying combinations of 10 instruments, we use perform orchestration by feeding target sounds to the networks and synthesizing a solution from the most likely samples output from the system.
+We then compare our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means.
+For quantitative comparison, we compute a specific distance metric on the amplitude spectrums of the target and solution, comparing the distance between Orchidea's solution to the target and our model's solution to the target.
 
-Next sections will be as follows: section \ref{sec:models} will describe the dataset we used for experiments and the models we trained as classifiers, showing the classification results.
+The next sections will be as follows: section \ref{sec:models} will describe the dataset we used for experiments and the models we trained as classifiers, showing the classification results.
 Sections \ref{sec:orchestration} and \ref{sec:conclusions} will present, respectively, the results of orchestration experiments using the trained models and the conclusions of our preliminary study on neural approaches for target-based computer-assisted orchestration.
 
-The codebase for this paper and some audio examples can be found at: \url{https://github.com/dzluke/DeepOrchestration}.
-
-%In this paper we perform preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}.
-We chose to use a CNN because of its success in audio classification \cite{Hershey17} and we decided to include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.
-The codebase for this paper can be found at: \url{https://github.com/dzluke/DeepOrchestration}.
+This paper is an extended version of a paper by the same title in \textit{The 2020 Joint Conference on AI Music Creativity} \cite{Cella2020c}.
+This version includes additional figures and extended information on our baseline testing and orchestration process.
+The codebase for this paper can be found at \url{https://github.com/dzluke/DeepOrchestration} and you can listen to audio examples at \url{https://dzluke.github.io/DeepOrchestration/}.
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[scale=0.22]{../ISMIR_2020/figs/CNN_LSTM_diagram.png}
-	\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
-\end{figure}
-
 \subsection{From Orchestration to Classification}
@@
 In this paper, we model assisted orchestration as a classification problem. The general methodology is as follows:
+
 \begin{enumerate}
@@
 \item we then pick the best classifier and we feed into it an unknown sound to be classified;
-\item since the output of the classifier will be in the form of the probability that specific instruments are present in the sound, we use this information to synthesize an orchestration for the target sound;
+\item since the output of the classifier will be in the form of the probability that specific instruments-note pairs are present in the sound, we use this information to synthesize an orchestration for the target sound;
 \item finally, we evaluate the generated orchestration against state-of-the-art systems for computer-assisted orchestrations.
@@
 In other words, the classifiers learn how to take a complex combination of pitches and timbres and deconstruct it into its original parts.
+
+In order to have a baseline to compare our classification results and to get a sense on the complexity of the problem, we used various parametric classifiers along with the neural models.
+
+As described in Sec.
+\ref{sec:baseline}, three baseline classifiers have been trained on simplified versions of our problem.
+They have been trained to either identifying only the instruments present, or the instruments and pitch classes present.
 
@@
 
-In order to have a baseline to compare our classification results and to get a sense on the complexity of the problem, we used various parametric classifiers along with the neural models.
-
-Three baseline classifiers have been trained to recognize either the instrument only for a total of 12 classes (the instruments in TinySOL) or the pair instrument-pitch \emph{class} (a pitch class is the pitch of a note without the corresponding octave, e.g.
-C, D\#, G, etc.) for a total of 25 classes (we will classifiy combinations of 2 instruments for 12 pitch classes each, plus one extra \emph{trash} class, see section \ref{sec:baseline}).
-
-Two neural classifiers, instead, have been trained on the more complex problem of recognizing instrument-pitch pairs, for a total of 424 classes (i.e.
-the neural classifiers would classify instruments and pitches with respective octaves, hence 12 instruments for 12 pitch classes and three dynamics; note that not all the samples in TinySOL have all dynamics).
+Two neural classifiers have been trained on the more complex problem of recognizing instrument-pitch pairs, for a total of 424 classes (i.e.
+424 unique instrument-pitch pairs).
 We want to point out, however, that this is still a simplified problem compared to assisted orchestration.
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.20]{../ISMIR_2020/figs/method.png}
+	\caption{An overview of the proposed method for assisted orchestration with neural models.
+Instruments and pitches are determined as peaks of the output probability distribution, while the dynamics are computed by quantizing the probabilities.
+\label{fig:method}}
+\end{figure}
+
 \subsection{Dataset}
@@
 Each sample is one instrument playing a single note in the \emph{ordinario} playing style, with one of three dynamics: \textit{pp}, \textit{mf}, or \textit{ff} (for example \texttt{Flute-C4-pp} or \texttt{Clarinet-D5-mf}).
-
 The instruments and ranges over which they were recorded are summarized in table \ref{tab:ranges}.
-
-For a given number of instruments $N$, each input to our models is a combination of $N$ TinySOL samples chosen among an orchestra of 10 instruments. The data is generated in the following way:
-
-\begin{itemize}
-\item $N$ random TinySOL samples are selected, leading to a variety of instruments, pitches, and dynamics; we did not allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures;
-\item the chosen samples are combined to be played simultaneously, and the resulting sum is normalized by the number of instruments;
-\item the Mel frequency cepstral coefficients (MFCCs) and the Mel spectrogram of the mixture are computed to be used as the input features for the baseline classifiers and the neural classifiers respectively.
-\end{itemize}
 
@@
 The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly 4 seconds long.
-The Mel spectrogram of the mixture is computed using an FFT hop length of 2048 samples (the window of each FFT is $46$ms wide) and 128 Mel bins.
-Therefore, the Mel features fed to the model are matrices of size $128\times 345$.
+
+We then computed the Mel spectrogram of the combination to be the features input to our models.
+We generated the Mel spectrogram using an FFT hop length of 2048 samples (the window of each FFT was $46$ms long), and a number of Mel bins equal to 128.
+Therefore, the Mel features fed to the model were matrices of size $128\times 345$.
+We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}.
+The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture temporal changes in the sound, assuming that musically relevant changes do not occur in a time frame less than $10$ms.
 
 The choice of using the Mel spectrogram as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an appropriate representation of sound and musical signals.
-The Mel spectrogram is generated using an FFT hop length of 2048 samples (the window of each FFT was $46$ms wide), and a number of Mel bins equal to 128.
-Therefore, the Mel features fed to the model were matrices of size $128\times 345$.
-We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}.
-The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture changes in the dynamic, assuming that no change faster than $10$ms would be allowed.
 Such setup proved to be successful in \cite{Salamon17}.
@@
 	Viola & Va & C3-C7\\
-Trumpet C & TpC & F\#3-C6\\
+	Trumpet in C & TpC & F\#3-C6\\
 	Trombone & Tbn & A\#0;A\#1-C5\\
@@
 	Flute & Fl & B3-D7\\
-Clarinet Bb & ClBb & D3-G6\\
+	Clarinet in B$\flat$ & ClBb & D3-G6\\
 	Contrabass & Cb & E1-C5\\
@@
 
-In order to get a better sense on the complexity of the problem, we tested three baseline classifiers: support vector machine (SVM), random forest (RF), and K-nearest neighbours (KNN).
+In order to get a better sense on the complexity of the problem, we tested three baseline classifiers: support vector machine (SVM), random forest (RF), and K-nearest neighbors (KNN).
 We used the implementations provided in the \emph{scikit-learn }library for Python \cite{scikit-learn}.
@@
 Each sample was made by a combination of one to ten instruments and was four seconds in length.
+
 In this case, differently from the neural models, the features used are the MFCCs of the resulting combination; we chose to use MFCCs for this setting, instead of the Mel spectrogram, to have a lower number of features, that is more manageable by parametric classifiers.
@@
 The complete problem of classification of the pairs instrument-pitch has 424 total classes.
-This problem is very hard for parametric classifiers, so we initially switched to the simpler problem of classifying only the instruments and not the pitch.
+This problem is very difficult for parametric classifiers, so we initially switched to the simpler problem of classifying only the instruments and not the pitch.
 This had the benefit of both reducing the number of classes and increasing the number of samples per class.
@@
 Flute, Oboe, and Violin reached an accuracy of 11.1\%, and Bass Tuba, Trumpet, and Trombone was 0.5\%.
-As we increased the number of instruments whose pitch classes was being identified, the accuracy continued to drop.Indeed, for classifying the pitch class of four instruments (Oboe, French Horn, Violin, and Flute) the accuracy was 2.7\%.
+As we increased the number of instruments whose pitch classes was being identified, the accuracy continued to drop.
+For classifying the pitch class of four instruments (Oboe, French Horn, Violin, and Flute) the accuracy was 2.7\%.
 
@@
 However, the parametric classifiers were unable to achieve accurate results as the number of instruments increased.
-Therefore, we did not attempt, with parametric classifiers, the full setting of the problem in which individual pitches are classified for all instruments.
+Therefore, we did not attempt, with parametric classifiers, the full setting of the problem in which individual pitches are classified for all instruments, and instead turned to our neural models.
 
@@
 Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class.
-%The model architecture is shown in Figure \ref{cnn_lstm}.
+The model architecture is shown in Figure \ref{cnn_lstm}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{../ISMIR_2020/figs/CNN_LSTM_diagram.png}
+	\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
+\end{figure}
 
@@
 \begin{figure}
-\begin{center}
-\includegraphics[scale=0.4]{../ISMIR_2020/figs/Acc_vs_N_CNN.png}
-\caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM depending on the number of instruments in the combinations. \label{best_acc_cnn}}
-\end{center}
-\end{figure}
-
-\begin{figure}
   \center{
-  \includegraphics[scale=0.4]{../ISMIR_2020/figs/CNN_vs_ResNet.png}}
+  \includegraphics[width=0.8\textwidth]{../ISMIR_2020/figs/CNN_vs_ResNet.png}}
   \caption{Best overall accuracy for CNN with LSTM (50 epochs, 200k samples per epoch) and ResNet (20 epochs, 400k samples per epoch) depending on the number of instruments in the combinations used for training.
@@
   \end{subfigure}
-  \caption{Best overall accuracy and for each instrument obtained by the CNN with LSTM and the ResNet depending on the number of instruments in the combinations}
+  \caption{Plot showing per instrument accuracies for CNN with LSTM and ResNet depending on the number of samples present in the combinations}
   \end{figure}
@@
 Some of the targets were made of instrumental samples: a sum of oboe and bassoon notes, single bassoon and bass clarinet notes, two multiphonics of bassoon and two symphonic chords.
-The other targets there are: the recordings of two bells, of a car horn, of a gong, of a screaming voice, of a wind harp, of the attack of a brass instrument and the recording of a boat docking.
-
-In order to test our models for orchestration, we used 15 targets from the Orchidea distribution.
-These targets represent a variety of signal types but are mostly static, in the sense that they do not change sensibly over time.
-Some of the targets were made of instrumental samples and chords, others are bells or gongs, and some do not feature any musical instruments
+Other targets were recordings that were not instrumental: bell sounds, car horn, gong, screaming voice, wind harp, the attack of a brass instrument, and the recording of a boat docking.
 
@@
 
-In order to have a fairer comparison, we did not allow Orchidea to use any of the advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it forced to use all 10 instruments in each solution.
+In order to have a better comparison, we did not allow Orchidea to use any of the advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it forced to use all 10 instruments in each solution.
 This creates a more fair comparison, since our models are unable to create constrained or sparse solutions.
@@
 
-We calculated the distance between the target samples and our orchestrated solutions.
-We then orchestrated the same targets with Orchidea and calculated the distance for Orchidea's solutions.
-A comparison of these results is in Table \ref{orch_eval}.
-
 \begin{equation}\label{distance}
@@
 
-A comparison of distances between our solutions and targets and Orchidea's solutions and targets is in Table \ref{orch_eval}.
+We calculated the distance between the target samples and our orchestrated solutions.
+We then orchestrated the same targets with Orchidea and calculated the distance for Orchidea's solutions.
+A comparison of these results is in Table \ref{orch_eval}.
 While our model is not able to outperform Orchidea, it shows consistent results.

### 2021-04-08 csmc_mume_2020/paper/JCMS_2021/paper.tex
More edits to text mostly

@@
 This paper is an extended version of a paper by the same title in \textit{The 2020 Joint Conference on AI Music Creativity} \cite{Cella2020c}.
-This version includes additional figures and extended information on our baseline testing and orchestration process.
+This version includes additional figures and extended information on our baseline testing and orchestration process, among other additions.
 The codebase for this paper can be found at \url{https://github.com/dzluke/DeepOrchestration} and you can listen to audio examples at \url{https://dzluke.github.io/DeepOrchestration/}.
@@
 
-In other words, the classifiers learn how to take a complex combination of pitches and timbres and deconstruct it into its original parts.
-
-In order to have a baseline to compare our classification results and to get a sense on the complexity of the problem, we used various parametric classifiers along with the neural models.
-
-As described in Sec.
-\ref{sec:baseline}, three baseline classifiers have been trained on simplified versions of our problem.
-They have been trained to either identifying only the instruments present, or the instruments and pitch classes present.
+In order to have a baseline to compare our classification results and to get a sense of the complexity of the problem, we used various parametric classifiers along with the neural models.
+As described in section \ref{sec:baseline}, three baseline parametric classifiers have been trained on simplified versions of our problem.
+They have been trained to either identify only the instruments present, or the instruments and pitch classes present, two simplified versions of our problem.
 
@@
 It is difficult to frame this problem as classification, since we would need to have a very high number of classes.
-Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in \ref{sec:dataset}), each class would be represented by a single sample.
+Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in section \ref{sec:dataset}), each class would be represented by a single sample.
 For these reasons, our models have been trained on simplified versions of the original problem and we used different strategies to provide the remaining information needed to generate orchestrations.
 
-Two neural classifiers have been trained on the more complex problem of recognizing instrument-pitch pairs, for a total of 424 classes (i.e.
+Two neural classifiers have been trained on the problem of recognizing instrument-pitch pairs, which has a total of 424 classes (i.e.
 424 unique instrument-pitch pairs).
@@
 
-The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit, whose structure is inspired by the success in \cite{Salamon17}.
+The first deep model we trained as a classifier for musical instruments and pitches was a CNN with a LSTM unit, whose structure is inspired by the success in~\cite{Salamon17}.
 CNNs show good performance on classification problems for their ability to extract spatial features, and have shown success in audio classification \cite{Hershey17}.

### 2021-04-08 csmc_mume_2020/paper/JCMS_2021/paper.tex
Add Carmine's changes

@@
 % first the title is needed
-\title{A Study on Neural Models for Target-Based Computer-Assisted Musical Orchestration}
+\title{Neural Models for Target-Based Computer-Assisted Musical Orchestration: A Preliminary Study}
 
@@
 Finally, we will highlight benefits and problems of neural approaches for assisted orchestration and we will propose possible future steps.
+This paper is an extended version of the paper ``A Study on Neural Models for Target-Based Computer-Assisted Musical Orchestration" published in the proceedings of \textit{The 2020 Joint Conference on AI Music Creativity}.
 
@@
 
-This paper is an extended version of a paper by the same title in \textit{The 2020 Joint Conference on AI Music Creativity} \cite{Cella2020c}.
-This version includes additional figures and extended information on our baseline testing and orchestration process, among other additions.
-The codebase for this paper can be found at \url{https://github.com/dzluke/DeepOrchestration} and you can listen to audio examples at \url{https://dzluke.github.io/DeepOrchestration/}.
+\sloppypar{The codebase for this paper can be found at \url{https://github.com/dzluke/DeepOrchestration} and you can listen to audio examples at \url{https://dzluke.github.io/DeepOrchestration/}.}
 
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{../ISMIR_2020/figs/CNN_LSTM_diagram.png}
+	\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
+\end{figure}
+
 Our architecture is made of four convolutional layers and two fully connected layers.
@@
 The model architecture is shown in Figure \ref{cnn_lstm}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{../ISMIR_2020/figs/CNN_LSTM_diagram.png}
-	\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
-\end{figure}
 

### 2021-04-08 csmc_mume_2020/paper/JCMS_2021/paper.tex
Final edits before submission

@@
 
-After training these models on varying combinations of 10 instruments, we use perform orchestration by feeding target sounds to the networks and synthesizing a solution from the most likely samples output from the system.
+After training these models on varying combinations of 10 instruments, we perform orchestration by feeding target sounds to the networks and synthesizing a solution from the samples output from the system.
 We then compare our orchestrations to Orchidea's solutions for the same targets, through both qualitative and quantitative means.
@@
 
-A complete orchestration solution, however, would normally be made by triples of instrument-pitch-dynamics.
-It is difficult to frame this problem as classification, since we would need to have a very high number of classes.
-Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in section \ref{sec:dataset}), each class would be represented by a single sample.
-For these reasons, our models have been trained on simplified versions of the original problem and we used different strategies to provide the remaining information needed to generate orchestrations.
-
 Two neural classifiers have been trained on the problem of recognizing instrument-pitch pairs, which has a total of 424 classes (i.e.
@@
 We want to point out, however, that this is still a simplified problem compared to assisted orchestration.
-Even in the case of instrument-pitch pair classification, indeed, we would miss the dynamics of the note.
-To cope with this problem, in the orchestration phase, we quantize the probability functions generated by the models.
+A complete orchestration solution would normally be made by triples of instrument-pitch-dynamics.
+It is difficult to frame this problem as classification, since we would need to have a very high number of classes.
+Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in section \ref{sec:dataset}), each class would be represented by a single sample.
+For these reasons, our models have been trained on simplified versions of the original problem, and we employed a different strategy to determine dynamics.
 When a network is given a target sound as input, it will attempt to apply the same classification rules, outputting a vector of the probabilities of each sample being in the target sound.
-Since the classes only encode instrument-pitch pairs and since we also need the dynamics, we quantize the probability function generated by the network into three echelons coresponding, respectively, to the dynamic levels \emph{pianissimo}, \emph{mezzoforte} and \emph{fortissimo} (see section \ref{sec:orchestration}).
+Since the classes only encode instrument-pitch pairs but we also need the dynamics, we quantize the probability function generated by the network into three echelons coresponding, respectively, to the dynamic levels \emph{pianissimo}, \emph{mezzoforte} and \emph{fortissimo} (see section \ref{sec:orchestration}).
 By taking the instruments and pitches that have the highest probability and by using quantized probabilities to retrieve dynamics, a full orchestrated solution can be created.
 Figure \ref{fig:method} illustrates the described ideas.
-
-%A complete orchestration solution, however, would normally be made by triples of \textit{instrument-pitch-dynamics}.
-It is difficult to frame this problem as classification, since we would need to have a very high number of classes.
-Moreover, for the nature of the samples we use (a typical sample is in the form \texttt{Flute-C4-pp}, as described in \ref{sec:dataset}), each class would be represented by a single sample.
-For this reasons, our models were not trained to determine \textit{instrument-pitch-dynamics} triples but instead \textit{instrument-pitch} pairs.
 
@@
 We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}.
-The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture temporal changes in the sound, assuming that musically relevant changes do not occur in a time frame less than $10$ms.
+The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture temporal changes in the sound.
 
@@
 \begin{center}
+	\caption{Table showing the pitch ranges present in the TinySOL database.\label{tab:ranges}}
 	\begin{tabular}{|c|c|c|}
@@
 	\end{tabular}
-\caption{Table showing the pitch ranges present in the TinySOL database.\label{tab:ranges}}
 \end{center}
@@
   \caption{Comparison of accuracies between SVM and RF.
-Each data point is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment.
-For the samples drawn from one of the two instruments, the pitch class of that sample is identified.
-For a sample not from one of the two instruments, the classifier simply attempted to identify that a sample from one of the non-specified instruments was present.
-In this setting, there are 25 classes: 24 from the 12 pitch classes from 2 instruments, and 1 class that specified whether an instrument that is not one of the two specified is present.
-\label{tab:baselines}}
+Each input to the models is a combination of two TinySOL samples, where at least one of the samples is from one of the two instruments specified for that experiment.}\label{tab:baselines}
     \begin{tabular}{|c|c|c|c|}
@@
 
-%In the rest of this section, we will show the classification results of the two neural models.
-
 \subsection{CNN with LSTM}
@@
 LSTM units provide a way to learn long term dependencies in the data \cite{Hochreiter97}, which is relevant given the sequential nature of audio.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{../ISMIR_2020/figs/CNN_LSTM_diagram.png}
-	\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
-\end{figure}
 
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{../ISMIR_2020/figs/CNN_LSTM_diagram.png}
+	\caption{Diagram of the CNN with LSTM architecture. \label{cnn_lstm}}
+\end{figure}
+
 \subsection{ResNet}
@@
   \center{
-  \includegraphics[width=0.8\textwidth]{../ISMIR_2020/figs/CNN_vs_ResNet.png}}
+  \includegraphics[width=0.7\textwidth]{../ISMIR_2020/figs/CNN_vs_ResNet.png}}
   \caption{Best overall accuracy for CNN with LSTM (50 epochs, 200k samples per epoch) and ResNet (20 epochs, 400k samples per epoch) depending on the number of instruments in the combinations used for training.
@@
 Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.
-You can listen to the targets and orchestrated solutions from Orchidea, ResNet, and the CNN with LSTM at \url{https://dzluke.github.io/DeepOrchestration/}.
 

### 2021-05-24 source_separation/paper/paper.tex
Small changes from SMC reviews

@@
 	Source separation is an important addition to improve orchestration as it allows for the orchestration of more complex sounds in a way that makes sense in an orchestral context.
-Orchestral music often uses multiple layers, so by separating the layers of a target and individually orchestrating them, the resulting solution is improved.
+Orchestral music often uses multiple layers, so by separating the layers of a target and individually orchestrating them, the resulting orchestration may improve.
 
@@
 
-	\begin{figure}[t]
+	\begin{figure*}[t]
 		\centering
-			\includegraphics[width=\columnwidth]{figures/diagram.png}
+			\includegraphics[width=\textwidth]{figures/diagram.png}
 			\caption{Diagram comparing the process of orchestration with and without separation.
@@
 At the bottom, the separation process is performed, and the resulting sub-targets are individually orchestrated with subsets of the orchestra before being recombined to obtain the full orchestration.}\label{fig:full_diagram}
-	\end{figure}
+	\end{figure*}
 
@@
 		\begin{enumerate}
-			\item Static sounds in which the harmonic average across time is a fitting representation of the sound
+			\item Static sounds that do not change harmonically over time
 			\item Sounds in which there is at least some pitched content and not only noise
@@
 Thus, these models output 4 stereo tracks that correspond to broad instrumental categories defined in the SiSEC 2018 Mus evaluation campaign: \textit{vocals, drums, bass} and \textit{other}~\cite{Stoter_SiSEC}.
-Although the audio sources we consider in this work are often non-musical, we hope that these source separation methods can still improve the orchestration by separating other types of sound in a musically meaningful way.
-	The second type of model we tested are ones trained for universal sound separation (TDCN++, NMF), which \cite{tdcnpp} defines as \textit{separating mixtures of arbitrary sounds of different types}.
+It is important here to make the distinction that we did not train these models ourselves.
+Therefore, we cannot guarantee that the data the models were trained on is optimal for solving our problem.
+However, one of the goals of this paper is to explore whether pre-trained music source separation models can separate non-musical sounds in a way that improves orchestration.
+	The second type of model we tested are ones that perform universal sound separation (TDCN++, NMF), which \cite{tdcnpp} defines as \textit{separating mixtures of arbitrary sounds of different types}.
+These methods more closely match our problem of separating non-musical target sounds.
 
@@
 			The \textbf{Demucs}~\cite{demucs} architecture is a deep learning model for musical source separation.
-It takes as input the stereo mixture, and outputs 4 stereo waveforms corresponding to the four categories: vocals, bass, drums and other.
+It takes the stereo mixture as input, and outputs 4 stereo waveforms corresponding to the four categories: vocals, bass, drums and other.
 Demucs is based on a convolutional encoder, a bidirectional LSTM, and a convolutional decoder.
@@
 
-		\subsection{Testing}
+		\subsection{Testing}\label{subsec:testing}
 		In order to compare the effectiveness of the separation methods, a full target orchestration and a ground truth orchestration are done for each target.
@@
 		\begin{enumerate}
-			\item The target is created as a combination of four randomly chosen sources, which are offset to begin playing at different times
-			\item The full target, without any separation performed, is orchestrated using the entire orchestra
-			\item Each separation method is performed on the target, each splitting the target into four sub-targets 
-			\item The full orchestra is randomly split into four equal sized subsets
-			\item For each method, the four sub-targets are separately orchestrated, each using a different subset of the orchestra, and then each orchestration is combined to play simultaneously
-			\item The ground truth orchestration is created by orchestrating the four sources, using the same orchestra splits from step 4, and then combined to play simultaneously
-			\item The distances between target and orchestration are calculated for the full target orchestration, the separated orchestration, and the ground truth orchestration 
+			\item The target is created as a combination of four randomly chosen sources, which are offset to begin playing at different times.
+			\item The full target, without any separation performed, is orchestrated using the entire orchestra. This is called the \textbf{full target orchestration}.
+			\item Each separation method is performed on the target, each splitting the target into four sub-targets. 
+			\item The full orchestra is randomly split into four equal sized subsets, each containing 7 to 8 instruments.
+			\item For each method, the four sub-targets are separately orchestrated, each using a different subset of the orchestra, and then each orchestration is combined to play simultaneously.
+This creates the \textbf{separated orchestrations}.
+			\item The \textbf{ground truth orchestration} is created by orchestrating the four sources, using the same orchestra subsets from step 4, and then combined to play simultaneously.
+			\item The distances between target and orchestration are calculated for the full target orchestration, the separated orchestration, and the ground truth orchestration. 
 		\end{enumerate}
 
-		\begin{figure}[t]
+		\begin{figure*}[t]
 			\centering
-				\includegraphics[width=\columnwidth]{figures/evaluation_diagram.png}
-				\caption{Diagram showing the testing process. Note that in both orchestrations that use a split orchestra, the split is the same.}\label{fig:eval_diagram}
-		\end{figure}
+				\includegraphics[width=\textwidth]{figures/evaluation_diagram.png}
+				\caption{Diagram showing the testing process.
+Note that in both orchestrations that use a split orchestra, they use the same division of the orchestra to orchestrate their four sub-targets.}\label{fig:eval_diagram}
+		\end{figure*}
 
@@
 The spectral distance metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization, and used in \cite{Cella2020b} to compare accuracies of orchestrations.
-The equation takes in the full FFT of the target $x$ and full FFT of the solution $\tilde{x}$.
+The equation takes in the FFT of the target $x$ and FFT of the solution $\tilde{x}$.
 Then for each bin $k$ of the FFT, it calculates the absolute difference between the values.
@@
 				\hline
-				Full target & 24.96 \\
+				Full target orchestration & 24.96 \\
 				\hline
-				Ground truth & 24.19 \\
+				Ground truth orchestration & 24.19 \\
 				\hline
@@
 		\end{center}
-		\caption{Average across 300 targets showing the distance between target and orchestration for various methods. ``Full target'' means no separation was performed. Lower values are better.}
+		\caption{Average across 300 targets showing the distance between target and orchestration for various methods.
+For Demucs, Open-Unmix, TDCN, NMF, and TDCN++, the value shown is the average distance between the target sound and the ``separated orchestration'' created from the output of the given separation method (see Sec.
+\ref{subsec:testing}).
+Similarly, ``Full target orchestration'' and ``Ground truth orchestration'', as defined in Sec.
+\ref{subsec:testing}, are the average values between target and orchestrated solution.
+Lower values are better.}
 		\label{tab:distances}

### 2021-05-25 source_separation/paper/paper.tex
Edit sec 3.3

@@
 		In order to compare the effectiveness of the separation methods, a full target orchestration and a ground truth orchestration are done for each target.
-The full target orchestration is the orchestration of the target without any separation performed.
-The ground truth orchestration takes the four sources that make up the target and orchestrates each one separately, then combines them.
+The \textbf{full target orchestration} is the orchestration of the target without any separation performed.
+The \textbf{ground truth orchestration} takes the four sources that make up the target and orchestrates each one separately, then combines the orchestrations.
 This ground truth represents the orchestration that could take place if the separation method was perfect and separated a target exactly into its constituent sources.
@@
 			\item The target is created as a combination of four randomly chosen sources, which are offset to begin playing at different times.
-			\item The full target, without any separation performed, is orchestrated using the entire orchestra. This is called the \textbf{full target orchestration}.
+			\item The full target, without any separation performed, is orchestrated using the entire orchestra. This is called the full target orchestration.
 			\item Each separation method is performed on the target, each splitting the target into four sub-targets. 
@@
 This creates the \textbf{separated orchestrations}.
-			\item The \textbf{ground truth orchestration} is created by orchestrating the four sources, using the same orchestra subsets from step 4, and then combined to play simultaneously.
+			\item The ground truth orchestration is created by orchestrating the four sources, using the same orchestra subsets from step 4, and then combined to play simultaneously.
 			\item The distances between target and orchestration are calculated for the full target orchestration, the separated orchestration, and the ground truth orchestration. 
@@
 Figure \ref{fig:orchestrations} illustrates the effect this offset has on the full target, separated, and ground truth orchestrations.
-The full target orchestration places all onsets at the beginning of the orchestration, since it is unable to differentiate between different sources, it cannot place the different orchestrations at the appropriate offsets.
-However, the separated orchestrations are able to correctly place the onsets of the sub-target orchestrations so far as the method has correctly separated out the sources.
-In the ground truth orchestration, the sources are offset by the same amount as they are in the creation of the target, leading to the ground truth having each source orchestration beginning at the correct time.
+In the full target orchestration, all onsets occur at the beginning of the orchestration.
+Since there is no knowledge of different sources, Orchidea cannot place the different orchestrations at the appropriate offsets.
+For the separated orchestrations, the onsets of the sub-target orchestrations can be placed at the correct offset if the separation method has correctly separated out the sources.
+In the ground truth orchestration, the sources are manually offset, leading to the ground truth having each source orchestration beginning at the correct time.
 For this reason, despite only performing static orchestrations, the resulting separated and ground truth orchestrations are offset in time compared to the full target orchestration.

### 2021-05-31 source_separation/paper/paper.tex
Add justification for target creation process
Edit description of Eq.1

@@
 The other three sources are then randomly assigned different times to begin playing such that they will start between the beginning of the target and half-way through the longest source sound.
+This is done to ensure that the sources overlap and to minimize the amount of time in which there is only a single source sounding.
 
@@
 The spectral distance metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization, and used in \cite{Cella2020b} to compare accuracies of orchestrations.
-The equation takes in the FFT of the target $x$ and FFT of the solution $\tilde{x}$.
-Then for each bin $k$ of the FFT, it calculates the absolute difference between the values.
+The equation takes in the magnitude spectrums of the FFTs of the target $x$ and of the solution $\tilde{x}$.
+Then for each bin $k$ of these spectrums, it calculates the absolute difference between the values.
 The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways.

### 2021-05-31 source_separation/paper/paper.tex
Add details on OpenUnmix's standardization

@@
 Each of these models is trained on the specific target using the same architecture, based on a three layer bidirectional LSTM network.
-Open-Unmix operates on the Short Time Fourier Transform (STFT) of the input mixture, which is first standardized.
+Open-Unmix operates on the Short Time Fourier Transform (STFT) of the input mixture, which is first standardized using the global mean and standard deviation for every frequency bin across all frames.
 It predicts the target by multiplying the output of the LSTM network with the magnitude spectrogram of the input, essentially applying a mask.

### 2021-06-02 source_separation/paper/paper.tex
Update file to use MLSP paper style

@@
-% -----------------------------------------------
-% Template for SMC 2021
-% Adapted from previous SMC paper templates
-% -----------------------------------------------
+% LaTeX template for MLSP papers. To be used with:
+%   * mlspconf.sty - ICASSP/ICIP LaTeX style file adapted for MLSP, and
+%   * IEEEbib.bst - IEEE bibliography style file.
+% --------------------------------------------------------------------------
 \documentclass{article}
-\usepackage{smc2021}
+\usepackage{amsmath,graphicx,mlspconf}
 %%%%%%%%%%%%%%%%%%%%%%%% Some useful packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \usepackage{paralist}% extended list environments
+\usepackage{url}
+\usepackage{hyperref}
 \usepackage[figure,table]{hypcap}% hyperref companion
 \usepackage{multirow}
-% Enable for Review only, remove for Camera Ready version
-\pagewiselinenumbers
 
-% Use this if english is the only language/alphabet used in the document
-\usepackage[english]{babel}
+\usepackage{amssymb}
+
+% Copyright notices.
+% ------------------
+% Select one of the four copyright notices below. Only required for the camera-ready paper submission.
+% 
+% * For papers in which all authors are employed by the US government:
+\copyrightnotice{U.S.\ Government work not protected by U.S.\ copyright}
+
+% * For papers in which all authors are employed by a Crown government (UK, Canada, and Australia):
+\copyrightnotice{978-1-7281-6338-3/21/\$31.00 {\copyright}2021 Crown}
+
+% * For papers in which all authors are employed by the European Union:
+\copyrightnotice{978-1-7281-6338-3/21/\$31.00
+{\copyright}2021 European Union}
+
+% * For all other papers:
+\copyrightnotice{978-1-7281-6338-3/21/\$31.00 {\copyright}2021 IEEE}
+
+% Header
+\toappear{2021 IEEE International Workshop on Machine Learning for Signal Processing, Oct.\ 25--28, 2021, Gold Coast, Australia}
+
+% Example definitions.
+% --------------------
+%\def\x{{\mathbf x}}
+%\def\L{{\cal L}}
 
@@
 % ------
-\def\papertitle{Source separation methods for computer-assisted orchestration}
+\title{SOURCE SEPARATION METHODS FOR COMPUTER-ASSISTED ORCHESTRATION}
+%
+% Double-blind peer review.
+% -------------------------
+% Anonymize your paper for the double-blind peer-review process using the 
+% following author and affiliation.
+\name{Anonymous\thanks{Anonymous.}}
+\address{Anonymous}
 
-% Authors
-% Please note that submissions are NOT anonymous, therefore 
-% authors' names have to be VISIBLE in your manuscript. 
-% Authors are entered as an ordered list, each one can be linked to multiple affiliations using the correct index.
-% Available tags for authors are: \firstname \middlename \lastname \generation \originalname \email \orcid
-% Available tags for affiliations are: \unit \department \institution \streetaddress \city \state \postcode \country \type
-% type can take as value: University, Company, Music, Independent, Other
+% Single address.
+% ---------------
+%\name{Author(s) Name(s)\thanks{Thanks to XYZ agency for funding.}}
+%\address{Author Affiliation(s)}
+
+% For example:
+% ------------
+%\address{%
+%    School \\
+%    Department \\
+%    Address
+%}
 %
-% \author[]{\mbox{\firstname{}\middlename{}\lastname{}\originalname{}\generation{}\email{}\orcid{}}}
-% mbox force an author not to be split over multiple lines
-\author[1]{\mbox{\firstname{Luke}\lastname{Dzwonczyk}\email{dz.luke@berkeley.edu}}}
-\author[1, 2, 3]{\mbox{\firstname{L\'eo}\lastname{Ch\'edin}}}
-\author[1]{\mbox{\firstname{Alejandro}\lastname{Saldarriaga-Fuertes}}}
-\author[1]{\mbox{\firstname{Max}\lastname{Sherr}}}
-\author[3]{\mbox{\firstname{H\'el\`ene-Camille}\lastname{Crayencour}}}
-\author[1]{\mbox{\firstname{Carmine-Emanuele}\lastname{Cella}}}
+% Two addresses.
+% --------------
+%\twoauthors{%
+%    A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}
+%}{%
+%    School A-B \\
+%    Department A-B \\
+%    Address A-B \\
+%    Email A-B
+%}{%
+%   C. Author-three, D. Author-four\sthanks{The fourth author performed the work while at ...}
+%}{%
+%    School C-D \\
+%    Department C-D \\
+%    Address C-D \\
+%    Email C-D
+%}
+% 
+% Two or more addresses (alternative form).
+% -----------------------------------------
+% If you need to list more than 2 authors or the option for two options above 
+% produces a poor author block, please use the following structure:
+%\name{%
+%    Author Name$^{\star \dagger}$%
+%    \qquad Author Name$^{\star}$%
+%    \qquad Author Name$^{\dagger}$\thanks{Thanks to XYZ agency for funding.}%
+%}
+%\address{%
+%    $^{\star}$ Affiliation Number One \\%
+%    $^{\dagger}$ Affiliation Number Two%
+%}
 
-%%Affiliations
-\affil[1]{\department{Center for New Music and Audio Technologies}\institution{University of California, Berkeley}\city{Berkeley}\state{California}\country{USA}\affiliationtype{University}}
-\affil[2]{\department{\'Ecole Normale Sup\'erieure Paris-Saclay}\institution{Universit\'e Paris-Saclay}\city{Paris}\state{}\country{France}\affiliationtype{University}}
-\affil[3]{\department{Laboratoire des Signaux et Syst\`emes}\institution{Centrale Sup\'elec, CNRS, Universit\'e Paris-Saclay}\city{Paris}\state{}\country{France}\affiliationtype{University}}
+\begin{document}
+%\ninept
 
-% Complete setup stage
-\completesetup
-
-% Title.
-% ------
-\title{\papertitle}
-% ***************************************** the document starts here ***************
-\begin{document}
-	%
-	\capstartfalse
 \maketitle
-	\capstarttrue
-	%
 
@@
 %
-
+\begin{keywords}
+One, two, three, four, five
+\end{keywords}
+%
 \section{Introduction}\label{sec:introduction}
-
   Computer-assisted composition is a field that focuses on the creation of computational tools to aid in the musical composition process \cite{FerVic2013, Ari2005}.
@@
 
-	The codebase for this paper can be found at: \mbox{\url{https://github.com/dzluke/smc2021}}, and you can listen to a selection of targets and orchestrations here: \mbox{\url{https://dzluke.github.io/smc2021/}}.
+%  The codebase for this paper can be found at: \url{https://github.com/dzluke/smc2021}, and you can listen to a selection of targets and orchestrations here: \url{https://dzluke.github.io/smc2021/}.
+  The codebase for this paper can be found at: \url{anonymous}, and you can listen to a selection of targets and orchestrations here: \url{anonymous}.
 
@@
 The random split of the orchestra is the same for each separation method of the same target.
-See Fig.
-\ref{fig:eval_diagram} for a diagram of this process.
+See Fig.~\ref{fig:eval_diagram} for a diagram of this process.
 
@@
     We created our own targets as combinations of four source sounds.
-The sources come from the NIGENS \cite{NIGENS} and BBC \cite{BBC} databases, and freesound.org \cite{freesound}.
+The sources come from the NIGENS~\cite{NIGENS} and BBC~\cite{BBC} databases, and freesound.org \cite{freesound}.
 We selected a total of 90 samples from these databases, choosing sounds that fit the following criteria:
@@
     As described in Section \ref{subsec:data}, each target is comprised of four sources that are offset in time to begin playing at different times.
-Figure \ref{fig:orchestrations} illustrates the effect this offset has on the full target, separated, and ground truth orchestrations.
+Figure~\ref{fig:orchestrations} illustrates the effect this offset has on the full target, separated, and ground truth orchestrations.
 In the full target orchestration, all onsets occur at the beginning of the orchestration.
@@
 
-	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-	%bibliography here
-	\bibliography{references}
+% References should be produced using the bibtex program from suitable
+% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
+% style file from IEEE produces unsorted bibliography list.
+% -------------------------------------------------------------------------
+\bibliographystyle{IEEEbib}
+\bibliography{refs}
 

### 2021-06-03 source_separation/paper/paper.tex
Comment out copyrights, update github link
Add index terms

@@
 % * For papers in which all authors are employed by the US government:
-\copyrightnotice{U.S.\ Government work not protected by U.S.\ copyright}
+%\copyrightnotice{U.S.\ Government work not protected by U.S.\ copyright}
 
 % * For papers in which all authors are employed by a Crown government (UK, Canada, and Australia):
-\copyrightnotice{978-1-7281-6338-3/21/\$31.00 {\copyright}2021 Crown}
+%\copyrightnotice{978-1-7281-6338-3/21/\$31.00 {\copyright}2021 Crown}
 
 % * For papers in which all authors are employed by the European Union:
-\copyrightnotice{978-1-7281-6338-3/21/\$31.00
-{\copyright}2021 European Union}
+%\copyrightnotice{978-1-7281-6338-3/21/\$31.00
+%{\copyright}2021 European Union}
 
@@
 \begin{keywords}
-One, two, three, four, five
+computer-assisted orchestration, sound source separation, Orchidea
 \end{keywords}
@@
 
-%  The codebase for this paper can be found at: \url{https://github.com/dzluke/smc2021}, and you can listen to a selection of targets and orchestrations here: \url{https://dzluke.github.io/smc2021/}.
+%  The codebase for this paper can be found at: \url{https://github.com/dzluke/MLSP2021}, and you can listen to a selection of targets and orchestrations here: \url{https://dzluke.github.io/MLSP2021/}.
   The codebase for this paper can be found at: \url{anonymous}, and you can listen to a selection of targets and orchestrations here: \url{anonymous}.
@@
       \item The target is created as a combination of four randomly chosen sources, which are offset to begin playing at different times.
-      \item The full target, without any separation performed, is orchestrated using the entire orchestra. This is called the full target orchestration.
+      \item The target, without any separation performed, is orchestrated using the entire orchestra. This is called the full target orchestration.
       \item Each separation method is performed on the target, each splitting the target into four sub-targets. 

### 2021-06-03 source_separation/create_separated_subtargets.py
Add names for clarity

@@ -11,5 +11,5 @@ for target_name in tqdm(os.listdir(targets_directory)):
     for model in ["TDCNN++", "TDCNN", "OpenUnmix", "Demucs"]:
         print(target_name, model)
         target = targets_directory + '/' + target_name
-        l = generate_separation_function(model, 2)
+        l = generate_separation_function(model, num_sub_targets=2)
         l[0](target, 2)

### 2021-06-03 source_separation/pipeline.py

@@ -165,7 +165,7 @@ def generate_separation_function(model_name, num_sub_targets):
         raise Exception("Model name must be one of those five : TDCN, TDCN++, OpenUnmix, Demucs, NMF")
 
     for perm in gen_perm_group(init_list, num_sub_targets):
-        l.append(lambda a, n: separate(a, model_name, n, perm))
+        l.append(lambda audio_path, num_subtargets: separate(audio_path, model_name, num_subtargets, perm))
     return l
 
 

### 2021-06-03 source_separation/paper/paper.tex
Move discussion of dynamic orch to future work

@@
 Formally, it is the process of creating an orchestral score that best matches an arbitrary target sound given a similarity metric and constraints~\cite{Maresz2003}.
-
-  Computer-assisted orchestrations can be divided into two categories: static and dynamic.
-Static orchestration considers the target as a single vector of timbral features and creates a single onset of note(s) as the solution.
-Dynamic orchestration considers a time series of features, which allows it to create an orchestration that evolves over time with multiple onsets.
-It uses time segmentation to cut the target in time into multiple segments, and orchestrates each segment.
+Static orchestration is a type of orchestration that considers the target as a single vector of timbral features and creates only a single onset of note(s) as the solution, no matter if the target contains multiple onsets of sounds.
+In our experiments, we only perform static orchestrations.
 
@@
 It embeds the target in a feature space and uses a jointly-optimized heuristic and constraint solver to find a combination of samples that best match the target~\cite{Cella18, Cella2020}.
-Orchidea is able to do both static and dynamic orchestration.
-In order to better observe the effects of source separation, we perform only static orchestrations in this paper.
-The reasoning for this is that we wish to disentangle the two difficult problems of source separation and time segmentation.
-By only performing static orchestrations, we reduce the complexity of the problem by removing time segmentation and can better observe the effects of source separation.
+All of our orchestrations are done using Orchidea's command line tools for batch processing, which is available for download at \url{http://www.orch-idea.org}
 
@@
 The orchestration would be greatly improved by removing the unnecessary onsets in the orchestration of the drone.
-Although this example is a dynamic orchestration, and we perform only static orchestrations, this paper represents a first step toward achieving dynamic orchestration with source separation.
 
@@
 
-    As we stated in Section \ref{sec:introduction}, we chose static orchestration instead of dynamic orchestration to avoid entangling the problems of time segmentation and source separation.
+    As we stated in Section \ref{sec:introduction}, we performed only static orchestration in our experiments.
+In contrast to static orchestration, dynamic orchestration considers a time series of features, which allow the orchestration to evolve over time and contain multiple onsets of notes.
+Orchidea is able to perform dynamic orchestration by cutting the target in time into multiple segments, and then orchestrating each segment.
+However we chose to use only static orchestrations in order to better observe the effects of source separation.
+The reasoning for this is that we wish to disentangle the two difficult problems of source separation and time segmentation.
+By only performing static orchestrations, we reduce the complexity of the problem by removing time segmentation and can better observe the effects of source separation.
 Future work could include applying source separation to dynamic targets and performing dynamic orchestration.

### 2021-06-03 source_separation/demucs/__init__.py
Update demucs model

@@ -1,0 +1,7 @@
+# Copyright (c) Facebook, Inc. and its affiliates.
+# All rights reserved.
+#
+# This source code is licensed under the license found in the
+# LICENSE file in the root directory of this source tree.
+
+__version__ = "2.0.2"

### 2021-06-03 source_separation/demucs/model.py

@@ -6,7 +6,7 @@
 
 import math
 
-import torch as th
+import julius
 from torch import nn
 
 from .utils import capture_init, center_trim
@@ -40,43 +40,28 @@ def rescale_module(module, reference):
             rescale_conv(sub, reference)
 
 
-def upsample(x, stride):
-    """
-    Linear upsampling, the output will be `stride` times longer.
-    """
-    batch, channels, time = x.size()
-    weight = th.arange(stride, device=x.device, dtype=th.float) / stride
-    x = x.view(batch, channels, time, 1)
-    out = x[..., :-1, :] * (1 - weight) + x[..., 1:, :] * weight
-    return out.reshape(batch, channels, -1)
-
-
-def downsample(x, stride):
-    """
-    Downsample x by decimation.
-    """
-    return x[:, :, ::stride]
-
-
 class Demucs(nn.Module):
     @capture_init
     def __init__(self,
-                 sources=4,
+                 sources,
                  audio_channels=2,
                  channels=64,
                  depth=6,
                  rewrite=True,
                  glu=True,
-                 upsample=False,
                  rescale=0.1,
+                 resample=True,
                  kernel_size=8,
                  stride=4,
                  growth=2.,
                  lstm_layers=2,
-                 context=3):
+                 context=3,
+                 normalize=False,
+                 samplerate=44100,
+                 segment_length=4 * 10 * 44100):
         """
         Args:
-            sources (int): number of sources to separate
+            sources (list[str]): list of source names
             audio_channels (int): stereo or mono
             channels (int): first convolution channels
             depth (int): number of encoder/decoder layers
@@ -84,8 +69,7 @@ class Demucs(nn.Module):
                 and a convolution to each decoder layer.
                 For the decoder layer, `context` gives the kernel size.
             glu (bool): use glu instead of ReLU
-            upsample (bool): use linear upsampling with convolutions
-                Wave-U-Net style, instead of transposed convolutions
+            resample_input (bool): upsample x2 the input and downsample /2 the output.
             rescale (int): rescale initial weights of convolutions
                 to get their standard deviation closer to `rescale`
             kernel_size (int): kernel size for convolutions
@@ -97,6 +81,11 @@ class Demucs(nn.Module):
                 decoder before the transposed convolution. If > 1,
                 will provide some context from neighboring time
                 steps.
+            samplerate (int): stored as meta information for easing
+                future evaluations of the model.
[…]

### 2021-06-03 source_separation/demucs/separate.py

@@ -5,159 +5,53 @@
 # LICENSE file in the root directory of this source tree.
 
 import argparse
-import hashlib
 import sys
 from pathlib import Path
+import subprocess
 
-import requests
+import julius
 import torch as th
-import tqdm
-from scipy.io import wavfile
+import torchaudio as ta
 
-# from .audio import AudioFile
-# from .utils import apply_model, load_model
-
-from .audio import AudioFile
+from .audio import AudioFile, convert_audio_channels
+from .pretrained import is_pretrained, load_pretrained
 from .utils import apply_model, load_model
 
-BASE_URL = "https://dl.fbaipublicfiles.com/demucs/v2.0/"
-PRETRAINED_MODELS = {
-    'demucs.th': 'f6c4148ba0dc92242d82d7b3f2af55c77bd7cb4ff1a0a3028a523986f36a3cfd',
-    'demucs.th.gz': 'e70767bfc9ce62c26c200477ea29a20290c708b210977e3ef2c75ace68ea4be1',
-    'demucs_extra.th': '3331bcc5d09ba1d791c3cf851970242b0bb229ce81dbada557b6d39e8c6a6a87',
-    'demucs_extra.th.gz': 'f9edcf7fe55ea5ac9161c813511991e4ba03188112fd26a9135bc9308902a094',
-    'light.th': '79d1ee3c1541c729c552327756954340a1a46a11ce0009dea77dc583e4b6269c',
-    'light.th.gz': '94c091021d8cdee0806b6df0afbeb59e73e989dbc2c16d2c1c370b2edce774fd',
-    'light_extra.th': '9e9b4af564229c80cc73c95d02d2058235bb054c6874b3cba4d5b26943a5ddcb',
-    'light_extra.th.gz': '48bb1a85f5ad0ca400512fcd0dcf91ec94e886a1602a552ee32133f5e09aeae0',
-    'tasnet.th': 'be56693f6a5c4854b124f95bb9dd043f3167614898493738ab52e25648bec8a2',
-    'tasnet_extra.th': '0ccbece3acd98785a367211c9c35b1eadae8d148b0d37fe5a5494d6d335269b5',
-}
 
-
-def separate(input_path,
-             output_path,
-             name='demucs',
-             quantized=False,
-             models=Path('models'),
-             dl=True,
-             device='cuda' if th.cuda.is_available() else 'cpu',
-             shifts=0,
-             split=0,
-             mp3=False,
-             mp3_bitrate=320,
[…]

### 2021-06-03 source_separation/demucs/utils.py

@@ -6,14 +6,18 @@
 
 import errno
 import functools
-import gzip
+import hashlib
+import inspect
+import io
 import os
 import random
 import socket
 import tempfile
 import warnings
+import zlib
 from contextlib import contextmanager
 
+from diffq import UniformQuantizer, DiffQuantizer
 import torch as th
 import tqdm
 from torch import distributed
@@ -92,7 +96,57 @@ def human_seconds(seconds, display='.2f'):
     return f"{format(value, display)} {last}"
 
 
-def apply_model(model, mix, shifts=None, split=False, progress=False):
+class TensorChunk:
+    def __init__(self, tensor, offset=0, length=None):
+        total_length = tensor.shape[-1]
+        assert offset >= 0
+        assert offset < total_length
+
+        if length is None:
+            length = total_length - offset
+        else:
+            length = min(total_length - offset, length)
+
+        self.tensor = tensor
+        self.offset = offset
+        self.length = length
+        self.device = tensor.device
+
+    @property
+    def shape(self):
+        shape = list(self.tensor.shape)
+        shape[-1] = self.length
+        return shape
+
+    def padded(self, target_length):
+        delta = target_length - self.length
+        total_length = self.tensor.shape[-1]
+        assert delta >= 0
+
+        start = self.offset - delta // 2
+        end = start + target_length
+
+        correct_start = max(0, start)
+        correct_end = min(total_length, end)
+
+        pad_left = correct_start - start
+        pad_right = end - correct_end
+
+        out = F.pad(self.tensor[..., correct_start:correct_end], (pad_left, pad_right))
+        assert out.shape[-1] == target_length
+        return out
+
+
+def tensor_chunk(tensor_or_chunk):
+    if isinstance(tensor_or_chunk, TensorChunk):
+        return tensor_or_chunk
+    else:
+        assert isinstance(tensor_or_chunk, th.Tensor)
+        return TensorChunk(tensor_or_chunk)
+
+
+def apply_model(model, mix, shifts=None, split=False,
+                overlap=0.25, transition_power=1., progress=False):
     """
     Apply model to a given mixture.
 
@@ -106,44 +160,60 @@ def apply_model(model, mix, shifts=None, split=False, progress=False):
             Useful for model with large memory footprint like Tasnet.
         progress (bool): if True, show a progress bar (requires split=True)
     """
-    channels, length = mix.size()
[…]

### 2021-06-03 source_separation/open_unmix/separate.py
Update openunmix model

@@ -1,172 +1,0 @@
-# -*- coding: utf-8 -*-
-
-import torch
-import tqdm
-import numpy as np
-import soundfile as sf
-import norbert
-import warnings
-import resampy
-from pathlib import Path
-import io
-from contextlib import redirect_stderr
-import json
-import scipy.signal
-
-from .utils import bandwidth_to_max_bin
-from .open_unmix import OpenUnmix
-
-
-def load_model(target, model_name='umxhq', device='cpu'):
-    """
-    target model path can be either <target>.pth, or <target>-sha256.pth
-    (as used on torchub)
-    """
-    model_path = Path(model_name).expanduser()
-    if not model_path.exists():
-        # model path does not exist, use hubconf model
-        try:
-            # disable progress bar
-            err = io.StringIO()
-            with redirect_stderr(err):
-                return torch.hub.load(
-                    'sigsep/open-unmix-pytorch',
-                    model_name,
-                    target=target,
-                    device=device,
-                    pretrained=True
-                )
-            print(err.getvalue())
-        except AttributeError:
-            raise NameError('Model does not exist on torchhub')
-            # assume model is a path to a local model_name direcotry
-    else:
-        # load model from disk
-        with open(Path(model_path, target + '.json'), 'r') as stream:
-            results = json.load(stream)
-
-        target_model_path = next(Path(model_path).glob("%s*.pth" % target))
-        state = torch.load(
-            target_model_path,
-            map_location=device
-        )
-
-        max_bin = bandwidth_to_max_bin(
-            state['sample_rate'],
-            results['args']['nfft'],
-            results['args']['bandwidth']
-        )
-
-        unmix = OpenUnmix(
-            n_fft=results['args']['nfft'],
-            n_hop=results['args']['nhop'],
-            nb_channels=results['args']['nb_channels'],
-            hidden_size=results['args']['hidden_size'],
-            max_bin=max_bin
-        )
-
-        unmix.load_state_dict(state)
-        unmix.stft.center = True
-        unmix.eval()
-        unmix.to(device)
-        return unmix
-
-
-def istft(X, rate=44100, n_fft=4096, n_hopsize=1024):
-    t, audio = scipy.signal.istft(
-        X / (n_fft / 2),
-        rate,
-        nperseg=n_fft,
-        noverlap=n_fft - n_hopsize,
[…]

### 2021-06-03 source_separation/open_unmix/test.py

@@ -1,369 +1,0 @@
-import torch
-import numpy as np
-import argparse
-import soundfile as sf
-import norbert
-import json
-from pathlib import Path
-import scipy.signal
-import resampy
-import model
-import utils
-import warnings
-import tqdm
-from contextlib import redirect_stderr
-import io
-
-
-def load_model(target, model_name='umxhq', device='cpu'):
-    """
-    target model path can be either <target>.pth, or <target>-sha256.pth
-    (as used on torchub)
-    """
-    model_path = Path(model_name).expanduser()
-    if not model_path.exists():
-        # model path does not exist, use hubconf model
-        try:
-            # disable progress bar
-            err = io.StringIO()
-            with redirect_stderr(err):
-                return torch.hub.load(
-                    'sigsep/open-unmix-pytorch',
-                    model_name,
-                    target=target,
-                    device=device,
-                    pretrained=True
-                )
-            print(err.getvalue())
-        except AttributeError:
-            raise NameError('Model does not exist on torchhub')
-            # assume model is a path to a local model_name direcotry
-    else:
-        # load model from disk
-        with open(Path(model_path, target + '.json'), 'r') as stream:
-            results = json.load(stream)
-
-        target_model_path = next(Path(model_path).glob("%s*.pth" % target))
-        state = torch.load(
-            target_model_path,
-            map_location=device
-        )
-
-        max_bin = utils.bandwidth_to_max_bin(
-            state['sample_rate'],
-            results['args']['nfft'],
-            results['args']['bandwidth']
-        )
-
-        unmix = model.OpenUnmix(
-            n_fft=results['args']['nfft'],
-            n_hop=results['args']['nhop'],
-            nb_channels=results['args']['nb_channels'],
-            hidden_size=results['args']['hidden_size'],
-            max_bin=max_bin
-        )
-
-        unmix.load_state_dict(state)
-        unmix.stft.center = True
-        unmix.eval()
-        unmix.to(device)
-        return unmix
-
-
-def istft(X, rate=44100, n_fft=4096, n_hopsize=1024):
-    t, audio = scipy.signal.istft(
-        X / (n_fft / 2),
-        rate,
-        nperseg=n_fft,
-        noverlap=n_fft - n_hopsize,
-        boundary=True
-    )
-    return audio
-
-
-def separate(
-    audio,
-    targets,
-    model_name='umxhq',
-    niter=1, softmask=False, alpha=1.0,
-    residual_model=False, device='cpu'
-):
[…]

### 2021-06-03 source_separation/open_unmix/utils.py

@@ -1,140 +1,58 @@
-import shutil
+from typing import Optional, Union
+
 import torch
 import os
 import numpy as np
+import torchaudio
+import warnings
+from pathlib import Path
+from contextlib import redirect_stderr
+import io
+import json
+
+import openunmix
+from openunmix import model
 
 
-def _sndfile_available():
-    try:
-        import soundfile
-    except ImportError:
-        return False
+def bandwidth_to_max_bin(rate: float, n_fft: int, bandwidth: float) -> np.ndarray:
+    """Convert bandwidth to maximum bin count
 
-    return True
+    Assuming lapped transforms such as STFT
 
+    Args:
+        rate (int): Sample rate
+        n_fft (int): FFT length
+        bandwidth (float): Target bandwidth in Hz
 
-def _torchaudio_available():
-    try:
-        import torchaudio
-    except ImportError:
-        return False
-
-    return True
-
-
-def get_loading_backend():
-    if _torchaudio_available():
-        return torchaudio_loader
-
-    if _sndfile_available():
-        return soundfile_loader
-
-
-def get_info_backend():
-    if _torchaudio_available():
-        return torchaudio_info
-
-    if _sndfile_available():
-        return soundfile_info
-
-
-def soundfile_info(path):
-    import soundfile
-    info = {}
-    sfi = soundfile.info(path)
-    info['samplerate'] = sfi.samplerate
-    info['samples'] = int(sfi.duration * sfi.samplerate)
-    info['duration'] = sfi.duration
-    return info
-
-
-def soundfile_loader(path, start=0, dur=None):
-    import soundfile
-    # get metadata
-    info = soundfile_info(path)
-    start = int(start * info['samplerate'])
-    # check if dur is none
-    if dur:
-        # stop in soundfile is calc in samples, not seconds
-        stop = start + int(dur * info['samplerate'])
-    else:
-        # set to None for reading complete file
-        stop = dur
-
-    audio, _ = soundfile.read(
-        path,
-        always_2d=True,
-        start=start,
-        stop=stop
-    )
-    return torch.FloatTensor(audio.T)
-
-
-def torchaudio_info(path):
-    import torchaudio
-    # get length of file in samples
-    info = {}
-    si, _ = torchaudio.info(str(path))
-    info['samplerate'] = si.rate
-    info['samples'] = si.length // si.channels
-    info['duration'] = info['samples'] / si.rate
-    return info
-
-
[…]

### 2021-06-03 source_separation/demucs/separate.py
Update demucs separate

@@ -74,7 +74,93 @@ def encode_mp3(wav, path, bitrate=320, samplerate=44100, channels=2, verbose=Fal
         f.write(mp3_data)
 
 
-def main():
+def separate(tracks, models=Path("models"), name="demucs_quantized",
+             device=None, out=Path("separated"), shifts=0, split=True,
+             overlap=0.25, mp3=False, float32=True, mp3_bitrate=320,
+             verbose=False):
+    """Separate the sources for the given tracks using demucs model.
+
+    Args:
+        tracks (Path): Path to tracks.
+        models (Path, optional): Folder where to put extracted tracks. A
+            subfolder with the model name will be created. Defaults to
+            Path("models").
+        name (str, optional): Model name. Defaults to "demucs_quantized".
+        device (torch.device, optional): Device to use, default is cuda
+            if available else cpu.
+        out (Path, optional): Path to trained models. Defaults to
+            Path("separated").
+        shifts (int, optional): Number of random shifts for equivariant
+            stabilization. INcrease separation time but imrpoves quality
+            for Demucs in the original paper. Defaults to 0.
+        split (bool, optional): [description]. Defaults to True.
+        overlap (float, optional): [description]. Defaults to 0.25.
+        mp3 (bool, optional): [description]. Defaults to False.
+        float32 (bool, optional): [description]. Defaults to True.
+        mp3_bitrate (int, optional): [description]. Defaults to 320.
+        verbose (bool, optional): [description]. Defaults to False.
+    """
+    # Convert paths to Path objects if not already
+    if type(tracks) == list:
+        map_object = map(Path, tracks)
+        tracks = list(map_object)
+    else:
+        tracks = [Path(tracks)]
+    models= Path(models)
+    out = Path(out)
+
+    if device is None:
+        device = "cuda" if th.cuda.is_available() else "cpu"
+
+    model_path = models / (name + ".th")
+    if model_path.is_file():
+        model = load_model(model_path)
+    else:
+        if is_pretrained(name):
+            model = load_pretrained(name)
+        else:
+            print(f"No pre-trained model {name}", file=sys.stderr)
+            sys.exit(1)
+    model.to(device)
+
+    out.mkdir(parents=True, exist_ok=True)
+    print(f"Separated tracks will be stored in {out.resolve()}")
+    for track in tracks:
+        if not track.exists():
+            print(
+                f"File {track} does not exist. If the path contains spaces, "
[…]

### 2021-06-03 source_separation/create_separated_subtargets.py
Update separation methods

@@ -1,15 +1,27 @@
 import os
+import argparse
+
 from tqdm import tqdm
 
 from pipeline import generate_separation_function
 
+# TEMP_OUTPUT_PATH = "./TEMP"
+# targets_directory = input("Targets_directory: ")
 
-TEMP_OUTPUT_PATH = "./TEMP"
-targets_directory = input("Targets_directory: ")
+def create_separated_subtargets(targets_directory, output_path):
+    for target_name in tqdm(os.listdir(targets_directory)):
+        # for model in ["TDCN++", "TDCN", "OpenUnmix", "Demucs"]:
+        for model in ["Demucs", "OpenUnmix", "TDCN++", "TDCN"]:
+            print(target_name, model)
+            target = targets_directory + '/' + target_name
+            l = generate_separation_function(model, num_subtargets=4)
+            l[0](target, num_subtargets=4)
 
-for target_name in tqdm(os.listdir(targets_directory)):
-    for model in ["TDCNN++", "TDCNN", "OpenUnmix", "Demucs"]:
-        print(target_name, model)
-        target = targets_directory + '/' + target_name
-        l = generate_separation_function(model, num_sub_targets=2)
-        l[0](target, 2)
+if __name__ == '__main__':
+    parser = argparse.ArgumentParser()
+    parser.add_argument('targets_directory', type=str,
+                        help="Path to the folder containing targets.")
+    parser.add_argument('output_path', type=str,
+                        help="Path to the output folder.")
+    args, _ = parser.parse_known_args()
+    create_separated_subtargets(args.targets_directory, args.output_path)

### 2021-06-03 source_separation/open_unmix/cli.py

@@ -1,199 +1,0 @@
-from pathlib import Path
-import torch
-import torchaudio
-import json
-import numpy as np
-
-
-from openunmix import utils
-from openunmix import predict
-from openunmix import data
-
-import argparse
-
-
-def separate():
-    parser = argparse.ArgumentParser(
-        description="UMX Inference",
-        add_help=True,
-        formatter_class=argparse.RawDescriptionHelpFormatter,
-    )
-
-    parser.add_argument("input", type=str, nargs="+", help="List of paths to wav/flac files.")
-
-    parser.add_argument(
-        "--model",
-        default="umxhq",
-        type=str,
-        help="path to mode base directory of pretrained models",
-    )
-
-    parser.add_argument(
-        "--targets",
-        nargs="+",
-        type=str,
-        help="provide targets to be processed. \
-              If none, all available targets will be computed",
-    )
-
-    parser.add_argument(
-        "--outdir",
-        type=str,
-        help="Results path where audio evaluation results are stored",
-    )
-
-    parser.add_argument(
-        "--ext",
-        type=str,
-        default=".wav",
-        help="Output extension which sets the audio format",
-    )
-
-    parser.add_argument("--start", type=float, default=0.0, help="Audio chunk start in seconds")
-
-    parser.add_argument(
-        "--duration",
-        type=float,
-        help="Audio chunk duration in seconds, negative values load full track",
-    )
-
-    parser.add_argument(
-        "--no-cuda", action="store_true", default=False, help="disables CUDA inference"
-    )
-
-    parser.add_argument(
-        "--audio-backend",
-        type=str,
-        default="sox_io",
-        help="Set torchaudio backend "
-        "(`sox_io`, `sox`, `soundfile` or `stempeg`), defaults to `sox_io`",
-    )
-
-    parser.add_argument(
-        "--niter",
-        type=int,
-        default=1,
-        help="number of iterations for refining results.",
-    )
-
-    parser.add_argument(
-        "--wiener-win-len",
-        type=int,
-        default=300,
-        help="Number of frames on which to apply filtering independently",
-    )
-
-    parser.add_argument(
-        "--residual",
-        type=str,
-        default=None,
-        help="if provided, build a source with given name"
-        "for the mix minus all estimated targets",
-    )
-
-    parser.add_argument(
-        "--aggregate",
-        type=str,
-        default=None,
-        help="if provided, must be a string containing a valid expression for "
-        "a dictionary, with keys as output target names, and values "
[…]

### 2021-06-03 source_separation/open_unmix/separate.py

@@ -1,0 +1,207 @@
+from pathlib import Path
+import torch
+import torchaudio
+import json
+import numpy as np
+
+
+from openunmix import utils
+from openunmix import predict
+from openunmix import data
+
+import argparse
+
+
+def separate(input, outdir=None, targets=None, model='umxhq', start=0.0,
+             duration=None, niter=1, residual=None, ext='.wav',
+             wiener_win_len=300, filterbank='torch', aggregate=None,
+             no_cuda=False, audio_backend='sox_io'):
+    if audio_backend != "stempeg":
+        torchaudio.set_audio_backend(audio_backend)
+
+    use_cuda = not no_cuda and torch.cuda.is_available()
+    device = torch.device("cuda" if use_cuda else "cpu")
+    print("Using ", device)
+    # parsing the output dict
+    aggregate_dict = None if aggregate is None else json.loads(aggregate)
+
+    # create separator only once to reduce model loading
+    # when using multiple files
+    separator = utils.load_separator(
+        model_str_or_path=model,
+        targets=targets,
+        niter=niter,
+        residual=residual,
+        wiener_win_len=wiener_win_len,
+        device=device,
+        pretrained=True,
+        filterbank=filterbank,
+    )
+
+    separator.freeze()
+    separator.to(device)
+
+    if audio_backend == "stempeg":
+        try:
+            import stempeg
+        except ImportError:
+            raise RuntimeError("Please install pip package `stempeg`")
+
+    # If we're dealing with a single file.
+    if type(input) != list:
+        input = [input]
+
+    # loop over the files
+    for input_file in input:
+        if audio_backend == "stempeg":
+            audio, rate = stempeg.read_stems(
+                input_file,
+                start=start,
+                duration=duration,
+                sample_rate=separator.sample_rate,
+                dtype=np.float32,
+            )
+            audio = torch.tensor(audio)
+        else:
+            audio, rate = data.load_audio(input_file, start=start, dur=duration)
+        estimates = predict.separate(
+            audio=audio,
+            rate=rate,
+            aggregate_dict=aggregate_dict,
+            separator=separator,
+            device=device,
+        )
+        if not outdir:
+            model_path = Path(model)
+            if not model_path.exists():
+                outdir = Path(Path(input_file).stem + "_" + model)
+            else:
+                outdir = Path(Path(input_file).stem + "_" + model_path.stem)
+        else:
+            outdir = Path(outdir) #/ Path(input_file).stem
[…]

### 2021-06-03 source_separation/pipeline.py

@@ -36,7 +36,7 @@ CONFIG_PATH = "orch_config.txt"  # path to the Orchidea configuration file (you 
 TEMP_OUTPUT_PATH = "./TEMP"
 RESULTS_PATH = "./results.json"
 TARGET_METADATA_PATH = os.path.join(TARGETS_PATH, 'metadata.json')
-TDCNpp_nb_sub_targets = 4
+TDCNpp_nb_subtargets = 4
 SAMPLING_RATE = 44100
 NUM_SUBTARGETS = 4
 full_orchestra = ['Fl', 'Fl', 'Ob', 'Ob', 'ClBb', 'ClBb', 'Bn', 'Bn', 'Tr', 'Tr', 'Tbn', 'Tbn', 'Hn', 'Hn',
@@ -93,15 +93,18 @@ def separate(audio_path, model_name, num_subtargets, *args):
         if model_name == "TDCN++":
             # ConvTasNet for Universal Sound Separation
             TDCNpp_separate.separate(TDCNpp_model_path + "/baseline_model",
-                                                       TDCNpp_model_path + "/baseline_inference.meta",
-                                                       audio_path,
-                                                       output_path)
+                                     TDCNpp_model_path + "/baseline_inference.meta",
+                                     audio_path,
+                                     output_path)
         elif model_name == "TDCN":
-            demucs.separate(audio_path, output_path, 'tasnet')
+            demucs.separate(audio_path, out=output_path, name='tasnet')
         elif model_name == "Demucs":
-            demucs.separate(audio_path, output_path, 'demucs')
+            # We use the quantized demucs model, which is much smaller and
+            # should have the same quality.
+            demucs.separate(audio_path, out=output_path,
+                            name='demucs_quantized')
         elif model_name == "OpenUnmix":
-            open_unmix.separate(audio_path, output_path)
+            open_unmix.separate(audio_path, outdir=output_path)
         elif model_name == "NMF":
             NMF.separate(audio_path, output_path)
         else:
@@ -149,10 +152,10 @@ def gen_perm_group(l, n):
     return r
 
 
-def generate_separation_function(model_name, num_sub_targets):
+def generate_separation_function(model_name, num_subtargets):
     l = []
     if model_name == "TDCN++":
-        init_list = ["sub_target{}".format(x) for x in range(TDCNpp_nb_sub_targets)]
[…]

### 2021-06-04 source_separation/open_unmix/__init__.py
Update to use local openunmix

@@ -2,10 +2,10 @@
 ![sigsep logo](https://sigsep.github.io/hero.png)
 Open-Unmix is a deep neural network reference implementation for music source separation, applicable for researchers, audio engineers and artists. Open-Unmix provides ready-to-use models that allow users to separate pop music into four stems: vocals, drums, bass and the remaining other instruments. The models were pre-trained on the MUSDB18 dataset. See details at apply pre-trained model.
 
-This is the python package API documentation. 
+This is the python package API documentation.
 Please checkout [the open-unmix website](https://sigsep.github.io/open-unmix) for more information.
 """
-from openunmix import utils
+from open_unmix import utils
 import torch.hub
 
 

### 2021-06-04 source_separation/open_unmix/evaluate.py

@@ -9,7 +9,7 @@ import museval
 import torch
 import tqdm
 
-from openunmix import utils
+from open_unmix import utils
 
 
 def separate_and_evaluate(

### 2021-06-04 source_separation/open_unmix/predict.py

@@ -1,4 +1,4 @@
-from openunmix import utils
+from open_unmix import utils
 
 
 def separate(

### 2021-06-04 source_separation/open_unmix/separate.py

@@ -5,9 +5,9 @@ import json
 import numpy as np
 
 
-from openunmix import utils
-from openunmix import predict
-from openunmix import data
+from open_unmix import utils
+from open_unmix import predict
+from open_unmix import data
 
 import argparse
 

### 2021-06-04 source_separation/open_unmix/utils.py

@@ -10,8 +10,8 @@ from contextlib import redirect_stderr
 import io
 import json
 
-import openunmix
-from openunmix import model
+import open_unmix
+from open_unmix import model
 
 
 def bandwidth_to_max_bin(rate: float, n_fft: int, bandwidth: float) -> np.ndarray:
@@ -233,7 +233,7 @@ def load_separator(
 
     # otherwise we load the separator from torchhub
     else:
-        hub_loader = getattr(openunmix, model_str_or_path)
+        hub_loader = getattr(open_unmix, model_str_or_path)
         separator = hub_loader(
             targets=targets,
             device=device,

### 2021-06-04 source_separation/paper/paper.tex
Add new paragraph for clarity
Add explanation for random orchestra subset

@@
 Then the distance between the target and solution is calculated, giving us a metric to compare the various separation methods.
-The random split of the orchestra is the same for each separation method of the same target.
 See Fig.~\ref{fig:eval_diagram} for a diagram of this process.
+
+  For each target, the same random split of the orchestra is used to orchestrate the output of each separation method.
+We randomly create these subsets of the orchestra in order to avoid any biases that could come from a hand-picking the instruments used to orchestrate each sub-target.
 
@@
 
-    \subsection{Separation Methods}
+    \subsection{Separation methods}
   We consider source separation models trained to tackle two distinct separation problems.
@@
 Since there is no knowledge of different sources, Orchidea cannot place the different orchestrations at the appropriate offsets.
+
     For the separated orchestrations, the onsets of the sub-target orchestrations can be placed at the correct offset if the separation method has correctly separated out the sources.
@@
 
-    \subsection{Future Work}\label{sec:futurework}  
+    \subsection{Future work}\label{sec:futurework}  
     As we stated previously, we believe that one of the reasons the supervised methods did not perform as well is because of the data they were trained on.

### 2021-06-04 source_separation/results.py
Update TDCN and TDCN+ naming

@@ -39,9 +39,9 @@ if __name__ == "__main__":
 
     make_plot_data(separated_target_distances['OpenUnmix'], 'Open-Unmix')
     make_plot_data(separated_target_distances['Demucs'], 'Demucs')
-    make_plot_data(separated_target_distances['TDCNN'], 'TDCN')
+    make_plot_data(separated_target_distances['TDCN'], 'TDCN')
     make_plot_data(separated_target_distances['NMF'], 'NMF')
-    make_plot_data(separated_target_distances['TDCNN++'], 'TDCN++')
+    make_plot_data(separated_target_distances['TDCN++'], 'TDCN++')
     make_plot_data(full_target_distances, 'Full target')
     make_plot_data(ground_truth_distances, 'Ground truth')
 

### 2021-06-04 source_separation/open_unmix/utils.py
More changes to environments

@@ -129,7 +129,7 @@ def load_target_models(targets, model_str_or_path="umxhq", device="cpu", pretrai
         # model path does not exist, use pretrained models
         try:
             # disable progress bar
-            hub_loader = getattr(openunmix, model_str_or_path + "_spec")
+            hub_loader = getattr(open_unmix, model_str_or_path + "_spec")
             err = io.StringIO()
             with redirect_stderr(err):
                 return hub_loader(targets=targets, device=device, pretrained=pretrained)

### 2021-06-05 source_separation/create_separated_subtargets.py
Add output_path

@@ -3,25 +3,24 @@ import argparse
 
 from tqdm import tqdm
 
-from pipeline import generate_separation_function
+from pipeline import generate_separation_function, TEMP_OUTPUT_PATH
 
-# TEMP_OUTPUT_PATH = "./TEMP"
 # targets_directory = input("Targets_directory: ")
 
 def create_separated_subtargets(targets_directory, output_path):
     for target_name in tqdm(os.listdir(targets_directory)):
-        # for model in ["TDCN++", "TDCN", "OpenUnmix", "Demucs"]:
-        for model in ["Demucs", "OpenUnmix", "TDCN++", "TDCN"]:
+        for model in ["TDCN++", "TDCN", "OpenUnmix", "Demucs"]:
             print(target_name, model)
             target = targets_directory + '/' + target_name
-            l = generate_separation_function(model, num_subtargets=4)
+            l = generate_separation_function(model, output_path,
+                                             num_subtargets=4)
             l[0](target, num_subtargets=4)
 
 if __name__ == '__main__':
     parser = argparse.ArgumentParser()
     parser.add_argument('targets_directory', type=str,
                         help="Path to the folder containing targets.")
-    parser.add_argument('output_path', type=str,
+    parser.add_argument('output_path', type=str, default=TEMP_OUTPUT_PATH,
                         help="Path to the output folder.")
     args, _ = parser.parse_known_args()
     create_separated_subtargets(args.targets_directory, args.output_path)

### 2021-06-05 source_separation/pipeline.py

@@ -74,7 +74,7 @@ def clear_directory(path):
             os.remove(full_path)
 
 
-def separate(audio_path, model_name, num_subtargets, *args):
+def separate(audio_path, output_path, model_name, num_subtargets, *args):
     """
     Separate the audio into the given
     :param audio_path: path to audio input (wav file)
@@ -87,7 +87,7 @@ def separate(audio_path, model_name, num_subtargets, *args):
     """
 
     file_name = audio_path.split("/")[-1].split(".")[0]
-    output_path = TEMP_OUTPUT_PATH + "/" + model_name + "/" + file_name
+    output_path = output_path + "/" + model_name + "/" + file_name
 
     if not os.path.exists(output_path):
         if model_name == "TDCN++":
@@ -152,7 +152,7 @@ def gen_perm_group(l, n):
     return r
 
 
-def generate_separation_function(model_name, num_subtargets):
+def generate_separation_function(model_name, output_path, num_subtargets):
     l = []
     if model_name == "TDCN++":
         init_list = ["sub_target{}".format(x) for x in range(TDCNpp_nb_subtargets)]
@@ -168,21 +168,23 @@ def generate_separation_function(model_name, num_subtargets):
         raise Exception("Model name must be one of those five : TDCN, TDCN++, OpenUnmix, Demucs, NMF")
 
     for perm in gen_perm_group(init_list, num_subtargets):
-        l.append(lambda audio_path, num_subtargets: separate(audio_path, model_name, num_subtargets, perm))
+        l.append(lambda audio_path, num_subtargets:
+            separate(audio_path, output_path, model_name, num_subtargets, perm))
     return l
 
 
-def generate_all_separation_functions():
+def generate_all_separation_functions(output_path):
     functions = {}
     for model in separation_models:
-        functions[model] = generate_separation_function(model, NUM_SUBTARGETS)[0]
+        functions[model] = generate_separation_function(model, output_path,
+                                                        NUM_SUBTARGETS)[0]
     return functions
 
 
 # a dict of functions where each function takes two parameters
 # the first parameter is audio
 # the second parameter is the number of subtargets to separate the target into
-separation_functions = generate_all_separation_functions()
+separation_functions = generate_all_separation_functions(TEMP_OUTPUT_PATH)
 num_separation_functions = len(separation_functions)
 
 

### 2021-06-05 source_separation/demucs/separate.py
Remove prints in demucs.separate

@@ -124,7 +124,7 @@ def separate(tracks, models=Path("models"), name="demucs_quantized",
     model.to(device)
 
     out.mkdir(parents=True, exist_ok=True)
-    print(f"Separated tracks will be stored in {out.resolve()}")
+    # print(f"Separated tracks will be stored in {out.resolve()}")
     for track in tracks:
         if not track.exists():
             print(
@@ -132,7 +132,7 @@ def separate(tracks, models=Path("models"), name="demucs_quantized",
                 "please try again after surrounding the entire path with quotes \"\".",
                 file=sys.stderr)
             continue
-        print(f"Separating track {track}")
+        # print(f"Separating track {track}")
         wav = load_track(track, device, model.audio_channels, model.samplerate)
 
         ref = wav.mean(0)

### 2021-06-05 source_separation/NMF.py
Add NMF to separation

@@ -8,7 +8,6 @@ from pathlib import Path
 
 def separate(audio_path,
              output_path,
-             model_name='nmf',
              n=4,
              blocksize=4096,
              hoplen=1024,

### 2021-06-05 source_separation/create_separated_subtargets.py

@@ -9,7 +9,7 @@ from pipeline import generate_separation_function, TEMP_OUTPUT_PATH
 
 def create_separated_subtargets(targets_directory, output_path):
     for target_name in tqdm(os.listdir(targets_directory)):
-        for model in ["TDCN++", "TDCN", "OpenUnmix", "Demucs"]:
+        for model in ["TDCN++", "TDCN", "OpenUnmix", "Demucs", "NMF"]:
             print(target_name, model)
             target = targets_directory + '/' + target_name
             l = generate_separation_function(model, output_path,

### 2021-06-08 source_separation/NMF.py
Add improved version of NMF

@@ -1,40 +1,0 @@
-import numpy as np
-import soundfile as sf
-import librosa as lr
-import cmath
-import math
-from sklearn.decomposition import NMF
-from pathlib import Path
-
-def separate(audio_path,
-             output_path,
-             n=4,
-             blocksize=4096,
-             hoplen=1024,
-             maxiter=10000):
-    y, sr = lr.load(audio_path)
-    #y, sr = sf.read('coque.wav')
-    #y = np.asfortranarray(y.T[0])
-    D = lr.stft(y, n_fft=blocksize, hop_length=hoplen)
-    fftf = lr.core.fft_frequencies(sr=sr, n_fft=blocksize)
-    S, _ = lr.magphase(D)
-
-    # Calculate NMF decomposition
-    model = NMF(n_components=n, init='random', random_state=0, max_iter=maxiter)
-    W = model.fit_transform(S)
-    H = model.components_
-
-    # Output individual features & full NMF appx.
-    output = {}
-    Path(output_path).mkdir(parents=True, exist_ok=True)
-    for i in range(n):
-        S = np.array(np.matrix(W.T[i]).T * np.matrix(H[i]))
-        # Recover phase data, algorithm idea from https://dsp.stackexchange.com/questions/9877/reconstruction-of-audio-signal-from-spectrogram
-        y_out = np.random.rand(len(y))
-        for _ in range(3):
-            D_out = lr.stft(y_out, n_fft=blocksize, hop_length=hoplen)
-            D_out = S * np.exp(complex(0.0, 1.0) * np.angle(D_out))
-            y_out = lr.istft(D_out, hop_length=hoplen)
-        output[i] = y_out
-        sf.write(output_path + "/feat{}.wav".format(i+1), y_out, sr)
-    return output

### 2021-06-08 source_separation/NMF/kl_nmf.py

@@ -1,0 +1,39 @@
+# kl_nmf.py
+# From https://github.com/CarmineCella/nmf_sources
+
+import numpy as np
+
+EPS = np.finfo(np.float32).eps  # as used in sklearn (nmf.py)
+
+def kl_nmf(X, k=10, it=100):
+    # VERIFIED with sklearn.decomposition.nmf
+
+    nr = X.shape[0]
+    nc = X.shape[1]
+
+    # init matrices (sklearn method):
+    avg = np.sqrt(X.mean() / k)
+    rng = np.random.RandomState(0)  # fixed random seed
+    H = avg * rng.randn(k, nc)
+    W = avg * rng.randn(nr, k)
+    np.abs(W, W)
+    np.abs(H, H)
+
+    I = np.ones([nr,nc])
+    for i in range(it):
+        # update W
+        WH = np.dot(W,H)
+        WH[WH == 0] = EPS
+        N = np.dot(np.divide(X,WH),H.T)
+        P = np.dot(I,H.T)
+        P[P == 0] = EPS
+        W *= np.divide(N,P)
+        #update H
+        WH = np.dot(W,H)
+        WH[WH == 0] = EPS
+        N = np.dot(W.T,np.divide(X,WH))
+        P = np.dot(W.T,I)
+        P[P == 0] = EPS
+        H *= np.divide(N,P)
+
+    return W,H

### 2021-06-08 source_separation/pipeline.py

@@ -11,7 +11,7 @@ import json
 import ConvTasNetUniversal.separate as TDCNpp_separate
 import open_unmix.separate as open_unmix
 import demucs.separate as demucs
-import NMF
+import NMF.separate as NMF
 
 
 ############################################################################
@@ -163,7 +163,7 @@ def generate_separation_function(model_name, output_path, num_subtargets):
     elif model_name == "OpenUnmix":
         init_list = ["drums", "bass", "other", "vocals"]
     elif model_name == "NMF":
-        init_list = ["feat1", "feat2", "feat3", "feat4"]
+        init_list = ["source_1", "source_2", "source_3", "source_4"]
     else:
         raise Exception("Model name must be one of those five : TDCN, TDCN++, OpenUnmix, Demucs, NMF")
 

### 2021-06-08 source_separation/paper/paper.tex
Add and comment out cosine similarity equation
Reduce size of figures

@@
 
-  Source separation is an important addition to improve orchestration as it allows for the orchestration of more complex sounds in a way that makes sense in an orchestral context.
+%  Source separation is an important addition to improve orchestration as it allows for the orchestration of more complex sounds in a way that makes sense in an orchestral context.
 Orchestral music often uses multiple layers, so by separating the layers of a target and individually orchestrating them, the resulting orchestration may improve.
@@
 
-  \begin{figure*}[t]
+  \begin{figure}[t]
     \centering
-      \includegraphics[width=\textwidth]{figures/diagram.png}
+      \includegraphics[width=\columnwidth]{figures/diagram.png}
       \caption{Diagram comparing the process of orchestration with and without separation.
@@
 At the bottom, the separation process is performed, and the resulting sub-targets are individually orchestrated with subsets of the orchestra before being recombined to obtain the full orchestration.}\label{fig:full_diagram}
-  \end{figure*}
+  \end{figure}
 
@@
 
-    \begin{figure*}[t]
+    \begin{figure}[t]
       \centering
-        \includegraphics[width=\textwidth]{figures/evaluation_diagram.png}
+        \includegraphics[width=\columnwidth]{figures/evaluation_diagram.png}
         \caption{Diagram showing the testing process.
 Note that in both orchestrations that use a split orchestra, they use the same division of the orchestra to orchestrate their four sub-targets.}\label{fig:eval_diagram}
-    \end{figure*}
+    \end{figure}
 
@@
     \centering
-      \includegraphics[width=\columnwidth]{figures/orch.jpg}
+      \includegraphics[width=0.7\columnwidth]{figures/orch.jpg}
       \caption{Example waveforms showing onsets of the four sources in the target, and possible onsets in the full, separated, and ground truth orchestrations.
@@
 where $\delta_{k1} = 1 \text{  if  } x_k \ge \tilde{x}_k, 0 \text{  otherwise}$; and $\delta_{k2} = 1 \text{  if  } x_k < \tilde{x}_k, 0 \text{  otherwise}$.
+
+%Equation explanation needed
+%Equation citation needed
+%
+%\begin{equation}
+%	M(P(t),P(c)) = 1 - \left( \frac{\langle P(t), P(c) \rangle}{\|P(t)\| \cdot \|P(c)\|} \right)
+%\end{equation}
 
@@
 
-    Another aspect we would like to work on is the assignment of the orchestras.
+%    Another aspect we would like to work on is the assignment of the orchestras.
 Currently, the subset of the orchestra that is assigned to each sub-target is randomly selected.

### 2021-06-08 source_separation/pipeline.py
Fix NMF estimations names

@@ -167,7 +167,7 @@ def generate_separation_function(model_name, output_path, num_subtargets):
     elif model_name == "OpenUnmix":
         init_list = ["drums", "bass", "other", "vocals"]
     elif model_name == "NMF":
-        init_list = ["source_1", "source_2", "source_3", "source_4"]
+        init_list = ["source_0", "source_1", "source_2", "source_3"]
     else:
         raise Exception("Model name must be one of those five : TDCN, TDCN++, OpenUnmix, Demucs, NMF")
 

### 2021-06-13 source_separation/paper/paper.tex
Update paper.tex

@@
 \usepackage{multirow}
+\usepackage{amssymb}
 
-\usepackage{amssymb}
+\renewcommand{\baselinestretch}{0.96} % use the smallest value to make our paper 6 pages long (including references!)
 
@@
 Within this field, target-based computer-assisted musical orchestration is an example of how musical creativity can be supported by tools such as machine learning.
+
+	Orchestration is hard to define, however one could say that orchestration is the set of musical writing techniques that focus on timbre, or \textit{tone color} of the sound.
+A simplistic  definition of musical orchestration would be to say that it consists of the distribution of the different musical voices of a piece to the instruments of the orchestra.
 Orchestration has been traditionally taught in an intuitive and non-formalized way, and computer-assisted orchestration seeks to help composers by accelerating certain creative processes.
-In particular, target-based computer-assisted orchestration helps composers explore new timbral possibilities by combining instrumental samples in such a way as to mimic the timbre of a given target sound.
-Formally, it is the process of creating an orchestral score that best matches an arbitrary target sound given a similarity metric and constraints~\cite{Maresz2003}.
+
+	 Among the different approaches and techniques of computer-assisted orchestration, one of them, \textit{target-based orchestration}~\cite{Maresz2003}.
+Target-based computer-assisted orchestration helps composers explore new timbral possibilities by combining instrumental samples in such a way as to mimic the timbre of a given target sound.
+This target corresponds to an audio extract, of the composer's choice, which can be musical or not.
+Formally, it is the process of creating an orchestral score that best matches an arbitrary target sound given a similarity metric and constraints.
+
   Static orchestration is a type of orchestration that considers the target as a single vector of timbral features and creates only a single onset of note(s) as the solution, no matter if the target contains multiple onsets of sounds.
@@
 
-%  Source separation is an important addition to improve orchestration as it allows for the orchestration of more complex sounds in a way that makes sense in an orchestral context.
+  We think source separation could be an important addition to improve orchestration as it allows for the orchestration of more complex sounds in a way that makes sense in an orchestral context.
 Orchestral music often uses multiple layers, so by separating the layers of a target and individually orchestrating them, the resulting orchestration may improve.
@@
 
-  \begin{figure}[t]
+  \begin{figure*}[t]
+%  \begin{figure}[t]
     \centering
-      \includegraphics[width=\columnwidth]{figures/diagram.png}
+      \includegraphics[width=0.7\textwidth]{figures/diagram.png}
       \caption{Diagram comparing the process of orchestration with and without separation.
@@
 At the bottom, the separation process is performed, and the resulting sub-targets are individually orchestrated with subsets of the orchestra before being recombined to obtain the full orchestration.}\label{fig:full_diagram}
-  \end{figure}
+%  \end{figure}  
+  \end{figure*}
 
@@
     \subsection{Testing}\label{subsec:testing}
-    In order to compare the effectiveness of the separation methods, a full target orchestration and a ground truth orchestration are done for each target.
+    In order to compare the effectiveness of the separation methods, we compare them with \textit{a full target orchestration}.
 The \textbf{full target orchestration} is the orchestration of the target without any separation performed.
-The \textbf{ground truth orchestration} takes the four sources that make up the target and orchestrates each one separately, then combines the orchestrations.
-This ground truth represents the orchestration that could take place if the separation method was perfect and separated a target exactly into its constituent sources.
 
@@
 This creates the \textbf{separated orchestrations}.
-      \item The ground truth orchestration is created by orchestrating the four sources, using the same orchestra subsets from step 4, and then combined to play simultaneously.
-      \item The distances between target and orchestration are calculated for the full target orchestration, the separated orchestration, and the ground truth orchestration. 
+      \item The distances between target and orchestration are calculated for the full target orchestration and the separated orchestration.
     \end{enumerate}
@@
 
-    \begin{figure}[t]
+    \begin{figure}
     \centering
       \includegraphics[width=0.7\columnwidth]{figures/orch.jpg}
-      \caption{Example waveforms showing onsets of the four sources in the target, and possible onsets in the full, separated, and ground truth orchestrations.
+      \caption{Example waveforms showing onsets of the four sources in the target, and possible onsets in the full and separated orchestrations.
 In this example, the separation method could not distinguish between sources 2 and 3 and identified them as a single sub-target.
@@
     As described in Section \ref{subsec:data}, each target is comprised of four sources that are offset in time to begin playing at different times.
-Figure~\ref{fig:orchestrations} illustrates the effect this offset has on the full target, separated, and ground truth orchestrations.
+Figure~\ref{fig:orchestrations} illustrates the effect this offset has on the full target and separated orchestrations.
 In the full target orchestration, all onsets occur at the beginning of the orchestration.
@@
     For the separated orchestrations, the onsets of the sub-target orchestrations can be placed at the correct offset if the separation method has correctly separated out the sources.
-In the ground truth orchestration, the sources are manually offset, leading to the ground truth having each source orchestration beginning at the correct time.
-For this reason, despite only performing static orchestrations, the resulting separated and ground truth orchestrations are offset in time compared to the full target orchestration.
+For this reason, despite only performing static orchestrations, the resulting separated orchestrations are offset in time compared to the full target orchestration.
 
@@
 
-  \begin{table}[t]
+  \begin{table}
     \begin{center}
@@
         \hline
-        NMF & 22.44 \\
+        \textbf{NMF} & \textbf{22.44} \\
         \hline
@@
         Full target orchestration & 24.96 \\
-        \hline
-        Ground truth orchestration & 24.19 \\
         \hline
@@
 \ref{subsec:testing}).
-Similarly, ``Full target orchestration'' and ``Ground truth orchestration'', as defined in Sec.
-\ref{subsec:testing}, are the average values between target and orchestrated solution.
+Similarly, ``Full target orchestration'', as defined in Sec.
+\ref{subsec:testing}, is the average values between target and orchestrated solution.
 Lower values are better.}
@@
     \subsection{Results}
-    The average distances for full target, separated, and ground truth orchestrations are displayed in Table \ref{tab:distances}.
+    The average distances for full target and separated orchestrations are displayed in Table \ref{tab:distances}.
 These results are the average distance between target and orchestration for 300 different targets.
-We found that NMF performed the best out of all separation methods, and even outperformed the ground truth orchestrations, on average.
+We found that NMF performed the best out of all separation methods.
 TDCN++, Open-Unmix, and NMF performed better than the full target orchestrations, meaning that separating targets using these methods leads to a better orchestration than if no separation is applied.
+However, Demucs and TDCN performed worse.
 
@@
 The results show that performing source separation followed by a static orchestration of each individual sub-target improved the spectral distance between the original target and its orchestrated version in three out of five of the methods tested.
-This shows that source separation is a useful pre-processing stage for computer-assisted orchestration.
+We believe this shows that source separation is a useful pre-processing stage for computer-assisted orchestration.
+We also conducted a qualitative assessment of the orchestrations by acoustic inspection and found that the separated orchestrations are better from a musical standpoint.
 
@@
 It is logical that unsupervised methods will perform well for problems in which there is limited data, as they require less data than supervised methods.
-NMF also outperformed the ground truth orchestrations, which could be due to the relative simplicity of NMF's output subtargets making subsequent orchestration easier for Orchidea, but more experimentation would need to be done in order to confirm this.
 It is important to notice that for the supervised methods, we used pre-trained versions of the neural models.
@@
 
-  Since unsupervised methods perform well, it is reasonable to think that supervised methods trained on a dataset of arbitrary sounds could yield better overall performance for orchestration.
+  Since unsupervised methods perform well, it is reasonable to think that supervised methods trained on an appropriate dataset of arbitrary sounds could yield better overall performance for orchestration.
 

### 2021-06-14 source_separation/paper/paper.tex
Add suggestions from Helene-Camille

@@
 
-	Orchestration is hard to define, however one could say that orchestration is the set of musical writing techniques that focus on timbre, or \textit{tone color} of the sound.
+	Orchestration~\cite{adler1989study} is hard to define, however one could say that orchestration is the set of musical writing techniques that focus on timbre, or \textit{tone color} of the sound.
 A simplistic  definition of musical orchestration would be to say that it consists of the distribution of the different musical voices of a piece to the instruments of the orchestra.
@@
 %  The codebase for this paper can be found at: \url{https://github.com/dzluke/MLSP2021}, and you can listen to a selection of targets and orchestrations here: \url{https://dzluke.github.io/MLSP2021/}.
-  The codebase for this paper can be found at: \url{anonymous}, and you can listen to a selection of targets and orchestrations here: \url{anonymous}.
+  The codebase for this paper can be found at: \url{anonymous}, and a selection of targets and orchestrations are available for listenig at: \url{anonymous}.
 
@@
 The separated orchestrations would then be combined to create the solution.
-This would then be compared to the orchestration if no separation had been performed.
+This would then be compared to the orchestration when no separation had been performed.
 

### 2021-06-14 source_separation/paper/paper.tex
Update paper.tex

@@
 
-	 Among the different approaches and techniques of computer-assisted orchestration, one of them, \textit{target-based orchestration}~\cite{Maresz2003}.
+	 Among the different approaches and techniques of computer-assisted orchestration, one of them is \textit{target-based orchestration}~\cite{Maresz2003}.
 Target-based computer-assisted orchestration helps composers explore new timbral possibilities by combining instrumental samples in such a way as to mimic the timbre of a given target sound.
@@
 Without source separation, Orchidea would detect each note of the melody as a new onset and cut the target in time at each new note.
-However, this segmentation in time would also effect the drone, cutting it in time and forcing it to have a new onset each time the melody changes notes.
+However, this segmentation in time would also affect the drone, cutting it in time and forcing it to have a new onset each time the melody changes notes.
 The resulting orchestration would have multiple onsets for the drone, even though only one onset is needed at the beginning of the drone.

### 2021-06-14 source_separation/pipeline.py
Add clearn_orch_config method

@@ -35,7 +35,7 @@ num_orchestrations_to_save = 0
 
 ALL_ORCHESTRATIONS_PATH = "./orchestrations"  # place to save every orchestration performed
 CONFIG_PATH = "orch_config.txt"  # path to the Orchidea configuration file (you shouldn't have to change this!)
-SEPARATIONS_PATH = "/volumes/Untitled/DeepOrchestration/MLSP21/separations"
+SEPARATIONS_PATH = "/volumes/Untitled/DeepOrchestration/MLSP21/separated"
 RESULTS_PATH = "./results.json"
 TARGET_METADATA_PATH = os.path.join(TARGETS_PATH, 'metadata.json')
 TDCNpp_nb_subtargets = 4
@@ -494,6 +494,12 @@ def orchestrate_with_threshold(target, save_path):
     return solutions
 
 
+def clean_orch_config():
+    copyfile('orch_config_template.txt', CONFIG_PATH)
+    set_config_parameter(CONFIG_PATH, 'sound_paths', SOL_PATH)
+    set_config_parameter(CONFIG_PATH, 'db_files', ANALYSIS_DB_PATH)
+
+
 if __name__ == "__main__":
     distance_metric = frame_distance(cosine_similarity)  # the distance metric to be used to evaluate solutions
 
@@ -536,13 +542,8 @@ if __name__ == "__main__":
 
     targets = librosa.util.find_files(TARGETS_PATH)
     print("Database contains {} targets".format(len(targets)))
-    copyfile('orch_config_template.txt', CONFIG_PATH)
 
-    set_config_parameter(CONFIG_PATH, 'sound_paths', SOL_PATH)
-    set_config_parameter(CONFIG_PATH, 'db_files', ANALYSIS_DB_PATH)
-
-    if num_orchestrations_to_save > 0:
-        if not os.path.exists(SAVED_ORCHESTRATIONS_PATH):
+    if num_orchestrations_to_save > 0 and not os.path.exists(SAVED_ORCHESTRATIONS_PATH):
             os.mkdir(SAVED_ORCHESTRATIONS_PATH)
 
     while num_completed < len(targets):
@@ -554,6 +555,8 @@ if __name__ == "__main__":
 
         print("Target:", target_name)
         print("num completed:", num_completed)
+
+        clean_orch_config()
 
         # orchestrate full (non-separated) target with Orchidea
         print("Orchestrating full target")

### 2021-06-14 source_separation/results.py
Ensure threshold indices are only valid for 3 thresholds

@@ -88,8 +88,8 @@ if __name__ == "__main__":
             threshold_distance = [x[i] for x in full_target_distances]
             avg_threshold_distance = mean(threshold_distance)
             print("Average full target distance for threshold {}: {}".format(thresholds[i], avg_threshold_distance))
-
-        threshold_indices = [0, 40, 80]
+        if len(thresholds) == 3:
+            threshold_indices = [0, 40, 80]
         print("Average distances across thresholds per method")
         for model, distances in separated_target_distances.items():
             print("\t" + model)

### 2021-06-15 source_separation/paper/paper.tex
Add qualitative evaluation section

@@
 
-	Orchestration~\cite{adler1989study} is hard to define, however one could say that orchestration is the set of musical writing techniques that focus on timbre, or \textit{tone color} of the sound.
+	Orchestration~\cite{Adler2016} is hard to define, however one could say that orchestration is the set of musical writing techniques that focus on timbre, or \textit{tone color} of the sound.
 A simplistic  definition of musical orchestration would be to say that it consists of the distribution of the different musical voices of a piece to the instruments of the orchestra.
@@
 %  The codebase for this paper can be found at: \url{https://github.com/dzluke/MLSP2021}, and you can listen to a selection of targets and orchestrations here: \url{https://dzluke.github.io/MLSP2021/}.
-  The codebase for this paper can be found at: \url{anonymous}, and a selection of targets and orchestrations are available for listenig at: \url{anonymous}.
+  The codebase for this paper can be found at: \url{anonymous}, and a selection of targets and orchestrations are available for listening at: \url{anonymous}.
 
@@
 
-  \begin{figure*}[t]
-%  \begin{figure}[t]
+  \begin{figure}[t]
     \centering
-      \includegraphics[width=0.7\textwidth]{figures/diagram.png}
+      \includegraphics[width=\columnwidth]{figures/diagram.png}
       \caption{Diagram comparing the process of orchestration with and without separation.
@@
 At the bottom, the separation process is performed, and the resulting sub-targets are individually orchestrated with subsets of the orchestra before being recombined to obtain the full orchestration.}\label{fig:full_diagram}
-%  \end{figure}  
-  \end{figure*}
+  \end{figure}
 
@@
     \subsection{Testing}\label{subsec:testing}
-    In order to compare the effectiveness of the separation methods, we compare them with \textit{a full target orchestration}.
-The \textbf{full target orchestration} is the orchestration of the target without any separation performed.
-
     The procedure we use for testing is as follows:
@@
       \item The target is created as a combination of four randomly chosen sources, which are offset to begin playing at different times.
-      \item The target, without any separation performed, is orchestrated using the entire orchestra. This is called the full target orchestration.
+      \item The target, \textit{without any separation performed}, is orchestrated using the entire orchestra. This creates what we call the \textit{full target orchestration}.
       \item Each separation method is performed on the target, each splitting the target into four sub-targets. 
@@
       \item For each method, the four sub-targets are separately orchestrated, each using a different subset of the orchestra, and then each orchestration is combined to play simultaneously.
-This creates the \textbf{separated orchestrations}.
+This creates the \textit{separated orchestrations}.
       \item The distances between target and orchestration are calculated for the full target orchestration and the separated orchestration.
@@
 We believe this shows that source separation is a useful pre-processing stage for computer-assisted orchestration.
-We also conducted a qualitative assessment of the orchestrations by acoustic inspection and found that the separated orchestrations are better from a musical standpoint.
 
-  Furthermore, the one unsupervised method we tested, NMF, outperformed all the supervised methods.
+  In order to fully understand the impact of source separation on the orchestration process, we also conducted a qualitative evaluation of the generated orchestrations by acoustic inspection.
+This evaluation showed that source separation significantly improves the quality or orchestrations from a musical standpoint. The next paragraphs will try to explain this in more detail.
+
+	A major disadvantage of the orchestrations generated without any source separation is that, when two or more sources overlap in time, the orchestration algorithm cannot distinguish between them.
+As such, the orchestration appears to be more rigid: the whole orchestra suddenly changes configuration when a new source dominates, without any blending with the previous dominating source.
+This is against a well know orchestration principle called \emph{dovetailing} \cite[pp.
+467-473]{Adler2016}.
+Dovetailing is a well known practice aimed at creating a blend of the instruments used during orchestration, in order to transition between timbres.
+Technically, dovetailing is the swapping and overlapping of musical lines between different instruments: musical lines are scattered across multiple instruments, connected with overlapping pitches.
+
+	An interesting result that we discovered after acoustic inspection is that the orchestrations generated after source separation tend to favor dovetailing.
+Since only a portion of the orchestra is assigned to each source and since the sources are inherently asynchronous with each other, the final orchestration is more fluid and appears more interesting, from a musical standpoint.
+
+  Quantitatively, the one unsupervised method we tested, NMF, outperformed all the supervised methods.
 This suggests that methods like NMF can still be useful for specific problems when available data is limited.
@@
 Therefore, we may infer from our results that models trained on generic sounds generate a better separation for orchestration when compared to models trained only on musical sounds.
-
 Since unsupervised methods perform well, it is reasonable to think that supervised methods trained on an appropriate dataset of arbitrary sounds could yield better overall performance for orchestration.

### 2021-06-24 csmc_mume_2020/paper/JCMS_2021/paper.tex
Clarify descr of picking random samples in sec 2.2

@@
 
-For a given number of instruments $N$, each input to our model is a combination of $N$ TinySOL samples chosen among an orchestra of 10 instruments.
-The data is generated by selecting $N$ random TinySOL samples, leading to a variety of instruments, pitches, and dynamics; we did not allow the same instrument to be chosen more than three times in order to ensure variety in the mixtures.
-Then the chosen samples are combined to be played simultaneously and normalized by the number of instruments.
+For a given number of instruments $N$, each input to our model is a combination of $N$ randomly selected TinySOL samples chosen from an orchestra of 10 instruments.
+The only constraint imposed on the random selection is that there cannot be more than 3 samples from the same instrument in any one combination.
+We do this to ensure instrumental variety in the combinations.
+By randomly selecting the samples, we achieve a variety of instruments, pitches, and dynamics while avoiding introducing any bias in the combinations that could result from hand-picking the samples.
+The selected samples are combined to be played simultaneously and the magnitude is normalized by the number of instruments.
 The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly 4 seconds long.

### 2021-06-24 csmc_mume_2020/paper/JCMS_2021/paper.tex
Revisions from reviews 1 and 2

@@
 In this paper, we try a different approach to this problem by experimenting with deep neural architectures.
-The main idea is to train a model to classify combinations of real instruments and then use it for orchestration.
+Unlike Orchidea, which uses an evolutionary optimization algorithm to find a solution, we frame computer-assisted orchestration as a classification task.
+The main idea is to train a model to classify combinations of real instruments and then use this model for orchestration.
+Where Orchidea has multiple pre- and post-processing steps, we create an end-to-end network that takes a target sound as input and outputs a list of samples that can be combined to create the orchestrated solution.
 A typical solution for assisted orchestration is a set of triples \emph{instrument-pitch-dynamics} such as \{\texttt{Flute C6 pp, Bassoon C4 mf, Bassoon G4 ff}\}.
@@
 
+We chose to perform data augmentation solely for increasing the robustness of the models, not for any musical reasons.
+Data augmentation is only applied during training, so no pitch shifting or feature dropout occurs when orchestrating a target.
+
 \subsection{Baselines}
@@
 
-In order to get a better sense on the complexity of the problem, we tested three baseline classifiers: support vector machine (SVM), random forest (RF), and K-nearest neighbors (KNN).
+In order to get a better sense of the complexity of the problem, we tested three baseline classifiers: support vector machine (SVM), random forest (RF), and K-nearest neighbors (KNN).
 We used the implementations provided in the \emph{scikit-learn }library for Python \cite{scikit-learn}.
@@
 
-In this case, differently from the neural models, the features used are the MFCCs of the resulting combination; we chose to use MFCCs for this setting, instead of the Mel spectrogram, to have a lower number of features, that is more manageable by parametric classifiers.
+In this case, differently from the neural models, the features used are the Mel-frequency cepstral coefficients (MFCCs) of the resulting combination.
+We chose to use MFCCs for this setting, instead of the Mel spectrogram, to have a lower number of features which is more manageable for parametric classifiers.
 We found SVM to have the highest accuracy of the three classifiers across all experiments.
@@
 After the LSTM layer, there is a final convolutional layer yielding a tensor of dimensions $32\times 8 \times 21$.
-We flatten the outputs and feed them into a fully connected layer with Dropout, then another another fully connected layer.
+We flatten the outputs and feed them into a fully connected layer with Dropout, then another fully connected layer.
 Finally, the sigmoid function is applied to the final layer.
@@
 
-\begin{figure}
+\begin{figure}[h]
   \centering
@@
      \hline
-      \textbf{Model} & \textbf{Ob + Bn} & \textbf{Bn} & \textbf{Bass cl.} & \textbf{Bell 1} & \textbf{Bell 2} & \textbf{Multiph. 1} & \textbf{Car horn} & \textbf{Boat} $\ldots$ \\
+      \textbf{Model} & \textbf{Ob + Bn} & \textbf{Bn} & \textbf{Bass cl.} & \textbf{Bell 1} & \textbf{Bell 2} & \textbf{Multiph. 1} & \textbf{Car horn} & \textbf{Boat} \\
       \hline
-      CNN with LSTM & 0.17 & 0.28 & 0.70 & 0.55 & 0.26 & \textbf{1.10} & 0.68 & \textbf{1.12} $\ldots$  \\
+      CNN with LSTM & 0.17 & 0.28 & 0.70 & 0.55 & 0.26 & \textbf{1.10} & 0.68 & \textbf{1.12} \\
       \hline
-      ResNet & 0.34 & 0.50 & 0.48 & 0.59 & 0.45 & 0.90 & 0.49 & \textbf{1.16} $\ldots$ \\
+      ResNet & 0.34 & 0.50 & 0.48 & 0.59 & 0.45 & 0.90 & 0.49 & \textbf{1.16} \\
       \hline
       \hline
-      \textbf{$\ldots$} & \textbf{Wind harp} & \textbf{Chord 1} & \textbf{Multiph. 2} & \textbf{Chord 2} & \textbf{Gong} & \textbf{Scream} & \textbf{Brass} & \textbf{Average}  \\
+      \textbf{Model} & \textbf{Wind harp} & \textbf{Chord 1} & \textbf{Multiph. 2} & \textbf{Chord 2} & \textbf{Gong} & \textbf{Scream} & \textbf{Brass} & \textbf{Average}  \\
       \hline
-      $\ldots$ & 0.55 & 0.79 & 0.70 & 0.57 & 0.73 & \textbf{1.14} & 0.79 & 0.71\\
+      CNN with LSTM & 0.55 & 0.79 & 0.70 & 0.57 & 0.73 & \textbf{1.14} & 0.79 & 0.71\\
       \hline
-      $\ldots$  & 0.61 & 0.86 & 0.51 & 0.37 & 0.71 & \textbf{1.03} & \textbf{1.05} & 0.66 \\
+      ResNet  & 0.61 & 0.86 & 0.51 & 0.37 & 0.71 & \textbf{1.03} & \textbf{1.05} & 0.66 \\
       \hline
@@
 
-In order to have a better comparison, we did not allow Orchidea to use any of the advanced features: we did not apply any symbolic constraints or harmonic analysis and we forced it forced to use all 10 instruments in each solution.
+In order to have a better comparison, we did not allow Orchidea to use any of the advanced features: we did not apply any symbolic constraints or harmonic analysis and forced it to use all 10 instruments in each solution.
 This creates a more fair comparison, since our models are unable to create constrained or sparse solutions.
@@
 For targets that had harmonic content, it was noted if the partials present in the target were also represented in the orchestrated solution.
-For example, one of the samples of a bell had partials that loosely represented a C minor chord, so we checked if the orchestration contained the notes of the chord.
-If a target had specific notes, we identified if the note or its partials were present.
-For example there is target of an Oboe playing an A4 and a Bassoon playing a C\#3, the solution from ResNet contained the partials of the Bassoon's note: C\#3, G\#4, C\#5, and E6 were all included on various instruments in the solution.
+For example: one of the samples of a bell had partials that loosely represented a C minor chord, so we checked whether the orchestration contained the notes of this chord.
+If a target included specific notes, we identified whether the note or its partials were present.
+For example there is a target of an Oboe playing an A4 and a Bassoon playing a C\#3.
+ResNet's solution for this target contained the partials of the Bassoon's note: C\#3, G\#4, C\#5, and E6 were all included on various instruments in the solution.
 
@@
 We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets.
-Overall, CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.
+Quantitatively, CNN performs better with an average ratio of $0.71$ compared to ResNet with an average ratio of $0.66$.
+Both models outperform Orchidea on three out of fifteen of the targets.
+Through our qualitative inspection, we find that CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.
 

### 2021-06-25 csmc_mume_2020/paper/JCMS_2021/paper.tex
Revise conclusion

@@
 For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$  columns and $3.45$ rows being zero-ed out.
-This method of data augmentation aimed to be more resilient to the possible variations in the recording of the instruments.
 
@@
     \caption{Quantitative comparison of orchestrations as ratios to Orchidea.
-Eqn.
-\ref{distance} was used to compute distances between orchestrations and targets.
+Eqn.~\ref{distance} was used to compute distances between orchestrations and targets.
 What is shown is the ratio between the distance of Orchidea's solution to the target and our solution's distance to the same target.
@@
 This leads to a solution whose harmonics are much closer to the target, which is an important part of aural similarity.
-Orchidea's solutions, moreover, can use any number of instruments included in the orchestra specified by the user, thus having variable-sized orchestrations from a single instrument to the whole set (this property is called \emph{sparsity} of the solution).
+Orchidea's solutions, moreover, can use any number of instruments included in the orchestra specified by the user, thus having variable-sized orchestrations from a single instrument to the whole set (this property is called the \emph{sparsity} of the solution).
 
@@
 We then orchestrated the same targets with Orchidea and calculated the distance for Orchidea's solutions.
-A comparison of these results is in Table \ref{orch_eval}.
-While our model is not able to outperform Orchidea, it shows consistent results.
-We find that the CNN and ResNet give similar accuracies during training, but perform differently when tasked with orchestrating targets.
-Quantitatively, CNN performs better with an average ratio of $0.71$ compared to ResNet with an average ratio of $0.66$.
-Both models outperform Orchidea on three out of fifteen of the targets.
-Through our qualitative inspection, we find that CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.
+A comparison of these results is in Table~\ref{orch_eval}.
 
@@
 \label{sec:conclusions}
+
+While our models are not able to outperform Orchidea, they shows consistent results.
+During training, CNN and ResNet both perform well, with ResNet consistently achieving higher training accuracies than CNN.
+For orchestration, CNN performs quantitatively better with an average ratio of $0.71$ compared to ResNet with an average ratio of $0.66$ (see Table \ref{orch_eval}).
+Both models outperform Orchidea on three out of fifteen of the targets.
+Through our qualitative inspection, we find that CNN seems to better emulate the timbre in its orchestrations, where ResNet is better for recreating the harmonic content of the target.
 
@@
 
+Finally, the OrchideaSOL dataset could replace TinySOL as the dataset used.
+OrchideaSOL could improve the system by adding extended playing techniques to the data, which can allow noisier targets to be better orchestrated.
+OrchideaSOL contains 13,265 samples from 14 instruments, almost ten times as many samples as TinySOL~\cite{Cella2020b}.
+This would greatly increase the number of classes, and therefore we would need to train the models on more samples and possibly even use deeper models.
+
 \bibliographystyle{apacite}

### 2021-06-28 csmc_mume_2020/paper/JCMS_2021/paper.tex
Add revisions from JCMS reviews

@@
 
-The approach studied in \cite{Carpentier2010} consists in finding a good orchestration for any given sound by searching combinations of sounds from a database with a multi-objective optimization heuristics and a constraint solver that are jointly optimized.
+The approach studied in \cite{Carpentier2010} consists in finding an orchestration for any given sound by searching combinations of sounds from a database with a multi-objective optimization heuristic and a constraint solver that are jointly optimized.
 Both the target sound and the sounds in the database are embedded in a feature space defined by a fixed feature function and each generated combination of sounds is evaluated by using a specific metric.
@@
 
-In this paper, we try a different approach to this problem by experimenting with deep neural architectures.
-Unlike Orchidea, which uses an evolutionary optimization algorithm to find a solution, we frame computer-assisted orchestration as a classification task.
+In this paper, we try a different approach to this problem by framing computer-assisted orchestration as a classification task.
 The main idea is to train a model to classify combinations of real instruments and then use this model for orchestration.
+To solve this classification problem, we turn to deep neural networks, which have shown success among a variety of classification tasks.
 Where Orchidea has multiple pre- and post-processing steps, we create an end-to-end network that takes a target sound as input and outputs a list of samples that can be combined to create the orchestrated solution.
+
 A typical solution for assisted orchestration is a set of triples \emph{instrument-pitch-dynamics} such as \{\texttt{Flute C6 pp, Bassoon C4 mf, Bassoon G4 ff}\}.
@@
 
-We perform preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit in the middle and \emph{ResNet}, a well known residual architecture that already yielded good results for image classification \cite{He15}.
+We perform preliminary experiments with two deep architectures: a convolutional neural network (CNN) with a long short-term memory (LSTM) unit in the middle and \emph{ResNet}, a well known residual architecture that has yielded positive results for image classification \cite{He15}.
 We choose to use a CNN because of its success in audio classification \cite{Hershey17} and include an LSTM unit in it because of its ability to learn long term dependencies in data \cite{Hochreiter97}, which is important given the temporal nature of audio.
@@
 
-\subsection{From Orchestration to Classification}
+\subsection{From Classification to Orchestration}
 
@@
 Such setup proved to be successful in \cite{Salamon17}.
+Specifically, the Mel spectrogram uses the Mel scale to create a perceptual representation of sound.
+Since we are attempting to recreate the timbre of a sound, and timbre is a perceptual descriptor, the Mel spectrogram is a fitting representation.
 
@@
 Each sample was made by a combination of one to ten instruments and was four seconds in length.
+At the beginning of each experiment, a new set of combinations was created.
 
@@
 
-The complete problem of classification of the pairs instrument-pitch has 424 total classes.
+The complete problem of classification of the instrument-pitch pairs has 424 total classes.
 This problem is very difficult for parametric classifiers, so we initially switched to the simpler problem of classifying only the instruments and not the pitch.
@@
 If another instrument was present in the input combination that was not Violin or Viola, the classifier would simply identify that an instrument that was not one of the two was present.
-The accuracies from this experiment are outlined in Table \ref{tab:baselines}.
+A selection of the results from this experiment are outlined in Table \ref{tab:baselines}.
 Since RF performed worse than SVM in every experiment, we stopped testing with it and used SVM from that point on.
@@
 However, the parametric classifiers were unable to achieve accurate results as the number of instruments increased.
-Therefore, we did not attempt, with parametric classifiers, the full setting of the problem in which individual pitches are classified for all instruments, and instead turned to our neural models.
+Therefore, we did not attempt, with parametric classifiers, the full setting of the problem in which individual pitches are classified for all instruments, and instead turned to neural models.
 
@@
 The CNN was trained on 200,000 generated samples over 50 epochs, and ResNet was trained on 400,000 generated samples over 20 epochs.
-Fig.
-\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$.
+ResNet, being a deeper model than the CNN, requires more data per epoch during training.
+Fig.~\ref{cnn_vs_resnet} shows the best overall test accuracy achieved by both models across the number of samples $N$.
 ResNet outperforms the CNN regardless of the number of samples used in the combination.
@@
 
-While our models are not able to outperform Orchidea, they shows consistent results.
-During training, CNN and ResNet both perform well, with ResNet consistently achieving higher training accuracies than CNN.
+While our models are not able to outperform Orchidea, they show consistent results.
+During training, CNN and ResNet both perform well, with ResNet achieving higher training accuracies than CNN.
 For orchestration, CNN performs quantitatively better with an average ratio of $0.71$ compared to ResNet with an average ratio of $0.66$ (see Table \ref{orch_eval}).

### 2021-06-29 csmc_mume_2020/paper/JCMS_2021/paper.tex
Add descr of Mel spectrogram

@@
 We used the Librosa package in Python to compute the features; more details on the exact computations can be found in \cite{mcfee15}.
-The choice of the hop length was made by doing a compromise between the amount of information extracted by each FFT window, and the ability to capture temporal changes in the sound.
+The choice of the hop length is a compromise between the amount of information extracted by each FFT window, and the ability to capture temporal changes in the sound.
 
@@
 In order to test our models for orchestration, we used 15 targets from the Orchidea distribution.
-These targets represent a variety of signal types but are mostly static, in the sense that the harmonic average across time is a good representation of the sample.
+These targets represent a variety of signal types but are mostly static, meaning they do not have significant harmonic change over time.
 Some of the targets were made of instrumental samples: a sum of oboe and bassoon notes, single bassoon and bass clarinet notes, two multiphonics of bassoon and two symphonic chords.
@@
 OrchideaSOL contains 13,265 samples from 14 instruments, almost ten times as many samples as TinySOL~\cite{Cella2020b}.
-This would greatly increase the number of classes, and therefore we would need to train the models on more samples and possibly even use deeper models.
+This would greatly increase the number of classes, and therefore the models would need to be trained on more samples and the number of layers in the models may need to increase.
 

### 2022-02-22 source_separation/evaluate_separated.py
Update evaluate_separated.py

@@ -1,19 +1,21 @@
-import os
 import argparse
-
-import numpy as np
+import csv
 import json
+import os
+
 import librosa
+import mir_eval
 import museval
+import numpy as np
 
-from pipeline import separation_models, SAMPLING_RATE
+from pipeline import SAMPLING_RATE, separation_models
 
 
-def main(targets_path, subtargets_path, separated_path, output_path):
+def main(metadata_path, subtargets_path, separated_path, outdir=None):
+
     # Get metadata
-    metadata_path = os.path.join(targets_path, 'metadata.json')
-    with open(metadata_path, 'r') as metadata_file:
-        metadata = json.load(metadata_file)
+    with open(metadata_path, 'r') as f:
+        metadata = json.load(f)
 
     # Initialize results dictionary.
     results = {}
@@ -21,8 +23,9 @@ def main(targets_path, subtargets_path, separated_path, output_path):
         results[model] = []
 
     # Evaluate each target
-    for target in metadata:
-        print('Evaluating results for {}'.format(target))
+    for idx, target in enumerate(metadata):
+        print('{}/{} Evaluating results for {}'.format(idx, len(metadata),
+                                                       target))
         # Get reference sources (= padded subtargets)
         target_metadata = metadata[target]
         ref_sources = []
@@ -30,8 +33,11 @@ def main(targets_path, subtargets_path, separated_path, output_path):
             audio, _ = librosa.load(os.path.join(subtargets_path,
                                                  data['name'] + '.wav'),
                                     sr=SAMPLING_RATE)
+            audio = librosa.util.normalize(audio)
             audio = np.pad(audio, data['padding'])
             ref_sources.append(audio)
+        ref_sources = np.array(ref_sources)
+        # ref_sources = ref_sources[..., np.newaxis]  #probably unnecessary
 
         # WORKAROUND (remove eventually)
         target = target.replace('*', '_')
@@ -41,30 +47,36 @@ def main(targets_path, subtargets_path, separated_path, output_path):
             # Get estimated sources(/subtargets)
             est_path = os.path.join(separated_path, model, target)
             est_sources_files = librosa.util.find_files(est_path)
-            if len(est_sources_files) != len(target_metadata):
-                raise ValueError('The number of estimated source should match'
-                                 'the number of references sources.')
+            assert len(est_sources_files) == len(target_metadata)
+
             est_sources = []
[…]

### 2022-02-22 csmc_mume_2020/paper/JCMS_2021/paper.tex
Final JCMS changes

@@
 Target-based computer-assisted orchestration takes a target sound as an input and attempts to find instrumental samples that best match the target given a specific similarity metric and a set of constraints.
+The target can be any arbitrary sound file, it does not have to be pitched, rhythmic, or ``musical'' in any way.
 A solution to this problem is a set of orchestral scores that represent the mixtures of audio samples in the database, ranked by similarity with the target sound.
@@
 
-The choice of using the Mel spectrogram as input features for classification models is common in music information retrieval \cite{McKinney2003} and can be considered to be an appropriate representation of sound and musical signals.
-Such setup proved to be successful in \cite{Salamon17}.
-Specifically, the Mel spectrogram uses the Mel scale to create a perceptual representation of sound.
+The Mel spectrogram is a representation of sound that mirrors the non-linear way in which humans perceive frequency, which makes it a perceptual representation of sound.
 Since we are attempting to recreate the timbre of a sound, and timbre is a perceptual descriptor, the Mel spectrogram is a fitting representation.
+It is common in music information retrieval \cite{McKinney2003} and has been successfully used in other tasks that employ neural networks for sound classification \cite{Salamon17}.
 

### 2022-03-26 csmc_mume_2020/paper/JCMS_2021/paper.tex
JCMS paper camera ready edits

@@
 
+\usepackage{hyperref}
 \usepackage{amssymb}
@@
 \usepackage{subcaption}
-\captionsetup{labelfont=bf, labelsep=period,compatibility=false}
+\captionsetup{labelfont=bf, labelsep=colon,compatibility=false}
 \usepackage{apacite}
+
+% allows URLs to be broken into new lines
+\def\UrlBreaks{\do\/\do-\do\&\do.\do:}
+
 \newcommand{\keywords}[1]{\par\addvspace\baselineskip
@@
 
-\sloppypar{The codebase for this paper can be found at \url{https://github.com/dzluke/DeepOrchestration} and you can listen to audio examples at \url{https://dzluke.github.io/DeepOrchestration/}.}
+The codebase for this paper can be found at \url{https://github.com/dzluke/DeepOrchestration} and you can listen to audio examples at \url{https://dzluke.github.io/DeepOrchestration/}.
 
@@
 \begin{enumerate}
-\item we train specific models to classify the instruments present in combinations of sounds from a database of instrument notes, up to ten simultaneous instruments;
-\item we then pick the best classifier and we feed into it an unknown sound to be classified;
-\item since the output of the classifier will be in the form of the probability that specific instruments-note pairs are present in the sound, we use this information to synthesize an orchestration for the target sound;
-\item finally, we evaluate the generated orchestration against state-of-the-art systems for computer-assisted orchestrations.
+\item We train specific models to classify the instruments present in combinations of sounds from a database of instrument notes, up to ten simultaneous instruments;
+\item We then pick the best classifier and we feed into it an unknown sound to be classified;
+\item Since the output of the classifier will be in the form of the probability that specific instruments-note pairs are present in the sound, we use this information to synthesize an orchestration for the target sound;
+\item Finally, we evaluate the generated orchestration against state-of-the-art systems for computer-assisted orchestrations.
 \end{enumerate}
@@
 By taking the instruments and pitches that have the highest probability and by using quantized probabilities to retrieve dynamics, a full orchestrated solution can be created.
-Figure \ref{fig:method} illustrates the described ideas.
+Fig.
+\ref{fig:method} illustrates the described ideas.
 
@@
 Each sample is one instrument playing a single note in the \emph{ordinario} playing style, with one of three dynamics: \textit{pp}, \textit{mf}, or \textit{ff} (for example \texttt{Flute-C4-pp} or \texttt{Clarinet-D5-mf}).
-The instruments and ranges over which they were recorded are summarized in table \ref{tab:ranges}.
+The instruments and ranges over which they were recorded are summarized in Tab.
+\ref{tab:ranges}.
 
-For a given number of instruments $N$, each input to our model is a combination of $N$ randomly selected TinySOL samples chosen from an orchestra of 10 instruments.
-The only constraint imposed on the random selection is that there cannot be more than 3 samples from the same instrument in any one combination.
+For a given number of instruments $N$, each input to our model is a combination of $N$ randomly selected TinySOL samples chosen from an orchestra of ten instruments.
+The only constraint imposed on the random selection is that there cannot be more than three samples from the same instrument in any one combination.
 We do this to ensure instrumental variety in the combinations.
@@
 The selected samples are combined to be played simultaneously and the magnitude is normalized by the number of instruments.
-The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly 4 seconds long.
+The resulting combination has a sample rate of $44100$Hz and is padded or trimmed to be exactly four seconds long.
 
@@
 We chose random columns and rows of the matrix to zero out.
-For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$  columns and $3.45$ rows being zero-ed out.
+For a given matrix, each column and each row had individually a $1\%$ chance to be set to zero, which yielded an average of $1.28$  columns and $3.45$ rows being zero-ed out.
 
@@
 If another instrument was present in the input combination that was not Violin or Viola, the classifier would simply identify that an instrument that was not one of the two was present.
-A selection of the results from this experiment are outlined in Table \ref{tab:baselines}.
+A selection of the results from this experiment are outlined in Tab.
+\ref{tab:baselines}.
 Since RF performed worse than SVM in every experiment, we stopped testing with it and used SVM from that point on.
@@
 Our architecture is made of four convolutional layers and two fully connected layers.
-Each convolutional layer is followed by a BatchNorm layer, a ReLU activation layer and a $2 \times 2$  MaxPool layer with a stride of 2.
+Each convolutional layer is followed by a BatchNorm layer, a ReLU activation layer and a $2 \times 2$  MaxPool layer with a stride of two.
 The kernel size is $3 \times 3$ with a stride of 1 and a padding of 1.
-The number of filters are 8, 16, 32, and 32.
+The number of filters are eight, 16, 32, and 32.
 
@@
 Since each class is independent, we are able to take the sigmoid activation and use binary classification for each class.
-The model architecture is shown in Figure \ref{cnn_lstm}.
+The model architecture is shown in Fig.
+\ref{cnn_lstm}.
 
@@
 and $I_N(\hat{X})$ is the set of indices containing the $N$ first maximums of the vector $\hat{X}$.
-More specifically, the function $M_N(\hat{X})$ takes as an input a vector of probabilities and outputs a vector where only the positions of the $N$ first maxima are set to 1.
+More specifically, the function $M_N(\hat{X})$ takes as an input a vector of probabilities and outputs a vector where only the positions of the $N$ first maxima are set to one.
 This new vector would be the orchestration of $N$ instruments given by the model.
@@
 Fig.
-\ref{best_acc_cnn} and Fig.
-\ref{best_acc_resnet} show the maximum test accuracy for each model.
-For ResNet, the variance in accuracy is much smaller until $N$ reaches 5, at which point it becomes similar to the CNN.
+\ref{best_acc} shows the maximum test accuracy for each model.
+For ResNet, the variance in accuracy is much smaller until $N$ reaches five, at which point it becomes similar to the CNN.
 The results on both figures show consistency on the relative accuracy of instruments, which was for us the first step towards the validation of this method.
@@
   \caption{Plot showing per instrument accuracies for CNN with LSTM and ResNet depending on the number of samples present in the combinations}
+  \label{best_acc}
   \end{figure}
@@
 
-To orchestrate, a target sound is input to the model, and the 10 classes with the highest probability are extracted.
-These 10 classes are the instrument-pitch pairs that are most represented in the target, and can be from any combination of the 10 instruments.
+To orchestrate, a target sound is input to the model, and the ten classes with the highest probability are extracted.
+These ten classes are the instrument-pitch pairs that are most represented in the target, and can be from any combination of the ten instruments.
 
@@
 
-In order to have a better comparison, we did not allow Orchidea to use any of the advanced features: we did not apply any symbolic constraints or harmonic analysis and forced it to use all 10 instruments in each solution.
+In order to have a better comparison, we did not allow Orchidea to use any of the advanced features: we did not apply any symbolic constraints or harmonic analysis and forced it to use all ten instruments in each solution.
 This creates a more fair comparison, since our models are unable to create constrained or sparse solutions.

### 2022-04-19 source_separation/code/evaluate_separation.py
Separation evaluation with cosine distance

@@ -1,205 +1,93 @@
-# From https://github.com/google-research/sound-separation
-# Copyright 2020 Google LLC
-#
-# Licensed under the Apache License, Version 2.0 (the "License");
-# you may not use this file except in compliance with the License.
-# You may obtain a copy of the License at
-#
-#         https://www.apache.org/licenses/LICENSE-2.0
-#
-# Unless required by applicable law or agreed to in writing, software
-# distributed under the License is distributed on an "AS IS" BASIS,
-# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
-# See the License for the specific language governing permissions and
-# limitations under the License.
 """Metrics for source separation."""
-import itertools
 import json
 import os
-from pathlib import Path
 from argparse import ArgumentParser
 from configparser import ConfigParser
-from multiprocessing import Manager, Process, RLock
+from itertools import product, permutations
+from pathlib import Path
 
 import librosa
 import museval
 import numpy as np
-import tensorflow.compat.v1 as tf
 from tqdm import tqdm
+
+from input_output import FileStruct, load_audio_files
+from tdcn.evaluate_lib import compute_metrics
+from utils import frame_distance, cosine_similarity
 
 # Get configuration
 config = ConfigParser(inline_comment_prefixes="#")
 config.read("config.ini")
-RESULTS_FNAME = "results_museval.json"
-
-def calculate_signal_to_noise_ratio_from_power(signal_power, noise_power,
-                                               epsilon):
-    """Computes the signal to noise ratio given signal_power and noise_power.
-
-    Args:
-        signal_power: A tensor of unknown shape and arbitrary rank.
-        noise_power: A tensor matching the signal tensor.
-        epsilon: An optional float for numerical stability, since silences
-            can lead to divide-by-zero.
-
-    Returns:
-        A tensor of size [...] with SNR computed between matching slices of the
-        input signal and noise tensors.
-    """
-    # Pre-multiplication and change of logarithm base.
-    constant = tf.cast(10.0 / tf.log(10.0), signal_power.dtype)
-
-    return constant * tf.log(
-            tf.truediv(signal_power + epsilon, noise_power + epsilon))
+RESULTS_FNAME = "cosine_distance_separated.json"
 
 
-def calculate_signal_to_noise_ratio(signal, noise, epsilon=1e-8):
-    """Computes the signal to noise ratio given signal and noise.
-
-    Args:
-        signal: A [..., samples] tensor of unknown shape and arbitrary rank.
[…]

### 2022-04-19 source_separation/code/input_output.py

@@ -1,0 +1,62 @@
+import os
+
+import librosa
+import numpy as np
+
+def load_audio_files(paths, sample_rate, normalize=True, paddings=None):
+    """ Load a list of paths to audio files and return an array.
+
+    Args:
+        paths (list): List of paths to the files.
+        normalize (bool, optional): If True, normalize the audio.
+    """
+    waveforms = []
+    for i, file in enumerate(paths):
+        audio, _ = librosa.load(file, sr=sample_rate)
+        if normalize:
+            audio = librosa.util.normalize(audio)
+        if paddings:
+            audio = np.pad(audio, paddings[i])
+        waveforms.append(audio)
+    return waveforms
+
+
+class FileStruct():
+    """Inspired by MSAF
+    Holds every paths stuff for a given target.
+    """
+
+    def __init__(self, ds_path, target, target_metadata):
+        self.ds_path = ds_path
+        self.target = target
+        self.target_metadata = target_metadata
+        self.n_sources = len(target_metadata)
+
+    def get_samples_paths(self):
+        files = []
+        for data in self.target_metadata:
+            f = os.path.join(self.ds_path, 'samples', data['name'] + '.wav')
+            files.append(f)
+        return files
+
+    def get_padding(self):
+        paddings = []
+        for data in self.target_metadata:
+            paddings.append(data['padding'])
+        return paddings
+
+    def get_separated_paths(self, model_name):
+        sep_folder = os.path.join(self.ds_path, 'separated',
+                                  f'{self.n_sources}sources',
+                                  model_name,
+                                  self.target)
+        files = librosa.util.find_files(sep_folder)
+        return files
+
+    def get_mixture_path(self):
+        """Mixture? Target? Which one should I use?
+        """
+        file = os.path.join(self.ds_path, 'targets',
+                            f'{self.n_sources}sources',
+                            self.target + '.wav')
+        return file

### 2022-04-19 source_separation/code/nmf/separate.py

@@ -7,7 +7,7 @@ import soundfile as sf
 from sklearn.cluster import KMeans
 from sklearn.preprocessing import StandardScaler
 
-from kl_nmf import kl_nmf
+from nmf.kl_nmf import kl_nmf
 
 FFT_SIZE = 2048
 HOPLEN = 256
@@ -32,10 +32,11 @@ def get_sources (mix, n_sources=4, components=32, fft_size=4096, hoplen=512):
     [W, H] = kl_nmf(A, components)
 
     # masking
-    masks = np.zeros((A.shape[0],A.shape[1],COMPONENTS), dtype=complex)
+    masks = np.zeros((A.shape[0],A.shape[1], components), dtype=complex)
     comps = np.zeros(masks.shape, dtype=complex)
-    for i in range(0, COMPONENTS):
+    for i in range(0, components):
         masks[:, :, i] = np.outer(W[:, i], H[i, :]) / np.dot(W, H)
+        masks[np.isnan(masks)] = 0
         comps[:, :, i] =  masks[:, :, i] * specgram
 
     # clustering
@@ -57,12 +58,12 @@ def get_sources (mix, n_sources=4, components=32, fft_size=4096, hoplen=512):
     return W, H, clusters.labels_, sources
 
 
-def separate(in_path, out_path, n_sources=4):
+def separate(in_path, out_path, n_sources=4, components=32):
     out_path = Path(out_path)
     mix, sr = sf.read(in_path)
     W, H, labels, sources = get_sources(mix,
                                         n_sources=n_sources,
-                                        components=COMPONENTS,
+                                        components=components,
                                         fft_size=FFT_SIZE,
                                         hoplen=HOPLEN)
     out_path.mkdir(parents=True, exist_ok=True)

### 2022-04-19 source_separation/code/orchestration.py

@@ -3,12 +3,13 @@ import os
 import random
 import subprocess
 from shutil import copyfile
+import json
 
 import librosa
 import soundfile as sf
 
+from pipeline import frame_distance, spectral_distance
 from utils import set_config_parameter
-# from pipeline import SAMPLING_RATE, CONFIG_PATH
 
 # Get configuration
 config = ConfigParser(inline_comment_prefixes="#")
@@ -71,11 +72,11 @@ def orchestrate_with_threshold(target, save_path, thresholds,
         save_path = save_path / save_name
         if os.path.exists(save_path):
             # print("Found orchestration on disk")
-            solution, _ = librosa.load(save_path, sr=SAMPLING_RATE)
+            solution, _ = librosa.load(save_path, sr=sample_rate)
         else:
-            set_config_parameter(CONFIG_PATH, 'onsets_threshold', onset_threshold)
+            set_config_parameter(config_path, 'onsets_threshold', onset_threshold)
             solution = orchestrate(target)
-            sf.write(save_path, solution, samplerate=SAMPLING_RATE)
+            sf.write(save_path, solution, samplerate=sample_rate)
         save_path = save_path.parent
         solutions.append(solution)
     return solutions
@@ -85,3 +86,221 @@ def clean_orch_config(config_path, sol_path, analysis_db_path):
     copyfile('orch_config_template.txt', config_path)
     set_config_parameter(config_path, 'sound_paths', sol_path)
     set_config_parameter(config_path, 'db_files', analysis_db_path)
+
+
+if __name__ == "__main__":
+    distance_metric = frame_distance(spectral_distance)  # the distance metric to be used to evaluate solutions
+
+    num_completed = 0
+    # full_target_distances is a nested list of distances of orchestrating full targets without separation
+    # full_target_distances[i][k] is the distance between target i and the orchestration of target i with threshold k
+    full_target_distances = []
+    # distances for separated targets
+    # separated_target_distances['model'][i] is a list of length 81 of the distances for all possible combinations (3^4)
+    # of target i after being separated by 'model', orchestrated, then combined
+    separated_target_distances = {}
+    for model in config['separation']['methods'].split(", "):
+        separated_target_distances[model] = []
+    # ground_truth_distances[i] is a list of length 81, it is the distances for all possible combinations of
+    # orchestrations of the ground truth source subtargets
+    ground_truth_distances = []
+
+    # a dictionary that represents the current state of the pipeline, including how many targets have
+    # been completed, and the distances.
[…]

### 2022-04-19 source_separation/code/pipeline.py

@@ -207,221 +207,3 @@ def save_best_orchestration(solutions, distances, file_path):
     index = distances.index(min(distances))
     best_solution = solutions[index]
     sf.write(file_path, best_solution, SAMPLING_RATE)
-
-
-if __name__ == "__main__":
-    distance_metric = frame_distance(spectral_distance)  # the distance metric to be used to evaluate solutions
-
-    num_completed = 0
-    # full_target_distances is a nested list of distances of orchestrating full targets without separation
-    # full_target_distances[i][k] is the distance between target i and the orchestration of target i with threshold k
-    full_target_distances = []
-    # distances for separated targets
-    # separated_target_distances['model'][i] is a list of length 81 of the distances for all possible combinations (3^4)
-    # of target i after being separated by 'model', orchestrated, then combined
-    separated_target_distances = {}
-    for model in separation_models:
-        separated_target_distances[model] = []
-    # ground_truth_distances[i] is a list of length 81, it is the distances for all possible combinations of
-    # orchestrations of the ground truth source subtargets
-    ground_truth_distances = []
-
-    # a dictionary that represents the current state of the pipeline, including how many targets have
-    # been completed, and the distances.
-    # this is stored/loaded as a json so the process can be stopped and restarted
-    results = {'num_completed': num_completed,
-               'full_target_distances': full_target_distances,
-               'separated_target_distances': separated_target_distances,
-               'ground_truth_distances': ground_truth_distances}
-
-    # load saved results from json, or create json if none exists
-    try:
-        results_file = open(RESULTS_PATH, 'r')
-        results = json.load(results_file)
-        if len(results) != 0:  # if it's not an empty json
-            num_completed = results['num_completed']
-            full_target_distances = results['full_target_distances']
-            separated_target_distances = results['separated_target_distances']
-            ground_truth_distances = results['ground_truth_distances']
-            print("Backing up results.json to results_backup.json")
-            copyfile(RESULTS_PATH, 'results_backup.json')
-    except FileNotFoundError:
-        results_file = open(RESULTS_PATH, 'w')
-    results_file.close()
-
-    targets = librosa.util.find_files(TARGETS_PATH)
-    print("Database contains {} targets".format(len(targets)))
-
-    if num_orchestrations_to_save > 0 and not os.path.exists(SAVED_ORCHESTRATIONS_PATH):
-        os.mkdir(SAVED_ORCHESTRATIONS_PATH)
-
-    while num_completed < len(targets):
[…]

### 2022-04-19 source_separation/code/separate_targets.py

@@ -4,6 +4,7 @@ from argparse import ArgumentParser
 from configparser import ConfigParser
 
 from librosa.util import find_files
+from tqdm import tqdm
 
 import demucs.separate as demucs
 import nmf.separate as nmf
@@ -25,7 +26,8 @@ def separate(method, target_fname, outdir, n_sources, **kwargs):
         open_unmix.separate(target_fname, outdir)
     elif method == "tdcn":
         ckpt = os.path.join(kwargs['tdcn_model_path'], "baseline_model")
-        mtgph = os.path.join(kwargs['tdcn_model_path'], "baseline_inference.meta")
+        mtgph = os.path.join(kwargs['tdcn_model_path'],
+                             "baseline_inference.meta")
         tdcn.separate(target_fname, outdir, ckpt, mtgph)
 
 
@@ -35,6 +37,7 @@ if __name__ == "__main__":
                         type=str,
                         help="")
     parser.add_argument("outdir")
+    parser.add_argument("--n-sources", type=int, default=4)
 
     args, _ = parser.parse_known_args()
 
@@ -42,9 +45,7 @@ if __name__ == "__main__":
     num_subtargets = config['separation'].getint('num_subtargets)')
     if num_subtargets !=4 and ("demucs" in separation_methods
                                or "open_unmix" in separation_methods):
-        warnings.warn("Music source separation methods cannot be used "
-                    "with a number of subtargets other than 4 (Demucs, "
-                    "Open-Unmix).")
+        warnings.warn("Music source separation methods cannot be used with a number of subtargets other than 4 (Demucs, Open-Unmix).")
         if "demucs" in separation_methods:
             separation_methods.remove("demucs")
         if "open_unmix" in separation_methods:
@@ -52,7 +53,7 @@ if __name__ == "__main__":
 
     targets = find_files(args.targets_path)
 
-    for target_fname in targets:
+    for target_fname in tqdm(targets):
         for method in separation_methods:
             separate(method, target_fname, args.outdir, args.n_sources,
                      tdcn_model_path=config['paths']['tdcn_model'])

### 2022-04-19 source_separation/code/tdcn/evaluate.py

@@ -4,7 +4,7 @@
 # you may not use this file except in compliance with the License.
 # You may obtain a copy of the License at
 #
-#     https://www.apache.org/licenses/LICENSE-2.0
+#         https://www.apache.org/licenses/LICENSE-2.0
 #
 # Unless required by applicable law or agreed to in writing, software
 # distributed under the License is distributed on an "AS IS" BASIS,
@@ -27,138 +27,138 @@ from train import permutation_invariant
 
 
 def _weights_for_nonzero_refs(source_waveforms):
-  """Return shape (batch, source) weights for signals that are nonzero."""
-  source_norms = tf.sqrt(tf.reduce_mean(tf.square(source_waveforms), axis=-1))
-  return tf.greater(source_norms, 1e-8)
+    """Return shape (batch, source) weights for signals that are nonzero."""
+    source_norms = tf.sqrt(tf.reduce_mean(tf.square(source_waveforms), axis=-1))
+    return tf.greater(source_norms, 1e-8)
 
 
 def compute_perminv_sisnri(source_waveforms, separated_waveforms,
                            mixture_waveform):
-  """Compute permutation-invariant, scale-invariant SNR."""
-  perm_inv_loss = permutation_invariant.wrap(
-      lambda tar, est: -metrics.signal_to_noise_ratio_gain_invariant(est, tar))
-  source_waveforms = np.asarray(source_waveforms).astype('float32')
-  separated_waveforms = np.asarray(separated_waveforms).astype('float32')
-  _, separated_waveforms = perm_inv_loss(source_waveforms[tf.newaxis],
-                                         separated_waveforms[tf.newaxis])
-  separated_waveforms = separated_waveforms[0]
-  weights_nonzero_refs = _weights_for_nonzero_refs(source_waveforms)
-  source_waveforms = tf.boolean_mask(source_waveforms,
-                                     weights_nonzero_refs, axis=0)
-  separated_waveforms = tf.boolean_mask(separated_waveforms,
-                                        weights_nonzero_refs, axis=0)
-  sisnr_separated = metrics.signal_to_noise_ratio_gain_invariant(
-      separated_waveforms, source_waveforms)
-  num_sources = tf.reduce_sum(tf.cast(weights_nonzero_refs, tf.int32))
-  sisnr_mixture = metrics.signal_to_noise_ratio_gain_invariant(
-      tf.tile(mixture_waveform[tf.newaxis], (num_sources, 1)),
-      source_waveforms)
-  return sisnr_separated, sisnr_mixture
[…]

### 2022-04-19 source_separation/code/tdcn/evaluate_lib.py

@@ -1,0 +1,248 @@
+# Copyright 2020 Google LLC
+#
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#     https://www.apache.org/licenses/LICENSE-2.0
+#
+# Unless required by applicable law or agreed to in writing, software
+# distributed under the License is distributed on an "AS IS" BASIS,
+# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
+# See the License for the specific language governing permissions and
+# limitations under the License.
+"""Evaluate separated audio from a DCASE 2020 task 4 separation model."""
+
+import os
+
+import numpy as np
+import pandas as pd
+import tensorflow.compat.v1 as tf
+
+from . import inference
+from .train import data_io
+from .train import metrics
+from .train import permutation_invariant
+
+
+def _weights_for_nonzero_refs(source_waveforms):
+  """Return shape (source,) weights for signals that are nonzero."""
+  source_norms = tf.sqrt(tf.reduce_mean(tf.square(source_waveforms), axis=-1))
+  return tf.greater(source_norms, 1e-8)
+
+
+def _weights_for_active_seps(power_sources, power_separated):
+  """Return (source,) weights for active separated signals."""
+  min_power = tf.reduce_min(power_sources, axis=-1, keepdims=True)
+  return tf.greater(power_separated, 0.01 * min_power)
+
+
+def compute_metrics(source_waveforms, separated_waveforms, mixture_waveform):
+  """Permutation-invariant SI-SNR, powers, and under/equal/over-separation."""
+
+  # Align separated sources to reference sources.
+  perm_inv_loss = permutation_invariant.wrap(
+      lambda tar, est: -metrics.signal_to_noise_ratio_gain_invariant(est, tar))
+  _, separated_waveforms = perm_inv_loss(source_waveforms[tf.newaxis],
+                                         separated_waveforms[tf.newaxis])
+  separated_waveforms = separated_waveforms[0]  # Remove batch axis.
+
+  # Compute separated and source powers.
+  power_separated = tf.reduce_mean(separated_waveforms ** 2, axis=-1)
+  power_sources = tf.reduce_mean(source_waveforms ** 2, axis=-1)
+
+  # Compute weights for active (separated, source) pairs where source is nonzero
[…]

### 2022-04-19 source_separation/code/utils.py

@@ -78,3 +78,127 @@ def combine_with_offset(metadata):
     """
     samples = [np.pad(sample['audio'], sample['padding']) for sample in metadata]
     return combine(samples)
+
+
+def normalized_magnitude_spectrum(target, solution):
+    """
+    Given a target and solution as audio files, returns the normalized magnitude spectrum of each as np arrays
+    The normalized magnitude spectrum is the absolute value of the real FFT, normalized by dividing by its
+    largest element
+    :param target: audio, shape (len,)
+    :param solution: audio, shape (len,)
+    :return: target normalized magnitude spectrum, solution normalized magnitude spectrum
+    """
+    # if the target is longer than the solution, must trim the target
+    if target.size > solution.size:
+        target = target[:solution.size]
+        N = next_power_of_2(target.size)
+    else:
+        N = next_power_of_2(solution.size)
+
+    target_spectrum = np.abs(np.fft.rfft(target, N))
+    solution_spectrum = np.abs(np.fft.rfft(solution, N))
+
+    target_max = np.max(target_spectrum) if np.max(target_spectrum) > 0 else 1
+    solution_max = np.max(solution_spectrum) if np.max(solution_spectrum) > 0 else 1
+    target_spectrum /= (target_max)
+    solution_spectrum /= (solution_max)
+
+    return target_spectrum, solution_spectrum
+
+
+def cosine_similarity(target, solution):
+    """
+    Calculates the cosine similarity between target and solution
+    cosine_similarity(X, Y) = <X, Y> / (||X||*||Y||)
+    :param target: audio, shape: (len,)
+    :param solution: audio, shape: (len,)
+    :return: scalar representing the similarity (distance)
+    """
+    target_spectrum, solution_spectrum = normalized_magnitude_spectrum(target, solution)
+    dot_product = np.dot(target_spectrum, solution_spectrum)
+    norms = np.linalg.norm(target_spectrum) * np.linalg.norm(solution_spectrum)
+    norms = norms if norms > 0 else 1
+    similarity = dot_product / norms
+    similarity = 1 - similarity
+    return similarity
+
+
+def spectral_distance(target, solution):
+    """
+    Calculates the spectral distance between target and solution
+    :param target: audio, shape: (len,)
+    :param solution: audio, shape: (len,)
+    :return: scalar representing the distance
+    """
+    # if the target is longer than the solution, must trim the target
+    if target.size > solution.size:
+        target = target[:solution.size]
[…]

### 2022-04-20 source_separation/paper/MLSP/paper.tex
Move old paper to MLSP folder

--- 2021-06-14 source_separation/paper/paper.tex
@@
 
-\renewcommand{\baselinestretch}{0.96} % use the smallest value to make our paper 6 pages long (including references!)
+\renewcommand{\baselinestretch}{0.97} % Chercher la valeur la plus petite
 
@@
 Within this field, target-based computer-assisted musical orchestration is an example of how musical creativity can be supported by tools such as machine learning.
-
-	Orchestration~\cite{adler1989study} is hard to define, however one could say that orchestration is the set of musical writing techniques that focus on timbre, or \textit{tone color} of the sound.
-A simplistic  definition of musical orchestration would be to say that it consists of the distribution of the different musical voices of a piece to the instruments of the orchestra.
 Orchestration has been traditionally taught in an intuitive and non-formalized way, and computer-assisted orchestration seeks to help composers by accelerating certain creative processes.
-
-	 Among the different approaches and techniques of computer-assisted orchestration, one of them, \textit{target-based orchestration}~\cite{Maresz2003}.
-Target-based computer-assisted orchestration helps composers explore new timbral possibilities by combining instrumental samples in such a way as to mimic the timbre of a given target sound.
-This target corresponds to an audio extract, of the composer's choice, which can be musical or not.
-Formally, it is the process of creating an orchestral score that best matches an arbitrary target sound given a similarity metric and constraints.
-
+In particular, target-based computer-assisted orchestration helps composers explore new timbral possibilities by combining instrumental samples in such a way as to mimic the timbre of a given target sound.
+Formally, it is the process of creating an orchestral score that best matches an arbitrary target sound given a similarity metric and constraints~\cite{Maresz2003}.
 Static orchestration is a type of orchestration that considers the target as a single vector of timbral features and creates only a single onset of note(s) as the solution, no matter if the target contains multiple onsets of sounds.
@@
 
-  We think source separation could be an important addition to improve orchestration as it allows for the orchestration of more complex sounds in a way that makes sense in an orchestral context.
+%  Source separation is an important addition to improve orchestration as it allows for the orchestration of more complex sounds in a way that makes sense in an orchestral context.
 Orchestral music often uses multiple layers, so by separating the layers of a target and individually orchestrating them, the resulting orchestration may improve.
@@
 %  The codebase for this paper can be found at: \url{https://github.com/dzluke/MLSP2021}, and you can listen to a selection of targets and orchestrations here: \url{https://dzluke.github.io/MLSP2021/}.
-  The codebase for this paper can be found at: \url{anonymous}, and a selection of targets and orchestrations are available for listenig at: \url{anonymous}.
+  The codebase for this paper can be found at: \url{anonymous}, and you can listen to a selection of targets and orchestrations here: \url{anonymous}.
 
@@
 The separated orchestrations would then be combined to create the solution.
-This would then be compared to the orchestration when no separation had been performed.
+This would then be compared to the orchestration if no separation had been performed.
 
-  \begin{figure*}[t]
-%  \begin{figure}[t]
+  \begin{figure}[t]
     \centering
-      \includegraphics[width=0.7\textwidth]{figures/diagram.png}
+      \includegraphics[width=\columnwidth]{figures/diagram.png}
       \caption{Diagram comparing the process of orchestration with and without separation.
@@
     \subsection{Testing}\label{subsec:testing}
-    In order to compare the effectiveness of the separation methods, we compare them with \textit{a full target orchestration}.
+    In order to compare the effectiveness of the separation methods, a full target orchestration and a ground truth orchestration are done for each target.
 The \textbf{full target orchestration} is the orchestration of the target without any separation performed.
+The \textbf{ground truth orchestration} takes the four sources that make up the target and orchestrates each one separately, then combines the orchestrations.
+This ground truth represents the orchestration that could take place if the separation method was perfect and separated a target exactly into its constituent sources.
 
@@
 This creates the \textbf{separated orchestrations}.
-      \item The distances between target and orchestration are calculated for the full target orchestration and the separated orchestration.
+      \item The ground truth orchestration is created by orchestrating the four sources, using the same orchestra subsets from step 4, and then combined to play simultaneously.
+      \item The distances between target and orchestration are calculated for the full target orchestration, the separated orchestration, and the ground truth orchestration. 
     \end{enumerate}
@@
 
-    \begin{figure}
+    \begin{figure}[t]
     \centering
       \includegraphics[width=0.7\columnwidth]{figures/orch.jpg}
-      \caption{Example waveforms showing onsets of the four sources in the target, and possible onsets in the full and separated orchestrations.
+      \caption{Example waveforms showing onsets of the four sources in the target, and possible onsets in the full, separated, and ground truth orchestrations.
 In this example, the separation method could not distinguish between sources 2 and 3 and identified them as a single sub-target.
@@
     As described in Section \ref{subsec:data}, each target is comprised of four sources that are offset in time to begin playing at different times.
-Figure~\ref{fig:orchestrations} illustrates the effect this offset has on the full target and separated orchestrations.
+Figure~\ref{fig:orchestrations} illustrates the effect this offset has on the full target, separated, and ground truth orchestrations.
 In the full target orchestration, all onsets occur at the beginning of the orchestration.
@@
     For the separated orchestrations, the onsets of the sub-target orchestrations can be placed at the correct offset if the separation method has correctly separated out the sources.
-For this reason, despite only performing static orchestrations, the resulting separated orchestrations are offset in time compared to the full target orchestration.
+In the ground truth orchestration, the sources are manually offset, leading to the ground truth having each source orchestration beginning at the correct time.
+For this reason, despite only performing static orchestrations, the resulting separated and ground truth orchestrations are offset in time compared to the full target orchestration.
 
@@
 
-  \begin{table}
+  \begin{table}[t]
     \begin{center}
@@
         \hline
-        \textbf{NMF} & \textbf{22.44} \\
+        NMF & 22.44 \\
         \hline
@@
         Full target orchestration & 24.96 \\
+        \hline
+        Ground truth orchestration & 24.19 \\
         \hline
@@
 \ref{subsec:testing}).
-Similarly, ``Full target orchestration'', as defined in Sec.
-\ref{subsec:testing}, is the average values between target and orchestrated solution.
+Similarly, ``Full target orchestration'' and ``Ground truth orchestration'', as defined in Sec.
+\ref{subsec:testing}, are the average values between target and orchestrated solution.
 Lower values are better.}
@@
     \subsection{Results}
-    The average distances for full target and separated orchestrations are displayed in Table \ref{tab:distances}.
+    The average distances for full target, separated, and ground truth orchestrations are displayed in Table \ref{tab:distances}.
 These results are the average distance between target and orchestration for 300 different targets.
-We found that NMF performed the best out of all separation methods.
+We found that NMF performed the best out of all separation methods, and even outperformed the ground truth orchestrations, on average.
 TDCN++, Open-Unmix, and NMF performed better than the full target orchestrations, meaning that separating targets using these methods leads to a better orchestration than if no separation is applied.
-However, Demucs and TDCN performed worse.
 
@@
 The results show that performing source separation followed by a static orchestration of each individual sub-target improved the spectral distance between the original target and its orchestrated version in three out of five of the methods tested.
-We believe this shows that source separation is a useful pre-processing stage for computer-assisted orchestration.
-We also conducted a qualitative assessment of the orchestrations by acoustic inspection and found that the separated orchestrations are better from a musical standpoint.
+This shows that source separation is a useful pre-processing stage for computer-assisted orchestration.
 
@@
 It is logical that unsupervised methods will perform well for problems in which there is limited data, as they require less data than supervised methods.
+NMF also outperformed the ground truth orchestrations, which could be due to the relative simplicity of NMF's output subtargets making subsequent orchestration easier for Orchidea, but more experimentation would need to be done in order to confirm this.
 It is important to notice that for the supervised methods, we used pre-trained versions of the neural models.
@@
 
-  Since unsupervised methods perform well, it is reasonable to think that supervised methods trained on an appropriate dataset of arbitrary sounds could yield better overall performance for orchestration.
+  Since unsupervised methods perform well, it is reasonable to think that supervised methods trained on a dataset of arbitrary sounds could yield better overall performance for orchestration.
 

### 2022-04-20 source_separation/paper/paper.tex
Paper with AIMC template

--- 2021-06-14 source_separation/paper/paper.tex
@@
-% LaTeX template for MLSP papers. To be used with:
-%   * mlspconf.sty - ICASSP/ICIP LaTeX style file adapted for MLSP, and
-%   * IEEEbib.bst - IEEE bibliography style file.
-% --------------------------------------------------------------------------
-\documentclass{article}
-\usepackage{amsmath,graphicx,mlspconf}
-%%%%%%%%%%%%%%%%%%%%%%%% Some useful packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%% See related documentation %%%%%%%%%%%%%%%%%%%%%%%%%%
-\usepackage[caption=false, font=footnotesize]{subfig}% Modern replacement for subfigure package
-\usepackage{paralist}% extended list environments
-\usepackage{url}
-\usepackage{hyperref}
-\usepackage[figure,table]{hypcap}% hyperref companion
-\usepackage{multirow}
-\usepackage{amssymb}
+% License: LaTeX Project Public License 1.3c
+% file aimc2022.tex
+% This is the LaTeX source for the instructions to authors using
+% the LaTeX2e package 'aimc22.sty' for contributions to
+% the Conference on AI Music Creativity.
 
-\renewcommand{\baselinestretch}{0.96} % use the smallest value to make our paper 6 pages long (including references!)
+\documentclass[letterpaper]{article}  %TODO: make sure this is the right format
 
-% Copyright notices.
-% ------------------
-% Select one of the four copyright notices below. Only required for the camera-ready paper submission.
+% if you need to pass options to natbib, use, e.g.:
+%     \PassOptionsToPackage{numbers, compress}{natbib}
+% before loading aimc2022
+
+% load package
+\usepackage{aimc2022}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{graphicx}       % figures
+
+\bibliographystyle{unsrtnat}
+
+\title{Source Separation Methods for Computer-Assisted Orchestration}
+
+% The \author macro works with any number of authors. There are two commands
+% used to separate the names and addresses of multiple authors: \And and \AND.
 %
-% * For papers in which all authors are employed by the US government:
-%\copyrightnotice{U.S.\ Government work not protected by U.S.\ copyright}
+% Using \And between authors leaves it to LaTeX to determine where to break the
+% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
+% authors names on the first line, and the last on the second line, try using
+% \AND instead of \And before the third author name.
 
-% * For papers in which all authors are employed by a Crown government (UK, Canada, and Australia):
-%\copyrightnotice{978-1-7281-6338-3/21/\$31.00 {\copyright}2021 Crown}
-
-% * For papers in which all authors are employed by the European Union:
-%\copyrightnotice{978-1-7281-6338-3/21/\$31.00
-%{\copyright}2021 European Union}
-
-% * For all other papers:
-\copyrightnotice{978-1-7281-6338-3/21/\$31.00 {\copyright}2021 IEEE}
-
-% Header
-\toappear{2021 IEEE International Workshop on Machine Learning for Signal Processing, Oct.\ 25--28, 2021, Gold Coast, Australia}
-
-% Example definitions.
-% --------------------
-%\def\x{{\mathbf x}}
-%\def\L{{\cal L}}
-
-% Title.
-% ------
-\title{SOURCE SEPARATION METHODS FOR COMPUTER-ASSISTED ORCHESTRATION}
-%
-% Double-blind peer review.
-% -------------------------
-% Anonymize your paper for the double-blind peer-review process using the 
-% following author and affiliation.
-\name{Anonymous\thanks{Anonymous.}}
-\address{Anonymous}
-
-% Single address.
-% ---------------
-%\name{Author(s) Name(s)\thanks{Thanks to XYZ agency for funding.}}
-%\address{Author Affiliation(s)}
-
-% For example:
-% ------------
-%\address{%
-%    School \\
-%    Department \\
-%    Address
-%}
-%
-% Two addresses.
-% --------------
-%\twoauthors{%
-%    A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}
-%}{%
-%    School A-B \\
-%    Department A-B \\
-%    Address A-B \\
-%    Email A-B
-%}{%
-%   C. Author-three, D. Author-four\sthanks{The fourth author performed the work while at ...}
-%}{%
-%    School C-D \\
-%    Department C-D \\
-%    Address C-D \\
-%    Email C-D
-%}
-% 
-% Two or more addresses (alternative form).
-% -----------------------------------------
-% If you need to list more than 2 authors or the option for two options above 
-% produces a poor author block, please use the following structure:
-%\name{%
-%    Author Name$^{\star \dagger}$%
-%    \qquad Author Name$^{\star}$%
-%    \qquad Author Name$^{\dagger}$\thanks{Thanks to XYZ agency for funding.}%
-%}
-%\address{%
-%    $^{\star}$ Affiliation Number One \\%
-%    $^{\dagger}$ Affiliation Number Two%
-%}
+\author{%
+  Single Author\thanks{Use footnote for providing further information
+    about author (webpage, alternative address)---\emph{not} for acknowledging
+    funding agencies.} \\
+  Department of Music\\
+  Strawberry-Raspberry University\\
+  Pittsburgh, PA 15213 \\
+  \texttt{rhino@music.strawberry-raspberry.edu} \\
+  % examples of more authors
+  % \And
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+  % \AND
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+  % \And
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+  % \And
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+}
 
 \begin{document}
-%\ninept
 
@@
 \end{abstract}
-%
-\begin{keywords}
-computer-assisted orchestration, sound source separation, Orchidea
-\end{keywords}
-%
+
 \section{Introduction}\label{sec:introduction}
@@
 This creates the \textbf{separated orchestrations}.
-      \item The distances between target and orchestration are calculated for the full target orchestration and the separated orchestration.
+      \item The ground truth orchestration is created by orchestrating the four sources, using the same orchestra subsets from step 4, and then combined to play simultaneously.
+      \item The distances between target and orchestration are calculated for the full target orchestration, the separated orchestration, and the ground truth orchestration.
     \end{enumerate}
@@
 \end{equation}
-where $\delta_{k1} = 1 \text{  if  } x_k \ge \tilde{x}_k, 0 \text{  otherwise}$; and $\delta_{k2} = 1 \text{  if  } x_k < \tilde{x}_k, 0 \text{  otherwise}$.
+where $\delta_{k1} = 1 \textrm{  if  } x_k \ge \tilde{x}_k, 0 \textrm{  otherwise}$; and $\delta_{k2} = 1 \textrm{  if  } x_k < \tilde{x}_k, 0 \textrm{  otherwise}$.
 
@@
 
-% References should be produced using the bibtex program from suitable
-% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
-% style file from IEEE produces unsorted bibliography list.
-% -------------------------------------------------------------------------
-\bibliographystyle{IEEEbib}
-\bibliography{refs}
-
+\bibliography{bibliography}
 \end{document}
